---
title: "Alec Jones Honours Thesis Code"
author: "Alec Jones"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
   toc: true
---

   <script src="hideOutput.js"></script> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading in and cleaning the Reef Life Survey Data from Barkley Sound, BC, Canada, in 2024

## Loading in the 2024 data

```{r Reading in the 2024 RLS data, echo=FALSE}
RLS_data_2024 <- read_csv(here("./01_raw_data/RLS_CANADA_Bamfield_2024.csv"))
```

## Cleaning up the data

```{r Cleaning up the dataframe to allow for use in R}
#Making a list of the variables that I want to convert to factors in the step
#below
factor_cols2 <- c("site_code", "site_name", "Code")

#This pipe is taking the original raw dataframe and cleaning it up and storing
#it as a cleaned dataframe
RLS_data_2024_clean <- RLS_data_2024%>%
  #Picking which columns I want to keep
  select(c(`Site No.`, `Site Name`, Latitude, Longitude, Date, vis, Time, Depth,
           Method, Block, Code, Species, Total))%>%
  #Removing columns/rows with N/As
  na.omit()%>%
  #Changing all of the species codes to be lowercase as there were some of the
  #same species with codes differing by one or two capitals
  mutate(Code = str_to_lower(Code))%>%
  #Renaming the variables to be easier to code with
  rename(site_code = `Site No.`,
         site_name = `Site Name`,
         latitude = Latitude,
         longitude = Longitude,
         survey_date = Date,
         depth = Depth,
         visibility = vis,
         hour = Time,
         total = Total)%>%
  #Changing the variables that I specified earlier to be factors
   mutate_at(factor_cols2, factor)%>%
  #One point at the Kirby Site had the latitude value for both latitude and
  #longitude columns so this is fixing that
   mutate(longitude = if_else(site_code == "BMSC2", -125.19876860, longitude))%>% 
  #In the dataset there are codes and counts of non-organisms (ex. dec = Debris
  # cloth). This is removing all the non-organisms from the dataset
  filter(!Code %in% c("dez", "deg", "dem", "deo", "def", "dew", "dec", "nsf", "snd"))%>%
  #This is grouping the data down to the species code within each individual
  #transect
   group_by(site_code, site_name, latitude, longitude, survey_date, depth, 
           visibility, hour, Code)%>%
  #This is summing all of each species spotted at each transect. Ex. adding up
  #all red sea urchins at the Wouwer channel site
  summarise(species_total = sum(total))
```

## Making a dataframe with the species of interest and ratios of interest

I made a dataframe with the counts of red turban snails and red sea urchins, 
along wih the log ratios of the two species to each other.

```{r Making dataframe with urchins, snails, and ratio of the two}
#Starting pipe by taking data from the clean df I made
 urch_sn_ratio_mod_2024 <- RLS_data_2024_clean%>%
  #This makes a column for each species with the numbers of each in each
  #transect (each row)
  pivot_wider(names_from = Code, values_from = species_total)%>%
  #Renaming red sea urchin and red turban snail columns
  mutate(Urchins = mfra,
         Turban_Snails = pgi)%>%
  #Selecting only the columns I wanted and getting rid of the other species
  select(site_code, site_name, latitude, longitude, survey_date, depth, 
           visibility, hour, Urchins, Turban_Snails)%>%
  #If urchins or snails were not counted at a site they were N/A at this stage
  #so this changes those counts to 0
  replace(is.na(.), 0)%>%
  #This calculates the ratios between species and makes a column for each. Date
  #is also converted to day month year form
  mutate(urch_sn_ratio = log10((Urchins +1) / (Turban_Snails + 1)),
         sn_urch_ratio = log10((Turban_Snails + 1) / (Urchins + 1)),
         urchin_dens = (Urchins / 100),
         snail_dens = (Turban_Snails / 100),
         survey_date = dmy(survey_date))
```


# Working with the potential predictor variables

For my project I incorporated various different predictor variables, some that I
obtained from csv files and some that needed to be downloaded from online
databases.

## Predictor 1: Wave exposure (via metric called Openness)

To get a measure of wave exposure for each of the RLS sites, I used a metric 
called openness from Hill et al. (2010)(10.3354/meps08815).

### Set up prior to calculating fetch for openness

```{r Making dfs of the unique sites and then just the lat and long, echo=FALSE, include=FALSE}
#Making a df with just the unique site info
site_data_2024 <- RLS_data_2024_clean%>%
  distinct(site_code, site_name, latitude, longitude)

#Making a df with just the unique lat and longs
site_coords_2024 <- site_data_2024%>%
  ungroup()%>%
  select(latitude, longitude)%>%
  distinct()

#This converts the site coordinates into a sf object and specifies the 
#coordinate reference system to be the same as the shape file
site_sp_2024 <- st_as_sf(site_coords_2024, coords = c("longitude", "latitude"), crs = 4326)
```

```{r Loading in the shapefiles to do the fetch calculations, include=FALSE}
#This is a high resolution shape file from Hakai. This code loads in the shape
#file and saves it as the proper shape file object
VI_base_layer_hi_res <- read_sf(dsn = here("./01_raw_data/hakai_zip"),
                                stringsAsFactors = F)
read_s
VI_base_layer_hi_res_sf <- st_as_sf(VI_base_layer_hi_res)

#specifying the projection to use for the sf
st_crs(VI_base_layer_hi_res_sf) <- "+proj=longlat +datum=WGS84 +no_defs"
```

```{r Subsetting the shape file for making plots later}
#This subsets the shape file to only be of Barkley Sound to make plotting more
#efficient

#This makes coordinate limits to subset the shapefile
shp_xmin <- -125.4
shp_xmax <- -124.95
shp_ymin <- 48.7
shp_ymax <- 49.0

#This makes a bounding of the coordinates that I just made
bbox <- st_bbox(c(xmin = shp_xmin, ymin = shp_ymin, xmax = shp_xmax, ymax = shp_ymax))

#This converts the boundary box to an SFC object
bbox_sf <- st_as_sfc(bbox)

#This crops the shape file to the area of the bounding box
VI_base_layer_hi_res_sf_subset <- st_crop(VI_base_layer_hi_res_sf, bbox_sf)
plot(VI_base_layer_hi_res_sf_subset)
```

### Calculating the fetch for the sites

```{r Calculating the fetch for the sites, eval=FALSE}
#Makes an object of all the bearings to calculate fetch along
bearings_openness <- seq(7.5, 360, by = 7.5)

#Specifies to use the GEOS engine
sf::sf_use_s2(FALSE)

#This calculates the fetch for each site along each of the bearings calculated
#above. Max possible fetch is 650,000 meters. It is using the shape file from 
#above to know where land is. This will likely take a very long time to run
#so do not worry.
fetch_df_2024_openness <- fetch_len_multi(site_sp_2024,
                                          bearings = bearings_openness, 
                                          dmax = 650000, 
                                          shoreline = VI_base_layer_hi_res_sf, 
                                          projected = FALSE)

#Converting the raw fetch data into a df
fetch_df_2024_openness <- as.data.frame(fetch_df_2024_openness)

#Saving the fetch data to a .csv so that the fetch function does not need to be
#run
write_csv(fetch_df_2024_openness, here("./01_raw_data/fetch_df_2024_openness.csv"))
```

### Downloading the fetch data that is already saved

```{r}
fetch_df_2024_openness <- read_csv(here("./01_raw_data/fetch_df_2024_openness.csv"))
```


### Turning the raw fetch data into a dataframe with column labels

```{r Turning the raw openness fetch data into a dataframe with column labels}
#Making an object with all of the site code names listed to be able to make a
#column of site codes in the fetch df below
site_code <- c("BMSC1", "BMSC10", "BMSC11", "BMSC12", "BMSC13", "BMSC14", "BMSC15", "BMSC16",
               "BMSC17", "BMSC18", "BMSC19", "BMSC2", "BMSC20", "BMSC21", "BMSC22", "BMSC23", "BMSC24",
               "BMSC25", "BMSC26", "BMSC27", "BMSC3", "BMSC4", "BMSC5", "BMSC6", "BMSC8", "BMSC9")

#Adding a column with the site codes
fetch_df_2024_openness$site_code <- site_code
#Adding a column with the site lat
fetch_df_2024_openness$latitude <- site_coords_2024$latitude
#Adding a column with the site long
fetch_df_2024_openness$longitude <- site_coords_2024$longitude

#Reorganizing the order of the columns so that site code, lat, and long are 
#first
fetch_df_2024_openness_final <- fetch_df_2024_openness[, c(49, 50, 51, seq(1, 48, by = 1))]

#Converting into long format for further analysis and adding to the RLS data
fetch_df_2024_openness_long <- fetch_df_2024_openness_final%>%
  #Makes column with bearing degree and column with corresponding fetch
  pivot_longer(cols = `7.5`: `360`, 
               names_to = "degree", 
               values_to = "fetch", 
               #Removes the prefix "degree" from the values going into the 
               #degree column
               names_prefix = "degree")
```


## Predictor 2: Mean SST

To get a measure of temperature I downloaded daily satellite sst data from the 
Copernicus Marine product for the year before the sampling in Barkley Sound 
occurred.

### Loading in the temperature data and calculating the average SST

```{r Loading in the temp .nc file downloaded from Copernicus, include=FALSE}
#Loading in the data
temp_2023_2024 <- brick(here("./01_raw_data/METOFFICE-GLO-SST-L4-NRT-OBS-SST-V2_1730223986748.nc"))

#Setting the coordinate system for the temperature data
crs(temp_2023_2024) <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
```

```{r Calculating the average temperature for each pixel, echo=TRUE, results='hide'}
average_temperature <- calc(temp_2023_2024, fun = mean, na.rm = TRUE)

writeRaster(average_temperature, 
            here("./01_raw_data/average_temperature.nc"), 
            format = "CDF", 
            overwrite = TRUE)
```


## Predictor 3: Protection from fishing status

In order to use the site protection status as a predictor, we needed to find
support for the fact that there was in fact a difference in the human impacts
between protected and unprotected sites.

To do this we used DFO flyover survey data that included the positions of where
recreational fishing boats were located in Barkley Sound. 

We used the DFO flyover counts of recreational boats actively fishing to 
calculate the annual mean number of recreational boats fishing within a given 
radius of each site to support the site protection status classifications and 
show that human impacts were higher for unprotected sites. To calculate the 
annual mean number of recreational boats fishing within a given radius of each 
site, we constructed circular polygons around each site (radii = 1 km, 1.5 km, 
2 km) using R (R Core Team, 2024). We merged these polygons with the DFO 
recreational boat counts to extract how many boats were spotted within the 
radius in each year. To get the annual mean, we averaged these counts across 
the three years of data from DFO (2021, 2022, 2023). We omitted aerial surveys 
where parts of Barkley Sound were obscured by fog.

### Loading in the DFO survey data

#### Loading in the 2023 data

```{r Loading in the data}
DFO_data_raw_2023 <- read_csv(here("./01_raw_data/DFO_data/Overflight2023_ObservationPoints_20241024.csv"))

#Makes list of the variables that need to be converted to factors below
dfo_factors <- c("Day Type", "Creel Subarea")

#takes the raw data and cleans it up
rec_boats_fishing2023 <- DFO_data_raw_2023%>%
  #converts the variables specified above to factors
  mutate_at(dfo_factors, factor)%>%
  #renaming the columns to make them easier to code with
  rename(rec_boats_fishing = `1. Rec Boat Fishing`,
         rec_boats_running = `2. Rec Boats Running`,
         comm_vessel_fishing = `3. Commercial Vessel Fishing`,
         comm_vessel_not_fishing = `4. Commercial Vessel Not Fishing`,
         small_comm_vessel = `5. Small Commercial Vessel`,
         large_comm_vessel = `6. Large Commercial Vessel`,
         whale_watching = `7.  Whale Watching Boat`,
         orcas = `8. Orcas`,
         whales = `9. Whales`,
         sea_state = `10. Sea State`)%>%
  #Replacing the cells that have "Null" with 0 and converting the count columns
  #to numeric
  mutate(across(c("rec_boats_fishing", "rec_boats_running", "comm_vessel_fishing", 
                  "comm_vessel_not_fishing", "small_comm_vessel", "large_comm_vessel",
                  "whale_watching", "orcas", "whales"), ~ ifelse(. == "<Null>", 0, .)),
         across(c("rec_boats_fishing", "rec_boats_running", "comm_vessel_fishing", 
                  "comm_vessel_not_fishing", "small_comm_vessel", "large_comm_vessel",
                  "whale_watching", "orcas", "whales"), as.numeric))%>%
  ##This code will remove the dates of the flights that had fog
  filter(!Date %in% c("2023-06-23", "2023-07-16", "2023-07-31", "2023-08-06", "2023-08-28",
                      "2023-08-29", "2023-09-09", "2023-09-15"))%>%
  #Selecting only the rows where recreational boats were fishing
  filter(rec_boats_fishing != 0)%>%
  #selecting all of the columns that are needed for further analysis
    select(c(UniqueID, `Filename/FlightID`, Year, Month, Day, Date, `Critical Period`, `Day Type`, `Creel Subarea`, GMA, PFMA, 
           `Pin ID`, Longitude, Latitude, rec_boats_fishing, rec_boats_running))
```


#### Loading in the 2022 data

```{r Loading in the data}
DFO_data_raw_2022 <- read_csv(here("./01_raw_data/DFO_data/Overflight2022_ObservationPoints_20240124_COUNTS.csv"))

#dfo_factors <- c("Day Type", "Creel Subarea")

rec_boats_fishing2022 <- DFO_data_raw_2022%>%
  mutate_at(dfo_factors, factor)%>%
  rename(rec_boats_fishing = `1. Rec Boat Fishing`,
         rec_boats_running = `2. Rec Boats Running`,
         comm_vessel_fishing = `3. Commercial Vessel Fishing`,
         comm_vessel_not_fishing = `4. Commercial Vessel Not Fishing`,
         small_comm_vessel = `5. Small Commercial Vessel`,
         large_comm_vessel = `6. Large Commercial Vessel`,
         whale_watching = `7.  Whale Watching Boat`,
         orcas = `8. Orcas`,
         whales = `9. Whales`,
         sea_state = `10. Sea State`)%>%
  mutate(across(c("rec_boats_fishing", "rec_boats_running", "comm_vessel_fishing", 
                  "comm_vessel_not_fishing", "small_comm_vessel", "large_comm_vessel",
                  "whale_watching", "orcas", "whales"), ~ ifelse(. == "<Null>", 0, .)),
         across(c("rec_boats_fishing", "rec_boats_running", "comm_vessel_fishing", 
                  "comm_vessel_not_fishing", "small_comm_vessel", "large_comm_vessel",
                  "whale_watching", "orcas", "whales"), as.numeric))%>%
   ##This code will remove the dates of the flights that had fog
  filter(!Date %in% c("2022-06-18", "2022-07-31", "2022-08-24", "2022-06-04"))%>%
    filter(rec_boats_fishing != 0)%>%
  #selecting all of the columns that are needed for further analysis
    select(c(UniqueID, `Filename/FlightID`, Year, Month, Day, Date, `Critical Period`, `Day Type`, `Creel Subarea`, GMA, PFMA, 
           `Pin ID`, Longitude, Latitude, rec_boats_fishing, rec_boats_running))
```


#### Loading in the 2021 data

```{r Loading in the data}
DFO_data_raw_2021 <- read_csv(here("./01_raw_data/DFO_data/Overflight2021_ObservationPoints_20240321_counts.csv"))

#dfo_factors <- c("Day Type", "Creel Subarea")

rec_boats_fishing2021 <- DFO_data_raw_2021%>%
  mutate_at(dfo_factors, factor)%>%
  rename(rec_boats_fishing = `1. Rec Boat Fishing`,
         rec_boats_running = `2. Rec Boats Running`,
         comm_vessel_fishing = `3. Commercial Vessel Fishing`,
         comm_vessel_not_fishing = `4. Commercial Vessel Not Fishing`,
         small_comm_vessel = `5. Small Commercial Vessel`,
         large_comm_vessel = `6. Large Commercial Vessel`,
         whale_watching = `7.  Whale Watching Boat`,
         orcas = `8. Orcas`,
         whales = `9. Whales`,
         sea_state = `10. Sea State`)%>%
  mutate(across(c("rec_boats_fishing", "rec_boats_running", "comm_vessel_fishing", 
                  "comm_vessel_not_fishing", "small_comm_vessel", "large_comm_vessel",
                  "whale_watching", "orcas", "whales"), ~ ifelse(. == "<Null>", 0, .)),
         across(c("rec_boats_fishing", "rec_boats_running", "comm_vessel_fishing", 
                  "comm_vessel_not_fishing", "small_comm_vessel", "large_comm_vessel",
                  "whale_watching", "orcas", "whales"), as.numeric))%>%
   ##This code will remove the dates of the flights that had fog
  filter(!Date %in% c("2021-07-31", "2021-07-19", "2021-07-20", "2021-07-24"))%>%
   filter(rec_boats_fishing != 0)%>%
  select(c(UniqueID, `Filename/FlightID`, Year, Month, Day, Date, `Critical Period`, `Day Type`, `Creel Subarea`, GMA, PFMA, 
           `Pin ID`, Longitude, Latitude, rec_boats_fishing, rec_boats_running))
```


#### Combining the three years of data into one data frame
```{r Combining the three years of data by row}
rec_boats_fishing_total <- rbind(rec_boats_fishing2021, rec_boats_fishing2022, rec_boats_fishing2023)
```


#### Making a preliminary plot to see if it worked

```{r Seting the ranges for the maps}
#first need to specify the range of the map
lat_min = min(env_mod_data$latitude) - 0.05
lat_max = max(env_mod_data$latitude) + 0.05

long_min = min(env_mod_data$longitude) - 0.05
long_max = max(env_mod_data$longitude) + 0.05
```


```{r Plotting the 2023 rec boats as a test, echo=TRUE}
 ggplot()+
    geom_sf(data = VI_base_layer_hi_res_sf_subset, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
    geom_point(data = rec_boats_fishing2023,
             mapping = aes(x = Longitude, y = Latitude, color = rec_boats_fishing),
             size = 1.5,) +
    scale_colour_gradient(low = "blue", high = "red")+
   geom_point(data = rec_boats_fishing2023,
             mapping = aes(x = Longitude, y = Latitude, colour = rec_boats_fishing), 
              shape = 1, size = 1.5, colour = "black") +
      geom_point(data = env_mod_data,
             mapping = aes(x = longitude, y = latitude), 
              shape = 21, size = 3, colour = "black", fill = "black")+
  labs(title = "2023", x = " \nLongitude", y = "Latitude\n ", colour = "# of Rec Fishing\nBoats Observed") +
  theme.figure() +
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br")
```



# Adding the predictors to a dataframe for modelling

## Adding the temperature data

```{r Extracting temperature data from the raster}
#This pipe makes a df with the lat and long of all of the different sites
site_points_df <- urch_sn_ratio_mod_2024%>%
   ungroup()%>%
  select(longitude, latitude)
 
#This sets the long and lat as coordinates
coordinates(site_points_df) <- ~longitude+latitude

#This sets the projection of the site locations to be on the same projection
#as the average temperature raster
proj4string(site_points_df) <- crs(average_temperature)

#This extracts the average temperature for where each site is located and then 
#subtracts the constant to convert from Kelvin to celsius
sst_values <- raster::extract(average_temperature, site_points_df) - 273.15

#Makes a column with the sst values in the modelling df
urch_sn_ratio_mod_2024$sst <- sst_values
```


## Adding the photoquadrat data

One of the predictors that I used was the dominant seafloor type of each
transect. I did this by using photoquadrats and stored the data from my 
analysis in an Excel sheet. I saved this as a .csv and simply loaded this into 
R to use in the models.

### Loading in the photoquadrat data

```{r Loading in the photoquadrat data}
pq_data <- read_csv(here("./01_raw_data/pq_data.csv"))

#This specifies the only columns that I wanted for analysis abd 
pq_data_final <- pq_data%>%
  select(c(site_code, depth, dominant_score, secondary_score))%>%
  #There were multiple rows for each site corresponding to each photoquadrat 
  #so this made just one row for each site
  distinct(site_code, depth, .keep_all = TRUE)

#This merges the photoquadrat data with the modelling data based on the specific
#transects. The all.x specifies a left join.
urch_sn_ratio_mod_2024 <- merge(urch_sn_ratio_mod_2024, pq_data_final, by = c("site_code", "depth"),
                                all.x = TRUE)
```

## Adding the openness data

```{r Summing the fetch vectors for each site then dividing by the largest openness value}
#Calculates the openness of each site
openness <- fetch_df_2024_openness_long%>%
  #grouping by each site
  group_by(site_code, latitude, longitude)%>%
  #adding up all of the fetch vectors for each site
  summarise(openness = sum(fetch))
  
#identifying the largest openness score within the dataset
max_openness <- max(openness$openness)

#Making a column of normalized openness by dividing by the max observed openness
openness <- openness%>%
mutate(norm_openness = openness / max_openness)

#Merging the modelling df with the openness df by each site
urch_sn_ratio_mod_2024 <- merge(urch_sn_ratio_mod_2024, openness, by = c("site_code", "latitude", "longitude"))
```

## Adding the recreational fishing boat data

To add the recreational fishin boat data we first calculated the mean number of
recreational fishing boats fishing within various radii of each RLS site. We did
this by making circular polygons around each RLS site and then merging those 
with the rec boat points and counting how many were in each circle each year, 
and then averaging across the three years of data.

### Calculating the mean recreational boats fishing within each radius

```{r Creating a circular radius around each site}
#Taking the site locations from the model
site_locations <- st_as_sf(urch_sn_ratio_mod_2024, coords = c("longitude", "latitude"), crs = 4326)

#Picking only the unique site locations as some sites have multiple transects
#with the same coordinates
site_locations_u <- site_locations%>%
  distinct(site_code, geometry)

#Changing the projection to handle meter distances better
site_locations_u_proj_m <- st_transform(site_locations_u, crs = 3857)
#Making circles with the different radii around each site
site_locations_radius_1km <- st_buffer(x = site_locations_u_proj_m, dist = 1000)
site_locations_radius_1.5km <- st_buffer(x = site_locations_u_proj_m, dist = 1500)
site_locations_radius_2km <- st_buffer(x = site_locations_u_proj_m, dist = 2000)
```

```{r Finding the overlapping points for each site}
#making a data frame with the coordinates of all the recreational fishing boats
rec_boats_fishing_total_points <- st_as_sf(rec_boats_fishing_total, coords = c("Longitude", "Latitude"), crs = 4326)
#Converting the CRS so it matches the other parts being used
rec_boats_fishing_total_points <- st_transform(rec_boats_fishing_total_points, crs = 3857)


#Finding the fishing boats that are within the given radius and making a new df
#with this info
rec_fishing_boat_total_radius_1km <- st_intersection(site_locations_radius_1km, rec_boats_fishing_total_points)

rec_fishing_boat_total_radius_1.5km <- st_intersection(site_locations_radius_1.5km, rec_boats_fishing_total_points)

rec_fishing_boat_total_radius_2km <- st_intersection(site_locations_radius_2km, rec_boats_fishing_total_points)
```

```{r Making dataframes with boat counts summed for each year at each site}
#Here I took the data frames with the counts and points of boats within the
#radii of the sites and I summed the number of boats counted around a site
#within each year. For example, I added all of the boats around Eagle bay in
#2021

rec_fishing_boat_total_radius_1km_summed <- rec_fishing_boat_total_radius_1km%>%
  #Grouping by each site in each year
  group_by(site_code, Year)%>%
  #adding up the number of boats around a site in each year
  summarise(yearly_rec_boats_fishing = sum(rec_boats_fishing))

rec_fishing_boat_total_radius_1.5km_summed <- rec_fishing_boat_total_radius_1.5km%>%
  group_by(site_code, Year)%>%
  summarise(yearly_rec_boats_fishing = sum(rec_boats_fishing))

rec_fishing_boat_total_radius_2km_summed <- rec_fishing_boat_total_radius_2km%>%
  group_by(site_code, Year)%>%
  summarise(yearly_rec_boats_fishing = sum(rec_boats_fishing))
```

```{r Making a total dataframe for each so that years with no boats have a 0 counts}
#For some sites that had boats counted in one particular year at a given radius
#did not have counts in all years. This step made data frames with a list of all
#of the sites that had at least some counts at the given radius in one of the 
#years and then made a row for each of the three years for that site.

#Makes a row for all three years for sites that had a least one boat counted
#in one of the years
all_sites_years_1km <- expand.grid(
  site_code = unique(rec_fishing_boat_total_radius_1km$site_code),
  Year = unique(rec_fishing_boat_total_radius_1km$Year))

all_sites_years_1.5km <- expand.grid(
  site_code = unique(rec_fishing_boat_total_radius_1.5km$site_code),
  Year = unique(rec_fishing_boat_total_radius_1.5km$Year))

all_sites_years_2km <- expand.grid(
  site_code = unique(rec_fishing_boat_total_radius_2km$site_code),
  Year = unique(rec_fishing_boat_total_radius_2km$Year))
```

```{r Merging the all years df with the rec boat count sums}
#This step merged the summed boat counts at each site with the dataframe that
# had all three years listed so that there was a row with 0 for the sites which
#did not have any counts in a particular year. This was necessary for making the
#average calculation accurate.

rec_fishing_boat_total_radius_1km_complete <- all_sites_years_1km %>%
  #Joins the rec fishing boat counts within radius to the df with all the years
  #for each site by site and year
  left_join(rec_fishing_boat_total_radius_1km_summed, by = c("site_code", "Year")) %>%
  #replaces the rows that have NAs with 0
  mutate(yearly_rec_boats_fishing = replace_na(yearly_rec_boats_fishing, 0))

rec_fishing_boat_total_radius_1.5km_complete <- all_sites_years_1.5km %>%
  left_join(rec_fishing_boat_total_radius_1.5km_summed, by = c("site_code", "Year")) %>%
  mutate(yearly_rec_boats_fishing = replace_na(yearly_rec_boats_fishing, 0))

rec_fishing_boat_total_radius_2km_complete <- all_sites_years_2km %>%
  left_join(rec_fishing_boat_total_radius_2km_summed, by = c("site_code", "Year")) %>%
  mutate(yearly_rec_boats_fishing = replace_na(yearly_rec_boats_fishing, 0))
```

```{r Making the final dataframe with the average boat counts per year for each site}
#This was the final step to average the recreational boat counts at each site 
#across the three years of data


rec_fishing_boat_total_radius_1km_final <- rec_fishing_boat_total_radius_1km_complete%>%
  #groups by each site
  group_by(site_code)%>%
  #makes a new column with the mean number of boats counted within each sites 
  #radius
  summarise(mean_rec_boat_counts_1km = mean(yearly_rec_boats_fishing))

rec_fishing_boat_total_radius_1.5km_final <- rec_fishing_boat_total_radius_1.5km_complete%>%
  group_by(site_code)%>%
  summarise(mean_rec_boat_counts_1.5km = mean(yearly_rec_boats_fishing))

rec_fishing_boat_total_radius_2km_final <- rec_fishing_boat_total_radius_2km_complete%>%
  group_by(site_code)%>%
  summarise(mean_rec_boat_counts_2km = mean(yearly_rec_boats_fishing))
```


### Adding the mean rec boat counts to the modelling data frame

```{r Adding the avg counts to the modelling dataset}
#This step added each of the data frames with the average rec boat counts
# to the modelling dataset. The resulting dataset has a column for avg 
#rec boat counts for each radius distance
urch_sn_ratio_mod_2024 <- urch_sn_ratio_mod_2024%>%
 left_join(rec_fishing_boat_total_radius_1km_final, by = c("site_code"))%>%
 left_join(rec_fishing_boat_total_radius_1.5km_final, by = c("site_code"))%>%
  left_join(rec_fishing_boat_total_radius_2km_final, by = c("site_code"))%>%
  replace(is.na(.), 0)
```


## Adding the region classifications for use as a random/fixed effect

In order to use region as a potential random or fixed effect during modelling
I made region classifications for the sites depending on which Island group/area
they were in. I made a few different region classifications, but ultimately we
went with the region1 classification because it made the most logical sense.

```{r Making a region column}
urch_sn_ratio_mod_2024 <- urch_sn_ratio_mod_2024%>%
  #makes a column for the region and then depending if the site code is
  #specified it puts the given region names below
  mutate(region1 = case_when(
    site_code %in% c("BMSC1", "BMSC10", "BMSC11", "BMSC12", "BMSC19", "BMSC2", "BMSC20", 
                     "BMSC21", "BMSC22", "BMSC23", "BMSC25", "BMSC3", "BMSC4", "BMSC5", "BMSC9") ~ "Deer group/Bamfield",
   site_code %in% c("BMSC13", "BMSC14", "BMSC15", "BMSC16", "BMSC17", "BMSC18") ~ "Broken Group", 
    site_code %in% c("BMSC24", "BMSC6", "BMSC8") ~ "Deep Imperial Eagle Channel",
    site_code %in% c("BMSC26", "BMSC27") ~ "Sarita"
  ))%>%
   mutate(region2 = case_when(
    site_code %in% c("BMSC1", "BMSC10", "BMSC11", "BMSC12", "BMSC19", "BMSC2", "BMSC20",
                     "BMSC3",  "BMSC5") ~ "Deer group",
    site_code %in% c("BMSC21", "BMSC22", "BMSC23", "BMSC25", "BMSC4", "BMSC9") ~ "Bamfield",
   site_code %in% c("BMSC13", "BMSC14", "BMSC15", "BMSC16", "BMSC17", "BMSC18") ~ "Broken Group", 
    site_code %in% c("BMSC24", "BMSC6", "BMSC8") ~ "Deep Imperial Eagle Channel",
    site_code %in% c("BMSC26", "BMSC27") ~ "Sarita"
  ))%>%
   mutate(region3 = case_when(
     site_code%in% c("BMSC19", "BMSC2", "BMSC20", "BMSC3", "BMSC24", "BMSC6", "BMSC8") ~ "Imperial Eagle",
    site_code %in% c("BMSC1", "BMSC10", "BMSC11", "BMSC12", 
                     "BMSC21", "BMSC22", "BMSC23", "BMSC25", "BMSC26", "BMSC27",
                     "BMSC4", "BMSC5", "BMSC9") ~ "Trevor Channel",
   site_code %in% c("BMSC13", "BMSC14", "BMSC15", "BMSC16", "BMSC17", "BMSC18") ~ "Broken Group", 
  ))%>%
  mutate(region4 = case_when(
    site_code %in% c("BMSC21", "BMSC22", "BMSC23", "BMSC25", "BMSC4", "BMSC9", "BMSC26", "BMSC27") ~ "Bamfield",
    site_code %in% c("BMSC13", "BMSC14", "BMSC15", "BMSC16", "BMSC17", "BMSC18") ~ "Broken Group",
    site_code %in% c("BMSC24", "BMSC6", "BMSC8") ~ "Deep Imperial Eagle Channel",
    site_code %in% c("BMSC1", "BMSC10", "BMSC11", "BMSC12", "BMSC19", "BMSC2", "BMSC20",
                     "BMSC3",  "BMSC5") ~ "Deer group"
  ))
```


## Writing the modelling data frame to a .csv for future use
```{r Writing the modelling df}
#The most recent version of this has the total boat counts added at the three radii
#Jan. 21
write_csv(urch_sn_ratio_mod_2024,
          here("./03_clean_data/urchin_snail_modelling_data.csv"))
```


# Doing modelling for the project

## Loading in the modelling data and cleaning it up

```{r Loading in the data, include=FALSE}
env_mod_data <- read_csv(here("./03_clean_data/urchin_snail_modelling_data.csv"))
```


```{r Cleaning up the data, include=FALSE}
#As the temp and openness data are on different scales we scaled them so as not
#to have the different scales influence the results
#These made columns with the scaled data
env_mod_data$sst_scale <- scale(env_mod_data$sst)
env_mod_data$openness_scale <- scale(env_mod_data$openness)
env_mod_data$norm_openness_scale <- scale(env_mod_data$norm_openness)

#We also tried some modelling with the recreational boats so made scaled columns
#for them as well
env_mod_data$mean_rec_boat_counts_1km_scale <- scale(env_mod_data$mean_rec_boat_counts_1km)
env_mod_data$mean_rec_boat_counts_1.5km_scale <- scale(env_mod_data$mean_rec_boat_counts_1.5km)
env_mod_data$mean_rec_boat_counts_2km_scale <- scale(env_mod_data$mean_rec_boat_counts_2km)

#making dominant seafloor score a factor
env_mod_data$dominant_score <- as.factor(env_mod_data$dominant_score)
#same for site code
env_mod_data$site_code <- as.factor(env_mod_data$site_code)
#tried using the different species as predictors for each other so made scaled
#columns of each
env_mod_data$Urchins_scale <- scale(env_mod_data$Urchins)
env_mod_data$Turban_Snails_scale <- scale(env_mod_data$Turban_Snails)

#Renaming the levels of the seafloor type to loose and solid
levels(env_mod_data$dominant_score) <- list(loose = "3", solid = "4")

#Making the region columns factors, first by making an object with their names
#and then using lapply to do it to all of them at once
regions <- c("region1", "region2", "region3", "region4")
env_mod_data[regions] <- lapply(env_mod_data[regions], factor)

## Reordering the factor levels of region1 so that Deer group bamfield is the
#intercept and renaming for plotting
levels(env_mod_data$region1) <- list(DGB = "Deer group/Bamfield",
                                     BG = "Broken Group",
                                     DIEC = "Deep Imperial Eagle Channel",
                                     SR = "Sarita" )
```


```{r Creating a factor variable of protection status for the modelling data, include=FALSE}
#Making an object with the protection status of the sites coded as 0 for not
#protected and 1 for protected from fishing
protect_status <- c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 
                    0, 0, 0, 1, 1, 0, 0, 0, 0, 0, 1, 1, 1, 1, 0)

#Making protect status into a column in the DF and as a factor
  env_mod_data$protect_status <- as.factor(protect_status)
  #Changing the levels to No Protection and Protection instead of 0 and 1
levels(env_mod_data$protect_status) <- c("No Protection", "Protected")

#Reordering so easier to see which sites are protected in the DF
########This needs to be fixed for the final markdown
env_mod_data <- env_mod_data[, c(seq(1, 11, by = 1), 46, 47, seq(12, 45, by = 1))]
env_mod_data <- env_mod_data[, c(seq(1, 10, by = 1), 46, seq(11, 45), 47 )]
```


## Exploratory plots of the data
Making some exploratory plots of the predictors against the response variables
(urchin dens, snail dens, snail to urchin ratio).

```{r Making theme for all subsequent figures, echo=FALSE}
theme.figure <- function (){
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    #axis.line = element_line(colour = "black"),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12, face = "plain"),
    legend.text = element_text(size = 10, face = "plain"),
    legend.title = element_text(size = 14, face = "bold")) 
}

```


### The EVs against the snail density

```{r snail dens against substrate score, echo=FALSE}
ggplot(env_mod_data, aes(x = dominant_score, y = snail_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r snail dens against openness,echo=FALSE}
ggplot(env_mod_data, aes(x = norm_openness, y = snail_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r snail dens against sst, echo=FALSE}
ggplot(env_mod_data, aes(x = sst, y = snail_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r snail dens against mean rec boat counts 1.5km, echo=FALSE}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_1.5km, y = snail_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r snail dens against mean rec boat counts 1km, echo=FALSE}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_1km, y = snail_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r snail dens against mean rec boat counts 2km, echo=FALSE}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_2km, y = snail_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

### The EVs against the urchin density

```{r urchin dens against substrate score, echo=FALSE}
ggplot(env_mod_data, aes(x = dominant_score, y = urchin_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r urchin dens against openness,echo=FALSE}
ggplot(env_mod_data, aes(x = norm_openness, y = urchin_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r urchin dens against sst, echo=FALSE}
ggplot(env_mod_data, aes(x = sst, y = urchin_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r urchin dens against mean rec boat counts 1.5km, echo=FALSE}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_1.5km, y = urchin_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r urchin dens against mean rec boat counts 1km, echo=FALSE}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_1km, y = urchin_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r urchin dens against mean rec boat counts 2km, echo=FALSE}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_2km, y = urchin_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

### The EVs against the snail to urchin ratio
```{r ratio against substrate score, echo=FALSE}
ggplot(env_mod_data, aes(x = dominant_score, y = sn_urch_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against openness,echo=FALSE}
ggplot(env_mod_data, aes(x = norm_openness, y = sn_urch_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against sst, echo=FALSE}
ggplot(env_mod_data, aes(x = sst, y = sn_urch_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against mean rec boat counts 1.5km, echo=FALSE}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_1.5km, y = sn_urch_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against mean rec boat counts 1km, echo=FALSE}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_1km, y = sn_urch_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against mean rec boat counts 2km, echo=FALSE}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_2km, y = sn_urch_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```



## Ecological models

The first set of models that we tried were generalized additive models to test
the ecological predictors on the densities of each species to see what is 
driving their populations.

### Red Turban Snail Ecology models

We started with the turban snail models. 

#### Model 1: Ecology model on the snail density with temperature, openness, and bottom type

```{r Turban snail ecology/environmental model}
#We initially tested this model with the untransformed snail density, but the 
#residuals were not good. We could not use Poisson or Negative binomial as there
#were decimals. We tried Gamma with adding a small constant to get rid of the 0s
#but this did not work well. Also tried Tweedie and scat but again did not look 
#good. At this point we decided to log transform
#the snail density. We needed to add a constant to the density because there
#were some zeroes so we instead of adding 1, we added 0.01 because the snail
#densities were the snail counts divided by 100.
ecology_snail_mod <- gam(log(snail_dens + 0.01) ~ s(sst_scale)+
                               s(norm_openness_scale)+
                           dominant_score,
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                         #we picked REML for smoothing parameter estimation as
                         #this is considered to be one of the best methods
                             method = "REML",
                         #we set select to TRUE meaning that the model can fully
                         #penalize smooth terms to zero effect if that is what 
                         #is suggested
                             select = TRUE)
```

```{r Turban snail ecology/environmental model output, echo=FALSE}
#This gives some summary output about the fitting of the smooth and how many edf
#are used. It also gives residual plots to look at the fit.
gam.check(ecology_snail_mod)

#this does a statistical test of the normality of the residuals
shapiro.test(ecology_snail_mod$residuals)

#This plots all of the partial effects plots so you can see the relationship
#with each of the variables
plot(ecology_snail_mod, all.terms = TRUE)

#This gives the summary of the model
summary(ecology_snail_mod)
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail ecology model residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_ecology_snail_mod <- residuals(ecology_snail_mod, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_ecology_snail_mod, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_ecology_snail_mod, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation for both methods when using a two sided test.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing turban snail ecology model for concurvity, echo=FALSE}
concurvity(ecology_snail_mod, full = FALSE)
concurvity(ecology_snail_mod, full = TRUE)
```
</div>

Temperature and openness seem to be somewhat concurve. Temp having no impact on
the model (being removed by select).