---
title: "Alec Jones Honours Thesis Code"
author: "Alec Jones"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
   toc: true
---

   <script src="hideOutput.js"></script> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(mgcv)
library(DHARMa)
library(mgcViz)
library(ggeffects)
library(gratia)
library(geoR)
library(geosphere)
library(gstat)
library(sf)
library(car)
library(MuMIn)
library(lme4)
library(spdep)
library(ggpubr)
library(glmmTMB)
library(emmeans)
library(conflicted)
library(performance)
library(ufs)
library(ggspatial)
library(ggrepel)
library(cowplot)
library(sp)
library(raster)
library(waver)
#library(patchwork)

conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
```

# Loading in and cleaning the Reef Life Survey Data from Barkley Sound, BC, Canada, in 2024

## Loading in the 2024 data

```{r Reading in the 2024 RLS data, echo=FALSE}
RLS_data_2024 <- read_csv(here("./01_raw_data/RLS_CANADA_Bamfield_2024.csv"))
```

## Cleaning up the data

```{r Cleaning up the dataframe to allow for use in R}
#Making a list of the variables that I want to convert to factors in the step
#below
factor_cols2 <- c("site_code", "site_name", "Code")

#This pipe is taking the original raw dataframe and cleaning it up and storing
#it as a cleaned dataframe
RLS_data_2024_clean <- RLS_data_2024%>%
  #Picking which columns I want to keep
  select(c(`Site No.`, `Site Name`, Latitude, Longitude, Date, vis, Time, Depth,
           Method, Block, Code, Species, Total))%>%
  #Removing columns/rows with N/As
  na.omit()%>%
  #Changing all of the species codes to be lowercase as there were some of the
  #same species with codes differing by one or two capitals
  mutate(Code = str_to_lower(Code))%>%
  #Renaming the variables to be easier to code with
  rename(site_code = `Site No.`,
         site_name = `Site Name`,
         latitude = Latitude,
         longitude = Longitude,
         survey_date = Date,
         depth = Depth,
         visibility = vis,
         hour = Time,
         total = Total)%>%
  #Changing the variables that I specified earlier to be factors
   mutate_at(factor_cols2, factor)%>%
  #One point at the Kirby Site had the latitude value for both latitude and
  #longitude columns so this is fixing that
   mutate(longitude = if_else(site_code == "BMSC2", -125.19876860, longitude))%>% 
  #In the dataset there are codes and counts of non-organisms (ex. dec = Debris
  # cloth). This is removing all the non-organisms from the dataset
  filter(!Code %in% c("dez", "deg", "dem", "deo", "def", "dew", "dec", "nsf", "snd"))%>%
  #This is grouping the data down to the species code within each individual
  #transect
   group_by(site_code, site_name, latitude, longitude, survey_date, depth, 
           visibility, hour, Code)%>%
  #This is summing all of each species spotted at each transect. Ex. adding up
  #all red sea urchins at the Wouwer channel site
  summarise(species_total = sum(total))
```

## Making a dataframe with the species of interest and ratios of interest

I made a dataframe with the counts of red turban snails and red sea urchins, 
along wih the log ratios of the two species to each other.

```{r Making dataframe with urchins, snails, and ratio of the two}
#Starting pipe by taking data from the clean df I made
 urch_sn_ratio_mod_2024 <- RLS_data_2024_clean%>%
  #This makes a column for each species with the numbers of each in each
  #transect (each row)
  pivot_wider(names_from = Code, values_from = species_total)%>%
  #Renaming red sea urchin and red turban snail columns
  mutate(Urchins = mfra,
         Turban_Snails = pgi)%>%
  #Selecting only the columns I wanted and getting rid of the other species
  select(site_code, site_name, latitude, longitude, survey_date, depth, 
           visibility, hour, Urchins, Turban_Snails)%>%
  #If urchins or snails were not counted at a site they were N/A at this stage
  #so this changes those counts to 0
  replace(is.na(.), 0)%>%
  #This calculates the ratios between species and makes a column for each. Date
  #is also converted to day month year form
  mutate(urch_sn_ratio = log10((Urchins +1) / (Turban_Snails + 1)),
         sn_urch_ratio = log10((Turban_Snails + 1) / (Urchins + 1)),
         urchin_dens = (Urchins / 100),
         snail_dens = (Turban_Snails / 100),
         survey_date = dmy(survey_date))
```


# Working with the potential predictor variables

For my project I incorporated various different predictor variables, some that I
obtained from csv files and some that needed to be downloaded from online
databases.

## Predictor 1: Wave exposure (via metric called Openness)

To get a measure of wave exposure for each of the RLS sites, I used a metric 
called openness from Hill et al. (2010)(10.3354/meps08815).

### Set up prior to calculating fetch for openness

```{r Making dfs of the unique sites and then just the lat and long, echo=FALSE, include=FALSE}
#Making a df with just the unique site info
site_data_2024 <- RLS_data_2024_clean%>%
  distinct(site_code, site_name, latitude, longitude)

#Making a df with just the unique lat and longs
site_coords_2024 <- site_data_2024%>%
  ungroup()%>%
  select(latitude, longitude)%>%
  distinct()

#This converts the site coordinates into a sf object and specifies the 
#coordinate reference system to be the same as the shape file
site_sp_2024 <- st_as_sf(site_coords_2024, coords = c("longitude", "latitude"), crs = 4326)
```

```{r Loading in the shapefiles to do the fetch calculations, include=FALSE}
#This is a high resolution shape file from Hakai. This code loads in the shape
#file and saves it as the proper shape file object
VI_base_layer_hi_res <- read_sf(dsn = here("./01_raw_data/hakai_zip"),
                                stringsAsFactors = F)

VI_base_layer_hi_res_sf <- st_as_sf(VI_base_layer_hi_res)

#specifying the projection to use for the sf
st_crs(VI_base_layer_hi_res_sf) <- "+proj=longlat +datum=WGS84 +no_defs"



#This subsets the shape file to only be of Barkley Sound to make plotting more
#efficient

#This makes coordinate limits to subset the shapefile
shp_xmin <- -125.4
shp_xmax <- -124.95
shp_ymin <- 48.7
shp_ymax <- 49.0

#This makes a bounding of the coordinates that I just made
bbox <- st_bbox(c(xmin = shp_xmin, ymin = shp_ymin, xmax = shp_xmax, ymax = shp_ymax))

#This converts the boundary box to an SFC object
bbox_sf <- st_as_sfc(bbox)

#This crops the shape file to the area of the bounding box
VI_base_layer_hi_res_sf_subset <- st_crop(VI_base_layer_hi_res_sf, bbox_sf)
#plot(VI_base_layer_hi_res_sf_subset)
```

### Calculating the fetch for the sites

```{r Calculating the fetch for the sites, eval=FALSE}
#Makes an object of all the bearings to calculate fetch along
bearings_openness <- seq(7.5, 360, by = 7.5)

#Specifies to use the GEOS engine
sf::sf_use_s2(FALSE)

#This calculates the fetch for each site along each of the bearings calculated
#above. Max possible fetch is 650,000 meters. It is using the shape file from 
#above to know where land is. This will likely take a very long time to run
#so do not worry.
#I have ## this code out because it takes way too long to run. There is code in
#the chunk below that will load in the data from a .csv that has already been
#made by running the fetch_len_multi function

#fetch_df_2024_openness <- fetch_len_multi(site_sp_2024,
                                        #  bearings = bearings_openness, 
                                        #  dmax = 650000, 
                                        #  shoreline = VI_base_layer_hi_res_sf, 
                                         # projected = FALSE)

#Converting the raw fetch data into a df
#fetch_df_2024_openness <- as.data.frame(fetch_df_2024_openness)

#Saving the fetch data to a .csv so that the fetch function does not need to be
#run
#write_csv(fetch_df_2024_openness, here("./01_raw_data/fetch_df_2024_openness.csv"))
```

### Downloading the fetch data that is already saved

```{r Loading in the pre calculated fetch data, echo=FALSE}
fetch_df_2024_openness <- read_csv(here("./01_raw_data/fetch_df_2024_openness.csv"))
```


### Turning the raw fetch data into a dataframe with column labels

```{r Turning the raw openness fetch data into a dataframe with column labels}
#Making an object with all of the site code names listed to be able to make a
#column of site codes in the fetch df below
site_code <- c("BMSC1", "BMSC10", "BMSC11", "BMSC12", "BMSC13", "BMSC14", "BMSC15", "BMSC16",
               "BMSC17", "BMSC18", "BMSC19", "BMSC2", "BMSC20", "BMSC21", "BMSC22", "BMSC23", "BMSC24",
               "BMSC25", "BMSC26", "BMSC27", "BMSC3", "BMSC4", "BMSC5", "BMSC6", "BMSC8", "BMSC9")

#Adding a column with the site codes
fetch_df_2024_openness$site_code <- site_code
#Adding a column with the site lat
fetch_df_2024_openness$latitude <- site_coords_2024$latitude
#Adding a column with the site long
fetch_df_2024_openness$longitude <- site_coords_2024$longitude

#Reorganizing the order of the columns so that site code, lat, and long are 
#first
fetch_df_2024_openness_final <- fetch_df_2024_openness[, c(49, 50, 51, seq(1, 48, by = 1))]

#Converting into long format for further analysis and adding to the RLS data
fetch_df_2024_openness_long <- fetch_df_2024_openness_final%>%
  #Makes column with bearing degree and column with corresponding fetch
  pivot_longer(cols = `7.5`: `360`, 
               names_to = "degree", 
               values_to = "fetch", 
               #Removes the prefix "degree" from the values going into the 
               #degree column
               names_prefix = "degree")
```


## Predictor 2: Mean SST

To get a measure of temperature I downloaded daily satellite sst data from the 
Copernicus Marine product for the year before the sampling in Barkley Sound 
occurred.

### Loading in the temperature data and calculating the average SST

```{r Loading in the temp .nc file downloaded from Copernicus, include=FALSE}
#Loading in the data
temp_2023_2024 <- brick(here("./01_raw_data/METOFFICE-GLO-SST-L4-NRT-OBS-SST-V2_1730223986748.nc"))

#Setting the coordinate system for the temperature data
crs(temp_2023_2024) <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
```

```{r Calculating the average temperature for each pixel, echo=TRUE, results='hide'}
average_temperature <- calc(temp_2023_2024, fun = mean, na.rm = TRUE)

writeRaster(average_temperature, 
            here("./01_raw_data/average_temperature.nc"), 
            format = "CDF", 
            overwrite = TRUE)
```


## Predictor 3: Protection from fishing status

In order to use the site protection status as a predictor, we needed to find
support for the fact that there was in fact a difference in the human impacts
between protected and unprotected sites.

To do this we used DFO flyover survey data that included the positions of where
recreational fishing boats were located in Barkley Sound. 

We used the DFO flyover counts of recreational boats actively fishing to 
calculate the annual mean number of recreational boats fishing within a given 
radius of each site to support the site protection status classifications and 
show that human impacts were higher for unprotected sites. To calculate the 
annual mean number of recreational boats fishing within a given radius of each 
site, we constructed circular polygons around each site (radii = 1 km, 1.5 km, 
2 km) using R (R Core Team, 2024). We merged these polygons with the DFO 
recreational boat counts to extract how many boats were spotted within the 
radius in each year. To get the annual mean, we averaged these counts across 
the three years of data from DFO (2021, 2022, 2023). We omitted aerial surveys 
where parts of Barkley Sound were obscured by fog.

### Loading in the DFO survey data

#### Loading in the 2023 data

```{r Loading in the  2023 DFO data}
DFO_data_raw_2023 <- read_csv(here("./01_raw_data/DFO_data/Overflight2023_ObservationPoints_20241024.csv"))

#Makes list of the variables that need to be converted to factors below
dfo_factors <- c("Day Type", "Creel Subarea")

#takes the raw data and cleans it up
rec_boats_fishing2023 <- DFO_data_raw_2023%>%
  #converts the variables specified above to factors
  mutate_at(dfo_factors, factor)%>%
  #renaming the columns to make them easier to code with
  rename(rec_boats_fishing = `1. Rec Boat Fishing`,
         rec_boats_running = `2. Rec Boats Running`,
         comm_vessel_fishing = `3. Commercial Vessel Fishing`,
         comm_vessel_not_fishing = `4. Commercial Vessel Not Fishing`,
         small_comm_vessel = `5. Small Commercial Vessel`,
         large_comm_vessel = `6. Large Commercial Vessel`,
         whale_watching = `7.  Whale Watching Boat`,
         orcas = `8. Orcas`,
         whales = `9. Whales`,
         sea_state = `10. Sea State`)%>%
  #Replacing the cells that have "Null" with 0 and converting the count columns
  #to numeric
  mutate(across(c("rec_boats_fishing", "rec_boats_running", "comm_vessel_fishing", 
                  "comm_vessel_not_fishing", "small_comm_vessel", "large_comm_vessel",
                  "whale_watching", "orcas", "whales"), ~ ifelse(. == "<Null>", 0, .)),
         across(c("rec_boats_fishing", "rec_boats_running", "comm_vessel_fishing", 
                  "comm_vessel_not_fishing", "small_comm_vessel", "large_comm_vessel",
                  "whale_watching", "orcas", "whales"), as.numeric))%>%
  ##This code will remove the dates of the flights that had fog
  filter(!Date %in% c("2023-06-23", "2023-07-16", "2023-07-31", "2023-08-06", "2023-08-28",
                      "2023-08-29", "2023-09-09", "2023-09-15"))%>%
  #Selecting only the rows where recreational boats were fishing
  filter(rec_boats_fishing != 0)%>%
  #selecting all of the columns that are needed for further analysis
    select(c(UniqueID, `Filename/FlightID`, Year, Month, Day, Date, `Critical Period`, `Day Type`, `Creel Subarea`, GMA, PFMA, 
           `Pin ID`, Longitude, Latitude, rec_boats_fishing, rec_boats_running))
```


#### Loading in the 2022 data

```{r Loading in the 2022 DFO data}
DFO_data_raw_2022 <- read_csv(here("./01_raw_data/DFO_data/Overflight2022_ObservationPoints_20240124_COUNTS.csv"))

#dfo_factors <- c("Day Type", "Creel Subarea")

rec_boats_fishing2022 <- DFO_data_raw_2022%>%
  mutate_at(dfo_factors, factor)%>%
  rename(rec_boats_fishing = `1. Rec Boat Fishing`,
         rec_boats_running = `2. Rec Boats Running`,
         comm_vessel_fishing = `3. Commercial Vessel Fishing`,
         comm_vessel_not_fishing = `4. Commercial Vessel Not Fishing`,
         small_comm_vessel = `5. Small Commercial Vessel`,
         large_comm_vessel = `6. Large Commercial Vessel`,
         whale_watching = `7.  Whale Watching Boat`,
         orcas = `8. Orcas`,
         whales = `9. Whales`,
         sea_state = `10. Sea State`)%>%
  mutate(across(c("rec_boats_fishing", "rec_boats_running", "comm_vessel_fishing", 
                  "comm_vessel_not_fishing", "small_comm_vessel", "large_comm_vessel",
                  "whale_watching", "orcas", "whales"), ~ ifelse(. == "<Null>", 0, .)),
         across(c("rec_boats_fishing", "rec_boats_running", "comm_vessel_fishing", 
                  "comm_vessel_not_fishing", "small_comm_vessel", "large_comm_vessel",
                  "whale_watching", "orcas", "whales"), as.numeric))%>%
   ##This code will remove the dates of the flights that had fog
  filter(!Date %in% c("2022-06-18", "2022-07-31", "2022-08-24", "2022-06-04"))%>%
    filter(rec_boats_fishing != 0)%>%
  #selecting all of the columns that are needed for further analysis
    select(c(UniqueID, `Filename/FlightID`, Year, Month, Day, Date, `Critical Period`, `Day Type`, `Creel Subarea`, GMA, PFMA, 
           `Pin ID`, Longitude, Latitude, rec_boats_fishing, rec_boats_running))
```


#### Loading in the 2021 data

```{r Loading in the 2021 DFO data}
DFO_data_raw_2021 <- read_csv(here("./01_raw_data/DFO_data/Overflight2021_ObservationPoints_20240321_counts.csv"))

#dfo_factors <- c("Day Type", "Creel Subarea")

rec_boats_fishing2021 <- DFO_data_raw_2021%>%
  mutate_at(dfo_factors, factor)%>%
  rename(rec_boats_fishing = `1. Rec Boat Fishing`,
         rec_boats_running = `2. Rec Boats Running`,
         comm_vessel_fishing = `3. Commercial Vessel Fishing`,
         comm_vessel_not_fishing = `4. Commercial Vessel Not Fishing`,
         small_comm_vessel = `5. Small Commercial Vessel`,
         large_comm_vessel = `6. Large Commercial Vessel`,
         whale_watching = `7.  Whale Watching Boat`,
         orcas = `8. Orcas`,
         whales = `9. Whales`,
         sea_state = `10. Sea State`)%>%
  mutate(across(c("rec_boats_fishing", "rec_boats_running", "comm_vessel_fishing", 
                  "comm_vessel_not_fishing", "small_comm_vessel", "large_comm_vessel",
                  "whale_watching", "orcas", "whales"), ~ ifelse(. == "<Null>", 0, .)),
         across(c("rec_boats_fishing", "rec_boats_running", "comm_vessel_fishing", 
                  "comm_vessel_not_fishing", "small_comm_vessel", "large_comm_vessel",
                  "whale_watching", "orcas", "whales"), as.numeric))%>%
   ##This code will remove the dates of the flights that had fog
  filter(!Date %in% c("2021-07-31", "2021-07-19", "2021-07-20", "2021-07-24"))%>%
   filter(rec_boats_fishing != 0)%>%
  select(c(UniqueID, `Filename/FlightID`, Year, Month, Day, Date, `Critical Period`, `Day Type`, `Creel Subarea`, GMA, PFMA, 
           `Pin ID`, Longitude, Latitude, rec_boats_fishing, rec_boats_running))
```


#### Combining the three years of data into one data frame
```{r Combining the three years of data by row}
rec_boats_fishing_total <- rbind(rec_boats_fishing2021, rec_boats_fishing2022, rec_boats_fishing2023)
```


#### Making a preliminary plot to see if it worked

```{r Seting the ranges for the maps}
#first need to specify the range of the map
lat_min = min(RLS_data_2024_clean$latitude) - 0.05
lat_max = max(RLS_data_2024_clean$latitude) + 0.05

long_min = min(RLS_data_2024_clean$longitude) - 0.05
long_max = max(RLS_data_2024_clean$longitude) + 0.05
```


```{r Making theme for all subsequent figures, echo=FALSE}
theme.figure <- function (){
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    #axis.line = element_line(colour = "black"),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12, face = "plain"),
    legend.text = element_text(size = 10, face = "plain"),
    legend.title = element_text(size = 14, face = "bold")) 
}

```

```{r Plotting the 2023 rec boats as a test, echo=TRUE}
 ggplot()+
    geom_sf(data = VI_base_layer_hi_res_sf_subset, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
    geom_point(data = rec_boats_fishing2023,
             mapping = aes(x = Longitude, y = Latitude, color = rec_boats_fishing),
             size = 1.5,) +
    scale_colour_gradient(low = "blue", high = "red")+
   geom_point(data = rec_boats_fishing2023,
             mapping = aes(x = Longitude, y = Latitude, colour = rec_boats_fishing), 
              shape = 1, size = 1.5, colour = "black") +
  labs(title = "2023", x = " \nLongitude", y = "Latitude\n ", colour = "# of Rec Fishing\nBoats Observed") +
  theme.figure() +
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br")
```



# Adding the predictors to a dataframe for modelling

## Adding the temperature data

```{r Extracting temperature data from the raster}
#This pipe makes a df with the lat and long of all of the different sites
site_points_df <- urch_sn_ratio_mod_2024%>%
   ungroup()%>%
  select(longitude, latitude)
 
#This sets the long and lat as coordinates
coordinates(site_points_df) <- ~longitude+latitude

#This sets the projection of the site locations to be on the same projection
#as the average temperature raster
proj4string(site_points_df) <- crs(average_temperature)

#This extracts the average temperature for where each site is located and then 
#subtracts the constant to convert from Kelvin to celsius
sst_values <- raster::extract(average_temperature, site_points_df) - 273.15

#Makes a column with the sst values in the modelling df
urch_sn_ratio_mod_2024$sst <- sst_values
```


## Adding the photoquadrat data

One of the predictors that I used was the dominant seafloor type of each
transect. I did this by using photoquadrats and stored the data from my 
analysis in an Excel sheet. I saved this as a .csv and simply loaded this into 
R to use in the models.

### Loading in the photoquadrat data

```{r Loading in the photoquadrat data}
pq_data <- read_csv(here("./01_raw_data/pq_data.csv"))

#This specifies the only columns that I wanted for analysis abd 
pq_data_final <- pq_data%>%
  select(c(site_code, depth, dominant_score, secondary_score))%>%
  #There were multiple rows for each site corresponding to each photoquadrat 
  #so this made just one row for each site
  distinct(site_code, depth, .keep_all = TRUE)

#This merges the photoquadrat data with the modelling data based on the specific
#transects. The all.x specifies a left join.
urch_sn_ratio_mod_2024 <- merge(urch_sn_ratio_mod_2024, pq_data_final, by = c("site_code", "depth"),
                                all.x = TRUE)
```

## Adding the openness data

```{r Summing the fetch vectors for each site then dividing by the largest openness value}
#Calculates the openness of each site
openness <- fetch_df_2024_openness_long%>%
  #grouping by each site
  group_by(site_code, latitude, longitude)%>%
  #adding up all of the fetch vectors for each site
  summarise(openness = sum(fetch))
  
#identifying the largest openness score within the dataset
max_openness <- max(openness$openness)

#Making a column of normalized openness by dividing by the max observed openness
openness <- openness%>%
mutate(norm_openness = openness / max_openness)

#Merging the modelling df with the openness df by each site
urch_sn_ratio_mod_2024 <- merge(urch_sn_ratio_mod_2024, openness, by = c("site_code", "latitude", "longitude"))
```

## Adding the recreational fishing boat data

To add the recreational fishin boat data we first calculated the mean number of
recreational fishing boats fishing within various radii of each RLS site. We did
this by making circular polygons around each RLS site and then merging those 
with the rec boat points and counting how many were in each circle each year, 
and then averaging across the three years of data.

### Calculating the mean recreational boats fishing within each radius

```{r Creating a circular radius around each site}
#Taking the site locations from the model
site_locations <- st_as_sf(urch_sn_ratio_mod_2024, coords = c("longitude", "latitude"), crs = 4326)

#Picking only the unique site locations as some sites have multiple transects
#with the same coordinates
site_locations_u <- site_locations%>%
  distinct(site_code, geometry)

#Changing the projection to handle meter distances better
site_locations_u_proj_m <- st_transform(site_locations_u, crs = 3857)
#Making circles with the different radii around each site
site_locations_radius_1km <- st_buffer(x = site_locations_u_proj_m, dist = 1000)
site_locations_radius_1.5km <- st_buffer(x = site_locations_u_proj_m, dist = 1500)
site_locations_radius_2km <- st_buffer(x = site_locations_u_proj_m, dist = 2000)
```

```{r Finding the overlapping points for each site}
#making a data frame with the coordinates of all the recreational fishing boats
rec_boats_fishing_total_points <- st_as_sf(rec_boats_fishing_total, coords = c("Longitude", "Latitude"), crs = 4326)
#Converting the CRS so it matches the other parts being used
rec_boats_fishing_total_points <- st_transform(rec_boats_fishing_total_points, crs = 3857)


#Finding the fishing boats that are within the given radius and making a new df
#with this info
rec_fishing_boat_total_radius_1km <- st_intersection(site_locations_radius_1km, rec_boats_fishing_total_points)

rec_fishing_boat_total_radius_1.5km <- st_intersection(site_locations_radius_1.5km, rec_boats_fishing_total_points)

rec_fishing_boat_total_radius_2km <- st_intersection(site_locations_radius_2km, rec_boats_fishing_total_points)
```

```{r Making dataframes with boat counts summed for each year at each site}
#Here I took the data frames with the counts and points of boats within the
#radii of the sites and I summed the number of boats counted around a site
#within each year. For example, I added all of the boats around Eagle bay in
#2021

rec_fishing_boat_total_radius_1km_summed <- rec_fishing_boat_total_radius_1km%>%
  #Grouping by each site in each year
  group_by(site_code, Year)%>%
  #adding up the number of boats around a site in each year
  summarise(yearly_rec_boats_fishing = sum(rec_boats_fishing))

rec_fishing_boat_total_radius_1.5km_summed <- rec_fishing_boat_total_radius_1.5km%>%
  group_by(site_code, Year)%>%
  summarise(yearly_rec_boats_fishing = sum(rec_boats_fishing))

rec_fishing_boat_total_radius_2km_summed <- rec_fishing_boat_total_radius_2km%>%
  group_by(site_code, Year)%>%
  summarise(yearly_rec_boats_fishing = sum(rec_boats_fishing))
```

```{r Making a total dataframe for each so that years with no boats have a 0 counts}
#For some sites that had boats counted in one particular year at a given radius
#did not have counts in all years. This step made data frames with a list of all
#of the sites that had at least some counts at the given radius in one of the 
#years and then made a row for each of the three years for that site.

#Makes a row for all three years for sites that had a least one boat counted
#in one of the years
all_sites_years_1km <- expand.grid(
  site_code = unique(rec_fishing_boat_total_radius_1km$site_code),
  Year = unique(rec_fishing_boat_total_radius_1km$Year))

all_sites_years_1.5km <- expand.grid(
  site_code = unique(rec_fishing_boat_total_radius_1.5km$site_code),
  Year = unique(rec_fishing_boat_total_radius_1.5km$Year))

all_sites_years_2km <- expand.grid(
  site_code = unique(rec_fishing_boat_total_radius_2km$site_code),
  Year = unique(rec_fishing_boat_total_radius_2km$Year))
```

```{r Merging the all years df with the rec boat count sums}
#This step merged the summed boat counts at each site with the dataframe that
# had all three years listed so that there was a row with 0 for the sites which
#did not have any counts in a particular year. This was necessary for making the
#average calculation accurate.

rec_fishing_boat_total_radius_1km_complete <- all_sites_years_1km %>%
  #Joins the rec fishing boat counts within radius to the df with all the years
  #for each site by site and year
  left_join(rec_fishing_boat_total_radius_1km_summed, by = c("site_code", "Year")) %>%
  #replaces the rows that have NAs with 0
  mutate(yearly_rec_boats_fishing = replace_na(yearly_rec_boats_fishing, 0))

rec_fishing_boat_total_radius_1.5km_complete <- all_sites_years_1.5km %>%
  left_join(rec_fishing_boat_total_radius_1.5km_summed, by = c("site_code", "Year")) %>%
  mutate(yearly_rec_boats_fishing = replace_na(yearly_rec_boats_fishing, 0))

rec_fishing_boat_total_radius_2km_complete <- all_sites_years_2km %>%
  left_join(rec_fishing_boat_total_radius_2km_summed, by = c("site_code", "Year")) %>%
  mutate(yearly_rec_boats_fishing = replace_na(yearly_rec_boats_fishing, 0))
```

```{r Making the final dataframe with the average boat counts per year for each site}
#This was the final step to average the recreational boat counts at each site 
#across the three years of data


rec_fishing_boat_total_radius_1km_final <- rec_fishing_boat_total_radius_1km_complete%>%
  #groups by each site
  group_by(site_code)%>%
  #makes a new column with the mean number of boats counted within each sites 
  #radius
  summarise(mean_rec_boat_counts_1km = mean(yearly_rec_boats_fishing))

rec_fishing_boat_total_radius_1.5km_final <- rec_fishing_boat_total_radius_1.5km_complete%>%
  group_by(site_code)%>%
  summarise(mean_rec_boat_counts_1.5km = mean(yearly_rec_boats_fishing))

rec_fishing_boat_total_radius_2km_final <- rec_fishing_boat_total_radius_2km_complete%>%
  group_by(site_code)%>%
  summarise(mean_rec_boat_counts_2km = mean(yearly_rec_boats_fishing))
```


### Adding the mean rec boat counts to the modelling data frame

```{r Adding the avg counts to the modelling dataset}
#This step added each of the data frames with the average rec boat counts
# to the modelling dataset. The resulting dataset has a column for avg 
#rec boat counts for each radius distance
urch_sn_ratio_mod_2024 <- urch_sn_ratio_mod_2024%>%
 left_join(rec_fishing_boat_total_radius_1km_final, by = c("site_code"))%>%
 left_join(rec_fishing_boat_total_radius_1.5km_final, by = c("site_code"))%>%
  left_join(rec_fishing_boat_total_radius_2km_final, by = c("site_code"))%>%
  replace(is.na(.), 0)
```


## Adding the region classifications for use as a random/fixed effect

In order to use region as a potential random or fixed effect during modelling
I made region classifications for the sites depending on which Island group/area
they were in. I made a few different region classifications, but ultimately we
went with the region1 classification because it made the most logical sense.

```{r Making a region column}
urch_sn_ratio_mod_2024 <- urch_sn_ratio_mod_2024%>%
  #makes a column for the region and then depending if the site code is
  #specified it puts the given region names below
  mutate(region1 = case_when(
    site_code %in% c("BMSC1", "BMSC10", "BMSC11", "BMSC12", "BMSC19", "BMSC2", "BMSC20", 
                     "BMSC21", "BMSC22", "BMSC23", "BMSC25", "BMSC3", "BMSC4", "BMSC5", "BMSC9") ~ "Deer group/Bamfield",
   site_code %in% c("BMSC13", "BMSC14", "BMSC15", "BMSC16", "BMSC17", "BMSC18") ~ "Broken Group", 
    site_code %in% c("BMSC24", "BMSC6", "BMSC8") ~ "Deep Imperial Eagle Channel",
    site_code %in% c("BMSC26", "BMSC27") ~ "Sarita"
  ))%>%
   mutate(region2 = case_when(
    site_code %in% c("BMSC1", "BMSC10", "BMSC11", "BMSC12", "BMSC19", "BMSC2", "BMSC20",
                     "BMSC3",  "BMSC5") ~ "Deer group",
    site_code %in% c("BMSC21", "BMSC22", "BMSC23", "BMSC25", "BMSC4", "BMSC9") ~ "Bamfield",
   site_code %in% c("BMSC13", "BMSC14", "BMSC15", "BMSC16", "BMSC17", "BMSC18") ~ "Broken Group", 
    site_code %in% c("BMSC24", "BMSC6", "BMSC8") ~ "Deep Imperial Eagle Channel",
    site_code %in% c("BMSC26", "BMSC27") ~ "Sarita"
  ))%>%
   mutate(region3 = case_when(
     site_code%in% c("BMSC19", "BMSC2", "BMSC20", "BMSC3", "BMSC24", "BMSC6", "BMSC8") ~ "Imperial Eagle",
    site_code %in% c("BMSC1", "BMSC10", "BMSC11", "BMSC12", 
                     "BMSC21", "BMSC22", "BMSC23", "BMSC25", "BMSC26", "BMSC27",
                     "BMSC4", "BMSC5", "BMSC9") ~ "Trevor Channel",
   site_code %in% c("BMSC13", "BMSC14", "BMSC15", "BMSC16", "BMSC17", "BMSC18") ~ "Broken Group", 
  ))%>%
  mutate(region4 = case_when(
    site_code %in% c("BMSC21", "BMSC22", "BMSC23", "BMSC25", "BMSC4", "BMSC9", "BMSC26", "BMSC27") ~ "Bamfield",
    site_code %in% c("BMSC13", "BMSC14", "BMSC15", "BMSC16", "BMSC17", "BMSC18") ~ "Broken Group",
    site_code %in% c("BMSC24", "BMSC6", "BMSC8") ~ "Deep Imperial Eagle Channel",
    site_code %in% c("BMSC1", "BMSC10", "BMSC11", "BMSC12", "BMSC19", "BMSC2", "BMSC20",
                     "BMSC3",  "BMSC5") ~ "Deer group"
  ))
```


## Writing the modelling data frame to a .csv for future use
```{r Writing the modelling df}
write_csv(urch_sn_ratio_mod_2024,
          here("./03_clean_data/urchin_snail_modelling_data.csv"))
```


# Doing modelling for the project

## Loading in the modelling data and cleaning it up

```{r Loading in the modelling data, include=FALSE}
env_mod_data_hon <- read_csv(here("./03_clean_data/urchin_snail_modelling_data.csv"))
```


```{r Cleaning up the data, include=FALSE}
#As the temp and openness data are on different scales we scaled them so as not
#to have the different scales influence the results
#These made columns with the scaled data
env_mod_data_hon$sst_scale <- scale(env_mod_data_hon$sst)
env_mod_data_hon$openness_scale <- scale(env_mod_data_hon$openness)
env_mod_data_hon$norm_openness_scale <- scale(env_mod_data_hon$norm_openness)

#We also tried some modelling with the recreational boats so made scaled columns
#for them as well
env_mod_data_hon$mean_rec_boat_counts_1km_scale <- scale(env_mod_data_hon$mean_rec_boat_counts_1km)
env_mod_data_hon$mean_rec_boat_counts_1.5km_scale <- scale(env_mod_data_hon$mean_rec_boat_counts_1.5km)
env_mod_data_hon$mean_rec_boat_counts_2km_scale <- scale(env_mod_data_hon$mean_rec_boat_counts_2km)

#making dominant seafloor score a factor
env_mod_data_hon$dominant_score <- as.factor(env_mod_data_hon$dominant_score)
#same for site code
env_mod_data_hon$site_code <- as.factor(env_mod_data_hon$site_code)
#tried using the different species as predictors for each other so made scaled
#columns of each
env_mod_data_hon$Urchins_scale <- scale(env_mod_data_hon$Urchins)
env_mod_data_hon$Turban_Snails_scale <- scale(env_mod_data_hon$Turban_Snails)

#Renaming the levels of the seafloor type to loose and solid
levels(env_mod_data_hon$dominant_score) <- list(loose = "3", solid = "4")

#Making the region columns factors, first by making an object with their names
#and then using lapply to do it to all of them at once
regions <- c("region1", "region2", "region3", "region4")
env_mod_data_hon[regions] <- lapply(env_mod_data_hon[regions], factor)

## Reordering the factor levels of region1 so that Deer group bamfield is the
#intercept and renaming for plotting
levels(env_mod_data_hon$region1) <- list(DGB = "Deer group/Bamfield",
                                     BG = "Broken Group",
                                     DIEC = "Deep Imperial Eagle Channel",
                                     SR = "Sarita" )
```


```{r Creating a factor variable of protection status for the modelling data, include=FALSE}
#Making an object with the protection status of the sites coded as 0 for not
#protected and 1 for protected from fishing
protect_status <- c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 
                    0, 0, 0, 1, 1, 0, 0, 0, 0, 0, 1, 1, 1, 1, 0)

#Making protect status into a column in the DF and as a factor
  env_mod_data_hon$protect_status <- as.factor(protect_status)
  #Changing the levels to No Protection and Protection instead of 0 and 1
levels(env_mod_data_hon$protect_status) <- c("No Protection", "Protected")

#Reordering so easier to see which sites are protected in the DF
env_mod_data_hon <- env_mod_data_hon[, c(seq(1, 8, by = 1), 35, seq(9, 34, by = 1))]
```


## Exploratory plots of the data
Making some exploratory plots of the predictors against the response variables
(urchin dens, snail dens, snail to urchin ratio).


### The EVs against the snail density

```{r snail dens against substrate score, echo=FALSE}
ggplot(env_mod_data_hon, aes(x = dominant_score, y = snail_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r snail dens against openness,echo=FALSE}
ggplot(env_mod_data_hon, aes(x = norm_openness, y = snail_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r snail dens against sst, echo=FALSE}
ggplot(env_mod_data_hon, aes(x = sst, y = snail_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r snail dens against mean rec boat counts 1.5km, echo=FALSE}
ggplot(env_mod_data_hon, aes(x = mean_rec_boat_counts_1.5km, y = snail_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r snail dens against mean rec boat counts 1km, echo=FALSE}
ggplot(env_mod_data_hon, aes(x = mean_rec_boat_counts_1km, y = snail_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r snail dens against mean rec boat counts 2km, echo=FALSE}
ggplot(env_mod_data_hon, aes(x = mean_rec_boat_counts_2km, y = snail_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

### The EVs against the urchin density

```{r urchin dens against substrate score, echo=FALSE}
ggplot(env_mod_data_hon, aes(x = dominant_score, y = urchin_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r urchin dens against openness,echo=FALSE}
ggplot(env_mod_data_hon, aes(x = norm_openness, y = urchin_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r urchin dens against sst, echo=FALSE}
ggplot(env_mod_data_hon, aes(x = sst, y = urchin_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r urchin dens against mean rec boat counts 1.5km, echo=FALSE}
ggplot(env_mod_data_hon, aes(x = mean_rec_boat_counts_1.5km, y = urchin_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r urchin dens against mean rec boat counts 1km, echo=FALSE}
ggplot(env_mod_data_hon, aes(x = mean_rec_boat_counts_1km, y = urchin_dens, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r urchin dens against mean rec boat counts 2km, echo=FALSE}
ggplot(env_mod_data_hon, aes(x = mean_rec_boat_counts_2km, y = urchin_dens, colour = region1))+
  geom_point()+
  theme.figure()
```



### The EVs against the snail to urchin ratio
```{r ratio against substrate score, echo=FALSE}
ggplot(env_mod_data_hon, aes(x = dominant_score, y = sn_urch_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against openness,echo=FALSE}
ggplot(env_mod_data_hon, aes(x = norm_openness, y = sn_urch_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against sst, echo=FALSE}
ggplot(env_mod_data_hon, aes(x = sst, y = sn_urch_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against mean rec boat counts 1.5km, echo=FALSE}
ggplot(env_mod_data_hon, aes(x = mean_rec_boat_counts_1.5km, y = sn_urch_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against mean rec boat counts 1km, echo=FALSE}
ggplot(env_mod_data_hon, aes(x = mean_rec_boat_counts_1km, y = sn_urch_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against mean rec boat counts 2km, echo=FALSE}
ggplot(env_mod_data_hon, aes(x = mean_rec_boat_counts_2km, y = sn_urch_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```



## Ecological models

The first set of models that we tried were generalized additive models to test
the ecological predictors on the densities of each species to see what is 
driving their populations.

### Red Turban Snail Ecology models

We started with the turban snail models. 

#### Model 1: Ecology model on the snail density with temperature, openness, and bottom type

```{r Turban snail ecology/environmental model}
#We initially tested this model with the untransformed snail density, but the 
#residuals were not good. We could not use Poisson or Negative binomial as there
#were decimals. We tried Gamma with adding a small constant to get rid of the 0s
#but this did not work well. Also tried Tweedie and scat but again did not look 
#good. At this point we decided to log transform in order to maximize model fit
#potentially at the expense of interpretability and with the drawbacks of the
#asymmetrical error structure on the untransformed scale.
#the snail density. We needed to add a constant to the density because there
#were some zeroes so we instead of adding 1, we added 0.01 because the snail
#densities were the snail counts divided by 100.
ecology_snail_mod <- gam(log(snail_dens + 0.01) ~ s(sst_scale)+
                               s(norm_openness_scale)+
                           dominant_score,
                             data = env_mod_data_hon,
                             family = gaussian(link = "identity"),
                         #we picked REML for smoothing parameter estimation as
                         #this is considered to be one of the best methods
                             method = "REML",
                         #we set select to TRUE meaning that the model can fully
                         #penalize smooth terms to zero effect if that is what 
                         #is suggested
                             select = TRUE)
```

```{r Turban snail ecology/environmental model output, echo=FALSE}
#This gives some summary output about the fitting of the smooth and how many edf
#are used. It also gives residual plots to look at the fit.
gam.check(ecology_snail_mod)

#this does a statistical test of the normality of the residuals
shapiro.test(ecology_snail_mod$residuals)

#This plots all of the partial effects plots so you can see the relationship
#with each of the variables
plot(ecology_snail_mod, all.terms = TRUE)

#This gives the summary of the model
summary(ecology_snail_mod)
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 

To test each of our models for spatial autocorrelation we conducted a Moran's
I test on the model residuals. To ensure even better confidence in the results
we used both a nearest neighbour weight matrix method and an inverse distance
weight matrix for the Moran's I weight matrix.

```{r Testing Turban snail ecology model residuals for spatial correlation, echo=FALSE, warning=FALSE}
#This extracts the residuals from the model and stores them as an object
residuals_ecology_snail_mod <- residuals(ecology_snail_mod, type = "response")


#This creates an object of the coordinates from the sites in the data and 
#specifies the coordinate system they are on
coords_for_auto_test <- st_as_sf(env_mod_data_hon, coords = c("longitude", "latitude"), crs = 4326)
#This converts the coordinate system to the 2d projection one
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
#This saves the coordinates in matrix form
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)

#This makes a neighbour matrix, specifying which sites are neighbours with
#which other sites. Whether or not sites are neighbours is based on if they are
#within the distance specified in d2.
nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#I picked 10 m for distance so that it just includes the transects at the 
#same sites as neighbours, but trying with larger distances kept the same result

#This takes the neighbours and makes a spatial weights matrix for calculating 
#the Moran's I statistic. The style W means that each row of the weights matrix 
#must sum to 1. For example if a site has 2 neighbours then each neighbour is 
#weighted as 0.5. The zero.policy as true means that if a site does not have
#any neighbours it is weighted as 0.
lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##This code performs a two sided moran's I test on the residuals using the 
#weights matrix that we just made.
moran.test(residuals_ecology_snail_mod, lw,
           alternative = "two.sided")

#The following code is calculating the Moran's I in the slightly different way
#of making an inverse distance weight matrix, where sites that are farther away
#become weighted less and less.

#This makes a matrix of the distances between the different sites.
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
#This makes an inverse distance matrix based on the original distance matrix. 
#A small constant is added to the distance matrix so that transects at the same
# are not given a zero weight.
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
#This makes the cells that are comparing the same exact transects to each other 
#zero.
diag(inverse_dist_matrix) <- 0
#This converts the inverse distance matrix into a weigts matrix for use in the
#Moran's I function. The style as W makes it so that each row must sum to 1.
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
#This computes the Moran's I two sided test on the model residuals.
moran.test(residuals_ecology_snail_mod, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation for both methods.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing turban snail ecology model for concurvity, echo=FALSE}
#These functions compute the concurvity for the model, a measure used in GAMs
#to see how closely related the different predictor variables are.
concurvity(ecology_snail_mod, full = FALSE)
concurvity(ecology_snail_mod, full = TRUE)
```
</div>

Temperature and openness seem to be somewhat concurve. Temp having no impact on
the model (being removed by select).


#### Model 2: Ecology model on the snail counts with temperature, openness, and bottom type (linear openness)

After initially seeing that the effect of openness was linear on the snail counts I tried doing another model
with the openness explicitly modelled as a linear term so that I got a slope for the effect. In the end I ended
up just going with the first model with the openness fit as a smooth term.

```{r Turban snail ecology/environmental model linear}
#As the openness in the previous model was shown to be linearly related we
#redid the model with it fitted as a linear predictor
ecology_snail_mod_linear <- gam(log(snail_dens + 0.01) ~ s(sst_scale)+
                               norm_openness_scale+
                           dominant_score,
                             data = env_mod_data_hon,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

```{r Turban snail ecology/environmental model linear output, echo=FALSE}
gam.check(ecology_snail_mod_linear)
shapiro.test(ecology_snail_mod_linear$residuals)
plot(ecology_snail_mod_linear, all.terms = TRUE)
summary(ecology_snail_mod_linear)
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail ecology model linear residuals for spatial correlation, echo=FALSE, warning=FALSE}

residuals_ecology_snail_mod_linear <- residuals(ecology_snail_mod_linear, type = "response")

coords_for_auto_test <- st_as_sf(env_mod_data_hon, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)


lw <- nb2listw(nb, style = "W", zero.policy = TRUE)



moran.test(residuals_ecology_snail_mod_linear, lw,
           alternative = "two.sided")


dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_ecology_snail_mod_linear, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation for both methods.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing turban snail ecology model linear for concurvity, echo=FALSE}
concurvity(ecology_snail_mod_linear, full = FALSE)
concurvity(ecology_snail_mod_linear, full = TRUE)
```
</div>

Concurvity seems to be improved when having openness as a linear term.

#### Model 3: Ecology model on the snail counts with temperature, openness, and bottom type and random effect (linear openness)

```{r Turban snail ecology/environmental model linear re}
#To see if a random effect of region was important to add we tried a model with
#this included
ecology_snail_mod_re <- gam(log(snail_dens + 0.01) ~ s(sst_scale)+
                               s(norm_openness_scale) +
                           dominant_score+
                             #By fitting the random effect in this form it is 
                             #fit as a random intercept effect
                                s(region1, bs = "re"),
                             data = env_mod_data_hon,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

```{r Turban snail ecology/environmental model linear re output, echo=FALSE}
gam.check(ecology_snail_mod_re)
shapiro.test(ecology_snail_mod_re$residuals)
plot(ecology_snail_mod_re, all.terms = TRUE)
summary(ecology_snail_mod_re)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail ecology model linear re residuals for spatial correlation, echo=FALSE, warning=FALSE}

residuals_ecology_snail_mod_re <- residuals(ecology_snail_mod_re, type = "response")


coords_for_auto_test <- st_as_sf(env_mod_data_hon, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)


lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


moran.test(residuals_ecology_snail_mod_re, lw,
           alternative = "two.sided")


dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_ecology_snail_mod_re, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation for both methods when using a two sided test.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing turban snail ecology model linear re for concurvity, echo=FALSE}
concurvity(ecology_snail_mod_re, full = FALSE)
concurvity(ecology_snail_mod_re, full = TRUE)
```
</div>

Random effect of region is highly concurve.


#### Model 4: Ecology model on the snail counts with temperature, openness, and bottom type and fixed effect of region (linear openness)

```{r Turban snail ecology/environmental model linear FE}
#We were also curious to look at region as a fixed effect so we tried a model 
#with this included
ecology_snail_mod_fe <- gam(log(snail_dens + 0.01) ~ s(sst_scale)+
                               s(norm_openness_scale) +
                           dominant_score+
                                region1,
                             data = env_mod_data_hon,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

```{r Turban snail ecology/environmental model linear fe output, echo=FALSE}
gam.check(ecology_snail_mod_fe)
shapiro.test(ecology_snail_mod_fe$residuals)
plot(ecology_snail_mod_fe, all.terms = TRUE)
summary(ecology_snail_mod_fe)
```
Region as a fixed effect does not seem to be significant.

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail ecology model fe residuals for spatial correlation, echo=FALSE, warning=FALSE}

residuals_ecology_snail_mod_fe <- residuals(ecology_snail_mod_fe, type = "response")


coords_for_auto_test <- st_as_sf(env_mod_data_hon, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)


lw <- nb2listw(nb, style = "W", zero.policy = TRUE)



moran.test(residuals_ecology_snail_mod_fe, lw,
           alternative = "two.sided")


dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_ecology_snail_mod_fe, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing turban snail ecology model linear fe for concurvity, echo=FALSE}
concurvity(ecology_snail_mod_fe, full = FALSE)
concurvity(ecology_snail_mod_fe, full = TRUE)
```
</div>

Temp and para are highly concurve.


#### Comparing Turban snail ecology models

```{r AIC of Turban snail ecology models}
#To compare between the models we had fit to identify the best model we used
#AIC analysis. Specifically AIC corrected as our sample size was low.

#This runs the AIC on the different models
AICc(ecology_snail_mod,
  ecology_snail_mod_linear,
    ecology_snail_mod_re,
    ecology_snail_mod_fe)%>%
  #This arranges by the smallest AICc
  arrange(AICc)%>%
  #This makes a column of the delta AIC for ease of comparison
  mutate(delta_AICc = AICc - AICc[1])
```

AIC indicates that it does not improve the model by adding a random effect of 
region. The random effect is basically being removed from the model by the 
select = TRUE argument. This means that for the snail ecology model I will 
pick the more parsimonius model without the random effect.

Also, having region as a fixed effect makes the AIC much worse.
Trying with both methods (REML and ML) the model with the fixed effect is much
worse.

For simplicity and because I did not know that the effect of openness would be
linear prior to fitting the original model I just went with the original model
over the one with openness fit as a linear term.


### Red sea urchin Ecology models

#### Model 5: Ecology model on the urchin counts with temperature, openness, and bottom type

```{r Urchin ecology/environmental model}
#For this model we tried with other distributions (tweedie, quasipoisson, scat)
#but the model fit appeared best with a simple gaussian model
ecology_urchin_mod <- gam(urchin_dens ~ s(sst_scale)+
                               s(norm_openness_scale)+
                           dominant_score,
                             data = env_mod_data_hon,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

```{r Urchin ecology/environmental model output, echo=FALSE}
set.seed(299)
gam.check(ecology_urchin_mod, rep = 500)
plot(ecology_urchin_mod, all.terms = TRUE)
summary(ecology_urchin_mod)
shapiro.test(ecology_urchin_mod$residuals)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Urchin ecology model residuals for spatial correlation, echo=FALSE, warning=FALSE}

residuals_ecology_urchin_mod <- residuals(ecology_urchin_mod, type = "response")


coords_for_auto_test <- st_as_sf(env_mod_data_hon, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)


lw <- nb2listw(nb, style = "W", zero.policy = TRUE)



moran.test(residuals_ecology_urchin_mod, lw,
           alternative = "two.sided")


dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_ecology_urchin_mod, lw_inverse,
           alternative = "two.sided")
```
</div>

Some positive autocorrelation but only with the inverse distance method. The 
neighbour method has no autocorrelation.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing urchin ecology model for concurvity, echo=FALSE}
concurvity(ecology_urchin_mod, full = FALSE)
concurvity(ecology_urchin_mod, full = TRUE)
```
</div>

Concurvity is high for temperature and openness. Both are being removed by the 
double penalty.


#### Model 6: Ecology model on the urchin counts with temperature, openness, and bottom type (random effect of region)

```{r Urchin ecology/environmental model region re}
#We wished to investigate using a random effect of region in the urchin model
#so we did that here
ecology_urchin_mod_re <- gam(urchin_dens ~ s(sst_scale)+
                               s(norm_openness_scale)+
                           dominant_score+
                                s(region1, bs = "re"),
                             data = env_mod_data_hon,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

```{r Urchin ecology/environmental model region re output, echo=FALSE}
set.seed(299)
gam.check(ecology_urchin_mod_re, rep = 500)
plot(ecology_urchin_mod_re, all.terms = TRUE)
summary(ecology_urchin_mod_re)
shapiro.test(ecology_urchin_mod_re$residuals)
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Urchin ecology model region re residuals for spatial correlation, echo=FALSE, warning=FALSE}

residuals_ecology_urchin_mod_re <- residuals(ecology_urchin_mod_re, type = "response")


coords_for_auto_test <- st_as_sf(env_mod_data_hon, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)



moran.test(residuals_ecology_urchin_mod_re, lw,
           alternative = "two.sided")

dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_ecology_urchin_mod_re, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation for either method.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing urchin ecology model re for concurvity, echo=FALSE}
concurvity(ecology_urchin_mod_re, full = FALSE)
concurvity(ecology_urchin_mod_re, full = TRUE)
```
</div>

Temp and region are basically fully concurve.


#### Model 7: Ecology model on the urchin counts with temperature, openness, and bottom type (fixed effect of region)

```{r Urchin ecology/environmental model region fe}
#We also wanted to investigate the use of region as a fixed effect in the 
#urchin model
ecology_urchin_mod_fe <- gam(urchin_dens ~ s(sst_scale)+
                               s(norm_openness_scale)+
                           dominant_score+
                                region1,
                             data = env_mod_data_hon,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

```{r Urchin ecology/environmental model region fe output, echo=FALSE}
set.seed(299)
gam.check(ecology_urchin_mod_fe, rep = 500)
plot(ecology_urchin_mod_fe, all.terms = TRUE)
summary(ecology_urchin_mod_fe)
shapiro.test(ecology_urchin_mod_fe$residuals)
```
Gaussian looks ok. When adding the fixed effect of region the openness becomes
borderline and negative. Also dominant substrate type becomes significant with 
more urchins on solid substrate.

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Urchin ecology model region fe residuals for spatial correlation, echo=FALSE, warning=FALSE}

residuals_ecology_urchin_mod_fe <- residuals(ecology_urchin_mod_fe, type = "pearson")


coords_for_auto_test <- st_as_sf(env_mod_data_hon, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)


lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


moran.test(residuals_ecology_urchin_mod_fe, lw,
           alternative = "two.sided")


dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_ecology_urchin_mod_fe, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing urchin ecology model fe for concurvity, echo=FALSE}
concurvity(ecology_urchin_mod_fe, full = FALSE)
concurvity(ecology_urchin_mod_fe, full = TRUE)
```
</div>

Temp is very concurve.


##### Testing the model for VIF stuff
```{r Testing the collinearity of the linear fixed effects in the model}
#This performs a check of the collinearity of the variables in the model
performance::check_collinearity(ecology_urchin_mod_fe,
                   component = c("all"))
```
The resulting values indicate that there is acceptable correlation between the 
predictor variables.

#### Comparing the urchin ecology models

```{r AIC of Urchin ecology models}
AICc(ecology_urchin_mod,
     ecology_urchin_mod_re,
     ecology_urchin_mod_fe)%>%
  arrange(AICc)%>%
  mutate(delta_AICc = AICc - AICc[1])
```
When testing with REML, the model with the random effect is much better than 
the model with no random effect. When comparing with ML, the model with the 
fixed effect is much better than the model without the fixed effect. The model
with the fixed and random effects are very similar in AIC and the random effect
model has slightly lower df.

When repeating with urchin density and REML the fixed and random are basically 
the same.

We decided to go with the random effect model as using region as a random effect
made more sense from a theoretical perspective.


## Fishing models

### Confirming the validity of protection status with DFO flyover data

For the protection status categories, I have grouped them as sites that 
experience at least a ban on fishing for part of the year as protected and all
others as non-protected. This means that the sites that are in the Broken group
but are not within the RCA are counted as not protected.

In order to confirm/support our assumptions of which sites are protected,
I used the DFO flyover data to compare the numbers of boats around the 
different sites.

Based on the boxplots below it seems evident that the protected sites selected
are experiencing less fishing pressure and less overall human/boat pressure. 

Ultimately, the comparison of fishing boat activity between protected and 
unprotected sites was used to support the idea of simply using a factor 
predictor of protection status as a predictor in anthropogenic models for these
species.

#### Boxplots of rec boat counts between protection levels

```{r Boxplots of the rec boats at the three distances between protection levels, echo=FALSE}
ggplot(env_mod_data_hon, aes(x = protect_status, y = mean_rec_boat_counts_1km))+
  geom_boxplot()+
  geom_jitter()+
  theme.figure()

ggplot(env_mod_data_hon, aes(x = protect_status, y = mean_rec_boat_counts_1.5km))+
  geom_boxplot()+
  geom_jitter()+
  theme.figure()

ggplot(env_mod_data_hon, aes(x = protect_status, y = mean_rec_boat_counts_2km))+
  geom_boxplot()+
  geom_jitter()+
  theme.figure()
```

#### Comparing the rec boat counts between sites in the broken group that have been labelled as protected or non-protected

To make sure that the effect of protection works to reduce boat counts within one region
we looked at only the sites within the Broken Group islands.

```{r Comparing the rec boat counts between sites in the broken group that have been labelled as protected or non-protected, echo=FALSE}
ggplot(env_mod_data_hon[c(8, 9, 10, 11, 12, 13),], aes(x = protect_status, y = mean_rec_boat_counts_1km))+
  geom_boxplot()+
  geom_point()+
  theme.figure()

ggplot(env_mod_data_hon[c(8, 9, 10, 11, 12, 13),], aes(x = protect_status, y = mean_rec_boat_counts_1.5km))+
  geom_boxplot()+
  geom_point()+
  theme.figure()

ggplot(env_mod_data_hon[c(8, 9, 10, 11, 12, 13),], aes(x = protect_status, y = mean_rec_boat_counts_2km))+
  geom_boxplot()+
  geom_point()+
  theme.figure()
```

The trend is clear even just for sites within the Broken Group.


#### Testing the differences in rec boat counts between protected and non-protected with stats

After seeing the clear trend of fewer boats on average around protected sites,
we wanted to use stats to actually see if there was a significant difference
between protected and unprotected sites.

##### Checking for assumptions of linear models

```{r Making a subset df with only one observation for each site}
#This chunk is making a df that has one row for each site, with the 
#corresponding mean number of rec boats fishing at the three radii tested
unique_site_df <- env_mod_data_hon%>%
  #This removes the existing groupings within the df
  ungroup()%>%
  #selecting the columns we want
  select(site_code, protect_status,mean_rec_boat_counts_1km, mean_rec_boat_counts_1.5km, mean_rec_boat_counts_2km, region1)%>%
  #this makes it so that there is only one row for each site as some sites had 
  #two transects
  distinct()
```


```{r Testing the variance between protection status}
#In order to run stats tests on the boat counts we need to make sure the data
#meets the assumptions of linear models. Here we use a levene's test to see
#if there is equal variance between protected and unprotected sites. We do this
#for each of the three radii.

leveneTest(mean_rec_boat_counts_1km ~ protect_status, data = unique_site_df)
leveneTest(mean_rec_boat_counts_1.5km ~ protect_status, data = unique_site_df)
leveneTest(mean_rec_boat_counts_2km ~ protect_status, data = unique_site_df)
```
Equal variance for 1km, but not for 1.5km or 2km.

```{r Normality of rec boat counts within protected sites}
#The second assumption that we tested for was normality of the data. This chunk
#makes a df that contains only sites within the protected areas to test for 
#normality within this group.

#Making data frame with only sites that are protected
rec_boat_protected <- subset(unique_site_df, protect_status == "Protected")

#Using shapiro wilks test to see if the mean rec boat counts are normal
shapiro.test(rec_boat_protected$mean_rec_boat_counts_1km)
#using histogram to look at normality visually
hist(rec_boat_protected$mean_rec_boat_counts_1km)

shapiro.test(rec_boat_protected$mean_rec_boat_counts_1.5km)
hist(rec_boat_protected$mean_rec_boat_counts_1.5km)

shapiro.test(rec_boat_protected$mean_rec_boat_counts_2km)
hist(rec_boat_protected$mean_rec_boat_counts_2km)
```
Only the 1.5 km boat counts are normal

```{r Normality of rec boat counts within non-protected sites}
#This chunk follows the same steps as the previous one however it is for the 
#unprotected sites.

#Making data frame with only sites that are non-protected
rec_boat_non_protected <- subset(unique_site_df, protect_status == "No Protection")

shapiro.test(rec_boat_non_protected$mean_rec_boat_counts_1km)
hist(rec_boat_non_protected$mean_rec_boat_counts_1km)

shapiro.test(rec_boat_non_protected$mean_rec_boat_counts_1.5km)
hist(rec_boat_non_protected$mean_rec_boat_counts_1.5km)

shapiro.test(rec_boat_non_protected$mean_rec_boat_counts_2km)
hist(rec_boat_non_protected$mean_rec_boat_counts_2km)
```
2km is normal but the others are not.

Because we have not met the assumptions of normality or homoscedasticity for 
all of our groups we are not able to use linear models. Instead we used 
Wilcoxon Rank sum tests.

##### Wilcoxon rank sum tests of the mean recreational boat counts

```{r Wilcoxon rank sum tests of mean rec boat counts between protected and non-protected sites}
#These lines of code run a separate wilcoxon rank sum test for each of the three
#radii that we looked at. Since alternative was set to "t" it is a two tailed
#test
wilcox_1km <- wilcox.test(mean_rec_boat_counts_1km ~ protect_status, data = unique_site_df, alternative = "t")
wilcox_1.5km <- wilcox.test(mean_rec_boat_counts_1.5km ~ protect_status, data = unique_site_df, alternative = "t")
wilcox_2km <- wilcox.test(mean_rec_boat_counts_2km ~ protect_status, data = unique_site_df, alternative = "t")

print(wilcox_1km)
print(wilcox_1.5km)
print(wilcox_2km)
```
All of the wilcoxon rank sum tests indicate that there is a highly significant
difference in the number of rec fishing boats between protected and non-protected
sites. Based on the boxplots we can say that there are significantly more rec
boats at unprotected sites than protected sites.

### Red Turban snail fishing models{.tabset}

After exploring the impacts of environmental drivers of the red turban snails,
we wanted to look into whether the protection from fishing would have any 
impacts on this species.Given the results of the mean rec boat comparison
between protected and unprotected sites we decided that simply using a predictor
of protection status would be a useful measure of fishing activity.

#### Plotting counts of snails between protected and unprotected

```{r Plotting counts of snails between protected and unprotected, echo=FALSE}
ggplot(env_mod_data_hon, aes(x = protect_status, y = Turban_Snails))+
  geom_boxplot()+
  geom_jitter(height = 0)+
  labs(x = "Protection Status", y = "Turban Snail counts")+
  theme.figure()
```

Seeing a weak trend of more snails in the protected sites when compared to the non-
protected sites. We now wanted to explore this statistically.

#### Model 8: Fishing model on the snail counts (glm)


```{r Snail count fishing model with protection linear}
#This chunk is fitting an lm on the snail density with the predictor of 
#protection status. We ended up squareroot transforming the snail density
#after initial testing as the fit of the residuals for the untransformed density
#was poor for all the distributions we tried. We first tried a log 
#transformation similar to the ecological model but this broke the assumption
#of homoscedasticity. Ultimately, the squareroot transformation with a gaussian
#family produced the best fit.

snail_fishing_mod <- glmmTMB(sqrt(snail_dens) ~
                                      protect_status,
                                    data = env_mod_data_hon,
                                    family = gaussian(link = "identity"))
```

```{r Snail count fishing model with protection linear output, echo=FALSE}
#This uses the DHARMa package to simulate the residuals from the model and test
#for any problems in the model fit.
snail_fishing_mod_resids <- simulateResiduals(snail_fishing_mod)
plot(snail_fishing_mod_resids)

#This made a histogram of the model residuals to look for normality
hist(residuals(snail_fishing_mod))
#Shaprio test for normality of residuals
shapiro.test(residuals(snail_fishing_mod))

#Looking at the model summary
summary(snail_fishing_mod)

#Getting an r2 value for the model
performance::r2(snail_fishing_mod)
```


```{r Plotting the linear snail fishing model with protection, echo=FALSE}
#This predicts values for mean snail density with standard errors and CIs for 
#each level of the predictor and saves as a df
pred_snail_fishing_mod <- ggpredict(snail_fishing_mod, terms = "protect_status")

#This is plotting the predicted values to visualize the model
ggplot(pred_snail_fishing_mod, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Turban Snail Density (n/m2)", x = "Protection Status")+
  theme.figure()
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail fishing model residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_snail_fishing_mod <- residuals(snail_fishing_mod, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data_hon, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_snail_fishing_mod, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_snail_fishing_mod, lw_inverse,
           alternative = "two.sided")
```
</div>
No autocorrelation.


#### Model 9: Fishing model on the snail counts with protection and a random effect (glmm)

Similar to the previous models, we wanted to explore if adding a random effect
of region would explain a substantial amount of the variation so we tested that
in this model.

```{r Snail count fishing model with protection linear re}
snail_fishing_mod_re <- glmmTMB(sqrt(snail_dens) ~
                                      protect_status +
                                  #This fits a random intercept effect
                                  (1|region1),
                                    data = env_mod_data_hon,
                                    family = gaussian(link = "identity"))
```

```{r Snail count fishing model re output, echo=FALSE}
#This chunk is similar to the previous model's
snail_fishing_mod_re_resids <- simulateResiduals(snail_fishing_mod_re)
plot(snail_fishing_mod_re_resids)

hist(residuals(snail_fishing_mod_re))
shapiro.test(residuals(snail_fishing_mod_re))

#Looking at model results
summary(snail_fishing_mod_re)
```


```{r Plotting the linear snail fishing model with protection re, echo=FALSE}
#As I have not specified for the predict function to predict the random effects
#it is predicting over the average of all the random effect levels.
pred_snail_fishing_mod_re <- ggpredict(snail_fishing_mod_re, terms = "protect_status")

ggplot(pred_snail_fishing_mod_re, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Turban Snail Counts", x = "Protection Status")+
  theme.figure()
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail fishing model re residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_snail_fishing_mod_re <- residuals(snail_fishing_mod_re, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data_hon, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_snail_fishing_mod_re, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_snail_fishing_mod_re, lw_inverse,
           alternative = "two.sided")
```
</div>
No autocorrelation.

#### Model 10: Fishing model on the snail counts with a fixed effect of region (glm)

Again, we wanted to see what a fixed effect of region would do instead of the
random effect.

```{r Snail count fishing model with protection linear fe}
snail_fishing_mod_fe <- glmmTMB(sqrt(Turban_Snails) ~
                                      protect_status + 
                                  region1,
                                    data = env_mod_data_hon,
                                    family = gaussian(link = "identity"))
```

```{r Snail count fishing model with protection linear fe output, echo=FALSE}
#Checking resids
snail_fishing_mod_fe_resids <- simulateResiduals(snail_fishing_mod_fe)
plot(snail_fishing_mod_fe_resids)

hist(residuals(snail_fishing_mod_fe))
shapiro.test(residuals(snail_fishing_mod_fe))

#Looking at model results
summary(snail_fishing_mod_fe)
```
Residuals not normal for this one and not looking great.

```{r Plotting the linear snail fishing model with protection fe, echo=FALSE}
#Again this is just looking at the overall effect of protection across regions
pred_snail_fishing_mod_fe <- ggpredict(snail_fishing_mod_fe, terms = "protect_status")

ggplot(pred_snail_fishing_mod_fe, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Turban Snail Counts", x = "Protection Status")+
  theme.figure()
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail fishing model linear no openness fe residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_snail_fishing_mod_fe <- residuals(snail_fishing_mod_fe, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data_hon, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_snail_fishing_mod_fe, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_snail_fishing_mod_fe, lw_inverse,
           alternative = "two.sided")
```
</div>
No autocorrelation.

#### Comparing Turban Snail fishing models

```{r AIC of the Turban snail fishing models}
AICc(snail_fishing_mod, 
    snail_fishing_mod_re,
    snail_fishing_mod_fe)%>%
  arrange(AICc)%>%
  mutate(delta_AICc = AICc - AICc[1])
```

AIC indicates that the simpler model without region as a random or fixed effect
is the best model.


### Red Sea Urchin fishing models

#### Plot of urchin counts between protection status

```{r Plot of number of urchins between protected and non-protected sites}
ggplot(env_mod_data_hon, aes(x = protect_status, y = Urchins))+
  geom_boxplot()+
  geom_jitter()+
  labs(x = "Protection Status", y = "Red Sea urchin counts")+
  theme.figure()
```


#### Model 11: Fishing model on the urchin counts (glm)

```{r Checking variance between protected and non-protected for urchins, echo=FALSE}
leveneTest(Urchins ~ protect_status, data = env_mod_data_hon)
```
Equal variance between protected and non-protected sites so assumption of 
homogeneity of variance is met.


```{r Urchin fishing model glm}
urchin_fishing_mod_glm <- glmmTMB(urchin_dens ~ protect_status,
                               family = gaussian(link = "identity"),
                                   data = env_mod_data_hon)
```

```{r Urchin fishing mod glm output, echo=FALSE}
#Checking resids
urchin_fishing_mod_glm_resids <- simulateResiduals(urchin_fishing_mod_glm)
plot(urchin_fishing_mod_glm_resids)

hist(residuals(urchin_fishing_mod_glm))
shapiro.test(residuals(urchin_fishing_mod_glm))

#Looking at model results
summary(urchin_fishing_mod_glm)
```
Residuals look ok for just using a gaussian distribution. Seeing a non-significant
negative effect of protection status on red sea urchins. That is, there is a 
trend of less red sea urchins in protected sites.

```{r Plotting the effect of protection status on urchin counts, echo=FALSE}
pred_urchin_fishing_mod_glm <- ggpredict(urchin_fishing_mod_glm, terms = "protect_status")

ggplot(pred_urchin_fishing_mod_glm, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(x = "Protection Status", y = "Predicted Urchin Counts")+
  theme.figure()
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Urchin fishing model glm residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_urchin_fishing_mod_glm <- residuals(urchin_fishing_mod_glm, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data_hon, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 325)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_urchin_fishing_mod_glm, lw,
           alternative = "two.sided")


#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_urchin_fishing_mod_glm, lw_inverse,
           alternative = "two.sided")
```
</div>
Seeing some autocorrelation in the inverse method but not the neighbour method.


#### Model 12: Fishing model on the urchin counts with random effect (glmm)

Similar to the previous sets of models we wanted to try the urchin fishing
models with a random effect of region as well.

```{r Urchin fishing model glmm}
urchin_fishing_mod_glmm <- glmmTMB(urchin_dens ~ protect_status + (1|region1),
                               family = gaussian(link = "identity"),
                                   data = env_mod_data_hon)
```

```{r Urchin fishing mod glmm output, echo=FALSE}
#Checking resids
urchin_fishing_mod_glmm_resids <- simulateResiduals(urchin_fishing_mod_glmm)
plot(urchin_fishing_mod_glmm_resids)

hist(residuals(urchin_fishing_mod_glmm))
shapiro.test(residuals(urchin_fishing_mod_glmm))

#Looking at model results
summary(urchin_fishing_mod_glmm)

#Looking at some of the variance metrics for the model
performance::r2(urchin_fishing_mod_glmm)
#This returns the adjusted and unadjusted intraclass correlation coefficients
#of the model
performance::icc(urchin_fishing_mod_glmm)

#Identifying the different intercepts of the random effect of region
coef(urchin_fishing_mod_glmm)
```
Residuals look ok for just using a gaussian distribution. Seeing a non-significant
negative effect of protection status on red sea urchins. That is, there are less
red sea urchins in protected sites.



```{r Plotting the effect of protection status on urchin counts glmm, echo=FALSE}
#This is predicting over all of the levels of the random effect
pred_urchin_fishing_mod_glmm <- ggpredict(urchin_fishing_mod_glmm, terms = c("protect_status"))

ggplot(pred_urchin_fishing_mod_glmm, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(x = "Protection Status", y = "Predicted Urchin Counts")+
  theme.figure()
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Urchin fishing model glmm residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_urchin_fishing_mod_glmm <- residuals(urchin_fishing_mod_glmm, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data_hon, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_urchin_fishing_mod_glmm, lw,
           alternative = "two.sided")


#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_urchin_fishing_mod_glmm, lw_inverse,
           alternative = "two.sided")
```
</div>
No autocorrelation with random effect included.

#### Model 13: Fishing model on the urchin counts with fixed effect (glm)

Investigating the impacts of including region as a fixed effect

```{r Urchin fishing model fe glm}
urchin_fishing_mod_glm_fe <- glmmTMB(urchin_dens ~ protect_status + region1,
                               family = gaussian(link = "identity"),
                                   data = env_mod_data_hon)
```


```{r Urchin fishing mod fe glm output, echo=FALSE}
#Checking resids
urchin_fishing_mod_glm_fe_resids <- simulateResiduals(urchin_fishing_mod_glm_fe)
plot(urchin_fishing_mod_glm_fe_resids)

hist(residuals(urchin_fishing_mod_glm_fe))
shapiro.test(residuals(urchin_fishing_mod_glm_fe))

#Looking at model results
summary(urchin_fishing_mod_glm_fe)
performance::r2(urchin_fishing_mod_glm_fe)
```
Residuals look ok for just using a gaussian distribution. Seeing a non-significant
negative effect of protection status on red sea urchins. That is, there are less
red sea urchins in protected sites. Main effects that are evident is regional 
effect with broken group and sarita having lower urchins than bamfield.


```{r Plotting the effect of protection status on urchin counts glm fe, echo=FALSE}
#This is predicting across both the protection status and between regions to 
#visualize the effects these predictors are having
pred_urchin_fishing_mod_glm_fe <- ggpredict(urchin_fishing_mod_glm_fe, terms = c("protect_status", "region1"))


ggplot(pred_urchin_fishing_mod_glm_fe, aes(x = group, y = predicted, colour = x))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(x = "Protection Status", y = "Predicted Urchin Counts")+
  theme.figure()
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Urchin fishing model fe glm residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_urchin_fishing_mod_glm_fe <- residuals(urchin_fishing_mod_glm_fe, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data_hon, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_urchin_fishing_mod_glm_fe, lw,
           alternative = "two.sided")


#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_urchin_fishing_mod_glm_fe, lw_inverse,
           alternative = "two.sided")
```
</div>
No autocorrelation with fixed effect included.

##### Testing the model for VIF stuff

```{r Checking the collinearity of the model}
#Used this function to test for collinearity in the model because there was 
#an issue with the vif function
performance::check_collinearity(urchin_fishing_mod_glm_fe,
                   component = c("all"))
```

#### Comparing urchin fishing models

```{r AIC of the urchin fishing models}
AICc(urchin_fishing_mod_glm,
    urchin_fishing_mod_glmm, 
    urchin_fishing_mod_glm_fe)%>%
  arrange(AICc)%>%
  mutate(delta_AICc = AICc - AICc[1])
```

Model with the random effect included has slightly worse AIC, but it is within
2 points even with an additional degree of freedom. The random effect model also
eliminates the autocorrelation. 
Model without the region included is the best based on AIC.
Because the model with the random effect is within 2 AIC units and it eliminates
the problem of spatial autocorrelation we went with this model.


## Snail to Urchin models{.tabset}

The relative densities of red turban snails and red sea urchins were assessed as a potential metric indicating the health of rocky reef ecosystems in coastal, temperate waters. The relative densities of these species can provide an estimate of reef health as higher densities of urchins are damaging (Breen & Mann, 1976), while turban snail densities give an estimate of the local kelp productivity (Shears & Babcock, 2003), which could balance out the effects of high urchin densities (Harrold & Reed, 1985). To use this metric, we calculated the log ratio of the number of red turban snails to the number of red sea urchins counted along a transect (Equation 1). The log ratio centers the metric around zero, with positive numbers indicating relatively greater snail density and negative numbers indicating relatively greater urchin density. Additionally, the log ratio means that an equivalent difference in the species’ densities in either direction results in the same ratio magnitude. As there were zero counts for each species at some sites, we added one to the count of each species (Equation 1) to ensure that all values were positive and non-zero. By having the ratio be snails to urchins, higher values indicated a healthier site.

We then modelled the differences in this metric between protected and unprotected sites to see if the protected sites were effective at making sites healthier.

### Model 14: Snail urchin ratio fishing model (glm)


```{r Checking variance between protected and non-protected for sn urch ratio, echo=FALSE}
leveneTest(sn_urch_ratio ~ protect_status, data = env_mod_data_hon)
```
Variance is equal between protection status.

```{r Snail ratio protection GLM model}
sn_urch_ratio_protection_model_glm <- glmmTMB(sn_urch_ratio ~ protect_status, 
                          data = env_mod_data_hon,
                          family = gaussian(link = "identity"))
```

```{r Sn urch ratio protection mod glm output, echo=FALSE}
#Checking resids
sn_urch_ratio_protection_model_glm_resids <- simulateResiduals(sn_urch_ratio_protection_model_glm)
plot(sn_urch_ratio_protection_model_glm_resids)


hist(residuals(sn_urch_ratio_protection_model_glm))
shapiro.test(residuals(sn_urch_ratio_protection_model_glm))

#Looking at model results
summary(sn_urch_ratio_protection_model_glm)
performance::r2(sn_urch_ratio_protection_model_glm)
```

```{r Plotting the effect of protection status on snail ratio in glm, echo=FALSE}
pred_sn_urch_ratio_protection_model_glm <- ggpredict(sn_urch_ratio_protection_model_glm, terms = "protect_status")

ggplot(pred_sn_urch_ratio_protection_model_glm, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Log Snail to Urchin ratio", x = "Protection Status")+
  theme.figure()
```

Seeing a strong positive effect of protection

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing sn urch ratio model linear residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_sn_urch_ratio_protection_model_glm <- residuals(sn_urch_ratio_protection_model_glm, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data_hon, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_sn_urch_ratio_protection_model_glm, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_sn_urch_ratio_protection_model_glm, lw_inverse,
           alternative = "two.sided")
```
</div>

Some correlation with the inverse distance method.


### Model 15: Snail urchin ratio fishing model (glmm)

```{r Snail ratio protection re GLMM model}
sn_urch_ratio_protection_model_glmm <- glmmTMB(sn_urch_ratio ~ protect_status +
                          (1|region1), 
                          data = env_mod_data_hon,
                          family = gaussian(link = "identity"))
```

```{r Snail urchin ratio protection mod re glmm output, echo=FALSE}
#Checking resids
sn_urch_ratio_protection_model_glmm_resids <- simulateResiduals(sn_urch_ratio_protection_model_glmm)
plot(sn_urch_ratio_protection_model_glmm_resids)

hist(residuals(sn_urch_ratio_protection_model_glmm))
shapiro.test(residuals(sn_urch_ratio_protection_model_glmm))

#Looking at model results
summary(sn_urch_ratio_protection_model_glmm)
performance::r2(sn_urch_ratio_protection_model_glmm)

#Checking for significance of random effect
performance::icc(sn_urch_ratio_protection_model_glmm)
```

Also seeing a significant positive effect of protection with region included as a random effect.

```{r Plotting the effect of protection status on snail ratio in re glmm, echo=FALSE}
#Predicting across all levels of the random effects
pred_sn_urch_ratio_protection_model_glmm <- ggpredict(sn_urch_ratio_protection_model_glmm, terms = "protect_status")

ggplot(pred_sn_urch_ratio_protection_model_glmm, aes(x = x, y = predicted))+
   geom_boxplot(data = env_mod_data_hon, aes(x = protect_status, y = sn_urch_ratio))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Log Snail to Urchin ratio", x = "Protection Status")+
  theme.figure()
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing sn urch ratio model linear re residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_sn_urch_ratio_protection_model_glmm <- residuals(sn_urch_ratio_protection_model_glmm, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data_hon, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_sn_urch_ratio_protection_model_glmm, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_sn_urch_ratio_protection_model_glmm, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation.

### Model 16: Snail urchin ratio fishing model with fixed effect (glm)

```{r Snail ratio protection fe GLM model}
sn_urch_ratio_protection_model_glm_fe <- glmmTMB(sn_urch_ratio ~ protect_status +
                          region1, 
                          data = env_mod_data_hon,
                          family = gaussian(link = "identity"))
```

```{r Snail urch ratio protection mod fe glm output, echo=FALSE}
#Checking resids
sn_urch_ratio_protection_model_glm_fe_resids <- simulateResiduals(sn_urch_ratio_protection_model_glm_fe)
plot(sn_urch_ratio_protection_model_glm_fe_resids)

hist(residuals(sn_urch_ratio_protection_model_glm_fe))
shapiro.test(residuals(sn_urch_ratio_protection_model_glm_fe))

#Looking at model results
summary(sn_urch_ratio_protection_model_glm_fe)
performance::r2(sn_urch_ratio_protection_model_glm_fe)
```

Effect of protection status is no longer significant but the trend is in the same direction, Sarita is significantly lower than Deer Group bamfield

```{r Plotting the effect of protection status on snail ratio in fe glm, echo=FALSE}
pred_sn_urch_ratio_protection_model_glm_fe <- ggpredict(sn_urch_ratio_protection_model_glm_fe, terms = c("protect_status", "region1"))

ggplot(pred_sn_urch_ratio_protection_model_glm_fe, aes(x = group, y = predicted, colour = x))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Log Snail to Urchin ratio", x = "Protection Status")+
  theme.figure()
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing sn urch ratio model linear fe residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_sn_urch_ratio_protection_model_glm_fe <- residuals(sn_urch_ratio_protection_model_glm_fe, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data_hon, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_sn_urch_ratio_protection_model_glm_fe, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_sn_urch_ratio_protection_model_glm_fe, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation.


##### Testing for the VIF of the model
```{r Testing for multicollinearity of the predictors snail to urchin}
#Using this function because there was a problem with the vif function
performance::check_collinearity(sn_urch_ratio_protection_model_glm_fe,
                   component = c("all"))
```

Correlation does not seem to be problematic.


### Comparing the snail to urchin ratio models

```{r AIC of the snail urchin ratio models without openness}
AICc(sn_urch_ratio_protection_model_glm,
     sn_urch_ratio_protection_model_glmm,
    sn_urch_ratio_protection_model_glm_fe)%>%
  arrange(AICc)%>%
  mutate(delta_AICc = AICc - AICc[1])
```


AIC indicated that the fixed effect region was the best model but we ultimately decided that it was not asking the question we wanted to look at. The random effect model was marginally better than the one without the random effect and it corrected for the spatial autocorrelation.


# Creating figures for the thesis

## Making a theme for all subsequent figures

```{r Making theme for all subsequent figures, echo=FALSE}
theme.figure.report <- function (){
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    #axis.line = element_line(colour = "black"),
    axis.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 14, face = "plain"),
    legend.text = element_text(size = 12, face = "plain"),
    legend.title = element_text(size = 14, face = "bold")) 
}
```

## Set up for site maps

### Making the map ranges

```{r Seting the ranges for the maps}
#In this chunk I made limits for the maps so that they show the view of 
#Barkley Sound that I want
lat_min = min(env_mod_data_hon$latitude) - 0.05
lat_max = max(env_mod_data_hon$latitude) + 0.05

long_min = min(env_mod_data_hon$longitude) - 0.05
long_max = max(env_mod_data_hon$longitude) + 0.05
```

### Loading in the low res shape file

```{r Loading in the base layer}
#I loaded in a lower res shape file that shows all of BC in order to make a
#large inset plot of Vancouver Island. Similar to the high res shape file I 
#subsetted it to only include the area around Vancouver Island.

VI_base_layer <- read_sf(dsn = here("./01_raw_data/canada_sf_data", ""), 
                         stringsAsFactors = F)
VI_base_layer_sf <- st_as_sf(VI_base_layer)

#specifying the projection for the sf
st_crs(VI_base_layer_sf) <- "+proj=longlat +datum=WGS84 +no_defs"


shp_xmin_low <- -129
shp_xmax_low <- -122.5
shp_ymin_low <- 48
shp_ymax_low <- 52

bbox_low <- st_bbox(c(xmin = shp_xmin_low, ymin = shp_ymin_low, xmax = shp_xmax_low, ymax = shp_ymax_low))

bbox_low_sf <- st_as_sfc(bbox_low)

VI_base_layer_sf_subset <- st_crop(VI_base_layer_sf, bbox_low_sf)
#plot(VI_base_layer_sf_subset)
```


### Loading in the shapefiles for the protected areas

In order to show the protected areas within Barkley Sound I used Google Earth to make kml files with polygons of the protected areas.

```{r loading in the RCA polygon}
#st_read loads in the kml file and stores it as a polygon
broken_group_RCA <- st_read(here("./01_raw_data/Broken_group_RCA.kml"))
```

```{r Loading in the Numukamis area}
numukamis_area <- st_read(here("./01_raw_data/Numukamis_closed_area.kml"))
```

```{r Loading in the Baeria rocks}
baeria_rocks_er <- st_read(here("./01_raw_data/baeria_rocks.kml"))
```

### Loading in the flight track for the DFO surveys

I also traced the DFO flight path used for the DFO flyover surveys in Barkley Sound using Google Earth and saved it as a kml file.

```{r Loading in the DFO flight track for Barkley Sound}
#This loads in the flight path so that it can be put on the plot.
dfo_flight_track <- st_read(here("./01_raw_data/dfo_new_flight_path.kml"))
```


###Shifting a couple of the sites to make it clearer

In order to not have a bunch of the sites overlapping each other when plotting maps I shifted some of the points slightly.

```{r Shifting some sites}
env_mod_data_hon_minor_shifted <- env_mod_data_hon%>%
  #These lines of code change the lat or long of the specified site to the given value
 mutate(longitude = if_else(site_code == "BMSC11", -125.153, longitude))%>%
  mutate(latitude = if_else(site_code == "BMSC5", 48.82333, latitude))%>%
  mutate(latitude = if_else(site_code == "BMSC9", 48.83179, latitude))%>%
  mutate(longitude = if_else(site_code == "BMSC9", -125.1490, longitude))
```

## Figure 1: Overall site map

```{r Making the inset map}
#This makes a map of Vancouver Island
(van_isl_map <- ggplot(VI_base_layer_sf_subset) +
   #allows visualization of sf objects
  geom_sf() +
   #Sets limits for the plot
  coord_sf(ylim = c(48.3, 50.8), xlim = c(-128.2, -123.2)) +
  theme.figure.report() +
   #Makes a red rectangle around Barkley Sound
  geom_rect(xmin = long_min, xmax = long_max, ymin = lat_min, ymax = lat_max,
            color = "red", fill = NA, linewidth = 1) +
  labs(x = "Longitude", y = "Latitude")+
   theme(axis.text = element_blank(),
         axis.title = element_blank(),
         axis.ticks = element_blank()))
```


```{r Overall site map with the site locations}
(overall_site_map <- ggplot()+
    geom_sf(data = VI_base_layer_hi_res_sf_subset, colour = "darkgrey", fill = "lightgrey")+
   #The lines below plot the pl=olygons of the different protected areas within
   #the sound
     geom_sf(data = broken_group_RCA, fill = "blue", colour ="blue", alpha = 0.1)+
   geom_sf(data = numukamis_area, fill = "blue", colour = "blue", alpha = 0.2)+
   geom_sf(data = baeria_rocks_er, fill = "blue", colour = "blue", alpha = 0.2)+
   #This adds in the points of the survey sites, shifted slightly to be not 
   #overlapping
     geom_point(env_mod_data_hon_minor_shifted, 
              mapping = aes(x = longitude, y = latitude, fill = protect_status,
                            shape = region1), size = 3)+
   #This specifies the different shapes to be used to differentiate
   #between the different regions
   scale_shape_manual(values = c(24, 21, 22, 23))+
   #Specifies the colours for unprotected and protected sites
   scale_fill_manual(values = c("#F46864", "#1CB8BC"))+
       coord_sf(xlim = c(long_min + 0.03, long_max - 0.035),
           ylim = c(lat_min + 0.01, lat_max - 0.04)) +
  labs(x = "Longitude", y = "Latitude", shape = "Region",
       fill = "Protection\nStatus") +
  theme.figure.report() +
  scale_x_continuous(breaks = c(-125.35, -125.20, -125.05)) +
 # scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  #Adds in an arrow indicating north
   annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
   #adds in a scale bar
  annotation_scale(location = "br", text_cex = 1.3)+
    theme(legend.position = "inside",
      legend.position.inside = c(0.84, 0.2),
         axis.text.y = element_text(angle = 45),
      legend.text = element_text(size = 12),
      legend.title = element_text(size = 14),
      legend.margin = margin(0.3, 0.3, 0.3, 0.3)))
```


```{r Making the site map with the van isl inset, fig.cap= Figure 1. Locations of Reef Life Survey study sites (n = 26) within Barkley Sound, BC, Canada, which were sampled in April and May of 2024. The colours indicate the protection-from-fishing status of each site. Some sites have been shifted slightly to prevent overlap. Region classifications are indicated by the shapes with the region codes corresponding to: BG = Broken Group, DIEC = Deep Imperial Eagle Channel, DGB = Deer Group/Bamfield, and NMB = Numukamis Bay. Areas protected from fishing are shown with blue shade: farthest west = Broken Group Rockfish Conservation Area, central north = Baeria Rocks Ecological Reserve, farthest east = Numukamis Bay Finfish Closure Area. Inset is a map of Vancouver Island illustrating the location of Barkley Sound (red square).}
(site_map_w_inset <- ggdraw(overall_site_map) +
  draw_plot(van_isl_map, x = 0.1, y = 0.15, width = 0.5, height = 0.3))
```

This was the rough version of the plot and then I saved it as a pdf and used inkscape to fine tune the locations of the legends and inset.

```{r Saving the site map}
ggsave("site_map_with_inset_regions.pdf", plot = site_map_w_inset,
       path = here("./04_plots/report_figures"), width = 165, height = 130, units = "mm")
```


## Figure 2: Temperature plot for Barkley Sound

```{r Converting the raster to a data frame so ggplot can use it}
avg_temp_df <- as.data.frame(average_temperature, xy = TRUE, na.rm = TRUE)
colnames(avg_temp_df) <- c("longitude", "latitude", "average_temp")

avg_temp_df <- avg_temp_df%>%
  mutate(avg_temp_c = average_temp - 273.15)
```

```{r Making the temperature map, fig.cap=Figure 2. Mean sea surface temperature (SST) in Barkley Sound, BC, Canada, from daily temperature measurements from the Global Ocean OSTIA Sea Surface Temperature and Sea Ice Analysis product (European Union-Copernicus Marine Service, 2015). Mean SST was calculated by averaging the daily temperatures within 0.05º latitude x 0.05º longitude cells for the year before Reef Life Survey subtidal surveys within Barkley Sound in April and May of 2024 (2023-04-22 to 2024-05-03).}
(temp_map_23_24 <- ggplot() +
   #Plotting the temp data as a raster with the shape file
  geom_raster(data = avg_temp_df, aes(x = longitude, y = latitude, fill = avg_temp_c)) +
   geom_sf(data = VI_base_layer_hi_res_sf_subset, colour = "#424A4E", fill = "lightgrey")+
  coord_sf(xlim = c(long_min + 0.029, long_max - 0.032),
           ylim = c(lat_min , lat_max - 0.025)) +
     scale_x_continuous(breaks = c(-125.35, -125.20, -125.05)) +
   #Setting a colour gradient from low to high temp
     scale_fill_gradient(
    low = "lightblue",          
    high = "#E95E48"
  ) +
  theme.figure.report() +
  labs(x = "Longitude",
       y = "Latitude",
       fill = "Mean SST (ºC)") +
  theme(legend.position = "inside",
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        axis.text.y = element_text(angle = 45))+
  annotation_scale(location = "br", text_cex = 1.3)+
   annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering))
```

Fine tuning of the legend location and text was done in inkscape.

```{r Saving the site map}
ggsave("temp_map_23_24.pdf", plot = temp_map_23_24,
       path = here("./04_plots/report_figures"), width = 165, height = 130, units = "mm")
```


## Figure 3: Example of Rec boats with radii

For this plot we used 2023 and 1.5 km radii as examples of the data we saw for the rec boat counts

```{r total rec boats within 1.5 km, fig.cap=Figure 3. Example of the raw data from DFO flyover surveys (June 1–September 30, 2023) indicating the locations of recreational boats fishing (black dots) with 1.5 km radii (red-filled circles) around each Reef Life Survey site (open black dots) in Barkley Sound, BC, Canada. Subtidal surveys were conducted at each site in April and May of 2024. The DFO surveys sometimes grouped boats that were close together as one point, so some points indicate multiple boats, but this has been omitted for ease of viewing. The flight path taken by DFO surveyors is shown (dashed blue line) with arrows showing the direction of flight. DFO surveys were completed in the summers of 2021-23. The mean numbers of recreational boats fishing within 1.0, 1.5, and 2.0 km radii per year were calculated to compare fishing activity between protected and unprotected sites.}
(total_rec_boats_map_2023_test <- ggplot()+
    geom_sf(data = VI_base_layer_hi_res_sf_subset, colour = "darkgrey", fill = "lightgrey")+
   #This plots the DFO plane flight path
   geom_sf(data = dfo_flight_track, colour =  "#355173", size = 3)+
  geom_point(data = rec_boats_fishing2023, 
             mapping = aes(x = Longitude, y = Latitude),
             size = 0.8, fill = "#070707")+
   #This plots the 1.5km radii around each RLS site
    geom_sf(data = site_locations_radius_1.5km, fill = "#E95E48", alpha = 0.3)+
   #this plots the RLS sites
    geom_point(env_mod_data_hon, 
              mapping = aes(x = longitude, y = latitude),
              shape = 21, size = 3, colour = "black")+
    coord_sf(xlim = c(long_min + 0.02, long_max - 0.035),
           ylim = c(lat_min + 0.01, lat_max - 0.03)) +
  labs(x = "Longitude", y = "Latitude") +
  theme.figure.report() +
   theme(axis.text.y = element_text(angle = 45))+
 scale_x_continuous(breaks = c(-125.35, -125.20, -125.05)) +
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br", text_cex = 1.3))
```

A legend was added and the flight path line styling was edited using inkscape.

```{r Saving the rec boats plot with dfo flight track}
ggsave("total_rec_boat_map_2023_1.5km_track.pdf", plot = total_rec_boats_map_2023_test,
       path = here("./04_plots/report_figures"), width = 160, height = 120, units = "mm")
```


## Figure 4: Boxplots of number of boats between protected and unprotected sites

```{r Boxplot of mean rec boats between protection levels 1.5 km}
(rec_boats_boxplot_1.5km <- ggplot(unique_site_df, aes(x = protect_status,
                                                     y = mean_rec_boat_counts_1.5km,
                                                     fill = protect_status))+
  geom_boxplot(alpha = 0.6, width = 0.4)+
   #Adding jittered raw data but only jittering on the x axis so it is the 
   #true boat count values
  geom_jitter(height = 0, width = 0.2, colour = "darkgrey", alpha = 0.5)+
  labs(x = "Protection Status", y = "Mean rec.\nboat counts")+
     scale_fill_manual(values = c("#E95E48", "#355173"))+
  theme.figure.report()+
   theme(legend.position = "none",
     #    axis.text.x = element_blank(),
         axis.title.y = element_blank()))
      #   axis.ticks.x = element_blank()))
```

```{r Boxplot of mean rec boats between protection levels 1 km}
(rec_boats_boxplot_1km <- ggplot(unique_site_df, aes(x = protect_status,
                                                     y = mean_rec_boat_counts_1km,
                                                     fill = protect_status))+
  geom_boxplot(alpha = 0.6, width = 0.4)+
  geom_jitter(height = 0, width = 0.2, colour = "darkgrey", alpha = 0.5)+
  labs(x = "Protection Status", y = "Mean rec.\nboat counts")+
     scale_fill_manual(values = c("#E95E48", "#355173"))+
  theme.figure.report()+
   theme(legend.position = "none"))
     #    axis.text.x = element_blank(),
         #axis.title.y = element_blank()))
      #   axis.ticks.x = element_blank()))
```

```{r Boxplot of mean rec boats between protection levels 2 km}
(rec_boats_boxplot_2km <- ggplot(unique_site_df, aes(x = protect_status,
                                                     y = mean_rec_boat_counts_2km,
                                                     fill = protect_status))+
  geom_boxplot(alpha = 0.6, width = 0.4)+
  geom_jitter(height = 0, width = 0.2, colour = "darkgrey", alpha = 0.5)+
  labs(x = "Protection Status", y = "Mean rec.\nboat counts",
       fill = "Protection Status")+
     scale_fill_manual(values = c("#E95E48", "#355173"))+
  theme.figure.report()+
   theme(legend.position = "none",
     axis.title.y = element_blank()))
```

```{r Combining the three plots into one, fig.cap=Figure 4. Distributions of the mean number of recreational boats fishing (per year) within A) 1.0, B) 1.5, and C) 2.0 km of Reef Life Survey sites in Barkley Sound, BC, Canada, separated by protection-from-fishing status (NP = No Protection, P = Protected). Raw data are shown with grey circles and have been jittered across the x-axis for ease of viewing. Means were calculated from DFO flyover survey data of recreational fishing boat locations from 2021-23. Asterisks indicate level of significance based on Wilcoxon rank sum tests (* = P < 0.05, ** = P < 0.01, *** = P < 0.001).}
rec_boats_overall_boxplots <- rec_boats_boxplot_1km + rec_boats_boxplot_1.5km + rec_boats_boxplot_2km

rec_boats_overall_boxplots
```

This was imported to inkscape where labels for the radii used were added (1, 1.5, 2, from left to right). Also the x axis labels were edited to just be Np and P

```{r Saving the rec boat counts boxplot figure}
ggsave("rec_boats_boxplot_overall.pdf", plot = rec_boats_overall_boxplots,
       path = here("./04_plots/report_figures"), width = 180, height = 80, units = "mm")
```


## Figure 5: Relationship of snail density with openness

```{r Making a plot of the effect of openness on log snails, fig.cap=Figure 5. The relationship of openness (normalized) with the log-transformed red turban snail (P. gibberosus) density from Reef Life Survey sites in Barkley Sound, BC, Canada, in April and May of 2024. Raw data (black points) are shown with a line of best fit (blue line) and 95% confidence intervals (blue shade).}
(pred_snail_openness_plot_log <- ggplot(env_mod_data_hon, aes(x = norm_openness, y = log(snail_dens + 0.01)))+
  geom_point()+
   #This fits a linear trend line with 95% CI
  geom_smooth(method = "lm", colour = "#637683", fill = "#8EBAD2")+
  labs(x = "Normalized Openness", y = "Snail density (log(n/m2 + 0.01))")+
  theme.figure.report())
```

Figure was further refined in inkscape

```{r Saving the snail openness plot}
ggsave("pred_snail_openness_plot_log.pdf", plot = pred_snail_openness_plot_log,
       path = here("./04_plots/report_figures"), width = 100, height = 100, units = "mm")
```


## Figure 6: Environmental trends for urchins

For this figure we were trying to show some of the trends in the urchin density data based on the environmental predictors of openness and substrate type. We did this by predicting data from the GAM run for the urchin density and then creating the plots with the predicted and raw data.


#### Plot of effect of seafloor type


##### Generating predicted data for plot

```{r predicting urchin density based on dominant seafloor}
#This block of code creates a df with data necessary to supply to the predict
#function in order to have it predict the fitted values of the urchin density
urchin_ecology_data_for_seafloor <- data.frame(
  #This supplies the two levels of the seafloor type
  dominant_score = levels(env_mod_data_hon$dominant_score),
  #This supplies the mean value of the scaled temperature data
  sst_scale = mean(env_mod_data_hon$sst_scale),
  #This supplies the mean value of the scaled openness data
  norm_openness_scale = mean(env_mod_data_hon$norm_openness_scale),
  #This supplies the Deer Group bamfield region as this was set as the reference
  #level in the model
  region1 = factor("DGB", levels = levels(env_mod_data_hon$region1))
)
#The resulting df has two rows, one each for loose and solid seafloor type, and
#then columns for seafloor type, temp, openness, and region


#This block of code is used to actually predict the fitted values of the urchin
#density data based on the dataframe we made above.

#This uses the predict function to predict from the GAM, using the data frame
#we supplied above as the data to predict upon. Setting se.fit to TRUE means
#that it will also estimate the standard error for each prediction.
predicted_sf_effect_urchin <- predict(ecology_urchin_mod_re, newdata = urchin_ecology_data_for_seafloor, se.fit = TRUE)

#This adds a column with the fitted values to the urchin ecology data for 
#seafloor that we made above
urchin_ecology_data_for_seafloor$fit <- predicted_sf_effect_urchin$fit

#This adds a column with the upper 95% CL for the predicted value based on the 
#estimated SE
urchin_ecology_data_for_seafloor$upper <- predicted_sf_effect_urchin$fit + 1.96 * predicted_sf_effect_urchin$se.fit

#This adds a column for the lower 95% Cl
urchin_ecology_data_for_seafloor$lower <- predicted_sf_effect_urchin$fit - 1.96 * predicted_sf_effect_urchin$se.fit
```

##### Creating Plot

```{r Plotting the predicted urchin density from seafloor type}
(predicted_urch_dens_dom_score <- ggplot(urchin_ecology_data_for_seafloor, aes(x = dominant_score, y = fit))+
   #This adds a greyed out boxplot in the background that shows the raw data
  geom_boxplot(data = env_mod_data_hon, mapping = aes(x = dominant_score, y = urchin_dens),
                alpha = 0.6, width = 0.5, colour = "grey")+
   #Adds 95% CI
    geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.15, linewidth = 1.2)+
   #Adds predicted points from the model predictions
  geom_point(aes(colour = dominant_score), size = 4)+
   scale_colour_manual(values = c("#EAB964", "#5E666F"))+
  labs(x = "Dominant seafloor type", y = "Urchin density (n/m2)")+
  theme.figure.report()+
   theme(legend.position = "none"))
```

So ultimately, for this plot the model predictions are the predicted urchin density between the two seafloor types within the Deer Group Bamfield region.



#### Plot of effect of openness

##### Generating predicted data for plot

```{r Trying to predict the urchin counts from openness}
#This block of code generates a df that is used in the predict function as the 
#data to predict upon.
urchin_ecology_data_for_openness <- data.frame(
  #This supplies the level of seafloor type as loose as this was the reference
  #level
  dominant_score = factor("loose", levels = levels(env_mod_data_hon$dominant_score)),
  #This supplies the mean of the scaled temperature data
  sst_scale = mean(env_mod_data_hon$sst_scale),
  #This generates a sequence of 100 equally spaced openness values between the 
  #max and min values in the scaled openness column
  norm_openness_scale = seq(min(env_mod_data_hon$norm_openness_scale), max(env_mod_data_hon$norm_openness_scale), length.out = 100),
  #This supplies the region level to be Deer Group Bamfield as this was the 
  #reference level
  region1 = factor("DGB", levels = levels(env_mod_data_hon$region1))
)
#The resulting DF has 100 rows of the different openness values and then also
#has a column for region, temp, and seafloor that are all the constants supplied
#above


#getting mean and sd for norm_openness
norm_openness_mean <- mean(env_mod_data_hon$norm_openness)
norm_openness_sd <- sd(env_mod_data_hon$norm_openness)

#This code uses the data supplied above to predict the urchin density at the 
#different openness values from the GAM. se.fit as TRUE means that SE estimates 
#are generated
predicted_openness_effect_urchin <- predict(ecology_urchin_mod_re, newdata = urchin_ecology_data_for_openness, se.fit = TRUE)
#This adds the fitted urchin density values to the df we made above
urchin_ecology_data_for_openness$fit <- predicted_openness_effect_urchin$fit
#This adds a column of upper 95% CL using the SE estimates
urchin_ecology_data_for_openness$upper <- predicted_openness_effect_urchin$fit + 1.96 * predicted_openness_effect_urchin$se.fit
#This adds a column of lower 95% CL using the SE estimates
urchin_ecology_data_for_openness$lower <- predicted_openness_effect_urchin$fit - 1.96 * predicted_openness_effect_urchin$se.fit

#These two lines return the mean and sd of the openness values used in order to 
#backcalculate the scaled openness values to the unscaled form for plotting
norm_openness_mean <- mean(env_mod_data_hon$norm_openness)
norm_openness_sd <- sd(env_mod_data_hon$norm_openness)

#This adds a column of the scaled openness sequence that has been back 
#transformed to the raw openness scale using the mean and SD
urchin_ecology_data_for_openness$norm_openness_original <- urchin_ecology_data_for_openness$norm_openness_scale * norm_openness_sd + norm_openness_mean
```


##### Creating Plot

```{r Plotting the predicted urchin density from openness}
(predicted_openness_effect_urchin <- ggplot(env_mod_data_hon, aes(x = norm_openness, y = urchin_dens))+
   #adds line of the predicted urchin density values against the backcalculated
   #openness sequence
  geom_line(data = urchin_ecology_data_for_openness, 
            mapping = aes(x = norm_openness_original, y = fit), colour = "black", linewidth = 1.25)+
   #adds a 95% confidence band for the predicted values
  geom_ribbon(data = urchin_ecology_data_for_openness, 
            mapping = aes(x = norm_openness_original, y = fit, ymin = lower, ymax = upper), alpha = 0.5, fill = "#E95E48")+
  geom_point()+
  labs(x = "Normalized openness", y = "Urchin density (n/m2)")+
  theme.figure.report())
```


Ultimately, this plot is showing the predicted urchin densities as a function of openness for urchins on loose seafloor in Deer Group Bamfield.

#### Combining to two plots into a multi-panel

```{r Combining the two plots into one, fig.cap=Figure 6. A) Distributions of red sea urchin (M. franciscanus) density between seafloor types (Loose = loose rubble substrate and solid = solid rock substrate) with model predictions (coloured points) and 95% confidence intervals (black error bars) from a generalized additive mixed model (GAMM). B) Relationship of red sea urchin density and openness (normalized) with raw data (black points) and GAMM predictions (black line) with a 95% confidence interval (red shade) shown. Predictions made by holding other continuous variables at their mean, and categorical and random variables at their reference level. For openness predictions, a sequence of 100 openness values was generated. Urchin densities were obtained from subtidal Reef Life Surveys at sites in Barkley Sound, BC, Canada, in April and May of 2024.}
(urchin_gam_plot_updated <- (predicted_urch_dens_dom_score)/(predicted_openness_effect_urchin))
```

This plot was then further cleaned up in inkscape (e.g., added panel letters and urchin symbols)

```{r Saving the plot}
ggsave("urchin_gam_plot_updated.pdf", plot = urchin_gam_plot_updated,
       path = here("./04_plots/report_figures"), width = 100, height = 180, units = "mm")
```



## Figure 7: Plots of denisities and ratios between protection levels

For this figure we were showing the effects of protection from fishing on the urchin and snail densities, and on the ratio of snails to urchins. We did this by predicting the output from the glmmTMB models and then plotting the predictions along with the raw data.

#### Plot of snail density

##### Creating the plot

The predicted values for this plot were already generated previously when we made the plots of the snail fishing models in order to visualize the effects of fishing from the models. 

```{r Making a boxplot of snail counts between protection status}
(snail_fishing_boxplot <- ggplot(env_mod_data_hon, aes(x = protect_status, y = snail_dens))+
   #adds greyed out boxplots in the background with the raw data. Outlier.shape
   #as NA means that outlier points are hidden but the axis is the full height 
   #of the data range
  geom_boxplot(env_mod_data_hon, mapping = aes(x = protect_status, y = snail_dens),
               alpha = 0.6, width = 0.5, colour = "grey", outlier.shape = NA)+
   #This adds the 95% CI. The CL from the prediction are squared to put them 
   #back onto the raw scale
  geom_errorbar(pred_snail_fishing_mod, mapping =
                  aes(ymin = (conf.low^2), ymax = (conf.high^2), x = x, y = predicted),
                width = 0.15, linewidth = 1.2)+
   #Adds the model predictions
   geom_point(pred_snail_fishing_mod, mapping = 
              aes(x = x, y = predicted, colour = x), size = 4)+
  labs(x = "Protection Status", y = expression(Snail~density~(n/m^2)))+
   scale_fill_manual(values = c("#E95E48", "#355173"))+
   scale_colour_manual(values = c("#E95E48", "#355173"))+
  coord_cartesian(ylim = c(0, 4.7))+
  theme.figure.report()+
    theme(legend.position = "none",
          axis.text.x = element_blank(),
         axis.title.x = element_blank(),
         axis.ticks.x = element_blank()))
```
#### Plot of urchin density 

##### Creating the plot

```{r Making a boxplot of urchin counts between protection status}
(urchin_fishing_boxplot <- ggplot(env_mod_data_hon, aes(x = protect_status, y = urchin_dens))+
   #Adds greyed out boxplots in the background with the raw data. 
  geom_boxplot(env_mod_data_hon, mapping = aes(x = protect_status, y = urchin_dens),
               alpha = 0.6, width = 0.5, colour = "grey")+
   #Adds the 95% CI from the model prediction
  geom_errorbar(pred_urchin_fishing_mod_glmm, mapping =
                  aes(ymin = conf.low, ymax = conf.high, x = x, y = predicted),
                width = 0.15, linewidth = 1.2)+
   #adds the points for the model predictions
   geom_point(pred_urchin_fishing_mod_glmm, mapping = 
              aes(x = x, y = predicted, colour = x), size = 4)+
  labs(x = "Protection Status", y = expression(Urchin~density~(n/m^2)))+
   scale_fill_manual(values = c("#E95E48", "#355173"))+
   scale_colour_manual(values = c("#E95E48", "#355173"))+
  theme.figure.report()+
    theme(legend.position = "none",
          axis.text = element_text(size = 16),
          axis.text.x = element_blank(),
         axis.title.x = element_blank(),
         axis.ticks.x = element_blank()))
```

#### Plot of snail to urchin ratio

##### Creating the plot 

```{r Making the snail to urchin ratio boxplot}
(snail_urchin_ratio_box <- ggplot(env_mod_data_hon, aes(x = protect_status, y = sn_urch_ratio))+
  geom_boxplot(env_mod_data_hon, mapping = aes(x = protect_status, y = sn_urch_ratio),
               alpha = 0.6, width = 0.5, colour = "grey")+
  geom_errorbar(pred_sn_urch_ratio_protection_model_glmm, mapping =
                  aes(ymin = conf.low, ymax = conf.high, x = x, y = predicted),
                width = 0.15, linewidth = 1.2)+
   geom_point(pred_sn_urch_ratio_protection_model_glmm, mapping = 
              aes(x = x, y = predicted, colour = x), size = 4)+
  labs(x = "Protection Status", y = "Log snail:urchin\nratio")+
   scale_fill_manual(values = c("#E95E48", "#355173"))+
      scale_colour_manual(values = c("#E95E48", "#355173"))+
  theme.figure.report()+
    theme(legend.position = "none",
          axis.text = element_text(size = 16)))
```

####Combining the three plots into one multi-panel

```{r Combining the snail, urchin, snail:urchin ratio boxplots, fig.cap=Figure 7. Distributions of the A) red turban snail (P. gibberosus) density, B) red sea urchin (M. franciscanus) density, and C) log snail-to-urchin ratios at Reef Life Survey sites in Barkley Sound, BC, Canada, separated by protection-from-fishing status (NP = No Protection, P = Protected). Coloured points are model predictions (A = linear model with square-root-transformed response, B and C = linear mixed-effects models) with 95% confidence intervals (black error bars), and the asterisk indicates P < 0.05. Predictions in B) and C) were averaged over the levels of the random effect of region. One outlier was removed in A).}
#This code assembles the three plots in a stacked fashion.
(snail_urchin_sn_urch_box_plot <- (snail_fishing_boxplot)/
  (urchin_fishing_boxplot)/
  (snail_urchin_ratio_box))
```

```{r Saving the protection status plots}
ggsave("snail_urchin_sn_urch_box_plot.pdf", plot = snail_urchin_sn_urch_box_plot,
       path = here("./04_plots/report_figures"), width = 100, height = 185, units = "mm")
```

This figure was further edited in inkscape to add the panel letters and adjust the text of the axis titles etc.
