---
title: "Alec Jones Honours Thesis Code"
author: "Alec Jones"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
   toc: true
---

   <script src="hideOutput.js"></script> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading in and cleaning the Reef Life Survey Data from Barkley Sound, BC, Canada, in 2024

## Loading in the 2024 data

```{r Reading in the 2024 RLS data, echo=FALSE}
RLS_data_2024 <- read_csv(here("./01_raw_data/RLS_CANADA_Bamfield_2024.csv"))
```

## Cleaning up the data

```{r Cleaning up the dataframe to allow for use in R}
#Making a list of the variables that I want to convert to factors in the step
#below
factor_cols2 <- c("site_code", "site_name", "Code")

#This pipe is taking the original raw dataframe and cleaning it up and storing
#it as a cleaned dataframed
RLS_data_2024_clean <- RLS_data_2024%>%
  #Picking which columns I want to keep
  select(c(`Site No.`, `Site Name`, Latitude, Longitude, Date, vis, Time, Depth,
           Method, Block, Code, Species, Total))%>%
  #Removing columns/rows with N/As
  na.omit()%>%
  #Changing all of the species codes to be lowercase as there were some of the
  #same species with codes differing by one or two capitals
  mutate(Code = str_to_lower(Code))%>%
  #Renaming the variables to be easier to code with
  rename(site_code = `Site No.`,
         site_name = `Site Name`,
         latitude = Latitude,
         longitude = Longitude,
         survey_date = Date,
         depth = Depth,
         visibility = vis,
         hour = Time,
         total = Total)%>%
  #Changing the variables that I specified earlier to be factors
   mutate_at(factor_cols2, factor)%>%
  #One point at the Kirby Site had the latitude value for both latitude and
  #longitude columns so this is fixing that
   mutate(longitude = if_else(site_code == "BMSC2", -125.19876860, longitude))%>% 
  #In the dataset there are codes and counts of non-organisms (ex. dec = Debris
  # cloth). This is removing all the non-organisms from the dataset
  filter(!Code %in% c("dez", "deg", "dem", "deo", "def", "dew", "dec", "nsf", "snd"))%>%
  #This is grouping the data down to the species code within each individual
  #transect
   group_by(site_code, site_name, latitude, longitude, survey_date, depth, 
           visibility, hour, Code)%>%
  #This is summing all of each species spotted at each transect. Ex. adding up
  #all red sea urchins at the Wouwer channel site
  summarise(species_total = sum(total))
```

## Making a dataframe with the species of interest and ratios of interest

I made a dataframe with the counts of red turban snails and red sea urchins, 
along wih the log ratios of the two species to each other.

```{r Making dataframe with urchins, snails, and ratio of the two}
#Starting pipe by taking data from the clean df I made
 urch_sn_ratio_mod_2024 <- RLS_data_2024_clean%>%
  #This makes a column for each species with the numbers of each in each
  #transect (each row)
  pivot_wider(names_from = Code, values_from = species_total)%>%
  #Renaming red sea urchin and red turban snail columns
  mutate(Urchins = mfra,
         Turban_Snails = pgi)%>%
  #Selecting only the columns I wanted and getting rid of the other species
  select(site_code, site_name, latitude, longitude, survey_date, depth, 
           visibility, hour, Urchins, Turban_Snails)%>%
  #If urchins or snails were not counted at a site they were N/A at this stage
  #so this changes those counts to 0
  replace(is.na(.), 0)%>%
  #This calculates the ratios between species and makes a column for each. Date
  #is also converted to day month year form
  mutate(urch_sn_ratio = log10((Urchins +1) / (Turban_Snails + 1)),
         sn_urch_ratio = log10((Turban_Snails + 1) / (Urchins + 1)),
         survey_date = dmy(survey_date))
```


# Working with the potential predictor variables

For my project I incorporated various different predictor variables, some that I
obtained from csv files and some that needed to be downloaded from online
databases.

## Predictor 1: Wave exposure (via metric called Openness)

To get a measure of wave exposure for each of the RLS sites, I used a metric 
called openness from Hill et al. (2010)(10.3354/meps08815).

### Set up prior to calculating fetch for openness

```{r Making dfs of the unique sites and then just the lat and long, echo=FALSE, include=FALSE}
#Making a df with just the unique site info
site_data_2024 <- RLS_data_2024_clean%>%
  distinct(site_code, site_name, latitude, longitude)

#Making a df with just the unique lat and longs
site_coords_2024 <- site_data_2024%>%
  ungroup()%>%
  select(latitude, longitude)%>%
  distinct()

#This converts the site coordinates into a sf object and specifies the 
#coordinate reference system to be the same as the shape file
site_sp_2024 <- st_as_sf(site_coords_2024, coords = c("longitude", "latitude"), crs = 4326)
```

```{r Loading in the shapefiles to do the fetch calculations, include=FALSE}
#This is a high resolution shape file from Hakai. This code loads in the shape
#file and saves it as the proper shape file object
VI_base_layer_hi_res <- read_sf(dsn = here("./01_raw_data/hakai_zip"),
                                stringsAsFactors = F)
read_s
VI_base_layer_hi_res_sf <- st_as_sf(VI_base_layer_hi_res)

#specifying the projection to use for the sf
st_crs(VI_base_layer_hi_res_sf) <- "+proj=longlat +datum=WGS84 +no_defs"
```

### Calculating the fetch for the sites

```{r Calculating the fetch for the sites}
#Makes an object of all the bearings to calculate fetch along
bearings_openness <- seq(7.5, 360, by = 7.5)

#Specifies to use the GEOS engine
sf::sf_use_s2(FALSE)

#This calculates the fetch for each site along each of the bearings calculated
#above. Max possible fetch is 650,000 meters. It is using the shape file from 
#above to know where land is. This will likely take a very long time to run
#so do not worry.
fetch_df_2024_openness <- fetch_len_multi(site_sp_2024, bearings = bearings_openness, dmax = 650000, shoreline = VI_base_layer_hi_res_sf, projected = FALSE)
```

### Turning the raw fetch data into a dataframe with column labels

```{r Turning the raw openness fetch data into a dataframe with column labels}
#Making an object with all of the site code names listed to be able to make a
#column of site codes in the fetch df below
site_code <- c("BMSC1", "BMSC10", "BMSC11", "BMSC12", "BMSC13", "BMSC14", "BMSC15", "BMSC16",
               "BMSC17", "BMSC18", "BMSC19", "BMSC2", "BMSC20", "BMSC21", "BMSC22", "BMSC23", "BMSC24",
               "BMSC25", "BMSC26", "BMSC27", "BMSC3", "BMSC4", "BMSC5", "BMSC6", "BMSC8", "BMSC9")

#Converting the raw fetch data into a df
fetch_df_2024_openness <- as.data.frame(fetch_df_2024_openness)
#Adding a column with the site codes
fetch_df_2024_openness$site_code <- site_code
#Adding a column with the site lat
fetch_df_2024_openness$latitude <- site_coords_2024$latitude
#Adding a column with the site long
fetch_df_2024_openness$longitude <- site_coords_2024$longitude

#Reorganizing the order of the columns so that site code, lat, and long are 
#first
fetch_df_2024_openness_final <- fetch_df_2024_openness[, c(49, 50, 51, seq(1, 48, by = 1))]

#Converting into long format for further analysis and adding to the RLS data
fetch_df_2024_openness_long <- fetch_df_2024_openness_final%>%
  #Makes column with bearing degree and column with corresponding fetch
  pivot_longer(cols = `7.5`: `360`, 
               names_to = "degree", 
               values_to = "fetch", 
               #Removes the prefix "degree" from the values going into the 
               #degree column
               names_prefix = "degree")
```

