---
title: "Extra chunks for Honours"
author: "Alec"
date: "2024-11-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Plotting the stability of each site across years

```{r Plotting the urch sn ratio across years for each site}
(stability_plot <- ggplot(all_data_test, aes(x = year(survey_date), y = urch_sn_ratio,
                                                 colour = depth, group = depth))+
  geom_point(size = 2)+
  theme.figure()+
  scale_color_viridis_c() +
   facet_wrap(~site_code, ncol = 10)+
    scale_y_continuous(breaks = c(-3, -2, -1, 0, 1, 2, 3)))
```


```{r Saving the plot so it is more visible, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/stability_plot.png"), width = 20, height = 10, 
    units = "in", pointsize = 15, res = 600)

stability_plot

dev.off()
```


##Checking data for normality etc.

###Urchin data
```{r Testing urchin counts for normality}
hist(urch_sn_ratio_RLS$Urchins)

shapiro.test(urch_sn_ratio_RLS$Urchins)
```
The urchin counts are not normal based on visual test and on shapiro-wilks test.

```{r Chi-square testing to see if urchin data follows Poisson}
urch_temp_poisson_test <- urch_sn_ratio_RLS%>%
  select(-c(urch_sn_ratio, Turban_Snails))%>%
  na.omit

hist(urch_temp_poisson_test$Urchins)

lambda <- mean(urch_temp_poisson_test$Urchins)

observed <- table(factor(urch_temp_poisson_test$Urchins,
                         levels = min(urch_temp_poisson_test$Urchins):max(urch_temp_poisson_test$Urchins)))

expected <- dpois(min(urch_temp_poisson_test$Urchins):max(urch_temp_poisson_test$Urchins),
                  lambda) * length(urch_temp_poisson_test$Urchins)

chisq.test(observed, p = expected, rescale.p = TRUE)
```

```{r Using vcd package to visually assess whether the data is poisson}
(urch_poissonness_plot <- distplot(urch_temp_poisson_test$Urchins, type = "poisson"))
```

The poissonness plot shows that the counts approximately follow the poisson
distribution.

###Snail data
```{r Testing snail counts for normality}
hist(urch_sn_ratio_RLS$Turban_Snails, n = 20)

shapiro.test(urch_sn_ratio_RLS$Turban_Snails)
```
The snail counts are not normal based on visual test and on shapiro-wilks test.

```{r Using vcd package to visually assess if the data is poisson}
snail_temp_poisson_test <- urch_sn_ratio_RLS%>%
  select(-c(urch_sn_ratio, Urchins))%>%
  na.omit

distplot(snail_temp_poisson_test$Turban_Snails)
```

The snail data appears to somewhat follow a poisson distribution.

###Urch sn ratio
```{r Making a dataset with no na for now}
urch_sn_ratio_temp_poisson <- urch_sn_ratio_RLS%>%
  na.omit()
```


```{r Histogram of urch sn ratio}
hist(urch_sn_ratio_temp_poisson$urch_sn_ratio, n = 20)
```

```{r Poissonness plot of urch sn ratio}
distplot(urch_sn_ratio_temp_poisson$urch_sn_ratio)
```




##Totalling up the fish counts

```{r Loading in the fish data}
fish_data_raw <- read_csv(here("./01_raw_data/IMOS_-_National_Reef_Monitoring_Network_Sub-Facility_-_Global_reef_fish_abundance_and_biomass_no_top.csv"))
```



```{r Cleaning up the fish data and totalling by species and transect}
#fish_factor_cols <- c("site_code", "site_name", "class", "order", "family", "species_name", "reporting_name")

fish_data_clean <- fish_data_raw%>%
     mutate_at(fish_factor_cols, factor)%>%
    group_by(site_code, site_name, latitude, longitude, survey_date, depth, 
        visibility, hour, reporting_name)%>%
 summarise(species_total = sum(total))
```


```{r Making a rockfish total for each site}
rockfish_data <- fish_data_raw%>%
   mutate_at(fish_factor_cols, factor)%>%
filter(order == "Scorpaeniformes")%>%
  group_by(site_code, site_name, latitude, longitude, survey_date, depth, 
           visibility, hour, reporting_name)%>%
  summarise(species_total = sum(total))

```

```{r}
rockfish_total_data <- rockfish_data%>%
    group_by(site_code, site_name, latitude, longitude, survey_date, depth, 
           visibility, hour)%>%
  summarise(total = sum(species_total))
```

##Plotting the fish counts

###Plotting the scorpaeniformes fish counts on the map

```{r Plotting the total scorpaeniformes for each transect in 2021}
(scorp_fish_count_2021 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = rockfish_total_data%>%
               filter(year(survey_date) == "2021"), mapping = aes(x = longitude, y = latitude, color = total),
             size = 3,) + 
  geom_point(data = rockfish_total_data%>%
               filter(year(survey_date) == "2021"), mapping = aes(x = longitude, y = latitude, color = total),
             shape = 1, size = 3,colour = "black")+
  scale_color_viridis_c() +
  labs(title = "2021", x = " \nLongitude", y = "Latitude\n ", color = "Scorpaeniformes counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
```

```{r Plotting the total scorpaeniformes for each transect in 2022}
(scorp_fish_count_2022 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = rockfish_total_data%>%
               filter(year(survey_date) == "2022"), mapping = aes(x = longitude, y = latitude, color = total),
             size = 3,) + 
  geom_point(data = rockfish_total_data%>%
               filter(year(survey_date) == "2022"), mapping = aes(x = longitude, y = latitude, color = total),
             shape = 1, size = 3,colour = "black")+
  scale_color_viridis_c() +
  labs(title = "2022", x = " \nLongitude", y = "Latitude\n ", color = "Scorpaeniformes counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
```

```{r Plotting the total scorpaeniformes for each transect in 2023}
(scorp_fish_count_2023 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = rockfish_total_data%>%
               filter(year(survey_date) == "2023"), mapping = aes(x = longitude, y = latitude, color = total),
             size = 3,) + 
  geom_point(data = rockfish_total_data%>%
               filter(year(survey_date) == "2023"), mapping = aes(x = longitude, y = latitude, color = total),
             shape = 1, size = 3,colour = "black")+
  scale_color_viridis_c() +
  labs(title = "2023", x = " \nLongitude", y = "Latitude\n ", color = "Scorpaeniformes counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
```


##Making plots for the presentation
```{r}
fake_data <- read_csv(here("./01_raw_data/fake_data.csv"))
```

```{r Making a plot with the fake data}
(fake_data_plot <- ggplot(fake_data, aes(x = Explanatory, y = Response))+
  geom_point()+
  geom_smooth(method = "lm")+
  theme.figure())
```


```{r Saving the plot}
png(filename = here("./04_plots/fake_data_plot.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

fake_data_plot

dev.off()
```


```{r Making fake data for predicted results of fetch}
x1 <- seq(100, 900, by = 100)
y1 <- seq(-1, 3 , by = 0.5)
predicted_data <- cbind(x1, y1)
```

```{r Plotting fake data for fetch}
(predicted_plot <- ggplot(predicted_data, aes(x = x1, y = y1))+
  geom_point()+
           geom_smooth(method = "lm")+
  theme.figure()+
   labs(x = "Kelp abundance", y = "Log urchin to snail ratio")+
   theme(axis.title = element_text(size = 18, face = "bold"),
    axis.text = element_text(size = 18, face = "plain")))
```



```{r Making fake data for substrate type, include=FALSE}
fake_substrate_type_data <- read_csv(here("./01_raw_data/example_year_data.csv"))
fake_substrate_type_data$substrate <- as.factor(fake_substrate_type_data$substrate)
levels(fake_substrate_type_data$substrate) <- list(Sand = "Sand", 
                                                   `Loose rocks` = "Loose rocks",
                                                   `Solid rock` = "Solid rock")
```

```{r Plotting the fake substrate data}
(fake_substrate_plot <- ggplot(fake_substrate_type_data, aes(x = substrate, y = urchin))+
   geom_boxplot()+
   theme.figure()+
   labs(x = "\nSubstrate type", y = "Log urchin to snail ratio\n")+
      theme(axis.title = element_text(size = 18, face = "bold"),
    axis.text = element_text(size = 18, face = "plain")))
```



```{r Making fake kelp data, include=FALSE}
fake_kelp_data <- read_csv(here("./01_raw_data/example_year_data.csv"))
```

```{r Plotting the fake kelp data}
(fake_kelp_plot <- ggplot(fake_kelp_data, aes(x = kelp, y = urchin))+
  geom_point()+
           geom_smooth(method = "lm")+
   labs(x = "\nKelp abundance (n/m2)", y = "Log urchin to snail ratio\n")+
   theme.figure()+
   theme(axis.title = element_text(size = 18, face = "bold"),
    axis.text = element_text(size = 18, face = "plain")))
```


###Making an example of yearly fluctuations
```{r Saving in the fake year data}
year_flux_data <- read_csv(here("./01_raw_data/example_year_data.csv"))
```

```{r}
(year_flux_snails <- ggplot()+
  geom_point(year_flux_data, mapping = aes(x = year, y = Urchins),
             size = 4)+
labs(x = " \nYear", y = "Counts\n ") +
  theme.figure()+
   theme(axis.title = element_text(size = 18, face = "bold"),
    axis.text = element_text(size = 18, face = "plain")))
```



#bearings <- seq(200, 230, by = 1)
#bearings_calculated <- c(179, 180, 181, 182, 183, 184, 185)

###Calculating fetch for all of the sites in 2024
```{r Calculating the fetch for the sites 2024}
#fetch_df_2024 <- fetch_len_multi(site_sp_2024, bearings, dmax = 40000, shoreline = VI_base_layer_hi_res_sf, projected = TRUE)

#fetch_df_2024_calc <- fetch_len_multi(site_sp_2024, bearings = bearings_calculated, dmax = 40000, shoreline = VI_base_layer_hi_res_sf, projected = TRUE)
```


```{r Turning the raw fetch data into a dataframe with column labels}
site_code <- c("BMSC1", "BMSC10", "BMSC11", "BMSC12", "BMSC13", "BMSC14", "BMSC15", "BMSC16",
               "BMSC17", "BMSC18", "BMSC19", "BMSC2", "BMSC20", "BMSC21", "BMSC22", "BMSC23", "BMSC24",
               "BMSC25", "BMSC26", "BMSC27", "BMSC3", "BMSC4", "BMSC5", "BMSC6", "BMSC8", "BMSC9")

fetch_df_2024 <- as.data.frame(fetch_df_2024)

fetch_df_2024$site_code <- site_code

fetch_df_2024$latitude <- site_coords_2024$latitude
fetch_df_2024$longitude <- site_coords_2024$longitude

fetch_df_2024_final <- fetch_df_2024[, c(32, 33, 34, seq(1, 31, by = 1))]
```

```{r Turning the raw fetch data using the calculated wind dir into a dataframe}
fetch_df_2024_calc <- as.data.frame(fetch_df_2024_calc)

fetch_df_2024_calc$site_code <- site_code

fetch_df_2024_calc$latitude <- site_coords_2024$latitude
fetch_df_2024_calc$longitude <- site_coords_2024$longitude

fetch_df_2024_calc_final <- fetch_df_2024_calc[, c(8, 9, 10, seq(1, 7, by = 1))]
```

```{r Converting the data into long format}
fetch_2024_long <- fetch_df_2024_final %>%
  pivot_longer(cols = `200`: `230`,  # Columns for compass degrees (assuming "degree1", "degree2", ...)
               names_to = "degree",           # Name of the new column for degrees
               values_to = "fetch",           # Name of the new column for fetch values
               names_prefix = "degree")       # Remove "degree" prefix from column names if needed

```

```{r Converting the data with the calculated wind dir into long format}
fetch_2024_calc_long <- fetch_df_2024_calc_final %>%
  pivot_longer(cols = `179`: `185`,  # Columns for compass degrees (assuming "degree1", "degree2", ...)
               names_to = "degree",           # Name of the new column for degrees
               values_to = "fetch",           # Name of the new column for fetch values
               names_prefix = "degree")       # Remove "degree" prefix from column names if needed

```

###Adding in the fetch data

```{r Calculating the average fetch from the range of values calculated previously for each site}
avg_fetch <- fetch_2024_long%>%
  group_by(site_code, latitude, longitude)%>%
  summarise(avg_fetch = mean(fetch))
```

```{r Adding the avg fetch values into the urch sn ratio df}
urch_sn_ratio_mod_2024 <- merge(urch_sn_ratio_mod_2024, avg_fetch, by = c("site_code", "latitude", "longitude"))
```


```{r Calculating the avg fetch values for the fetch data using the calc wind dir}
avg_fetch_calc <- fetch_2024_calc_long%>%
  group_by(site_code, latitude, longitude)%>%
  summarise(avg_fetch = mean(fetch))
```

```{r Adding the avg fetch values into a new urch sn ratio df}
urch_sn_ratio_mod_calc_2024 <- merge(urch_sn_ratio_mod_calc_2024, avg_fetch_calc, by = c("site_code", "latitude", "longitude"))
```

#Modelling with new fetch data

##Loading in the data
```{r Loading in the data}
env_mod_calc_data <- read_csv(here("./03_clean_data/urch_sn_ratio_mod_calc_data.csv"))
```

```{r Scaling the explanatory variables and converting substrate type to factor}
env_mod_calc_data$sst_scale <- scale(env_mod_calc_data$sst)
env_mod_calc_data$avg_fetch_scale <- scale(env_mod_calc_data$avg_fetch)
env_mod_calc_data$dominant_score <- as.factor(env_mod_calc_data$dominant_score)
levels(env_mod_calc_data$dominant_score) <- list(loose = "3", solid = "4")
```

##Plotting EVs against urch sn ratio

```{r ratio against fetch}
ggplot(env_mod_calc_data, aes(x = avg_fetch_scale, y = urch_sn_ratio))+
  geom_point()+
  theme.figure()
```


##Building the GAM

###No spatial autocorrelation term
```{r Making a gam with no spatial autocorrelation term to start with calc data}
env_mod_calc <- gam(urch_sn_ratio ~ s(sst_scale) + s(avg_fetch_scale) + dominant_score,
               data = env_mod_calc_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(env_mod, rep = 500)

plot(env_mod)
```


```{r Checking the model fit with DHARMa}
env_mod_calc_resids <- simulateResiduals(env_mod_calc)
plot(env_mod_calc_resids)
```
All good




```{r Building the model, this time with fetch fully removed}
env_mod_no_fetch_calc <- gam(urch_sn_ratio ~ s(sst_scale) + dominant_score,
                        data = env_mod_calc_data,
                        method = 'REML',
                        family = gaussian(link = "identity"),
                        select = FALSE)

gam.check(env_mod_no_fetch_calc, rep = 500)
```

```{r}
env_mod_no_fetch_calc_resid <- simulateResiduals(env_mod_no_fetch_calc)
plot(env_mod_no_fetch_calc_resid)
```
Possible error based on DHARMa output

###With autocorrelation effect

```{r Building the gam with the spatial autocorrelation term}
env_mod_sp_gp_calc <- gam(urch_sn_ratio ~ s(sst_scale) + s(avg_fetch_scale) + dominant_score + s(latitude, longitude, bs = "gp", k = 10),
                  data = env_mod_calc_data, 
                  family = gaussian(link = "identity"),
                  method = 'REML',
                  select = TRUE)   #In this case I have added the lat long as an interaction term directly in the model

set.seed(10)
gam.check(env_mod_sp_gp_calc, rep = 500)
plot(env_mod_sp_gp_calc)
```

```{r DHARMa for model fit}
env_mod_sp_gp_calc_resid <- simulateResiduals(env_mod_sp_gp_calc)
plot(env_mod_sp_gp_calc_resid)
```
All good


```{r Making a dataframe with sites at the same location shifted slightly}
env_mod_data_shifted_calc <- env_mod_calc_data%>%
    mutate(latitude = if_else(row_number() == 2, 48.82896, latitude))%>%
  mutate(latitude = if_else(row_number() == 5, 48.85746, latitude))%>%
  mutate(latitude = if_else(row_number() == 7, 48.85827, latitude))%>%
  mutate(latitude = if_else(row_number() == 21, 48.91608, latitude))%>%
   mutate(latitude = if_else(row_number() == 23, 48.83858, latitude))%>%
   mutate(latitude = if_else(row_number() == 25, 48.90711, latitude))%>%
  mutate(latitude = if_else(row_number() == 30, 48.82734, latitude))%>%
  mutate(latitude = if_else(row_number() == 32, 48.95024, latitude))%>%
   mutate(latitude = if_else(row_number() == 34, 48.95508, latitude))
```


```{r Building the gam with the spatial autocorrelation term with calc and shifted}
env_mod_sp_gp_calc_shifted <- gam(urch_sn_ratio ~ s(sst_scale) + s(avg_fetch_scale) + dominant_score + s(latitude, longitude, bs = "gp", k = 10),
                  data = env_mod_data_shifted_calc, 
                  family = gaussian(link = "identity"),
                  method = 'REML',
                  select = TRUE)   #In this case I have added the lat long as an interaction term directly in the model

set.seed(10)
gam.check(env_mod_sp_gp_calc_shifted, rep = 500)
plot(env_mod_sp_gp_calc_shifted)
```
```{r DHARMa for model fit}
env_mod_sp_gp_calc_shifted_resid <- simulateResiduals(env_mod_sp_gp_calc_shifted)
plot(env_mod_sp_gp_calc_shifted_resid)
```
All good


```{r Building gamm with spatial autocorrelation term corSpher with calc}
env_mod_sp_cor_Spher_calc <- gamm(urch_sn_ratio ~ s(sst_scale) + s(avg_fetch_scale) + dominant_score, 
                        data = env_mod_data_shifted_calc,
                       family = gaussian(link = "identity"),
                       correlation = corSpher(form = ~ longitude + latitude), 
                       method = 'REML',
                       select = TRUE)

gam.check(env_mod_sp_cor_Spher_calc$gam, rep = 500)
plot(env_mod_sp_cor_Spher_calc$gam)
```



```{r Building gamm with spatial autocorrelation term corExp with calc}
env_mod_sp_cor_exp_calc <- gamm(urch_sn_ratio ~ s(sst_scale) + s(avg_fetch_scale) + dominant_score, 
                        data = env_mod_data_shifted_calc,
                       family = gaussian(link = "identity"),
                       correlation = corExp(form = ~ longitude + latitude), 
                       method = 'REML',
                       select = TRUE)


gam.check(env_mod_sp_cor_exp_calc$gam, rep = 500)
plot(env_mod_sp_cor_exp_calc$gam)
```

###AIC with the different models for calc

```{r Comparing the calc models with an AIC}
AIC(env_mod_calc, env_mod_no_fetch_calc, env_mod_sp_gp_calc, env_mod_sp_gp_calc_shifted,
    env_mod_sp_cor_Spher_calc$lme, 
    env_mod_sp_cor_exp_calc$lme
    )%>%
arrange(AIC)%>%
  mutate(delta_AIC = AIC - AIC[1])
```


```{r Comparing the models with unshifted points in an AIC}
AIC(env_mod_calc, env_mod_no_fetch_calc, env_mod_sp_gp_calc)%>%
arrange(AIC)%>%
  mutate(delta_AIC = AIC - AIC[1])
```


```{r Comparing the calc models with shifted points in an AIC}
AIC(env_mod_calc, env_mod_no_fetch_calc, env_mod_sp_gp_calc_shifted,
    env_mod_sp_cor_Spher_calc$lme, 
    env_mod_sp_cor_exp_calc$lme
    )%>%
arrange(AIC)%>%
  mutate(delta_AIC = AIC - AIC[1])
```

