---
title: "Openness"
author: "Alec Jones"
date: "2026-08-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(waver)
library(sf)
```

## Predictor 1: Wave exposure (via metric called Openness)

To get a measure of wave exposure for each of the RLS sites, I used a metric 
called openness from Hill et al. (2010)(10.3354/meps08815).

### Set up prior to calculating fetch for openness



****THIS CHUNK YOU NEED TO USE THE COORDINATES FROM YOUR OWN SITES****

```{r Making dfs of the unique sites and then just the lat and long, echo=FALSE, include=FALSE}
#Making a df with just the unique site info
site_data_2024 <- RLS_data_2024_clean%>%
  distinct(site_code, site_name, latitude, longitude)

#Making a df with just the unique lat and longs
site_coords_2024 <- site_data_2024%>%
  ungroup()%>%
  select(latitude, longitude)%>%
  distinct()

#This converts the site coordinates into a sf object and specifies the 
#coordinate reference system to be the same as the shape file
site_sp_2024 <- st_as_sf(site_coords_2024, coords = c("longitude", "latitude"), crs = 4326)
```

```{r Loading in the shapefiles to do the fetch calculations, include=FALSE}
#This is a high resolution shape file from Hakai. This code loads in the shape
#file and saves it as the proper shape file object
VI_base_layer_hi_res <- read_sf(dsn = here("./01_raw_data/hakai_zip"),
                                stringsAsFactors = F)

VI_base_layer_hi_res_sf <- st_as_sf(VI_base_layer_hi_res)

#specifying the projection to use for the sf
st_crs(VI_base_layer_hi_res_sf) <- "+proj=longlat +datum=WGS84 +no_defs"


#This subsets the shape file to only be of Barkley Sound to make plotting more
#efficient

#This makes coordinate limits to subset the shapefile
shp_xmin <- -125.4
shp_xmax <- -124.95
shp_ymin <- 48.7
shp_ymax <- 49.0

#This makes a bounding of the coordinates that I just made
bbox <- st_bbox(c(xmin = shp_xmin, ymin = shp_ymin, xmax = shp_xmax, ymax = shp_ymax))

#This converts the boundary box to an SFC object
bbox_sf <- st_as_sfc(bbox)

#This crops the shape file to the area of the bounding box
VI_base_layer_hi_res_sf_subset <- st_crop(VI_base_layer_hi_res_sf, bbox_sf)

#This plots as a sanity check
plot(VI_base_layer_hi_res_sf_subset)
```

### Calculating the fetch for the sites

```{r Calculating the fetch for the sites, eval=FALSE}
#Makes an object of all the bearings to calculate fetch along
bearings_openness <- seq(7.5, 360, by = 7.5)

#Specifies to use the GEOS engine
sf::sf_use_s2(FALSE)

#This calculates the fetch for each site along each of the bearings calculated
#above. Max possible fetch is 650,000 meters. It is using the shape file from 
#above to know where land is. This will likely take a very long time to run
#so do not worry.
#I have ## this code out because it takes way too long to run. There is code in
#the chunk below that will load in the data from a .csv that has already been
#made by running the fetch_len_multi function

fetch_df_2024_openness <- fetch_len_multi(site_sp_2024,
                                          #the bearings you specified above
                                          bearings = bearings_openness,
                                          #the max distance you want to calculate fetch for
                                          dmax = 650000,
                                          #the shape file used for the shore
                                          shoreline = VI_base_layer_hi_res_sf, 
                                          #whether the shapefile is on a projected coordinate system
                                          projected = FALSE)

#Converting the raw fetch data into a df
fetch_df_2024_openness <- as.data.frame(fetch_df_2024_openness)

#Saving the fetch data to a .csv so that the fetch function does not need to be
#run
write_csv(fetch_df_2024_openness, here("./01_raw_data/fetch_df_2024_openness.csv"))
```


### Downloading the fetch data that is already saved

```{r Loading in the pre calculated fetch data, echo=FALSE}
fetch_df_2024_openness <- read_csv(here("./01_raw_data/fetch_df_2024_openness.csv"))
```


### Turning the raw fetch data into a dataframe with column labels

```{r Turning the raw openness fetch data into a dataframe with column labels}
#Making an object with all of the site code names listed to be able to make a
#column of site codes in the fetch df below
site_code <- c("BMSC1", "BMSC10", "BMSC11", "BMSC12", "BMSC13", "BMSC14", "BMSC15", "BMSC16",
               "BMSC17", "BMSC18", "BMSC19", "BMSC2", "BMSC20", "BMSC21", "BMSC22", "BMSC23", "BMSC24",
               "BMSC25", "BMSC26", "BMSC27", "BMSC3", "BMSC4", "BMSC5", "BMSC6", "BMSC8", "BMSC9")

#Adding a column with the site codes
fetch_df_2024_openness$site_code <- site_code
#Adding a column with the site lat
fetch_df_2024_openness$latitude <- site_coords_2024$latitude
#Adding a column with the site long
fetch_df_2024_openness$longitude <- site_coords_2024$longitude

#Reorganizing the order of the columns so that site code, lat, and long are 
#first
fetch_df_2024_openness_final <- fetch_df_2024_openness[, c(49, 50, 51, seq(1, 48, by = 1))]

#Converting into long format for further analysis and adding to the RLS data
fetch_df_2024_openness_long <- fetch_df_2024_openness_final%>%
  #Makes column with bearing degree and column with corresponding fetch
  pivot_longer(cols = `7.5`: `360`, 
               names_to = "degree", 
               values_to = "fetch", 
               #Removes the prefix "degree" from the values going into the 
               #degree column
               names_prefix = "degree")
```


```{r Summing the fetch vectors for each site then dividing by the largest openness value}
#Calculates the openness of each site
openness <- fetch_df_2024_openness_long%>%
  #grouping by each site
  group_by(site_code, latitude, longitude)%>%
  #adding up all of the fetch vectors for each site
  summarise(openness = sum(fetch))
  
#identifying the largest openness score within the dataset
max_openness <- max(openness$openness)

#Making a column of normalized openness by dividing by the max observed openness
openness <- openness%>%
mutate(norm_openness = openness / max_openness)

#Merging the modelling df with the openness df by each site
urch_sn_ratio_mod_2024 <- merge(urch_sn_ratio_mod_2024, openness, by = c("site_code", "latitude", "longitude"))
```