---
title: "Preliminary Analysis with RLS data"
author: "Alec"
date: "2024-09-13"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(lubridate)
library(sf) #for spatial data
library(cowplot)
library(rnaturalearth)
library(ggspatial)
library(ggrepel)
library(waver)
library(vcd)
library(sp)
library(viridisLite)
```

#Preliminary analyses of the RLS data from Barkley Sound

##Loading in the data and organizing it
```{r Setting up the here package}
here::here()
```

```{r Loading in the invertebrate file, include=FALSE}
invert_data_raw <- read_csv(here("./01_raw_data/IMOS_-_National_Reef_Monitoring_Network_Sub-Facility_-_Global_mobile_macroinvertebrate_abundance_no_top.csv"))
```

```{r Cleaning up columns and converting certain columns to factors}
factor_cols <- c("site_code", "site_name", "reporting_name")

invert_data_clean <- invert_data_raw%>%
  select(-c(FID, survey_id, country, area, ecoregion, realm, location,
            survey_latitude, survey_longitude, biomass, phylum, class, order,
            family, program))%>%
  mutate_at(factor_cols, factor)
```


```{r Summing the total numbers of animals per transect, include=FALSE}
invert_data_final_clean <- invert_data_clean%>%
  group_by(site_code, site_name, latitude, longitude, survey_date, depth, 
           visibility, hour, reporting_name)%>%
  summarise(species_total = sum(total))
```

```{r Correcting the lat and long to be correct}
#invert_data_final_clean <- invert_data_final_clean%>%
  
```


```{r Making a separate df for each year}
invert_data_clean_2021 <- invert_data_final_clean%>%
  filter(year(survey_date) == "2021")

invert_data_clean_2022 <- invert_data_final_clean%>%
  filter(year(survey_date) == "2022")

invert_data_clean_2023 <- invert_data_final_clean%>%
  filter(year(survey_date) == "2023")
```

```{r Writing a .csv for the clean data with the counts of each species/transect}
write_csv(invert_data_final_clean, here("./03_clean_data/invert_data_final_clean.csv"))
```




##Calculating urch to sn ratio and getting rid of other species
```{r Loading in the new clean df and making sure the right columns are factors, include=FALSE}
invert_data_final_clean <- read_csv(here("./03_clean_data/invert_data_final_clean.csv"))

factor_cols <- c("site_code", "site_name", "reporting_name")

invert_data_final_clean <- invert_data_final_clean%>%
  mutate_at(factor_cols, factor)
```



```{r removing other species and calculating urchin ratio}
 urch_sn_ratio_RLS <- invert_data_final_clean%>%
  pivot_wider(names_from = reporting_name, values_from = species_total)%>%
  mutate(Urchins = `Mesocentrotus franciscanus`,
         Turban_Snails = `Pomaulax gibberosus`)%>%
  select(site_code, site_name, latitude, longitude, survey_date, depth, 
           visibility, hour, Urchins, Turban_Snails)%>%
  replace(is.na(.), 0)%>%
  mutate(urch_sn_ratio = log10((Urchins +1) / (Turban_Snails + 1)))
```

##Totalling up the fish counts

```{r Loading in the fish data}
fish_data_raw <- read_csv(here("./01_raw_data/IMOS_-_National_Reef_Monitoring_Network_Sub-Facility_-_Global_reef_fish_abundance_and_biomass_no_top.csv"))
```



```{r Cleaning up the fish data and totalling by species and transect}
#fish_factor_cols <- c("site_code", "site_name", "class", "order", "family", "species_name", "reporting_name")

fish_data_clean <- fish_data_raw%>%
     mutate_at(fish_factor_cols, factor)%>%
    group_by(site_code, site_name, latitude, longitude, survey_date, depth, 
        visibility, hour, reporting_name)%>%
 summarise(species_total = sum(total))
```

```{r Making a rockfish total for each site}
rockfish_data <- fish_data_raw%>%
   mutate_at(fish_factor_cols, factor)%>%
filter(order == "Scorpaeniformes")%>%
  group_by(site_code, site_name, latitude, longitude, survey_date, depth, 
           visibility, hour, reporting_name)%>%
  summarise(species_total = sum(total))

```

```{r}
rockfish_total_data <- rockfish_data%>%
    group_by(site_code, site_name, latitude, longitude, survey_date, depth, 
           visibility, hour)%>%
  summarise(total = sum(species_total))
```



##Making some visualizations of the urch_sn ratio and correlation between urch and sn

```{r Making theme for all subsequent figures}
theme.figure <- function (){
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(colour = "black"),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12, face = "plain"),
    legend.text = element_text(size = 10, face = "plain"),
    legend.title = element_text(size = 14, face = "bold")) 
}

```

###Overall across years
```{r Plotting urch against sn overall and colouring by year}
(urch_sn_plot <- ggplot(urch_sn_ratio_RLS, aes(x = Turban_Snails, y = Urchins,
                                               colour = as.factor(year(survey_date))))+
  geom_point()+
  theme.figure()+
  labs(colour = "Year"))
 #geom_smooth(method = "lm")
```

###In 2021
```{r Plotting number of urchins vs snails in 2021}
(urch_sn_plot_2021 <-ggplot(urch_sn_ratio_RLS%>%
                              filter(year(survey_date) == "2021"),
  aes(x = Turban_Snails, y = Urchins))+
  geom_point()+
  theme.figure())
 #geom_smooth(method = "lm")
```

###In 2022
```{r Plotting number of urchins vs snails in 2022}
(urch_sn_plot_2022 <-ggplot(urch_sn_ratio_RLS%>%
                              filter(year(survey_date) == "2022"),
  aes(x = Turban_Snails, y = Urchins))+
  geom_point()+
  theme.figure())
 #geom_smooth(method = "lm")
```

###In 2023
```{r Plotting number of urchins vs snails in 2023}
(urch_sn_plot_2023 <-ggplot(urch_sn_ratio_RLS%>%
                              filter(year(survey_date) == "2023"),
  aes(x = Turban_Snails, y = Urchins))+
  geom_point()+
  theme.figure())
 #geom_smooth(method = "lm")
```




##Making maps of the data in Barkley Sound

###Doing the set up for the base layer and map characteristics
```{r Loading in the base layer}
VI_base_layer <- read_sf(dsn = here("./01_raw_data/canada_sf_data", ""), 
                         stringsAsFactors = F)
VI_base_layer_sf <- st_as_sf(VI_base_layer)

#ggplot(VI_base_layer_sf)+
  #geom_sf()

```

```{r Specifying the projection we want to use for our data}
st_crs(VI_base_layer_sf) <- "+proj=longlat +datum=WGS84 +no_defs"
```

```{r Making the range for the map}
#first need to specify the range of the map
lat_min = min(urch_sn_ratio_RLS$latitude) - 0.01
lat_max = max(urch_sn_ratio_RLS$latitude) + 0.01

long_min = min(urch_sn_ratio_RLS$longitude) - 0.01
long_max = max(urch_sn_ratio_RLS$longitude) + 0.01

lat_min_launch = min(urch_sn_ratio_RLS$latitude) - 0.01
lat_max_launch = max(urch_sn_ratio_RLS$latitude) + 0.05

long_min_launch = min(urch_sn_ratio_RLS$longitude) - 0.2
long_max_launch = max(urch_sn_ratio_RLS$longitude) + 0.1
```


###Making a site map

```{r Making a df with just one of each site}
site_map_data <- invert_data_final_clean%>%
  distinct(site_code, site_name, latitude, longitude)
```


```{r Making the map with the site names, fig.width=7.5, fig.height=5.5}
(RLS_site_name_map <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  labs(x = " \nLongitude", y = "Latitude\n ") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1)) +
  scale_y_continuous(breaks = c(48.8, 48.85, 48.9))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
geom_label_repel(
    data = site_map_data, 
    mapping = aes(x = longitude, y = latitude, label = site_name), 
    size = 2,
   # box.padding = 0.35,  # Add padding between labels and points
  #  point.padding = 0.5, # Add padding between points and text
    segment.color = 'grey50', # Line from label to point
    max.overlaps = 15      # Adjust maximum number of overlaps
  ))
```

```{r Making the map with the site codes, fig.width=7.5, fig.height=5.5}
(RLS_site_code_map <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  labs(x = " \nLongitude", y = "Latitude\n ") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
geom_label_repel(
    data = site_map_data, 
    mapping = aes(x = longitude, y = latitude, label = site_code), 
    size = 2,
   # box.padding = 0.35,  # Add padding between labels and points
  #  point.padding = 0.5, # Add padding between points and text
    segment.color = 'grey50', # Line from label to point
    max.overlaps = 15      # Adjust maximum number of overlaps
  ))
```


```{r Making a data frame of the boat launch sites}
launch_lat <- c(48.878415, 48.828809, 49.012965, 48.946169)
launch_long <- c(-125.046749, -125.127513, -125.360924, -125.547637)
names <- c("Poat Nook", "Grappler Inlet", "Toquart Bay", "Ucluelet")
launch_df <- tibble(launch_long, launch_lat, names)
```



```{r Making map with just the points and the boat launches}
(RLS_site_point_map <- ggplot()+
       geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min_launch, long_max_launch),
           ylim = c(lat_min_launch, lat_max_launch)) +
  labs(x = " \nLongitude", y = "Latitude\n ") +
  theme.figure() +
  #scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
 # scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering)+ 
    geom_point(data = site_map_data, mapping = aes(x = longitude, y = latitude),
             size = 3, colour = "#29AF7FFF") + 
  geom_point(data = site_map_data, mapping = aes(x = longitude, y = latitude),
             shape = 1, size = 3,colour = "black")+
      geom_point(data = launch_df, mapping = aes(x = launch_long, y = launch_lat),
                 size = 3, colour = "#FDE725FF")+
      geom_point(data = launch_df, mapping = aes(x = launch_long, y = launch_lat),
                shape = 1, size = 3, colour = "black"))
```


```{r Making the inset map}
(van_isl_map <- ggplot(VI_base_layer_sf) +
  geom_sf() +
  coord_sf(ylim = c(48, 52), xlim = c(-129, -122)) +
  theme.figure() +
 # geom_rect(xmin = long_min, xmax = long_max, ymin = lat_min, ymax = lat_max,
            #color = "red", fill = NA, linewidth = 1) +
  labs(x = "Longitude", y = "Latitude") +
    theme(axis.text = element_text(size = 14))+
  scale_x_continuous(breaks = c(-128, -125, -122)) +
  scale_y_continuous(breaks = c(48, 50, 52)))
```



###Making maps of the urch to sn ratio for each year

```{r Making the plot with the urch to sn ratio on the map 2021}
(urch_sn_ratio_map_2021 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2021"), mapping = aes(x = longitude, y = latitude, color = urch_sn_ratio),
             size = 3,) + 
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2021"), mapping = aes(x = longitude, y = latitude, color = urch_sn_ratio),
             shape = 1, size = 3,colour = "black")+
  scale_color_viridis_c() +
  labs(title = "2021", x = " \nLongitude", y = "Latitude\n ", color = "Log Urchin to Snail Ratio") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
```

```{r Making the plot with the urch to sn ratio on the map 2022}
(urch_sn_ratio_map_2022 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2022"), mapping = aes(x = longitude, y = latitude, color = urch_sn_ratio),
             size = 3,) + 
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2022"), mapping = aes(x = longitude, y = latitude, color = urch_sn_ratio),
             shape = 1, size = 3,colour = "black")+
  scale_color_viridis_c() +
  labs(title = "2022", x = " \nLongitude", y = "Latitude\n ", color = " Log Urchin to Snail Ratio") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
 # geom_label(data = urch_sn_ratio_RLS, 
             #mapping = aes(x = -125.134813, y = 48.834945, label = "BMSC")))

```


```{r Making the plot with the urch to sn ratio on the map 2023}
(urch_sn_ratio_map_2023 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2023"), mapping = aes(x = longitude, y = latitude, color = urch_sn_ratio),
             size = 3,) + 
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2023"), mapping = aes(x = longitude, y = latitude, color = urch_sn_ratio),
             shape = 1, size = 3,colour = "black")+
  scale_color_viridis_c() +
  labs(title = "2023", x = " \nLongitude", y = "Latitude\n ", color = " Log Urchin to Snail Ratio") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
 # geom_label(data = urch_sn_ratio_RLS, 
             #mapping = aes(x = -125.134813, y = 48.834945, label = "BMSC")))

```


```{r Making the plot with the urch to sn ratio on the map 2024}
(urch_sn_ratio_map_2024 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = all_data_test%>%
               filter(year(survey_date) == "2024"), mapping = aes(x = longitude, y = latitude, color = urch_sn_ratio),
             size = 3,) + 
  geom_point(data = all_data_test%>%
               filter(year(survey_date) == "2024"), mapping = aes(x = longitude, y = latitude, color = urch_sn_ratio),
             shape = 1, size = 3,colour = "black")+
  scale_color_viridis_c() +
  labs(title = "2024", x = " \nLongitude", y = "Latitude\n ", color = " Log Urchin to Snail Ratio") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
 # geom_label(data = urch_sn_ratio_RLS, 
             #mapping = aes(x = -125.134813, y = 48.834945, label = "BMSC")))

```

###Plotting the urchin counts on the map
```{r Plot with urchin counts on map 2021}
(urch_counts_map_2021 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2021"), mapping = aes(x = longitude, y = latitude, color = Urchins),
             size = 2,) + 
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2021"), mapping = aes(x = longitude, y = latitude, color = Urchins),
             shape = 1, size = 2,colour = "black")+
  scale_color_gradient(low = "#35A7FF", high = "#FF5A5F" ) +
  labs(title = "2021", x = " \nLongitude", y = "Latitude\n ", color = "Urchin Counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1)) +
  scale_y_continuous(breaks = c(48.8, 48.85, 48.9))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
 # geom_label(data = urch_sn_ratio_RLS, 
             #mapping = aes(x = -125.134813, y = 48.834945, label = "BMSC")))
```


```{r Plot with Urchin counts 2022}
(urch_counts_map_2022 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2022"), mapping = aes(x = longitude, y = latitude, color = Urchins),
             size = 2,) + 
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2022"), mapping = aes(x = longitude, y = latitude, color = Urchins),
             shape = 1, size = 2,colour = "black")+
  scale_color_gradient(low = "#35A7FF", high = "#FF5A5F" ) +
  labs(title = "2022", x = " \nLongitude", y = "Latitude\n ", color = "Urchin Counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1)) +
  scale_y_continuous(breaks = c(48.8, 48.85, 48.9))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
 # geom_label(data = urch_sn_ratio_RLS, 
             #mapping = aes(x = -125.134813, y = 48.834945, label = "BMSC")))
```



```{r Plot of Urchin counts 2023}
(urch_counts_map_2023 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2023"), mapping = aes(x = longitude, y = latitude, color = Urchins),
             size = 2,) + 
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2023"), mapping = aes(x = longitude, y = latitude, color = Urchins),
             shape = 1, size = 2,colour = "black")+
  scale_color_gradient(low = "#35A7FF", high = "#FF5A5F" ) +
  labs(title = "2023", x = " \nLongitude", y = "Latitude\n ", color = "Urchin Counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1)) +
  scale_y_continuous(breaks = c(48.8, 48.85, 48.9))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
 # geom_label(data = urch_sn_ratio_RLS, 
             #mapping = aes(x = -125.134813, y = 48.834945, label = "BMSC")))
```



###Plotting the snail counts on the map

```{r Plot of snail counts 2021}
(snail_counts_map_2021 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2021"), mapping = aes(x = longitude, y = latitude, color = Turban_Snails),
             size = 2,) + 
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2021"), mapping = aes(x = longitude, y = latitude, color = Turban_Snails),
             shape = 1, size = 2,colour = "black")+
  scale_color_gradient(low = "#35A7FF", high = "#FF5A5F" ) +
  labs(title = "2021", x = " \nLongitude", y = "Latitude\n ", color = "Red Turban Snail Counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1)) +
  scale_y_continuous(breaks = c(48.8, 48.85, 48.9))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))

```


```{r Plot of Snail counts 2022}
(snail_counts_map_2022 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2022"), mapping = aes(x = longitude, y = latitude, color = Turban_Snails),
             size = 2,) + 
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2022"), mapping = aes(x = longitude, y = latitude, color = Turban_Snails),
             shape = 1, size = 2,colour = "black")+
  scale_color_gradient(low = "#35A7FF", high = "#FF5A5F" ) +
  labs(title = "2022", x = " \nLongitude", y = "Latitude\n ", color = "Red Turban Snail Counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1)) +
  scale_y_continuous(breaks = c(48.8, 48.85, 48.9))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))

```


```{r Plot of Snail counts 2023}
(snail_counts_map_2023 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2023"), mapping = aes(x = longitude, y = latitude, color = Turban_Snails),
             size = 2,) + 
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2023"), mapping = aes(x = longitude, y = latitude, color = Turban_Snails),
             shape = 1, size = 2,colour = "black")+
  scale_color_gradient(low = "#35A7FF", high = "#FF5A5F" ) +
  labs(title = "2023", x = " \nLongitude", y = "Latitude\n ", color = "Red Turban Snail Counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1)) +
  scale_y_continuous(breaks = c(48.8, 48.85, 48.9))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
```



##Plotting the fish counts

###Plotting the scorpaeniformes fish counts on the map

```{r Plotting the total scorpaeniformes for each transect in 2021}
(scorp_fish_count_2021 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = rockfish_total_data%>%
               filter(year(survey_date) == "2021"), mapping = aes(x = longitude, y = latitude, color = total),
             size = 3,) + 
  geom_point(data = rockfish_total_data%>%
               filter(year(survey_date) == "2021"), mapping = aes(x = longitude, y = latitude, color = total),
             shape = 1, size = 3,colour = "black")+
  scale_color_viridis_c() +
  labs(title = "2021", x = " \nLongitude", y = "Latitude\n ", color = "Scorpaeniformes counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
```

```{r Plotting the total scorpaeniformes for each transect in 2022}
(scorp_fish_count_2022 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = rockfish_total_data%>%
               filter(year(survey_date) == "2022"), mapping = aes(x = longitude, y = latitude, color = total),
             size = 3,) + 
  geom_point(data = rockfish_total_data%>%
               filter(year(survey_date) == "2022"), mapping = aes(x = longitude, y = latitude, color = total),
             shape = 1, size = 3,colour = "black")+
  scale_color_viridis_c() +
  labs(title = "2022", x = " \nLongitude", y = "Latitude\n ", color = "Scorpaeniformes counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
```

```{r Plotting the total scorpaeniformes for each transect in 2023}
(scorp_fish_count_2023 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = rockfish_total_data%>%
               filter(year(survey_date) == "2023"), mapping = aes(x = longitude, y = latitude, color = total),
             size = 3,) + 
  geom_point(data = rockfish_total_data%>%
               filter(year(survey_date) == "2023"), mapping = aes(x = longitude, y = latitude, color = total),
             shape = 1, size = 3,colour = "black")+
  scale_color_viridis_c() +
  labs(title = "2023", x = " \nLongitude", y = "Latitude\n ", color = "Scorpaeniformes counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
```
###Plotting the goby counts

```{r Plotting the goby counts in 2021}
(goby_fish_count_2021 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = fish_data_clean%>%
               filter(reporting_name == "Rhinogobiops nicholsii",
                      year(survey_date) == "2023"), mapping = aes(x = longitude, y = latitude, color = species_total),
             size = 3,) + 
  geom_point(data = fish_data_clean%>%
               filter(reporting_name == "Rhinogobiops nicholsii",
                      year(survey_date) == "2023"), mapping = aes(x = longitude, y = latitude, color = species_total),
             shape = 1, size = 3,colour = "black")+
  scale_color_viridis_c() +
  labs(title = "2021", x = " \nLongitude", y = "Latitude\n ", color = "Blackeye goby counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
```


###Saving all of the figures so far

####Site maps
```{r Saving the site name map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/site_name_map.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

RLS_site_name_map

dev.off()
```

```{r Saving the site code map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/site_code_map.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

RLS_site_code_map

dev.off()
```


```{r Saving the site point map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/site_point_map.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

RLS_site_point_map

dev.off()
```

```{r Saving the Vancouver Island map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/van_isl_map.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

van_isl_map

dev.off()
```

####Urch Sn ratio maps
```{r Saving the 2021 urch sn ratio map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/urch_sn_ratio_map_2021.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

urch_sn_ratio_map_2021

dev.off()
```

```{r Saving the 2022 urch sn ratio map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/urch_sn_ratio_map_2022.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

urch_sn_ratio_map_2022

dev.off()
```

```{r Saving the 2023 urch sn ratio map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/urch_sn_ratio_map_2023.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

urch_sn_ratio_map_2023

dev.off()
```

```{r Saving the 2024 urch sn ratio map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/urch_sn_ratio_map_2024.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

urch_sn_ratio_map_2024

dev.off()
```

####Urch counts maps
```{r Saving the 2021 urch counts map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/urch_counts_map_2021.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

urch_counts_map_2021

dev.off()
```

```{r Saving the 2022 urch counts map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/urch_counts_map_2022.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

urch_counts_map_2022

dev.off()
```

```{r Saving the 2023 urch counts map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/urch_counts_map_2023.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

urch_counts_map_2023

dev.off()
```


####Snail counts maps
```{r Saving the 2021 snail counts map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/snail_counts_map_2021.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

snail_counts_map_2021

dev.off()
```

```{r Saving the 2022 snail counts map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/snail_counts_map_2022.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

snail_counts_map_2022

dev.off()
```


```{r Saving the 2023 snail counts map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/snail_counts_map_2023.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

snail_counts_map_2023

dev.off()
```



##Checking data for normality etc.

###Urchin data
```{r Testing urchin counts for normality}
hist(urch_sn_ratio_RLS$Urchins)

shapiro.test(urch_sn_ratio_RLS$Urchins)
```
The urchin counts are not normal based on visual test and on shapiro-wilks test.

```{r Chi-square testing to see if urchin data follows Poisson}
urch_temp_poisson_test <- urch_sn_ratio_RLS%>%
  select(-c(urch_sn_ratio, Turban_Snails))%>%
  na.omit

hist(urch_temp_poisson_test$Urchins)

lambda <- mean(urch_temp_poisson_test$Urchins)

observed <- table(factor(urch_temp_poisson_test$Urchins,
                         levels = min(urch_temp_poisson_test$Urchins):max(urch_temp_poisson_test$Urchins)))

expected <- dpois(min(urch_temp_poisson_test$Urchins):max(urch_temp_poisson_test$Urchins),
                  lambda) * length(urch_temp_poisson_test$Urchins)

chisq.test(observed, p = expected, rescale.p = TRUE)
```

```{r Using vcd package to visually assess whether the data is poisson}
(urch_poissonness_plot <- distplot(urch_temp_poisson_test$Urchins, type = "poisson"))
```

The poissonness plot shows that the counts approximately follow the poisson
distribution.


###Snail data
```{r Testing snail counts for normality}
hist(urch_sn_ratio_RLS$Turban_Snails, n = 20)

shapiro.test(urch_sn_ratio_RLS$Turban_Snails)
```
The snail counts are not normal based on visual test and on shapiro-wilks test.

```{r Using vcd package to visually assess if the data is poisson}
snail_temp_poisson_test <- urch_sn_ratio_RLS%>%
  select(-c(urch_sn_ratio, Urchins))%>%
  na.omit

distplot(snail_temp_poisson_test$Turban_Snails)
```

The snail data appears to somewhat follow a poisson distribution.

###Urch sn ratio
```{r Making a dataset with no na for now}
urch_sn_ratio_temp_poisson <- urch_sn_ratio_RLS%>%
  na.omit()
```


```{r Histogram of urch sn ratio}
hist(urch_sn_ratio_temp_poisson$urch_sn_ratio, n = 20)
```

```{r Poissonness plot of urch sn ratio}
distplot(urch_sn_ratio_temp_poisson$urch_sn_ratio)
```




##Plotting the stability of each site across years

```{r Plotting the urch sn ratio across years for each site}
(stability_plot <- ggplot(all_data_test, aes(x = year(survey_date), y = urch_sn_ratio,
                                                 colour = depth, group = depth))+
  geom_point(size = 2)+
  theme.figure()+
  scale_color_viridis_c() +
   facet_wrap(~site_code, ncol = 10)+
    scale_y_continuous(breaks = c(-3, -2, -1, 0, 1, 2, 3)))
```


```{r Saving the plot so it is more visible, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/stability_plot.png"), width = 20, height = 10, 
    units = "in", pointsize = 15, res = 600)

stability_plot

dev.off()
```


##Adding in the 2024 data
```{r Loading in the 2024 csv}
RLS_data_2024 <- read_csv(here("./01_raw_data/RLS_CANADA_Bamfield_2024.csv"))
```

```{r Removing first row and unnecessary columns}
#factor_cols <- c("site_code", "site_name", "Code")

RLS_data_2024_clean <- RLS_data_2024%>%
  select(c(`Site No.`, `Site Name`, Latitude, Longitude, Date, vis, Time, Depth, Method, Block, Code, Species, Total))%>%
  na.omit()%>%
  rename(site_code = `Site No.`,
         site_name = `Site Name`,
         latitude = Latitude,
         longitude = Longitude,
         survey_date = Date,
         depth = Depth,
         visibility = vis,
         hour = Time,
         total = Total)%>%
   mutate_at(factor_cols, factor)%>%
   group_by(site_code, site_name, latitude, longitude, survey_date, depth, 
           visibility, hour, Code)%>%
  summarise(species_total = sum(total))
```

```{r removing other species and calculating urchin ratio}
 urch_sn_ratio_RLS_2024 <- RLS_data_2024_clean%>%
  pivot_wider(names_from = Code, values_from = species_total)%>%
  mutate(Urchins = mfra,
         Turban_Snails = pgi)%>%
  select(site_code, site_name, latitude, longitude, survey_date, depth, 
           visibility, hour, Urchins, Turban_Snails)%>%
  replace(is.na(.), 0)%>%
  mutate(urch_sn_ratio = log10((Urchins +1) / (Turban_Snails + 1)))%>%
  mutate(survey_date = dmy(survey_date))
```

```{r Making a new data frame with the 2024 data added}
all_data_test <- rbind(urch_sn_ratio_RLS, urch_sn_ratio_RLS_2024)
```


##Wave exposure calculations

Add in the proper lat and long for each of the sites

###Calculating fetch for each of the sites
```{r Making a df of the site coordinates and adjusting the coordinates of sites on land}
site_coordinates_corrected <- site_map_data%>%
  select(c(latitude, longitude))%>%
 mutate(latitude = if_else(row_number() == 7, 48.869, latitude))%>%
 mutate(longitude = if_else(row_number() == 7, -125.319, longitude))%>%
      mutate(latitude = if_else(row_number() == 9, 48.88, latitude))%>%
      mutate(longitude = if_else(row_number() == 9, -125.312, longitude))%>%
  mutate(latitude = if_else(row_number() == 12, 48.911, latitude))%>%
  mutate(longitude = if_else(row_number() == 12, -125.270, longitude))%>%
     mutate(latitude = if_else(row_number() == 14, 48.851, latitude))%>%
     mutate(longitude = if_else(row_number() == 14, -125.199, longitude))%>%
  mutate(latitude = if_else(row_number() == 24, 48.829, latitude))%>%
  mutate(longitude = if_else(row_number() == 24, -125.19, longitude))%>%
     mutate(latitude = if_else(row_number() == 32, 48.831, latitude))%>%
     mutate(longitude = if_else(row_number() == 32, -125.152, longitude))
```

Sites 7, 9, 12, 14, 24, 32 need to be shifted


```{r Converting site coordinates into spatial points file type}
site_sp <- st_as_sf(site_coordinates_corrected, coords = c("longitude", "latitude"), crs = 4326)
bearings <- seq(1, 360, by = 1)
```

```{r Calculating the fetch for the points}
fetch_df <- fetch_len_multi(site_sp, bearings, dmax = 30000, shoreline = VI_base_layer_sf, projected = TRUE)
```


###Using the compass rose to visualize the fetch for each site

##Making plots for the presentation
```{r}
fake_data <- read_csv(here("./01_raw_data/fake_data.csv"))
```

```{r Making a plot with the fake data}
(fake_data_plot <- ggplot(fake_data, aes(x = Explanatory, y = Response))+
  geom_point()+
  geom_smooth(method = "lm")+
  theme.figure())
```


```{r Saving the plot}
png(filename = here("./04_plots/fake_data_plot.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

fake_data_plot

dev.off()
```


```{r Making fake data for predicted results of fetch}
x1 <- seq(100, 900, by = 100)
y1 <- seq(-1, 3 , by = 0.5)
predicted_data <- cbind(x1, y1)
```

```{r Plotting fake data for fetch}
(predicted_plot <- ggplot(predicted_data, aes(x = x1, y = y1))+
  geom_point()+
           geom_smooth(method = "lm")+
  theme.figure()+
   labs(x = "Kelp abundance", y = "Log urchin to snail ratio")+
   theme(axis.title = element_text(size = 18, face = "bold"),
    axis.text = element_text(size = 18, face = "plain")))
```



```{r Making fake data for substrate type, include=FALSE}
fake_substrate_type_data <- read_csv(here("./01_raw_data/example_year_data.csv"))
fake_substrate_type_data$substrate <- as.factor(fake_substrate_type_data$substrate)
levels(fake_substrate_type_data$substrate) <- list(Sand = "Sand", 
                                                   `Loose rocks` = "Loose rocks",
                                                   `Solid rock` = "Solid rock")
```

```{r Plotting the fake substrate data}
(fake_substrate_plot <- ggplot(fake_substrate_type_data, aes(x = substrate, y = urchin))+
   geom_boxplot()+
   theme.figure()+
   labs(x = "\nSubstrate type", y = "Log urchin to snail ratio\n")+
      theme(axis.title = element_text(size = 18, face = "bold"),
    axis.text = element_text(size = 18, face = "plain")))
```



```{r Making fake kelp data, include=FALSE}
fake_kelp_data <- read_csv(here("./01_raw_data/example_year_data.csv"))
```

```{r Plotting the fake kelp data}
(fake_kelp_plot <- ggplot(fake_kelp_data, aes(x = kelp, y = urchin))+
  geom_point()+
           geom_smooth(method = "lm")+
   labs(x = "\nKelp abundance (n/m2)", y = "Log urchin to snail ratio\n")+
   theme.figure()+
   theme(axis.title = element_text(size = 18, face = "bold"),
    axis.text = element_text(size = 18, face = "plain")))
```


###Making an example of yearly fluctuations
```{r Saving in the fake year data}
year_flux_data <- read_csv(here("./01_raw_data/example_year_data.csv"))
```

```{r}
(year_flux_snails <- ggplot()+
  geom_point(year_flux_data, mapping = aes(x = year, y = Urchins),
             size = 4)+
labs(x = " \nYear", y = "Counts\n ") +
  theme.figure()+
   theme(axis.title = element_text(size = 18, face = "bold"),
    axis.text = element_text(size = 18, face = "plain")))
```

