---
title: "Preliminary Analyses of the urchins/snails"
author: "Alec"
date: "2024-09-11"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)

```

#Preliminary analyses of the RLS data from Barkley Sound

##Loading in the data and organizing it
```{r Setting up the here package}
here::here()
```

###Loading in the 2021 data

```{r loading in the csv}
data_2021_raw <- read_csv(here("./01_raw_data/RLS_CANADA_Bamfield_2021.csv"))
```

```{r removing blank row and column}
data_2021_no_blanks <- data_2021_raw%>%
  select(Diver, Buddy, `Site No.`, `Site Name`, Latitude, Longitude, Date, vis, 
         Direction, Time, `P-Qs`, Depth, Method, Block, Code, Species,
         `Common name`, Total)%>%
  na.omit()
```

```{r Changing certain variables to factors}
cols_2021 <- c("Site No.", "Site Name", "Method", "Block", "Code", "Species")

data_2021_clean <- data_2021_no_blanks %>%
  mutate_at(cols_2021, factor)
```

```{r adding totals together from each block and filtering for just M2}
data_2021_clean_for_testing <- data_2021_clean%>%
  filter(Method == 2)%>%
  group_by(`Site No.`, `Site Name`, Latitude, Longitude, Date, vis, Direction, 
           Time, Depth, Code)%>%
  summarise(species_total = sum(Total))
```

###Loading in the 2022 data

```{r loading in the csv}
data_2022_raw <- read_csv(here("./01_raw_data/RLS_CANADA_Bamfield_2022.csv"))
```

```{r removing blank row and column}
data_2022_no_blanks <- data_2022_raw%>%
  select(Diver, Buddy, `Site No.`, `Site Name`, Latitude, Longitude, Date, vis, 
         Direction, Time, `P-Qs`, Depth, Method, Block, Code, Species,
         `Common name`, Total)
  #na.omit()
```

```{r Changing certain variables to factors}
cols_2022 <- c("Site No.", "Site Name", "Method", "Block", "Code", "Species")

data_2022_clean <- data_2022_no_blanks %>%
  mutate_at(cols_2022, factor)
```

```{r adding totals together from each block and filtering for just M2}
data_2022_clean_for_testing <- data_2022_clean%>%
  filter(Method == 2)%>%
  group_by(`Site No.`, `Site Name`, Latitude, Longitude, Date, vis, Direction,
           Depth, Code)%>%
  summarise(species_total = sum(Total))
```


###Loading in the 2023 data
```{r loading in the csv}
data_2023_raw <- read_csv(here("./01_raw_data/RLS_CANADA_Bamfield_2023.csv"))
```

```{r removing blank row and column}
data_2023_no_blanks <- data_2023_raw%>%
  select(Diver, Buddy, `Site No.`, `Site Name`, Latitude, Longitude, Date, vis, 
         Direction, Time, `P-Qs`, Depth, Method, Block, Code, Species,
         `Common name`, Total)
  #na.omit()
```

```{r Changing certain variables to factors}
cols_2023 <- c("Site No.", "Site Name", "Method", "Block", "Code", "Species")

data_2023_clean <- data_2023_no_blanks %>%
  mutate_at(cols_2023, factor)
```

```{r replacing the code for M. franciscanus and P. gibberosus so it matches}
data_2023_clean$Code <- recode_factor(data_2023_clean$Code, mfra = "Mfr",
                                      pgi = "Pgi")
```

```{r adding totals together from each block and filtering for just M2}
data_2023_clean_for_testing <- data_2023_clean%>%
  filter(Method == 2)%>%
  group_by(`Site No.`, `Site Name`, Latitude, Longitude, Date, vis, Direction, 
           Time, Depth, Code)%>%
  summarise(species_total = sum(Total))
```

##Early analysis of the data

###Comparing the ratios of snails to urchins at each site

This analysis involves eliminating the counts from methods 1 and 0 and then
summing the urchins and snails counted between Block 1 and 2 for each transect.
Then the total number of snails was divided by the total number of urchins for
each transect.

####In 2021

```{r calculating and making a column showing the snail to urchin ratio at each site}
sn_urch_ratio_2021 <- data_2021_clean_for_testing%>%
  pivot_wider(names_from = Code, values_from = species_total)%>%
  mutate(sn_urch_ratio = Pgi / Mfr,
         Urchins = Mfr,
         Snails = Pgi)%>%
  select(`Site No.`, `Site Name`, Latitude, Longitude,
         Date, vis, Direction, Time, Depth, sn_urch_ratio, Snails, Urchins)
```


####In 2022

Eliminated time to get all 25 transects correct. Check with Dom about this.

```{r calculating and making a column showing the snail to urchin ratio at each site}
sn_urch_ratio_2022 <- data_2022_clean_for_testing%>%
  pivot_wider(names_from = Code, values_from = species_total)%>%
  mutate(sn_urch_ratio = Pgi / Mfr,
         Urchins = Mfr,
         Snails = Pgi)%>%
  select(`Site No.`, `Site Name`, Latitude, Longitude,
         Date, vis, Direction, Depth, sn_urch_ratio, Snails, Urchins)
```

####In 2023

Needs to be checked for the rows that have species missing and for duplicates 
of the same transect

```{r calculating and making a column showing the snail to urchin ratio at each site}
sn_urch_ratio_2023 <- data_2023_clean_for_testing%>%
  pivot_wider(names_from = Code, values_from = species_total)%>%
  mutate(sn_urch_ratio = Pgi / Mfr,
         Urchins = Mfr,
         Snails = Pgi)%>%
  select(`Site No.`, `Site Name`, Latitude, Longitude,
         Date, vis, Direction, Time, Depth, sn_urch_ratio, Snails, Urchins)
```



###Saving the ratio data frames as .csv

####2021
```{r}
write_csv(sn_urch_ratio_2021, "./03_clean_data/sn_urch_ratio_2021.csv")
```

####2022
```{r}
write_csv(sn_urch_ratio_2022, "./03_clean_data/sn_urch_ratio_2022.csv")
```

####2023
```{r}
write_csv(sn_urch_ratio_2023, "./03_clean_data/sn_urch_ratio_2023.csv")
```

###Plotting the sn urch data

####2021

```{r Urchins to snails}
sn_urch_plot_2021 <- ggplot(sn_urch_ratio_2021, aes(x = Snails, y = Urchins))+
  geom_point()
 # geom_smooth(method = "lm")

sn_urch_plot_2021
```

```{r Plotting the sn_urch ratio against longitude}
sn_urch_plot_long_2021 <- ggplot(sn_urch_ratio_2021, aes(x = Longitude, 
                                                         y = sn_urch_ratio))+
  geom_point()+
 geom_smooth(method = "lm")

sn_urch_plot_long_2021
```

```{r Plotting the sn_urch ratio against latitude}
sn_urch_plot_lat_2021 <- ggplot(sn_urch_ratio_2021, aes(x = Latitude, 
                                                         y = sn_urch_ratio))+
  geom_point()
 # geom_smooth(method = "lm")

sn_urch_plot_lat_2021
```

#####Testing 2021 without effingham west
```{r FIltering out effingham west}
sn_urch_2021_no_EW <- sn_urch_ratio_2021%>%
  filter(`Site No.` != "BMSC13")

sn_urch_2021_no_EW_plot <- ggplot(sn_urch_2021_no_EW, aes(x = Snails, y = Urchins))+
  geom_point()+
 geom_smooth(method = "lm")

sn_urch_2021_no_EW_plot
```


####2022
```{r Urchins to snails}
sn_urch_plot_2022 <- ggplot(sn_urch_ratio_2022, aes(x = Snails, y = Urchins))+
  geom_point()
 # geom_smooth(method = "lm")

sn_urch_plot_2022
```

```{r Plotting the sn_urch ratio against longitude}
sn_urch_plot_long_2022 <- ggplot(sn_urch_ratio_2022, aes(x = Longitude, 
                                                         y = sn_urch_ratio))+
  geom_point()+
 geom_smooth(method = "lm")

sn_urch_plot_long_2022
```

```{r Plotting the sn_urch ratio against latitude}
sn_urch_plot_lat_2022 <- ggplot(sn_urch_ratio_2022, aes(x = Latitude, 
                                                         y = sn_urch_ratio))+
  geom_point()
 # geom_smooth(method = "lm")

sn_urch_plot_lat_2021
```

####2023

This needs to be done again once the 2023 data is cleaned up
```{r Urchins to snails}
sn_urch_plot_2023 <- ggplot(sn_urch_ratio_2023, aes(x = Snails, y = Urchins))+
  geom_point()
 # geom_smooth(method = "lm")

sn_urch_plot_2023
```


###