---
title: "Preliminary Analysis with RLS data"
author: "Alec"
date: "2024-09-13"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(lubridate)
library(sf) #for spatial data
library(cowplot)
library(rnaturalearth)
library(ggspatial)
library(ggrepel)
library(waver)
library(vcd)
library(sp)
library(viridisLite)
library(dplyr)
library(raster)
library(RColorBrewer)
library(conflicted)

conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
```

#Preliminary analyses of the RLS data from Barkley Sound

#2023 Analyses
##Loading in the data and organizing it
```{r Setting up the here package}
here::here()
```

```{r Loading in the invertebrate file, include=FALSE}
invert_data_raw <- read_csv(here("./01_raw_data/IMOS_-_National_Reef_Monitoring_Network_Sub-Facility_-_Global_mobile_macroinvertebrate_abundance_no_top.csv"))
```

```{r Cleaning up columns and converting certain columns to factors}
factor_cols <- c("site_code", "site_name", "reporting_name")

invert_data_clean <- invert_data_raw%>%
  select(-c(FID, survey_id, country, area, ecoregion, realm, location,
            survey_latitude, survey_longitude, biomass, phylum, class, order,
            family, program))%>%
  mutate_at(factor_cols, factor)
```


```{r Summing the total numbers of animals per transect, include=FALSE}
invert_data_final_clean <- invert_data_clean%>%
  group_by(site_code, site_name, latitude, longitude, survey_date, depth, 
           visibility, hour, reporting_name)%>%
  summarise(species_total = sum(total))
```


####Correcting the lat and longs of the sites
```{r Correcting the lat and long to be correct}
invert_data_final_clean_updated_lat_long <- invert_data_final_clean%>%
 mutate(latitude = if_else(site_code == "BMSC1", 48.82894897, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC1", -125.1975708, longitude))%>%
  
   mutate(latitude = if_else(site_code == "BMSC2", 48.85039902, latitude))%>%
   mutate(longitude = if_else(site_code == "BMSC2", -125.1987686, longitude))%>%
  
 mutate(latitude = if_else(site_code == "BMSC3", 48.85558319, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC3", -125.1837997, longitude))%>%
  
   mutate(latitude = if_else(site_code == "BMSC4", 48.81511688, latitude))%>%
   mutate(longitude = if_else(site_code == "BMSC4", -125.1753311, longitude))%>%
  
 mutate(latitude = if_else(site_code == "BMSC5", 48.82733154, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC5", -125.1966019, longitude))%>%
  
   mutate(latitude = if_else(site_code == "BMSC6", 48.95023346, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC6", -125.1555481, longitude))%>%
  
  mutate(latitude = if_else(site_code == "BMSC7", 48.954644, latitude))%>%
  mutate(longitude = if_else(site_code == "BMSC7", -125.153993, longitude))%>%
  
    mutate(latitude = if_else(site_code == "BMSC8", 48.95508194, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC8", -125.1533737, longitude))%>%
  
    mutate(latitude = if_else(site_code == "BMSC9", 48.83478928, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC9", -125.1470261, longitude))%>%
  
 mutate(latitude = if_else(site_code == "BMSC10", 48.87051773, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC10", -125.160347, longitude))%>%
  
 mutate(latitude = if_else(site_code == "BMSC11", 48.85746765, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC11", -125.1582336, longitude))%>%
  
  mutate(latitude = if_else(site_code == "BMSC12", 48.858284, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC12", -125.1609192, longitude))%>%
  
mutate(latitude = if_else(site_code == "BMSC13", 48.8650322, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC13", -125.3137207, longitude))%>%
  
mutate(latitude = if_else(site_code == "BMSC14", 48.87908173, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC14", -125.2974014, longitude))%>%
  
 mutate(latitude = if_else(site_code == "BMSC15", 48.88028336, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC15", -125.3128815, longitude))%>% 
  
 mutate(latitude = if_else(site_code == "BMSC16", 48.89070129, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC16", -125.300499, longitude))%>% 
  
 mutate(latitude = if_else(site_code == "BMSC17", 48.86548233, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC17", -125.3614807, longitude))%>% 
  
mutate(latitude = if_else(site_code == "BMSC18", 48.91161728, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC18", -125.2670364, longitude))%>% 
  
 mutate(latitude = if_else(site_code == "BMSC19", 48.82860184, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC19", -125.2212982, longitude))%>%  
  
 mutate(latitude = if_else(site_code == "BMSC20", 48.83566666, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC20", -125.214798, longitude))%>%  
  
mutate(latitude = if_else(site_code == "BMSC21", 48.85205078, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC21", -125.1235657, longitude))%>%    
  
 mutate(latitude = if_else(site_code == "BMSC22", 48.85426712, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC22", -125.1170349, longitude))%>%    
  
  mutate(latitude = if_else(site_code == "BMSC23", 48.837589, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC23", -125.144145, longitude))%>%     
  
    mutate(latitude = if_else(site_code == "BMSC24", 48.916073, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC24", -125.131174, longitude))%>%
  
     mutate(latitude = if_else(site_code == "BMSC24", 48.916073, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC24", -125.131174, longitude))%>%
  
      mutate(latitude = if_else(site_code == "BMSC25", 48.838595, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC25", -125.135015, longitude))%>% 
  
        mutate(latitude = if_else(site_code == "BMSC26", 48.9071, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC26", -125.037017, longitude))%>% 
  
     mutate(latitude = if_else(site_code == "BMSC27", 48.901183, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC27", -125.060433, longitude))
  
```


```{r Writing a .csv for the clean data with the counts of each species/transect}
write_csv(invert_data_final_clean_updated_lat_long,
          here("./03_clean_data/invert_data_final_clean_updated_lat_long.csv"))
```




##Calculating urch to sn ratio and getting rid of other species
```{r Loading in the new clean df and making sure the right columns are factors, include=FALSE}
invert_data_final_clean_updated_lat_long <- read_csv(here("./03_clean_data/invert_data_final_clean_updated_lat_long.csv"))

factor_cols <- c("site_code", "site_name", "reporting_name")

invert_data_final_clean_updated_lat_long <- invert_data_final_clean_updated_lat_long%>%
  mutate_at(factor_cols, factor)
```



```{r removing other species and calculating urchin ratio}
 urch_sn_ratio_RLS <- invert_data_final_clean_updated_lat_long%>%
  pivot_wider(names_from = reporting_name, values_from = species_total)%>%
  mutate(Urchins = `Mesocentrotus franciscanus`,
         Turban_Snails = `Pomaulax gibberosus`)%>%
  select(site_code, site_name, latitude, longitude, survey_date, depth, 
           visibility, hour, Urchins, Turban_Snails)%>%
  replace(is.na(.), 0)%>%
  mutate(urch_sn_ratio = log10((Urchins +1) / (Turban_Snails + 1)))
```

##Making some visualizations of the urch_sn ratio and correlation between urch and sn
###Overall across years
```{r Plotting urch against sn overall and colouring by year}
(urch_sn_plot <- ggplot(urch_sn_ratio_RLS, aes(x = Turban_Snails, y = Urchins,
                                               colour = as.factor(year(survey_date))))+
  geom_point()+
  theme.figure()+
  labs(colour = "Year"))
 #geom_smooth(method = "lm")
```

###In 2021
```{r Plotting number of urchins vs snails in 2021}
(urch_sn_plot_2021 <-ggplot(urch_sn_ratio_RLS%>%
                              filter(year(survey_date) == "2021"),
  aes(x = Turban_Snails, y = Urchins))+
  geom_point()+
  theme.figure())
 #geom_smooth(method = "lm")
```

###In 2022
```{r Plotting number of urchins vs snails in 2022}
(urch_sn_plot_2022 <-ggplot(urch_sn_ratio_RLS%>%
                              filter(year(survey_date) == "2022"),
  aes(x = Turban_Snails, y = Urchins))+
  geom_point()+
  theme.figure())
 #geom_smooth(method = "lm")
```

###In 2023
```{r Plotting number of urchins vs snails in 2023}
(urch_sn_plot_2023 <-ggplot(urch_sn_ratio_RLS%>%
                              filter(year(survey_date) == "2023"),
  aes(x = Turban_Snails, y = Urchins))+
  geom_point()+
  theme.figure())
 #geom_smooth(method = "lm")
```




##Making maps of the data in Barkley Sound

###Making a theme for all subsequent figures
```{r Making theme for all subsequent figures}
theme.figure <- function (){
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(colour = "black"),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12, face = "plain"),
    legend.text = element_text(size = 10, face = "plain"),
    legend.title = element_text(size = 14, face = "bold")) 
}

```

###Doing the set up for the base layer and map characteristics
```{r Loading in the base layer}
VI_base_layer <- read_sf(dsn = here("./01_raw_data/canada_sf_data", ""), 
                         stringsAsFactors = F)
VI_base_layer_sf <- st_as_sf(VI_base_layer)

#ggplot(VI_base_layer_sf)+
  #geom_sf()

```

```{r Specifying the projection we want to use for our data}
st_crs(VI_base_layer_sf) <- "+proj=longlat +datum=WGS84 +no_defs"
```

```{r Making the range for the map}
#first need to specify the range of the map
lat_min = min(urch_sn_ratio_RLS$latitude) - 0.05
lat_max = max(urch_sn_ratio_RLS$latitude) + 0.05

long_min = min(urch_sn_ratio_RLS$longitude) - 0.05
long_max = max(urch_sn_ratio_RLS$longitude) + 0.05

lat_min_launch = min(urch_sn_ratio_RLS$latitude) - 0.01    #The launch ones are for the boat launch plot
lat_max_launch = max(urch_sn_ratio_RLS$latitude) + 0.06

long_min_launch = min(urch_sn_ratio_RLS$longitude) - 0.2
long_max_launch = max(urch_sn_ratio_RLS$longitude) + 0.1
```


###Making a site map

```{r Making a df with just one of each site}
site_map_data <- invert_data_final_clean_updated_lat_long%>%
  distinct(site_code, site_name, latitude, longitude)
```


```{r Making the map with the site names, fig.width=7.5, fig.height=5.5}
(RLS_site_name_map <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  labs(x = " \nLongitude", y = "Latitude\n ") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1)) +
  scale_y_continuous(breaks = c(48.8, 48.85, 48.9))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
geom_label_repel(
    data = site_map_data, 
    mapping = aes(x = longitude, y = latitude, label = site_name), 
    size = 2,
   # box.padding = 0.35,  # Add padding between labels and points
  #  point.padding = 0.5, # Add padding between points and text
    segment.color = 'grey50', # Line from label to point
    max.overlaps = 15      # Adjust maximum number of overlaps
  ))
```


```{r Making the map with the site codes, fig.width=7.5, fig.height=5.5}
(RLS_site_code_map <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  labs(x = " \nLongitude", y = "Latitude\n ") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
geom_label_repel(
    data = site_map_data, 
    mapping = aes(x = longitude, y = latitude, label = site_code), 
    size = 2,
   # box.padding = 0.35,  # Add padding between labels and points
  #  point.padding = 0.5, # Add padding between points and text
    segment.color = 'grey50', # Line from label to point
    max.overlaps = 15      # Adjust maximum number of overlaps
  ))
```


```{r Making a data frame of the boat launch sites}
launch_lat <- c(48.878415, 48.828809, 49.012965, 48.946169)
launch_long <- c(-125.046749, -125.127513, -125.360924, -125.547637)
names <- c("Poat Nook", "Grappler Inlet", "Toquart Bay", "Ucluelet")
launch_df <- tibble(launch_long, launch_lat, names)
```



```{r Making map with just the points and the boat launches}
(RLS_site_point_map <- ggplot()+
       geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min_launch, long_max_launch),
           ylim = c(lat_min_launch, lat_max_launch)) +
  labs(x = " \nLongitude", y = "Latitude\n ") +
  theme.figure() +
  #scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
 # scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering)+ 
    geom_point(data = site_map_data, mapping = aes(x = longitude, y = latitude),
             size = 3, colour = "#29AF7FFF") + 
  geom_point(data = site_map_data, mapping = aes(x = longitude, y = latitude),
             shape = 1, size = 3,colour = "black")+
      geom_point(data = launch_df, mapping = aes(x = launch_long, y = launch_lat),
                 size = 3, colour = "#FDE725FF")+
      geom_point(data = launch_df, mapping = aes(x = launch_long, y = launch_lat),
                shape = 1, size = 3, colour = "black"))
```


```{r Making the inset map}
(van_isl_map <- ggplot(VI_base_layer_sf) +
  geom_sf() +
  coord_sf(ylim = c(48, 52), xlim = c(-129, -122)) +
  theme.figure() +
 # geom_rect(xmin = long_min, xmax = long_max, ymin = lat_min, ymax = lat_max,
            #color = "red", fill = NA, linewidth = 1) +
  labs(x = "Longitude", y = "Latitude") +
    theme(axis.text = element_text(size = 14))+
  scale_x_continuous(breaks = c(-128, -125, -122)) +
  scale_y_continuous(breaks = c(48, 50, 52)))
```



###Making maps of the urch to sn ratio for each year

```{r Making the plot with the urch to sn ratio on the map 2021}
(urch_sn_ratio_map_2021 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2021"), mapping = aes(x = longitude, y = latitude, color = urch_sn_ratio),
             size = 3) + 
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2021"), mapping = aes(x = longitude, y = latitude, color = urch_sn_ratio),
             shape = 1, size = 3,colour = "black")+
  scale_color_viridis_c() +
  labs(title = "2021", x = " \nLongitude", y = "Latitude\n ", color = "Log Urchin to Snail Ratio") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))

```

```{r Making the plot with the urch to sn ratio on the map 2022}
(urch_sn_ratio_map_2022 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2022"), mapping = aes(x = longitude, y = latitude, color = urch_sn_ratio),
             size = 3,) + 
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2022"), mapping = aes(x = longitude, y = latitude, color = urch_sn_ratio),
             shape = 1, size = 3,colour = "black")+
  scale_color_viridis_c() +
  labs(title = "2022", x = " \nLongitude", y = "Latitude\n ", color = " Log Urchin to Snail Ratio") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
 # geom_label(data = urch_sn_ratio_RLS, 
             #mapping = aes(x = -125.134813, y = 48.834945, label = "BMSC")))

```


```{r Making the plot with the urch to sn ratio on the map 2023}
(urch_sn_ratio_map_2023 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2023"), mapping = aes(x = longitude, y = latitude, color = urch_sn_ratio),
             size = 3,) + 
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2023"), mapping = aes(x = longitude, y = latitude, color = urch_sn_ratio),
             shape = 1, size = 3,colour = "black")+
  scale_color_viridis_c() +
  labs(title = "2023", x = " \nLongitude", y = "Latitude\n ", color = " Log Urchin to Snail Ratio") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
 # geom_label(data = urch_sn_ratio_RLS, 
             #mapping = aes(x = -125.134813, y = 48.834945, label = "BMSC")))

```




###Plotting the urchin counts on the map
```{r Plot with urchin counts on map 2021}
(urch_counts_map_2021 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2021"), mapping = aes(x = longitude, y = latitude, color = Urchins),
             size = 2,) + 
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2021"), mapping = aes(x = longitude, y = latitude, color = Urchins),
             shape = 1, size = 2,colour = "black")+
  scale_color_gradient(low = "#35A7FF", high = "#FF5A5F" ) +
  labs(title = "2021", x = " \nLongitude", y = "Latitude\n ", color = "Urchin Counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1)) +
  scale_y_continuous(breaks = c(48.8, 48.85, 48.9))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
 # geom_label(data = urch_sn_ratio_RLS, 
             #mapping = aes(x = -125.134813, y = 48.834945, label = "BMSC")))
```


```{r Plot with Urchin counts 2022}
(urch_counts_map_2022 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2022"), mapping = aes(x = longitude, y = latitude, color = Urchins),
             size = 2,) + 
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2022"), mapping = aes(x = longitude, y = latitude, color = Urchins),
             shape = 1, size = 2,colour = "black")+
  scale_color_gradient(low = "#35A7FF", high = "#FF5A5F" ) +
  labs(title = "2022", x = " \nLongitude", y = "Latitude\n ", color = "Urchin Counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1)) +
  scale_y_continuous(breaks = c(48.8, 48.85, 48.9))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
 # geom_label(data = urch_sn_ratio_RLS, 
             #mapping = aes(x = -125.134813, y = 48.834945, label = "BMSC")))
```



```{r Plot of Urchin counts 2023}
(urch_counts_map_2023 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2023"), mapping = aes(x = longitude, y = latitude, color = Urchins),
             size = 2,) + 
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2023"), mapping = aes(x = longitude, y = latitude, color = Urchins),
             shape = 1, size = 2,colour = "black")+
  scale_color_gradient(low = "#35A7FF", high = "#FF5A5F" ) +
  labs(title = "2023", x = " \nLongitude", y = "Latitude\n ", color = "Urchin Counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1)) +
  scale_y_continuous(breaks = c(48.8, 48.85, 48.9))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
 # geom_label(data = urch_sn_ratio_RLS, 
             #mapping = aes(x = -125.134813, y = 48.834945, label = "BMSC")))
```



###Plotting the snail counts on the map

```{r Plot of snail counts 2021}
(snail_counts_map_2021 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2021"), mapping = aes(x = longitude, y = latitude, color = Turban_Snails),
             size = 2,) + 
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2021"), mapping = aes(x = longitude, y = latitude, color = Turban_Snails),
             shape = 1, size = 2,colour = "black")+
  scale_color_gradient(low = "#35A7FF", high = "#FF5A5F" ) +
  labs(title = "2021", x = " \nLongitude", y = "Latitude\n ", color = "Red Turban Snail Counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1)) +
  scale_y_continuous(breaks = c(48.8, 48.85, 48.9))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))

```


```{r Plot of Snail counts 2022}
(snail_counts_map_2022 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2022"), mapping = aes(x = longitude, y = latitude, color = Turban_Snails),
             size = 2,) + 
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2022"), mapping = aes(x = longitude, y = latitude, color = Turban_Snails),
             shape = 1, size = 2,colour = "black")+
  scale_color_gradient(low = "#35A7FF", high = "#FF5A5F" ) +
  labs(title = "2022", x = " \nLongitude", y = "Latitude\n ", color = "Red Turban Snail Counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1)) +
  scale_y_continuous(breaks = c(48.8, 48.85, 48.9))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))

```


```{r Plot of Snail counts 2023}
(snail_counts_map_2023 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2023"), mapping = aes(x = longitude, y = latitude, color = Turban_Snails),
             size = 2,) + 
  geom_point(data = urch_sn_ratio_RLS%>%
               filter(year(survey_date) == "2023"), mapping = aes(x = longitude, y = latitude, color = Turban_Snails),
             shape = 1, size = 2,colour = "black")+
  scale_color_gradient(low = "#35A7FF", high = "#FF5A5F" ) +
  labs(title = "2023", x = " \nLongitude", y = "Latitude\n ", color = "Red Turban Snail Counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1)) +
  scale_y_continuous(breaks = c(48.8, 48.85, 48.9))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
```


###Saving all of the figures so far

####Site maps
```{r Saving the site name map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/site_name_map.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

RLS_site_name_map

dev.off()
```

```{r Saving the site code map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/site_code_map.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

RLS_site_code_map

dev.off()
```


```{r Saving the site point map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/site_point_map.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

RLS_site_point_map

dev.off()
```

```{r Saving the Vancouver Island map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/van_isl_map.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

van_isl_map

dev.off()
```

####Urch Sn ratio maps
```{r Saving the 2021 urch sn ratio map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/urch_sn_ratio_map_2021.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

urch_sn_ratio_map_2021

dev.off()
```

```{r Saving the 2022 urch sn ratio map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/urch_sn_ratio_map_2022.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

urch_sn_ratio_map_2022

dev.off()
```

```{r Saving the 2023 urch sn ratio map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/urch_sn_ratio_map_2023.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

urch_sn_ratio_map_2023

dev.off()
```

```{r Saving the 2024 urch sn ratio map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/urch_sn_ratio_map_2024.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

urch_sn_ratio_map_2024

dev.off()
```

####Urch counts maps
```{r Saving the 2021 urch counts map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/urch_counts_map_2021.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

urch_counts_map_2021

dev.off()
```

```{r Saving the 2022 urch counts map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/urch_counts_map_2022.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

urch_counts_map_2022

dev.off()
```

```{r Saving the 2023 urch counts map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/urch_counts_map_2023.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

urch_counts_map_2023

dev.off()
```


####Snail counts maps
```{r Saving the 2021 snail counts map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/snail_counts_map_2021.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

snail_counts_map_2021

dev.off()
```

```{r Saving the 2022 snail counts map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/snail_counts_map_2022.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

snail_counts_map_2022

dev.off()
```


```{r Saving the 2023 snail counts map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/snail_counts_map_2023.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

snail_counts_map_2023

dev.off()
```




##Wave exposure calculations

Add in the proper lat and long for each of the sites

###Calculating fetch for each of the sites
```{r Making a df of the site coordinates and adjusting the coordinates of sites on land}
site_coordinates_corrected <- site_map_data%>%
  select(c(latitude, longitude))
 #mutate(latitude = if_else(row_number() == 7, 48.869, latitude))%>%
# mutate(longitude = if_else(row_number() == 7, -125.319, longitude))%>%
 #     mutate(latitude = if_else(row_number() == 9, 48.88, latitude))%>%
  #    mutate(longitude = if_else(row_number() == 9, -125.312, longitude))%>%
  #mutate(latitude = if_else(row_number() == 12, 48.911, latitude))%>%
  #mutate(longitude = if_else(row_number() == 12, -125.270, longitude))%>%
   #  mutate(latitude = if_else(row_number() == 14, 48.851, latitude))%>%
    # mutate(longitude = if_else(row_number() == 14, -125.199, longitude))%>%
#  mutate(latitude = if_else(row_number() == 24, 48.829, latitude))%>%
 # mutate(longitude = if_else(row_number() == 24, -125.19, longitude))%>%
  #   mutate(latitude = if_else(row_number() == 32, 48.831, latitude))%>%
   #  mutate(longitude = if_else(row_number() == 32, -125.152, longitude))
```

Sites 3, 13, 14, 15, 16, 24, 27, 29, 32 need to be shifted


```{r Converting site coordinates into spatial points file type}
site_sp <- st_as_sf(site_coordinates_corrected, coords = c("longitude", "latitude"), crs = 4326)
bearings <- seq(1, 1, by = 1)
```

```{r Calculating the fetch for the points}
fetch_df <- fetch_len_multi(site_sp, bearings, dmax = 30000, shoreline = VI_base_layer_sf, projected = TRUE)
```


###Using the compass rose to visualize the fetch for each site


#2024 Analyses
##Adding in the 2024 data
```{r Loading in the 2024 csv}
RLS_data_2024 <- read_csv(here("./01_raw_data/RLS_CANADA_Bamfield_2024.csv"))
```

```{r Removing first row and unnecessary columns}
factor_cols2 <- c("site_code", "site_name", "Code")

RLS_data_2024_clean <- RLS_data_2024%>%
  select(c(`Site No.`, `Site Name`, Latitude, Longitude, Date, vis, Time, Depth,
           Method, Block, Code, Species, Total))%>%
  na.omit()%>%
  mutate(Code = str_to_lower(Code))%>%
  rename(site_code = `Site No.`,
         site_name = `Site Name`,
         latitude = Latitude,
         longitude = Longitude,
         survey_date = Date,
         depth = Depth,
         visibility = vis,
         hour = Time,
         total = Total)%>%
   mutate_at(factor_cols2, factor)%>%
   mutate(longitude = if_else(site_code == "BMSC2", -125.19876860, longitude))%>% # one point at kirby had the lat for both lat and long
  filter(!Code %in% c("dez", "deg", "dem", "deo", "def", "dew", "dec", "nsf", "snd"))%>%
   group_by(site_code, site_name, latitude, longitude, survey_date, depth, 
           visibility, hour, Code)%>%
  summarise(species_total = sum(total))
```

```{r Making a data frame with different lat/longs for transects at same sites}
RLS_data_2024_clean_updated_lat_long <- RLS_data_2024_clean%>%
   mutate(latitude = if_else(site_code == "BMSC1" & depth == 6.5, 48.8290000, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC1" & depth == 6.5, -125.191000, longitude))%>%

mutate(latitude = if_else(site_code == "BMSC11" & depth == 5, 48.8560000, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC11" & depth == 5, -125.1520000, longitude))%>%

mutate(latitude = if_else(site_code == "BMSC12" & depth == 4.5, 48.8620000, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC12" & depth == 4.5, -125.1600000, longitude))%>%

mutate(latitude = if_else(site_code == "BMSC12" & depth == 7, 48.858284, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC12" & depth == 7, -125.1650000, longitude))%>%

mutate(latitude = if_else(site_code == "BMSC24" & depth == 5.5, 48.921000, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC24" & depth == 5.5, -125.131174, longitude))%>%

mutate(latitude = if_else(site_code == "BMSC25" & depth == 5.5, 48.843000, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC25" & depth == 5.5, -125.135015, longitude))%>%

mutate(latitude = if_else(site_code == "BMSC26" & depth == 6.5, 48.91500, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC26" & depth == 6.5, -125.037017, longitude))%>%

mutate(latitude = if_else(site_code == "BMSC5" & depth == 6.0, 48.8230000, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC5" & depth == 6.0, -125.1980000, longitude))%>%

mutate(latitude = if_else(site_code == "BMSC5" & depth == 6.5, 48.82733154, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC5" & depth == 6.5, -125.2010000, longitude))%>%

mutate(latitude = if_else(site_code == "BMSC6" & depth == 8, 48.95023346, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC6" & depth == 8, -125.1620000, longitude))%>%

mutate(latitude = if_else(site_code == "BMSC8" & depth == 7.5, 48.95508194, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC8" & depth == 7.5, -125.1460000, longitude))
```


###Urch sn ratio 2024
```{r removing other species and calculating urchin ratio}
 urch_sn_ratio_RLS_2024 <- RLS_data_2024_clean_updated_lat_long%>%
  pivot_wider(names_from = Code, values_from = species_total)%>%
  mutate(Urchins = mfra,
         Turban_Snails = pgi)%>%
  select(site_code, site_name, latitude, longitude, survey_date, depth, 
           visibility, hour, Urchins, Turban_Snails)%>%
  replace(is.na(.), 0)%>%
  mutate(urch_sn_ratio = log10((Urchins +1) / (Turban_Snails + 1)),
         survey_date = dmy(survey_date))
```

###Urch to all others ratio 2024
```{r Making a data frame of urchin to all other macrofauna ratio}
Urchins <- urch_sn_ratio_RLS_2024$Urchins

urch_all_ratio_RLS_2024 <- RLS_data_2024_clean_updated_lat_long%>%
  dplyr::filter(Code != "mfra")%>%
  group_by(site_code, site_name, latitude, longitude, survey_date, depth)%>%
  summarise(total_counts = sum(species_total))%>%
  cbind(Urchins = Urchins)%>%
  mutate(urch_all_ratio = log10((Urchins + 1) / (total_counts + 1)))
```

##2024 Wave exposure calculations

###Making dataframes for the site lat and longs 2024
```{r making 2024 Site data dataframe}
site_data_2024 <- RLS_data_2024_clean%>%
  distinct(site_code, site_name, latitude, longitude)
```

```{r Site coordinates df}
site_coords_2024 <- site_data_2024%>%
  ungroup()%>%
  select(latitude, longitude)%>%
  distinct()
```

###Adjusting the points to be in the water
```{r}
site_coordinates_corrected_2024 <- site_coords_2024%>%
 mutate(latitude = if_else(row_number() == 1, 48.828689, latitude))%>%
 mutate(longitude = if_else(row_number() == 1, -125.197557, longitude))%>%
  
     mutate(latitude = if_else(row_number() == 11, 48.829336, latitude))%>%
     mutate(longitude = if_else(row_number() == 11, -125.221597, longitude))%>%
  
  mutate(latitude = if_else(row_number() == 12, 48.850541, latitude))%>%
  mutate(longitude = if_else(row_number() == 12, -125.198564, longitude))%>%
  
     mutate(latitude = if_else(row_number() == 13, 48.836203, latitude))%>%
     mutate(longitude = if_else(row_number() == 13, -125.216354, longitude))%>%
  
 mutate(latitude = if_else(row_number() == 14, 48.852253, latitude))%>%
 mutate(longitude = if_else(row_number() == 14, -125.124429, longitude))%>%
  
   mutate(latitude = if_else(row_number() == 17, 48.915916, latitude))%>%
   mutate(longitude = if_else(row_number() == 17, -125.131160, longitude))%>%
  
 mutate(latitude = if_else(row_number() == 22, 48.815872, latitude))%>%
 mutate(longitude = if_else(row_number() == 22, -125.176694, longitude))%>%
  
   mutate(latitude = if_else(row_number() == 24, 48.950505, latitude))%>%
 mutate(longitude = if_else(row_number() == 24, -125.155822, longitude))%>%
  
   mutate(latitude = if_else(row_number() == 26, 48.835115, latitude))%>%
 mutate(longitude = if_else(row_number() == 26, -125.146946, longitude))
```


###Setup for using the waver package




```{r Converting site coordinates into spatial points file type 2024}
site_sp_2024 <- st_as_sf(site_coordinates_corrected_2024, coords = c("longitude", "latitude"), crs = 4326)
bearings <- seq(1, 360, by = 1)
```

###Calculating fetch for all of the sites in 2024
```{r Calculating the fetch for the sites 2024}
fetch_df_2024 <- fetch_len_multi(site_sp_2024, bearings, dmax = 30000, shoreline = VI_base_layer_sf, projected = TRUE)
```

###Turning the raw fetch data into a dataframe with column labels
```{r}
site_code <- c("BMSC1", "BMSC10", "BMSC11", "BMSC12", "BMSC13", "BMSC14", "BMSC15", "BMSC16",
               "BMSC17", "BMSC18", "BMSC19", "BMSC2", "BMSC20", "BMSC21", "BMSC22", "BMSC23", "BMSC24",
               "BMSC25", "BMSC26", "BMSC27", "BMSC3", "BMSC4", "BMSC5", "BMSC6", "BMSC8", "BMSC9")

fetch_df_2024 <- as.data.frame(fetch_df_2024)

fetch_df_2024$site_code <- site_code

fetch_df_2024$latitude <- site_coordinates_corrected_2024$latitude
fetch_df_2024$longitude <- site_coordinates_corrected_2024$longitude

fetch_df_2024_final <- fetch_df_2024[, c(361, 362, 363, seq(1, 360, by = 1))]
```

```{r Converting the data into long format}
fetch_2024_long <- fetch_df_2024_final %>%
  pivot_longer(cols = `1`: `360`,  # Columns for compass degrees (assuming "degree1", "degree2", ...)
               names_to = "degree",           # Name of the new column for degrees
               values_to = "fetch",           # Name of the new column for fetch values
               names_prefix = "degree")       # Remove "degree" prefix from column names if needed

```


###Plotting the fetch for just one site as a test
```{r}
ggplot(fetch_2024_long %>% filter(site_code == "BMSC1"), aes(x = as.numeric(degree), y = fetch,
                                                             fill = fetch)) +
  geom_bar(stat = "identity", position = "identity") +
  coord_polar(start = 0) +  # Convert to polar coordinates
  scale_x_continuous(limits = c(0, 360.5), breaks = seq(45, 360, by = 45)) +  # Optional: Compass degree labels
  scale_fill_viridis_c() +
  labs(title = "Fetch Plot for BMSC1", x = "Compass Degrees", y = "Fetch (m)") +
  theme_minimal()
```


###Plotting the fetch for the sites
```{r Plotting fetch for all the sites, fig.width=50, fig.height=50}
(test_fetch_2024_all_sites_plot <- ggplot(fetch_2024_long, aes(x = as.numeric(degree), y = fetch,
                            fill = fetch)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(start = 0) +
  scale_x_continuous(limits = c(0, 360.5), breaks = seq(45, 360, by = 45)) +
  scale_fill_viridis_c()+
  labs(title = "Fetch Plot for All Sites", x = "Compass Degrees", y = "Fetch (m)",
       fill = "Fetch (m)") +
  theme_minimal() +
  facet_wrap(~ site_code))
```


```{r Saving the 2021 snail counts map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/test_fetch_2024_all_sites_plot.png"), width = 15, height = 15, 
    units = "in", pointsize = 15, res = 700)

test_fetch_2024_all_sites_plot

dev.off()
```


###Plotting the fetch for all of the sites in a way to better visualize where they are
```{r Creating a function to make a fetch plot for each site}
create_fetch_plot <- function(site_data, site_code) {
  ggplot(site_data, aes(x = as.numeric(degree), y = fetch,
                        fill = fetch)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(start = 0) +
  scale_x_continuous(limits = c(0, 360.5), breaks = seq(45, 360, by = 45)) +
  scale_fill_viridis_c()+
  theme_minimal()+
    theme(legend.position = "none")
  ggtitle(site_code)
}
```


```{r Extracting the plots for each individual site}
site_codes <- unique(fetch_2024_long$site_code) #extracting a list of unique site codes

#Create a named list of fetch plots for each site
fetch_plots <- lapply(site_codes, function(site) {
  #filter fetch_2024_long to get data for this specific site
  site_data <- fetch_2024_long%>%
    filter(site_code == site)
  
  #call the function to create the fetch plot for this site
  create_fetch_plot(site_data, site)
})

names(fetch_plots) <- site_codes
```


####Making the base map and site code map

#####Base map
```{r Making the base map plot}
(map_plot <- ggplot() +
  geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey") +
  coord_sf(xlim = c(long_min, long_max), ylim = c(lat_min, lat_max)) +
  labs(x = "Longitude", y = "Latitude")+
   theme.figure()+
     scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
     annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))

```


```{r Saving the base map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/base_map_plot.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

map_plot

dev.off()
```


#####Site code map

#####Making data frame with just one site for each site
```{r}
site_data_2024_final <- site_data_2024%>%
  ungroup%>%
  select(site_code, latitude, longitude)%>%
  distinct()

```



```{r Making the map with the site codes 2024, fig.width=7.5, fig.height=5.5}
(RLS_site_code_map_2024 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  labs(x = " \nLongitude", y = "Latitude\n ") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  )+
geom_label(
    data = site_data_2024_final, 
    mapping = aes(x = longitude, y = latitude, label = site_code), 
    size = 2
   # box.padding = 0.35,  # Add padding between labels and points
  #  point.padding = 0.5, # Add padding between points and text
   # segment.color = 'grey50', # Line from label to point
   # max.overlaps = 15      # Adjust maximum number of overlaps
  ))
```

#####Saving the 2024 site code map

```{r Saving the base map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/RLS_site_code_map_2024.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

RLS_site_code_map_2024

dev.off()
```

####Making the combined plot with the fetch plots on the map

#####Making all the individual fetch plots
```{r}

fetch_plots <- list()

# Example for generating fetch plots for each site
for (site in unique(fetch_2024_long$site_code)) {
  
  site_data <- fetch_2024_long%>%
    filter(site_code == site)
  
 fetch_plot <- ggplot(site_data, 
                      aes(x = as.numeric(degree), y = fetch)) +
  geom_bar(stat = "identity", width = 1, position = "identity", fill = "#3CBB75FF") +
  coord_polar(start = 0) +  # Convert to polar coordinates
  scale_x_continuous(limits = c(0, 360.5), breaks = seq(45, 360, by = 45)) +  # Optional: Compass degree labels
  #scale_fill_viridis_c() +
  #labs(title = paste("Fetch for", site)) +
  theme_minimal() +
   theme(axis.title.y = element_blank(),
         axis.title.x = element_blank(),  # Remove x-axis label (degrees)
      axis.text.x = element_blank(),   # Remove x-axis text
      axis.text.y = element_blank(),   # Remove y-axis text
      axis.ticks.x = element_blank(),  # Remove x-axis ticks
      axis.ticks.y = element_blank()   # Remove y-axis ticks
    )
  
  
  # Store the plot in the list
  fetch_plots[[site]] <- fetch_plot
}

print(fetch_plots[[site]])

# Loop through each site and save the plot
for (site in names(fetch_plots)) {
  # Generate the file name based on the site name
  file_name <- paste0("fetch_plot_", site, ".png")  # You can change the file format to .pdf or .jpeg if needed
  
  # Save the plot for each site
  ggsave(filename = file_name, 
         plot = fetch_plots[[site]],  # The plot to save
         path = here("./04_plots"),
         width = 6, height = 6,  # Adjust the size of the saved plot
         units = "in",  # Size units (inches in this case)
         dpi = 300)  # Adjust the resolution
}
```







```{r Making the plot with all of the fetch plots on the map}
fetch_plots <- list()

# Example for generating fetch plots for each site
for (site in unique(fetch_2024_long$site_code)) {
  print(paste("Processing site:", site))
  
  site_data <- fetch_2024_long%>%
    filter(site_code == site)
  
 fetch_plot <- ggplot(site_data, 
                      aes(x = as.numeric(degree), y = fetch)) +
  geom_bar(stat = "identity", width = 0.9, position = "identity") +
  coord_polar(start = 0) +  # Convert to polar coordinates
  scale_x_continuous(limits = c(0, 360.5), breaks = seq(45, 360, by = 45)) +  # Optional: Compass degree labels
  #scale_fill_viridis_c() +
  labs(title = paste("Fetch for", site)) +
  theme_minimal()
  
  # Store the plot in the list
  fetch_plots[[site]] <- fetch_plot
}

# Overlay each fetch plot on the main map
for (site in names(fetch_plots)) {
      print(fetch_plots[[site]]) 
  longitude <- fetch_2024_long$longitude[fetch_2024_long$site_code == site][1]  # Example to get the longitude
  latitude <- fetch_2024_long$latitude[fetch_2024_long$site_code == site][1]    # Example to get the latitude
  
  map_plot <- map_plot + 
    annotation_custom(
      ggplotGrob(fetch_plots[[site]]),
      xmin = longitude - 0.01,  # Adjust as needed for spacing
      xmax = longitude + 0.01,
      ymin = latitude - 0.01,
      ymax = latitude + 0.01
    )
}
```

```{r, fig.width=20, fig.height=20}
# Finalize and display the combined plot
print(map_plot)
```



##2024 urch sn ratio plot

```{r Making the plot with the urch to sn ratio on the map 2024}
(urch_sn_ratio_map_2024 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_sn_ratio_RLS_2024,
             mapping = aes(x = longitude, y = latitude, color = urch_sn_ratio),
             size = 3,) +
    geom_point(data = urch_sn_ratio_RLS_2024,
             mapping = aes(x = longitude, y = latitude, colour = urch_sn_ratio), 
              shape = 1, size = 3, colour = "black") + 
  #scale_fill_viridis_c() +
  scale_colour_gradient2(
    low = "blue",          # Color for negative values
    mid = "white",         # Color for zero
    high = "red",          # Color for positive values
    midpoint = 0           # Set the midpoint to zero
  ) +
  labs(title = "2024", x = " \nLongitude", y = "Latitude\n ", colour = " Log Urchin to\n Snail Ratio") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
```

```{r Making the plot with the standardized urch to sn ratio on the map in 2024, fig.height= 12, fig.width= 10}
(stand_urch_sn_ratio_map_2024 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_jitter(data = urch_sn_ratio_RLS_2024%>%
               filter(year(survey_date) == "2024"), mapping = aes(x = longitude, y = latitude, fill = std_urch_sn_ratio), 
              shape = 21, size = 3, height = 0.005, width = 0.005) + 
  scale_fill_viridis_c() +
  labs(title = "2024 Jittered Points", x = " \nLongitude", y = "Latitude\n ", fill = "Log St. Urchin to Snail Ratio") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
```

```{r Making the plot of urchin to all other macrofauna ratio}
(urch_all_ratio_map_2024 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_point(data = urch_all_ratio_RLS_2024,
             mapping = aes(x = longitude, y = latitude, color = urch_all_ratio),
             size = 3,) +
    geom_point(data = urch_all_ratio_RLS_2024,
             mapping = aes(x = longitude, y = latitude, colour = urch_all_ratio), 
              shape = 1, size = 3, colour = "black") +
  #scale_colour_viridis_c() +
  scale_colour_gradient2(
    low = "blue",          # Color for negative values
   mid = "white",         # Color for zero
    high = "red",          # Color for positive values
    midpoint = 0           # Set the midpoint to zero
  ) +
  labs(title = "2024", x = " \nLongitude", y = "Latitude\n ",
       fill = " Log Urchin to\n Macrofauna Ratio") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
```


###Saving the 2024 urch sn ratio plot
```{r Saving the 2024 urch sn ratio plot, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/urch_sn_ratio_map_2024.png"), width = 8, height = 6, 
    units = "in", pointsize = 15, res = 600)

urch_sn_ratio_map_2024

dev.off()
```

```{r Saving the standardized 2024 urch sn ratio plot, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/stand_urch_sn_ratio_map_2024.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

stand_urch_sn_ratio_map_2024

dev.off()
```

##Urchin and snail counts plots 2024

###Urchin plots
```{r Making the plot with the urch counts on the map 2024, fig.height= 12, fig.width= 10}
(urch_count_map_2024 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_jitter(data = urch_sn_ratio_RLS_2024%>%
               filter(year(survey_date) == "2024"), mapping = aes(x = longitude, y = latitude, fill = Urchins), 
              shape = 21, size = 3, height = 0.005, width = 0.005) + 
  scale_fill_viridis_c() +
  labs(title = "2024 Jittered Points", x = " \nLongitude", y = "Latitude\n ", fill = " Urchin Counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
```


```{r Making the plot with the standardized urch counts on the map 2024, fig.height= 12, fig.width= 10}
(stand_urch_count_map_2024 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_jitter(data = urch_sn_ratio_RLS_2024%>%
               filter(year(survey_date) == "2024"), mapping = aes(x = longitude, y = latitude, fill = stand_urchins), 
              shape = 21, size = 3, height = 0.005, width = 0.005) + 
  scale_fill_viridis_c() +
  labs(title = "2024 Jittered Points", x = " \nLongitude", y = "Latitude\n ", fill = " Standardized Urchin Counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
```


####Saving the urchin counts plots
```{r Saving the urchin counts plot, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/urch_count_map_2024.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

urch_count_map_2024

dev.off()
```

```{r Saving the standardized urchin counts plot, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/stand_urch_count_map_2024.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

stand_urch_count_map_2024

dev.off()
```

###Snail plots

```{r Making the plot with the snail counts on the map 2024, fig.height= 12, fig.width= 10}
(snail_count_map_2024 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_jitter(data = urch_sn_ratio_RLS_2024%>%
               filter(year(survey_date) == "2024"), mapping = aes(x = longitude, y = latitude, fill = Turban_Snails), 
              shape = 21, size = 3, height = 0.005, width = 0.005) + 
  scale_fill_viridis_c() +
  labs(title = "2024 Jittered Points", x = " \nLongitude", y = "Latitude\n ", fill = " Turban Snail Counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
```


```{r Making the plot with the standardized snail counts on the map 2024, fig.height= 12, fig.width= 10}
(stand_snail_count_map_2024 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_jitter(data = urch_sn_ratio_RLS_2024%>%
               filter(year(survey_date) == "2024"), mapping = aes(x = longitude, y = latitude, fill = stand_snails), 
              shape = 21, size = 3, height = 0.005, width = 0.005) + 
  scale_fill_viridis_c() +
  labs(title = "2024 Jittered Points", x = " \nLongitude", y = "Latitude\n ", fill = " Standardized Turban Snail Counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
```

####Saving the snail counts plot
```{r Saving the snail counts plot, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/snail_count_map_2024.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

snail_count_map_2024

dev.off()
```

```{r Saving the standardized snail counts plot, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/stand_snail_count_map_2024.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

stand_snail_count_map_2024

dev.off()
```



##Making plots with the gobies

###Making a dataframe with the gobies
```{r}
goby_urch_sn_counts_2024 <- RLS_data_2024_clean%>%
  pivot_wider(names_from = Code, values_from = species_total)%>%
  mutate(Urchins = mfra,
         Turban_Snails = pgi,
         Goby = rni)%>%
  select(site_code, site_name, latitude, longitude, survey_date, depth, 
           visibility, hour, Urchins, Turban_Snails, Goby)%>%
  replace(is.na(.), 0)%>%
   mutate(sn_goby_ratio = log10((Turban_Snails +1) / (Goby + 1)))
```


###Goby Counts plot

```{r Making the plot with the goby counts on the map 2024, fig.height= 12, fig.width= 10}
(goby_count_map_2024 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_jitter(data = goby_urch_sn_counts_2024, mapping = aes(x = longitude, y = latitude, fill = Goby), 
              shape = 21, size = 3, height = 0.005, width = 0.005) + 
  scale_fill_viridis_c() +
  labs(title = "2024", x = " \nLongitude", y = "Latitude\n ", fill = " Blackeye Goby Counts") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
```

###Plotting sn to goby ratio on map 2024
```{r}
(sn_goby_ratio_map_2024 <- ggplot()+
    geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  geom_jitter(data = goby_urch_sn_counts_2024, mapping = aes(x = longitude, y = latitude, fill = sn_goby_ratio), 
              shape = 21, size = 3, height = 0.005, width = 0.005) + 
  scale_fill_viridis_c() +
  labs(title = "2024", x = " \nLongitude", y = "Latitude\n ", fill = " Log Snails to Goby Ratio") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.4, -125.3, -125.2, -125.1, -125.0)) +
  scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br"))
```



###Plotting goby counts vs snail counts
```{r Goby counts against snail counts}
(goby_vs_snail_counts_2024 <- ggplot(goby_urch_sn_counts_2024, aes(x = Goby, y = Turban_Snails))+
   geom_point()+
   theme.figure())
```




##Working with temperature data

###Loading in the temperature data
```{r Loading in the temp .nc file downloaded from Copernicus}
temp_2023_2024 <- brick(here("./01_raw_data/METOFFICE-GLO-SST-L4-NRT-OBS-SST-V2_1730223986748.nc"))
crs(temp_2023_2024) <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
```

```{r Calculating the average temperature for each pixel}
average_temperature <- calc(temp_2023_2024, fun = mean, na.rm = TRUE)

writeRaster(average_temperature, 
            here("./01_raw_data/average_temperature.nc"), 
            format = "CDF", 
            overwrite = TRUE)
```

```{r Checking output}
avg_temp_check <- brick(here("./01_raw_data/average_temperature.nc"))
plot(avg_temp_check) 
```


###Plotting the temp data onto the barkley sound map
```{r Converting the raster to a data frame so ggplot can use it}
avg_temp_df <- as.data.frame(average_temperature, xy = TRUE, na.rm = TRUE)
colnames(avg_temp_df) <- c("longitude", "latitude", "average_temp")

avg_temp_df <- avg_temp_df%>%
  mutate(avg_temp_c = average_temp - 273.15)
```

```{r Making the combined plot}
(avg_temp_map_2023_24 <- ggplot() +
  geom_raster(data = avg_temp_df, aes(x = longitude, y = latitude, fill = avg_temp_c)) +
   geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max)) +
  scale_fill_viridis_c(option = "C", name = "Avg Temperature (°C)") +  # Optional color scale
  theme.figure() +
  labs(title = "Sea Surface Average Temperature",
       x = "Longitude",
       y = "Latitude"))  # This ensures correct projection
```


####Saving the temp map plot
```{r Saving the 2023-24 average temp plot, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/avg_temp_map_2023_24.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

avg_temp_map_2023_24

dev.off()
```


##Working with the global fishing watch data
```{r Loading in the global fishing watch data}
#gfw_data_raw <- read_csv(here("./01_raw_data/public-global-fishing-effort-v3.0.csv"))

gfw_data_clean <- gfw_data_raw%>%
  select(-c(`Time Range`))%>%
  rename(latitude = Lat,
         longitude = Lon,
         vessel_IDs = `Vessel IDs`,
         app_fish_hrs = `Apparent Fishing Hours`)%>%
  mutate(cell_lat_min = latitude - 0.005,
         cell_lat_max = latitude + 0.005,
         cell_long_min = longitude - 0.005,
         cell_long_max = longitude + 0.005)
  
```


```{r Converting the lat long min/maxes into polygons}
gfw_data_clean_poly <- gfw_data_clean %>%
  rowwise() %>%
  mutate(
    geometry = list(st_polygon(list(rbind(
      c(cell_long_min, cell_lat_min),
      c(cell_long_min, cell_lat_max),
      c(cell_long_max, cell_lat_max),
      c(cell_long_max, cell_lat_min),
      c(cell_long_min, cell_lat_min)
    ))))
  ) %>%
  ungroup() %>%
  st_as_sf(crs = 4326)
```


```{r Plotting the apparent fishing hour data on the map}
(app_fishing_hours_plot <- ggplot(gfw_data_clean_poly) +
  geom_sf(aes(fill = app_fish_hrs)) +
  scale_fill_gradient(low = "blue", high = "red")+
   geom_sf(data = VI_base_layer_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max))+
  theme.figure() +
  labs(fill = "Fishing Hours",
       x = "Longitude",
       y = "Latitude"))
```



```{r Saving the fishing hours plot, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/app_fishing_hours_plot.png"), width = 7.5, height = 5.5, 
    units = "in", pointsize = 15, res = 600)

app_fishing_hours_plot

dev.off()
```

