---
title: "Report Figures"
author: "Alec"
date: "2024-12-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(sf)
library(ggspatial)
library(ggrepel)
library(cowplot)
library(sp)
library(raster)
library(rphylopic)
library(patchwork)
```

#Loading in the data

##True unshifted data

```{r Loading in the data}
plotting_data <- read_csv(here("./03_clean_data/urch_sn_ratio_mod_data.csv"))
```

###Shifting a couple of the sites to make it clearer

```{r Shifting some sites}
plotting_data_minor_shifted <- plotting_data%>%
    #mutate(latitude = if_else(site_code == "BMSC11", 48.85828, latitude))%>%
 mutate(longitude = if_else(site_code == "BMSC11", -125.153, longitude))%>%
  mutate(latitude = if_else(site_code == "BMSC5", 48.82333, latitude))%>%
  mutate(latitude = if_else(site_code == "BMSC9", 48.83179, latitude))%>%
  mutate(longitude = if_else(site_code == "BMSC9", -125.1490, longitude))
  
  
  #adding protected status
  protect_status <- c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 
                    0, 0, 0, 1, 1, 0, 0, 0, 0, 0, 1, 1, 1, 1, 0)

  plotting_data_minor_shifted$protect_status <- as.factor(protect_status)
levels(plotting_data_minor_shifted$protect_status) <- c("No Protection", "Protected")
```


##Shifted data for better visualization

```{r}
plotting_data_shifted <- read_csv(here("./03_clean_data/plotting_data_shifted.csv"))
```

### Loading in the temperature data
```{r Loading in the temp .nc file downloaded from Copernicus}
temp_2023_2024 <- brick(here("./01_raw_data/METOFFICE-GLO-SST-L4-NRT-OBS-SST-V2_1730223986748.nc"))
crs(temp_2023_2024) <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
```

```{r Calculating the average temperature for each pixel}
average_temperature <- calc(temp_2023_2024, fun = mean, na.rm = TRUE)

writeRaster(average_temperature, 
            here("./01_raw_data/average_temperature.nc"), 
            format = "CDF", 
            overwrite = TRUE)
```

### Making a theme for all subsequent figures

```{r Making theme for all subsequent figures, echo=FALSE}
theme.figure.report <- function (){
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    #axis.line = element_line(colour = "black"),
    axis.title = element_text(size = 20, face = "bold"),
    axis.text = element_text(size = 18, face = "plain"),
    legend.text = element_text(size = 14, face = "plain"),
    legend.title = element_text(size = 16, face = "bold")) 
}

```


## Site maps

### Making the map ranges

```{r Seting the ranges for the maps}
#first need to specify the range of the map
lat_min = min(plotting_data$latitude) - 0.05
lat_max = max(plotting_data$latitude) + 0.05

long_min = min(plotting_data$longitude) - 0.05
long_max = max(plotting_data$longitude) + 0.05
```

### Loading in the shape file

```{r Loading in the new high res shape file}
VI_base_layer_hi_res <- read_sf(dsn = here("./01_raw_data/hakai_zip"),
                                stringsAsFactors = F)

VI_base_layer_hi_res_sf <- st_as_sf(VI_base_layer_hi_res)

#specifying the projection to use for the sf
st_crs(VI_base_layer_hi_res_sf) <- "+proj=longlat +datum=WGS84 +no_defs"
```

```{r Loading in the base layer}
VI_base_layer <- read_sf(dsn = here("./01_raw_data/canada_sf_data", ""), 
                         stringsAsFactors = F)
VI_base_layer_sf <- st_as_sf(VI_base_layer)

#specifying the projection for the sf
st_crs(VI_base_layer_sf) <- "+proj=longlat +datum=WGS84 +no_defs"
```

### Loading in the shapefiles for the protected areas

```{r loading in the RCA polygon}
broken_group_RCA <- st_read(here("./01_raw_data/Broken_group_RCA.kml"))
```

```{r Loading in the Numukamis area}
numukamis_area <- st_read(here("./01_raw_data/Numukamis_closed_area.kml"))
```

```{r Loading in the Baeria rocks}
baeria_rocks_er <- st_read(here("./01_raw_data/baeria_rocks.kml"))
```


### Overall site map

```{r Overall Barkley Sound Map}
(overall_Barkley_sound_map <- ggplot()+
    geom_sf(data = VI_base_layer_hi_res_sf, colour = "darkgrey", fill = "lightgrey")+
      geom_point(plotting_data, 
              mapping = aes(x = longitude, y = latitude),
              shape = 21, size = 2, fill = "red")+
   coord_sf(xlim = c(long_min - 0.1, long_max),
           ylim = c(lat_min - 0.05, lat_max)) +
  labs(x = " \nLongitude", y = "Latitude\n ") +
  theme.figure.report() +
 scale_x_continuous(breaks = c(-125.5, -125.25, -125.0)) +
  scale_y_continuous(breaks = c(48.75, 48.85, 48.95))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br")+
      geom_point( 
             mapping = aes(x = -125.134813, y = 48.83), shape = 23,
                           fill = "yellow", size = 3))
```

```{r Making the inset map}
(van_isl_map <- ggplot(VI_base_layer_sf) +
  geom_sf() +
  coord_sf(ylim = c(48.3, 50.8), xlim = c(-128.2, -123.2)) +
  theme.figure.report() +
  geom_rect(xmin = long_min, xmax = long_max, ymin = lat_min, ymax = lat_max,
            color = "red", fill = NA, linewidth = 1) +
  labs(x = "Longitude", y = "Latitude") +
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_blank()))
```

####Saving the van isl map

```{r Saving the van isl map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/report_figures/van_isl_map.png"), width = 4, height = 3, 
    units = "in", pointsize = 15, res = 600)

van_isl_map

dev.off()
```

```{r basemap with inset, fig.width=7.5, fig.height=5.5}
(map_w_inset <- ggdraw(overall_Barkley_sound_map) +
  draw_plot(van_isl_map, x = 0.125, y = 0.138, width = 0.5, height = 0.3))
```

####Saving the overall site map

```{r Saving the overall site map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/report_figures/overall_Barkley_sound_map.png"), width = 8, height = 6, 
    units = "in", pointsize = 15, res = 600)

map_w_inset

dev.off()
```


### Site map with points

```{r Overall site map with the site locations}
(overall_site_map <- ggplot()+
    geom_sf(data = VI_base_layer_hi_res_sf, colour = "darkgrey", fill = "lightgrey")+
     geom_sf(data = broken_group_RCA, fill = "blue", colour ="blue", alpha = 0.1)+
   geom_sf(data = numukamis_area, fill = "blue", colour = "blue", alpha = 0.2)+
   geom_sf(data = baeria_rocks_er, fill = "blue", colour = "blue", alpha = 0.2)+
     geom_point(plotting_data_minor_shifted, 
              mapping = aes(x = longitude, y = latitude, fill = protect_status,
                            shape = protect_status), size = 3)+
   scale_shape_manual(values = c(24, 21))+
   scale_fill_manual(values = c("#F46864", "#1CB8BC"))+
       coord_sf(xlim = c(long_min + 0.03, long_max - 0.035),
           ylim = c(lat_min + 0.01, lat_max - 0.04)) +
  labs(x = "Longitude", y = "Latitude", shape = "Protection\nStatus",
       fill = "Protection\nStatus") +
  theme.figure.report() +
   theme(legend.position = "none",
         axis.text.y = element_text(angle = 45))+
  scale_x_continuous(breaks = c(-125.35, -125.20, -125.05)) +
 # scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br", text_cex = 1.3))
```


####Saving the site map

```{r Saving the overall site map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/report_figures/overall_site_map.png"), width = 6, height = 4.5, 
    units = "in", pointsize = 15, res = 600)

overall_site_map

dev.off()
```


### Site map with inset

```{r Making the site map with the van isl inset}
(site_map_w_inset <- ggdraw(overall_site_map) +
  draw_plot(van_isl_map, x = 0.13, y = 0.19, width = 0.5, height = 0.3))

#can alternatively make the width 0.4 and the height 0.25 and it wont cover any of the islands
```

####Saving the site map with inset

```{r Saving the overall site map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/report_figures/site_map_with_inset.png"), width = 6, height = 4.5, 
    units = "in", pointsize = 15, res = 600)

site_map_w_inset

dev.off()
```

## Urchin Ratio map

```{r Making the plot with the urch to sn ratio on the map 2024}
(urch_sn_ratio_map_figures <- ggplot()+
    geom_sf(data = VI_base_layer_hi_res_sf, colour = "darkgrey", fill = "lightgrey")+
    coord_sf(xlim = c(long_min + 0.04, long_max - 0.04),
           ylim = c(lat_min + 0.04, lat_max - 0.04)) +
  geom_point(data = plotting_data_shifted,
             mapping = aes(x = longitude, y = latitude, color = urch_sn_ratio),
             size = 3,) +
    geom_point(data = plotting_data_shifted,
             mapping = aes(x = longitude, y = latitude, colour = urch_sn_ratio), 
              shape = 1, size = 3, colour = "black") + 
  scale_colour_gradient2(
    low = "blue",          # Color for negative values
    mid = "white",         # Color for zero
    high = "red",          # Color for positive values
    midpoint = 0           # Set the midpoint to zero
  ) +
  labs(x = " \nLongitude", y = "Latitude\n ", colour = " Log Urchin to\n Snail Ratio") +
  theme.figure() +
  scale_x_continuous(breaks = c(-125.35, -125.20, -125.05)) +
 # scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br")+
     geom_point( 
             mapping = aes(x = -125.134813, y = 48.83), shape = 23,
                           fill = "yellow", size = 4))
```

####Saving the site map

```{r Saving the overall site map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/report_figures/urch_sn_ratio_map.png"), width = 7.5, height = 4, 
    units = "in", pointsize = 15, res = 600)

urch_sn_ratio_map_figures

dev.off()
```

## Urchin Counts

```{r Making a non-linear colour palette to effectively show the urchin counts}
colour_breaks_urchins <- c(0, 100, 250, 350, 529, 709)
openness_colours_urchins <- c("#440154", "#414487", "#2a788e", "#22a884", "#7ad151", "#fde725")
```

```{r Urchin Counts plot}
(urchin_counts_map_figs <- ggplot()+
    geom_sf(data = VI_base_layer_hi_res_sf, colour = "darkgrey", fill = "lightgrey")+
     geom_point(plotting_data_shifted, 
              mapping = aes(x = longitude, y = latitude, fill = Urchins),
              shape = 21, size = 3, colour = "black")+
       coord_sf(xlim = c(long_min + 0.03, long_max - 0.035),
           ylim = c(lat_min + 0.01, lat_max - 0.04)) +
  labs(x = " \nLongitude", y = "Latitude\n ",
       fill = "Urchin\nCounts") +
  theme.figure() +
   scale_fill_gradientn(colours = openness_colours_urchins,
                        values = scales::rescale(colour_breaks_urchins),
                        limits = c(0, 709))+
  # scale_fill_gradient(low = "blue", high = "red")+
  scale_x_continuous(breaks = c(-125.35, -125.20, -125.05)) +
 # scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br")+
      geom_point( 
             mapping = aes(x = -125.134813, y = 48.83), shape = 23,
                           fill = "yellow", size = 4))
```

```{r Saving the urchin counts map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/report_figures/urchin_counts_map.png"), width = 6.5, height = 4, 
    units = "in", pointsize = 15, res = 600)

urchin_counts_map_figs

dev.off()
```

## Snail Counts

```{r Making a non-linear colour palette to effectively show the snail counts}
colour_breaks_snails <- c(0, 50, 100, 200, 300, 645)
openness_colours_snails <- c("#440154", "#414487", "#2a788e", "#22a884", "#7ad151", "#fde725")
```

```{r Snail Counts plot}
(snail_counts_map_figs <- ggplot()+
    geom_sf(data = VI_base_layer_hi_res_sf, colour = "darkgrey", fill = "lightgrey")+
     geom_point(plotting_data_shifted, 
              mapping = aes(x = longitude, y = latitude, fill = Turban_Snails),
              shape = 21, size = 3, colour = "black")+
       coord_sf(xlim = c(long_min + 0.03, long_max - 0.035),
           ylim = c(lat_min + 0.01, lat_max - 0.04)) +
  labs(x = " \nLongitude", y = "Latitude\n ",
       fill = "Turban Snail\nCounts") +
  theme.figure() +
   scale_fill_gradientn(colours = openness_colours_snails,
                        values = scales::rescale(colour_breaks_snails),
                        limits = c(0, 645))+
   #scale_fill_gradient(low = "blue", high = "red")+
  scale_x_continuous(breaks = c(-125.35, -125.20, -125.05)) +
 # scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br")+
      geom_point( 
             mapping = aes(x = -125.134813, y = 48.83), shape = 23,
                           fill = "yellow", size = 4))
```

```{r Saving the snail counts map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/report_figures/snail_counts_map.png"), width = 7, height = 4, 
    units = "in", pointsize = 15, res = 600)

snail_counts_map_figs

dev.off()
```


## Site openness map

```{r Making a non-linear colour palette to effectively show the normalized openness}
colour_breaks <- c(0, 0.1, 0.2, 0.5, 1)
openness_colours <- c("#440154", "#3b528b", "#21918c", "#5ec962", "#fde725")
```


```{r Openness values at each site map}
(site_openness_map <- ggplot()+
    geom_sf(data = VI_base_layer_hi_res_sf, colour = "darkgrey", fill = "lightgrey")+
     geom_point(plotting_data_minor_shifted, 
              mapping = aes(x = longitude, y = latitude, fill = norm_openness),
              shape = 21, size = 3, colour = "black")+
       coord_sf(xlim = c(long_min + 0.03, long_max - 0.035),
           ylim = c(lat_min + 0.01, lat_max - 0.04)) +
  labs(x = "Longitude", y = "Latitude",
       fill = "Normalized\nOpenness") +
  theme.figure.report() +
   theme(axis.text.y = element_text(angle = 45))+
        # legend.position = "none")+
   scale_fill_gradientn(colours = openness_colours,
                        values = scales::rescale(colour_breaks),
                        limits = c(0, 1))+
  scale_x_continuous(breaks = c(-125.35, -125.20, -125.05)) +
 # scale_y_continuous(breaks = c(48.80, 48.85, 48.90, 48.95, 49.00))+
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br", text_cex = 1.3))
```

#### Saving the openness map

```{r Saving the openness map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/report_figures/site_openness_map.png"), width = 6, height = 4.5, 
    units = "in", pointsize = 15, res = 600)

site_openness_map

dev.off()
```



## Temperature map

### Plotting the temp data onto the barkley sound map
```{r Converting the raster to a data frame so ggplot can use it}
avg_temp_df <- as.data.frame(average_temperature, xy = TRUE, na.rm = TRUE)
colnames(avg_temp_df) <- c("longitude", "latitude", "average_temp")

avg_temp_df <- avg_temp_df%>%
  mutate(avg_temp_c = average_temp - 273.15)
```

```{r Making the temperature map}
(temp_map_23_24 <- ggplot() +
  geom_raster(data = avg_temp_df, aes(x = longitude, y = latitude, fill = avg_temp_c)) +
   geom_sf(data = VI_base_layer_hi_res_sf, colour = "#424A4E", fill = "lightgrey")+
  coord_sf(xlim = c(long_min + 0.029, long_max - 0.032),
           ylim = c(lat_min , lat_max - 0.025)) +
     scale_x_continuous(breaks = c(-125.35, -125.20, -125.05)) +
  #scale_fill_viridis_c(option = "C", name = "Avg Temperature (°C)") +  # Optional color scale
     scale_fill_gradient(
    low = "lightblue",          
    high = "#E95E48"
  ) +
  theme.figure.report() +
  labs(x = "Longitude",
       y = "Latitude",
       fill = "Mean SST (ºC)") +
  theme(#legend.position = "none",
        axis.text.y = element_text(angle = 45))+
  annotation_scale(location = "br", text_cex = 1.3)+
   annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering))
```


#### Saving the temperature map

```{r Saving the temperature map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/report_figures/temperature_map_23_24.png"), width = 6, height = 4.5, 
    units = "in", pointsize = 15, res = 600)

temp_map_23_24

dev.off()
```



## Example Recreational boat plot

### Using 1.5 km radius

```{r total rec boats within 1.5 km}
(total_rec_boats_map_2023_test <- ggplot()+
    geom_sf(data = VI_base_layer_hi_res_sf, colour = "darkgrey", fill = "lightgrey")+
    geom_sf(data = site_locations_radius_1.5km, fill = "#E95E48", alpha = 0.3)+
 coord_sf(xlim = c(long_min + 0.03, long_max - 0.035),
           ylim = c(lat_min + 0.01, lat_max - 0.04)) +
   geom_point(plotting_data, 
              mapping = aes(x = longitude, y = latitude),
              shape = 21, size = 3, colour = "black")+
  geom_point(data = rec_boats_fishing2023, 
             mapping = aes(x = Longitude, y = Latitude))+
  labs(x = "Longitude", y = "Latitude") +
  theme.figure.report() +
   theme(axis.text.y = element_text(angle = 45))+
 scale_x_continuous(breaks = c(-125.35, -125.20, -125.05)) +
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br", text_cex = 1.3))
```

#### Saving the first rec map for the poster

```{r Saving the temperature map, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/report_figures/total_rec_boat_map_2023_with_site_circles.png"), width = 6, height = 4.5, 
    units = "in", pointsize = 15, res = 600)

total_rec_boats_map_2023_test

dev.off()
```

### Total rec boats

```{r total rec boat }
(total_rec_boats_map_2023 <- ggplot()+
    geom_sf(data = VI_base_layer_hi_res_sf, colour = "darkgrey", fill = "lightgrey")+
      coord_sf(xlim = c(long_min, long_max),
           ylim = c(lat_min, lat_max))+
   
 coord_sf(xlim = c(long_min + 0.03, long_max - 0.035),
           ylim = c(lat_min + 0.01, lat_max - 0.04)) +
  geom_point(data = rec_boats_fishing2023, 
             mapping = aes(x = Longitude, y = Latitude))+
  labs(x = "Longitude", y = "Latitude") +
  theme.figure.report() +
   theme(axis.text.y = element_text(angle = 45))+
 scale_x_continuous(breaks = c(-125.35, -125.20, -125.05)) +
  annotation_north_arrow(
    location = "tr",
    which_north = "true",
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(location = "br", text_cex = 1.3) +
 geom_rect(aes(xmin = (long_min + 0.085), xmax = (long_max - 0.26), ymin = (lat_min + 0.086), ymax = (lat_max - 0.084)),
            color = "#E95E48", fill = NA, linewidth = 2))
```

### Broken group rec baots

```{r Inset view of smaller area with the radii around the sites}
(broken_group_rec_boats <- ggplot()+
    geom_sf(data = VI_base_layer_hi_res_sf, colour = "darkgrey", fill = "lightgrey")+
  geom_sf(data = site_locations_radius_1.5km, fill = "#E95E48", alpha = 0.3)+
      coord_sf(xlim = c(long_min + 0.085, long_max - 0.26),
           ylim = c(lat_min + 0.086, lat_max - 0.084))+
      geom_point(plotting_data, 
              mapping = aes(x = longitude, y = latitude),
              shape = 21, size = 3, colour = "black")+
  geom_point(data = rec_boats_fishing2023, 
             mapping = aes(x = Longitude, y = Latitude),
             size = 3)+
  labs(x = "Longitude", y = "Latitude") +
  theme.figure.report() +
   theme(axis.text = element_blank(), 
         axis.title = element_blank(),
         axis.ticks = element_blank()))
 # annotation_north_arrow(
  #  location = "tr",
   # which_north = "true",
    #style = north_arrow_fancy_orienteering
  #) +
 # annotation_scale(location = "br", text_cex = 1.3))
```

### Combining the maps into an inset

```{r Making the rec fishing boats map with inset}
(rec_boat_map_2023_with_inset <- ggdraw(total_rec_boats_map_2023) +
  draw_plot(broken_group_rec_boats, x = 0.2, y = 0.19, width = 0.3, height = 0.5))
```


## Plot of mean recreational boat counts between protection levels

```{r Boxplot of mean rec boats between protection levels}
(rec_boats_boxplot_1.5km <- ggplot(unique_site_df, aes(x = protect_status,
                                                     y = mean_rec_boat_counts_1.5km,
                                                     fill = protect_status))+
  geom_boxplot(alpha = 0.6, width = 0.15)+
  geom_jitter(height = 0, width = 0.1, colour = "darkgrey")+
  labs(x = "Protection Status", y = "Mean rec. boat counts")+
     scale_fill_manual(values = c("#E95E48", "#355173"))+
  theme.figure.report()+
   theme(legend.position = "none",
         axis.text.x = element_blank(),
         axis.title.x = element_blank(),
         axis.ticks.x = element_blank()))
```


#### Saving the rec boats boxplot

```{r Saving the rec boats boxplot, fig.width=7.5, fig.height=5.5, include=FALSE}
png(filename = here("./04_plots/report_figures/rec_boats_boxplot_1.5km.png"), width = 3.5, height = 3.5, 
    units = "in", pointsize = 15, res = 600)

rec_boats_boxplot_1.5km

dev.off()
```


### Plot of effect of openness on snails

#### Non-log plot

```{r Getting snail silhouette with phylopic}
#Run this function and then type into the console which one you want then copy paste the uuid into the next part
#pick_phylopic(name = "Trochidae")
#The uuid for the snail silhouette is "af3b0a39-4ba9-4134-800e-d181a7288689"

snail_picture_uuid <- "af3b0a39-4ba9-4134-800e-d181a7288689"
snail_picture_img <- get_phylopic(uuid = snail_picture_uuid)
#snail_picture_red <- recolor_phylopic(img = snail_picture_img, fill = "#E95E48")
```

```{r Making a plot of the effect of openness on snails}
(pred_snail_openness_plot <- ggplot(env_mod_data, aes(x = norm_openness, y = Turban_Snails))+
  geom_point()+
  geom_line(data = snail_op_pred_data, aes(
                                           y = exp(predicted_snails) - 1))+
  geom_ribbon(data = snail_op_pred_data,
              aes(ymin = exp(predicted_snails - 1.96 * 0.2691) - 1,
                  ymax = exp(predicted_snails + 1.96 * 0.2691) - 1),
              alpha = 0.5, fill = "#8EBAD2")+
  add_phylopic(img = snail_picture_img, x = 0.9, y = 550, height = 150)+
  labs(x = "Normalized Openness", y = "Red turban snail counts")+
  theme.figure.report())
```

##### Saving the snail openness plot

```{r Saving the snail openness plot}
png(filename = here("./04_plots/report_figures/pred_snail_openness_plot.png"), width = 6, height = 4.5, 
    units = "in", pointsize = 15, res = 600)

pred_snail_openness_plot

dev.off()
```


#### Log plot

```{r Making a plot of the effect of openness on log snails}
(pred_snail_openness_plot_log <- ggplot(env_mod_data, aes(x = norm_openness, y = log(Turban_Snails + 1)))+
  geom_point()+
  geom_smooth(method = "lm", colour = "#637683", fill = "#8EBAD2")+
  add_phylopic(img = snail_picture_img, x = 0.9, y = 5.5, height = 2)+
  labs(x = "Normalized Openness", y = "Log red turban snail counts")+
  theme.figure.report())
```

##### Saving the snail openness plot

```{r Saving the snail openness plot}
png(filename = here("./04_plots/report_figures/pred_snail_openness_plot_log.png"), width = 6, height = 4.6, 
    units = "in", pointsize = 15, res = 600)

pred_snail_openness_plot_log

dev.off()
```

### Boxplot of snails between protection status

```{r Making a boxplot of snail counts between protection status}
(snail_fishing_boxplot <- ggplot(env_mod_data, aes(x = protect_status, y = Turban_Snails))+
  geom_boxplot(env_mod_data, mapping = aes(fill = protect_status))+
  geom_jitter(height = 0, width = 0.1, colour = "darkgrey")+
  labs(x = "Protection Status", y = "Red turban snail counts")+
   scale_fill_manual(values = c("#E95E48", "#355173"))+
  add_phylopic(img = snail_picture_img, x = 2.4, y = 580, height = 150)+
  theme.figure.report()+
    theme(legend.position = "none"))
```

### Boxplot of urchins between protection status

```{r Getting urchin silhouette with phylopic}
#urchin_picture_img <- pick_phylopic(name = "Strongylocentrotus purpuratus", n = 4, view = 4)
#The uuid for the image i am using is "a6125e0d-c627-40e3-9966-5ad288277545"

urchin_picture_uuid <- "a6125e0d-c627-40e3-9966-5ad288277545"
urchin_picture_img <- get_phylopic(uuid = urchin_picture_uuid)
#urchin_picture_img_red <- recolor_phylopic(img = urchin_picture_img, alpha = 0.5, fill = "#E95E48")
```


```{r Making a boxplot of urchin counts between protection status}
(urchin_fishing_boxplot <- ggplot(env_mod_data, aes(x = protect_status, y = Urchins))+
  geom_boxplot(env_mod_data, mapping = aes(fill = protect_status),
               alpha = 0.6, width = 0.15, position = position_nudge(x = -0.1))+
  geom_errorbar(pred_urchin_fishing_mod_glmm, mapping =
                  aes(ymin = conf.low, ymax = conf.high, x = x, y = predicted),
                width = 0.15, position = position_nudge(0.1))+
   geom_point(pred_urchin_fishing_mod_glmm, mapping = 
              aes(x = x, y = predicted, colour = x), size = 4, position = position_nudge(0.1))+
  labs(x = "Protection Status", y = "Red sea urchin counts")+
   scale_fill_manual(values = c("#E95E48", "#355173"))+
   scale_colour_manual(values = c("#E95E48", "#355173"))+
  add_phylopic(img = urchin_picture_img, x = 2.3, y = 620, height = 150)+
  theme.figure.report()+
    theme(legend.position = "none",
          axis.text = element_text(size = 16),
          axis.text.x = element_blank(),
         axis.title.x = element_blank(),
         axis.ticks.x = element_blank()))
```


```{r Saving the urchin fishing boxplot}
png(filename = here("./04_plots/report_figures/urchin_fishing_boxplot.png"), width = 3.5, height = 3.5, 
    units = "in", pointsize = 15, res = 600)

urchin_fishing_boxplot

dev.off()
```

### Urchin fishing boxplots with the model outputs (with fixed effect of region)

```{r Making a plot of the urchin counts based on protection and region}
(urchin_fishing_boxplot_fe <- ggplot(env_mod_data, aes(x = region1,
                                                           y = Urchins
                                                           ))+
    geom_boxplot(env_mod_data%>%
                   filter(protect_status == "Protected"), 
                 mapping = aes(x = region1, y = Urchins) ,
              alpha = 0.6, width = 0.15, fill = "#355173", position = position_nudge(x = 0.15))+
   geom_boxplot(env_mod_data%>%
                 filter(protect_status == "No Protection"), 
               mapping = aes(x = region1, y = Urchins) ,
              alpha = 0.6, width = 0.15, fill = "#E95E48", position = position_nudge(x = -0.35))+
 
  geom_errorbar(urchin_counts_non_protected_preds, mapping =
                  aes(ymin = conf.low, ymax = conf.high, x = group, y = predicted),
                 width = 0.15, position = position_nudge(-0.15))+
    geom_errorbar(urchin_counts_protected_preds, mapping =
                  aes(ymin = conf.low, ymax = conf.high, x = group, y = predicted),
                 width = 0.15, position = position_nudge(0.35))+
     geom_point(urchin_counts_non_protected_preds, mapping = 
              aes(x = group, y = predicted), size = 4, alpha = 1,
             colour = "#E95E48",
             position = position_nudge(-0.15))+
    geom_point(urchin_counts_protected_preds, mapping = 
              aes(x = group, y = predicted), size = 4, alpha = 1,
             colour = "#355173",
             position = position_nudge(0.35))+
  #geom_boxplot(alpha = 0.6, width = 0.15)+
 # geom_jitter(height = 0, width = 0.1, colour = "darkgrey")+
  labs(x = "Region", y = "Red sea urchin counts")+
     scale_fill_manual(values = c("#E95E48", "#355173"))+
        scale_colour_manual(values = c("#E95E48", "#355173"))+
  add_phylopic(img = urchin_picture_img, x = 4.3, y = 680, height = 120)+
  theme.figure.report()+
   theme(legend.position = "none"))
```

```{r Saving the urchin fishing boxplots with model points fe}
png(filename = here("./04_plots/report_figures/urchin_fishing_boxplot_fe.png"), width = 8, height = 4, 
    units = "in", pointsize = 15, res = 600)

urchin_fishing_boxplot_fe

dev.off()
```


### Urchin fishing boxplots with the model outputs (with fixed effect of region) with facetting

```{r Making a plot of the urchin counts based on protection and region}
(urchin_fishing_boxplot_facet <- ggplot(plotting_pred_urchin_fishing_mod_glm_fe, aes(x = x,
                                                           y = predicted
                                                           ))+
  geom_errorbar(plotting_pred_urchin_fishing_mod_glm_fe, mapping =
                  aes(ymin = pmax(conf.low, 0), ymax = conf.high, x = x, y = predicted),
                 width = 0.15, position = position_nudge(-0.15))+
    #geom_errorbar(plotting_pred_urchin_fishing_mod_glm_fe, mapping =
     #             aes(ymin = pmax(conf.low, 0), ymax = conf.high, x = x, y = predicted),
      #           width = 0.15, position = position_nudge(0.35))+
     geom_point(plotting_pred_urchin_fishing_mod_glm_fe, mapping = 
              aes(x = x, y = predicted, colour = x), size = 4, alpha = 1,
            
             position = position_nudge(-0.15))+
 #   geom_point(plotting_pred_urchin_fishing_mod_glm_fe, mapping = 
  #            aes(x = x, y = predicted), size = 4, alpha = 1,
   #          colour = "#355173",
     #        position = position_nudge(0.35))+
   
       geom_boxplot(plotting_pred_urchin_fishing_mod_glm_fe, 
                 mapping = aes(x = x, y = Urchins, fill = x) ,
              alpha = 0.6, width = 0.15, position = position_nudge(x = 0.15))+
  # geom_boxplot(plotting_pred_urchin_fishing_mod_glm_fe, 
    #           mapping = aes(x = x, y = Urchins) ,
       #       alpha = 0.6, width = 0.15, fill = "#E95E48", position = position_nudge(x = -0.35))+
   
   
  #geom_boxplot(alpha = 0.6, width = 0.15)+
 # geom_jitter(height = 0, width = 0.1, colour = "darkgrey")+
  labs(x = "Region", y = "Red sea urchin counts")+
     scale_fill_manual(values = c("#E95E48", "#355173"))+
        scale_colour_manual(values = c("#E95E48", "#355173"))+
 # add_phylopic(img = urchin_picture_img, x = 4.3, y = 680, height = 120)+
  theme.figure.report()+
   theme(legend.position = "none",
         axis.text.x = element_blank(),
         axis.ticks.x = element_blank(),
         strip.text.x = element_text(size = 14))+
   facet_wrap(~region1, nrow = 1))
```

```{r Saving the urchin fishing boxplots with model points fe}
png(filename = here("./04_plots/report_figures/urchin_fishing_boxplot_facet.png"), width = 6, height = 4, 
    units = "in", pointsize = 15, res = 600)

urchin_fishing_boxplot_facet

dev.off()
```

### Urchin to snail ratio boxplot with just the original random effect of region
```{r}
(urchin_snail_ratio_box_re <- ggplot(env_mod_data, aes(x = protect_status, y = urch_sn_ratio))+
  geom_boxplot(env_mod_data, mapping = aes(fill = protect_status),
               alpha = 0.6, width = 0.15, position = position_nudge(x = -0.1))+
  geom_errorbar(pred_urch_sn_ratio_protection_model_glmm, mapping =
                  aes(ymin = conf.low, ymax = conf.high, x = x, y = predicted),
                width = 0.15, position = position_nudge(0.1))+
   geom_point(pred_urch_sn_ratio_protection_model_glmm, mapping = 
              aes(x = x, y = predicted, colour = x), size = 4, position = position_nudge(0.1))+
  labs(x = "Protection Status", y = "Log urchin:snail ratio")+
   scale_fill_manual(values = c("#E95E48", "#355173"))+
      scale_colour_manual(values = c("#E95E48", "#355173"))+
   add_phylopic(img = urchin_picture_img, x = 2.35, y = 2.1, height = 0.8)+
   add_phylopic(img = snail_picture_img, x = 2.35, y = 1.2, height = 0.6)+
  theme.figure.report()+
    theme(legend.position = "none",
          axis.text = element_text(size = 16),
           axis.text.x = element_blank(),
         axis.title.x = element_blank()
        ))
```

```{r Saving the urchin fishing boxplot}
png(filename = here("./04_plots/report_figures/urchin_snail_ratio_box_re.png"), width = 3.5, height = 3.5, 
    units = "in", pointsize = 15, res = 600)

urchin_snail_ratio_box_re

dev.off()
```

### Urchin snail ratio boxplot


```{r Making a boxplot of urchin counts between protection status}
(urch_sn_ratio_fishing_boxplot <- ggplot(env_mod_data, aes(x = region1,
                                                           y = urch_sn_ratio
                                                           ))+
    geom_boxplot(env_mod_data%>%
                   filter(protect_status == "Protected"), 
                 mapping = aes(x = region1, y = urch_sn_ratio) ,
              alpha = 0.6, width = 0.15, fill = "#355173", position = position_nudge(x = 0.2))+
   geom_boxplot(env_mod_data%>%
                 filter(protect_status == "No Protection"), 
               mapping = aes(x = region1, y = urch_sn_ratio) ,
              alpha = 0.6, width = 0.15, fill = "#E95E48", position = position_nudge(x = -0.4))+
 
  geom_errorbar(urch_sn_ratio_non_protected_preds, mapping =
                  aes(ymin = conf.low, ymax = conf.high, x = group, y = predicted),
                 width = 0.15, position = position_nudge(-0.2))+
    geom_errorbar(urch_sn_ratio_protected_preds, mapping =
                  aes(ymin = conf.low, ymax = conf.high, x = group, y = predicted),
                 width = 0.15, position = position_nudge(0.4))+
     geom_point(urch_sn_ratio_non_protected_preds, mapping = 
              aes(x = group, y = predicted), size = 4, alpha = 1,
             colour = "#E95E48",
             position = position_nudge(-0.2))+
    geom_point(urch_sn_ratio_protected_preds, mapping = 
              aes(x = group, y = predicted), size = 4, alpha = 1,
             colour = "#355173",
             position = position_nudge(0.4))+
  #geom_boxplot(alpha = 0.6, width = 0.15)+
 # geom_jitter(height = 0, width = 0.1, colour = "darkgrey")+
  labs(x = "Region", y = "Log urchin to snail ratio")+
     scale_fill_manual(values = c("#E95E48", "#355173"))+
        scale_colour_manual(values = c("#E95E48", "#355173"))+
  add_phylopic(img = urchin_picture_img, x = 4.4, y = 2.4, height = 0.5)+
   add_phylopic(img = snail_picture_img, x = 4.4, y = 1.9, height = 0.4)+
  theme.figure.report()+
   theme(legend.position = "none"))
```


```{r Saving the urchin to snail ratio fishing boxplot}
png(filename = here("./04_plots/report_figures/urch_sn_ratio_fishing_boxplot.png"), width = 8, height = 4.5, 
    units = "in", pointsize = 15, res = 600)

urch_sn_ratio_fishing_boxplot

dev.off()
```

### Trying with just the model output points

```{r Making a boxplot of urch sn ratio between protection status}
(urch_sn_ratio_fishing_model_points_only <- ggplot(env_mod_data, aes(x = region1,
                                                           y = urch_sn_ratio
                                                           ))+
 
  geom_errorbar(urch_sn_ratio_non_protected_preds, mapping =
                  aes(ymin = conf.low, ymax = conf.high, x = group, y = predicted),
                 width = 0.15, position = position_nudge(-0.1))+
    geom_errorbar(urch_sn_ratio_protected_preds, mapping =
                  aes(ymin = conf.low, ymax = conf.high, x = group, y = predicted),
                 width = 0.15, position = position_nudge(0.1))+
     geom_point(urch_sn_ratio_non_protected_preds, mapping = 
              aes(x = group, y = predicted), size = 4, alpha = 1,
             colour = "#E95E48",
             position = position_nudge(-0.1))+
    geom_point(urch_sn_ratio_protected_preds, mapping = 
              aes(x = group, y = predicted), size = 4, alpha = 1,
             colour = "#355173",
             position = position_nudge(0.1))+
  #geom_boxplot(alpha = 0.6, width = 0.15)+
 # geom_jitter(height = 0, width = 0.1, colour = "darkgrey")+
  labs(x = "Region", y = "Log urchin to snail ratio")+
     scale_fill_manual(values = c("#E95E48", "#355173"))+
        scale_colour_manual(values = c("#E95E48", "#355173"))+
 # add_phylopic(img = urchin_picture_img, x = 2.45, y = 2.2, height = 1)+
 #  add_phylopic(img = snail_picture_img, x = 2.45, y = 1.2, height = 0.8)+
  theme.figure.report()+
   theme(legend.position = "none"))
```

```{r Saving the urchin to snail ratio fishing model points plot}
png(filename = here("./04_plots/report_figures/urch_sn_ratio_fishing_model_points_only.png"), width = 4, height = 4, 
    units = "in", pointsize = 15, res = 600)

urch_sn_ratio_fishing_model_points_only

dev.off()
```

### Doing the urchin to snail ratio plots with facetting

```{r Making a plot of the urch sn ratio based on protection and region}
(urch_sn_ratio_boxplot_facet <- ggplot(plotting_pred_urch_sn_ratio_protection_model_glm_fe, aes(x = x,
                                                           y = predicted
                                                           ))+
  geom_errorbar(plotting_pred_urch_sn_ratio_protection_model_glm_fe, mapping =
                  aes(ymin = conf.low, ymax = conf.high, x = x, y = predicted),
                 width = 0.15, position = position_nudge(-0.15))+
    #geom_errorbar(plotting_pred_urchin_fishing_mod_glm_fe, mapping =
     #             aes(ymin = pmax(conf.low, 0), ymax = conf.high, x = x, y = predicted),
      #           width = 0.15, position = position_nudge(0.35))+
     geom_point(plotting_pred_urch_sn_ratio_protection_model_glm_fe, mapping = 
              aes(x = x, y = predicted, colour = x), size = 4, alpha = 1,
            
             position = position_nudge(-0.15))+
 #   geom_point(plotting_pred_urchin_fishing_mod_glm_fe, mapping = 
  #            aes(x = x, y = predicted), size = 4, alpha = 1,
   #          colour = "#355173",
     #        position = position_nudge(0.35))+
   
       geom_boxplot(plotting_pred_urch_sn_ratio_protection_model_glm_fe, 
                 mapping = aes(x = x, y = urch_sn_ratio, fill = x) ,
              alpha = 0.6, width = 0.15, position = position_nudge(x = 0.15))+
  # geom_boxplot(plotting_pred_urchin_fishing_mod_glm_fe, 
    #           mapping = aes(x = x, y = Urchins) ,
       #       alpha = 0.6, width = 0.15, fill = "#E95E48", position = position_nudge(x = -0.35))+
   
   
  #geom_boxplot(alpha = 0.6, width = 0.15)+
 # geom_jitter(height = 0, width = 0.1, colour = "darkgrey")+
  labs(x = "Region", y = "Log urchin to snail ratio")+
     scale_fill_manual(values = c("#E95E48", "#355173"))+
        scale_colour_manual(values = c("#E95E48", "#355173"))+
 # add_phylopic(img = urchin_picture_img, x = 4.3, y = 680, height = 120)+
  theme.figure.report()+
   theme(legend.position = "none",
         axis.text.x = element_blank(),
         axis.ticks.x = element_blank(),
         strip.text.x = element_text(size = 14))+
   facet_wrap(~region1, nrow = 1))
```
```{r Saving the urchin to snail ratio fishing model points plot}
png(filename = here("./04_plots/report_figures/urch_sn_ratio_boxplot_facet.png"), width = 6, height = 4, 
    units = "in", pointsize = 15, res = 600)

urch_sn_ratio_boxplot_facet

dev.off()
```



### Plotting the three plots for the poster in one ggplot

```{r Arranging the three plots in one ggplot}
(patchwork_plot <- (rec_boats_boxplot_1.5km | urchin_fishing_boxplot) /
   (patchwork::plot_spacer() | urchin_snail_ratio_box_re))


```


```{r Saving the patchwork plot}
png(filename = here("./04_plots/report_figures/patchwork_plot.png"), width = 7, height = 7, 
    units = "in", pointsize = 15, res = 1000)

patchwork_plot

dev.off()
```


### Plotting the effect of dominant substrate on the urchin counts

```{r Plotting dom substrate on urchin counts}
(dominant_substrate_boxplot <- ggplot(env_mod_data, aes(x = dominant_score, y = Urchins, fill = dominant_score))+
    geom_boxplot(alpha = 0.6, width = 0.15)+
  geom_jitter(height = 0, width = 0.1, colour = "darkgrey")+
  labs(x = "Dominant bottom type", y = "Red sea urchin counts")+
  theme.figure.report()+
     theme(legend.position = "none"))
```
```{r Saving the urchin counts dom score}
png(filename = here("./04_plots/report_figures/dominant_substrate_boxplot.png"), width = 4, height = 4, 
    units = "in", pointsize = 15, res = 600)

dominant_substrate_boxplot

dev.off()
```
