---
title: "RLS modelling results"
author: "Alec"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
   toc: true
---
   <script src="hideOutput.js"></script> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(mgcv)
library(DHARMa)
library(mgcViz)
library(ggeffects)
library(gratia)
library(geoR)
library(geosphere)
library(gstat)
library(sf)
library(car)
library(MuMIn)
library(lme4)
library(flextable)
library(spdep)
```

<!--Jan. 22. I have updated the data with the flights with fog ommitted from the data.
If this results in large problems the original data is available in the backup csv.
Except for the total boats columns but if these are needed then i can just reload the 
DFO data and not exclude the flights with fog.-->

# Modelling of red sea urchins, red turban snails, and their ratio in Barkley Sound {.tabset}

## Loading in the data and other setup

For using the data I have downloaded clean data from my preliminary explorations
that contains all of the exlanantory and response variables that I would like to
use. I have also made a separate dataset where I have averaged the values for 
urchin, snail counts etc. for transects occurring at the same sites. I have also 
made a theme for all subsequent figures

```{r Loading in the data, echo=FALSE}
env_mod_data <- read_csv(here("./03_clean_data/urch_sn_ratio_mod_data.csv"))
```

```{r Scaling the explanatory variables and converting substrate type to factor, include=FALSE}
env_mod_data$sst_scale <- scale(env_mod_data$sst)
env_mod_data$openness_scale <- scale(env_mod_data$openness)
env_mod_data$norm_openness_scale <- scale(env_mod_data$norm_openness)
env_mod_data$mean_app_fishing_hours_scale <- scale(env_mod_data$mean_app_fishing_hours)
env_mod_data$mean_rec_boat_counts_500m_scale <- scale(env_mod_data$mean_rec_boat_counts_500m)
env_mod_data$mean_rec_boat_counts_1km_scale <- scale(env_mod_data$mean_rec_boat_counts_1km)
env_mod_data$mean_rec_boat_counts_1.5km_scale <- scale(env_mod_data$mean_rec_boat_counts_1.5km)
env_mod_data$mean_rec_boat_counts_750m_scale <- scale(env_mod_data$mean_rec_boat_counts_750m)
env_mod_data$mean_rec_boat_counts_2km_scale <- scale(env_mod_data$mean_rec_boat_counts_2km)
env_mod_data$dominant_score <- as.factor(env_mod_data$dominant_score)
env_mod_data$site_code <- as.factor(env_mod_data$site_code)
env_mod_data$Urchins_scale <- scale(env_mod_data$Urchins)
env_mod_data$Turban_Snails_scale <- scale(env_mod_data$Turban_Snails)
#Total boat counts
env_mod_data$mean_total_boat_counts_1km_scale <- scale(env_mod_data$mean_total_boat_counts_1km)
env_mod_data$mean_total_boat_counts_1.5km_scale <- scale(env_mod_data$mean_total_boat_counts_1.5km)
env_mod_data$mean_total_boat_counts_2km_scale <- scale(env_mod_data$mean_total_boat_counts_2km)

levels(env_mod_data$dominant_score) <- list(loose = "3", solid = "4")

regions <- c("region1", "region2", "region3", "region4")
env_mod_data[regions] <- lapply(env_mod_data[regions], factor)
```


```{r Creating a dataset with averaged values for urchin/snail counts, and urch sn ratio, include=FALSE}
env_mod_data_avg <- env_mod_data%>%
  group_by(site_code, site_name, latitude, longitude, sst,
           openness, norm_openness, region1, region2, region3, region4, 
           mean_app_fishing_hours, mean_rec_boat_counts_500m, mean_rec_boat_counts_1km, 
           mean_rec_boat_counts_1.5km, mean_rec_boat_counts_750m, mean_rec_boat_counts_2km,
           sst_scale, openness_scale, norm_openness_scale, mean_app_fishing_hours_scale,
           mean_rec_boat_counts_500m_scale, mean_rec_boat_counts_1km_scale, 
           mean_rec_boat_counts_1.5km_scale, mean_rec_boat_counts_750m_scale,
           mean_rec_boat_counts_2km_scale, mean_total_boat_counts_1km,
           mean_total_boat_counts_1.5km, mean_total_boat_counts_2km, mean_total_boat_counts_1km_scale,
           mean_total_boat_counts_1.5km_scale, mean_total_boat_counts_2km_scale)%>%
  summarise(urchins_avg = mean(Urchins),
            snails_avg = mean(Turban_Snails),
            urch_sn_ratio_avg = mean(urch_sn_ratio),
            depth_avg = mean(depth))

env_mod_data_avg$urchins_avg_scale <- scale(env_mod_data_avg$urchins_avg)
env_mod_data_avg$turban_snails_avg_scale <- scale(env_mod_data_avg$snails_avg)
```


```{r Making theme for all subsequent figures, echo=FALSE}
theme.figure <- function (){
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    #axis.line = element_line(colour = "black"),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12, face = "plain"),
    legend.text = element_text(size = 10, face = "plain"),
    legend.title = element_text(size = 14, face = "bold")) 
}

```

## Modelling the log urchin to snail ratio {.tabset}

### Plotting the EVs against urchin to snail ratio

```{r ratio against substrate score, echo=FALSE}
ggplot(env_mod_data, aes(x = dominant_score, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against openness,echo=FALSE}
ggplot(env_mod_data, aes(x = norm_openness, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against sst, echo=FALSE}
ggplot(env_mod_data, aes(x = sst, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against mean rec boat counts 1.5km, echo=FALSE}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_1.5km, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against mean rec boat counts 1km, echo=FALSE}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_1km, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against mean rec boat counts 2km, echo=FALSE}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_2km, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against mean apparent fishing hours, echo=FALSE}
ggplot(env_mod_data, aes(x = mean_app_fishing_hours, y = urch_sn_ratio))+
  geom_point()+
  theme.figure()
```


### Plotting the EVs against each other to look into concurvity

```{r SST vs openness, echo=FALSE}
ggplot(env_mod_data, aes(x = sst, y = norm_openness, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r SST vs mean rec fishing 1 km, echo=FALSE}
ggplot(env_mod_data, aes(x = sst, y = mean_total_boat_counts_1km, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r SST vs mean apparent fishing hours, echo=FALSE}
ggplot(env_mod_data, aes(x = sst, y = mean_app_fishing_hours, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r Openness vs mean rec boats 1km, echo=FALSE}
ggplot(env_mod_data, aes(x = norm_openness, y = mean_rec_boat_counts_1km, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r Openness vs mean apparent fishing hours, echo=FALSE}
ggplot(env_mod_data, aes(x = norm_openness, y = mean_app_fishing_hours, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r Mean app hours vs mean rec boats 1km, echo=FALSE}
ggplot(env_mod_data, aes(x = mean_app_fishing_hours, y = mean_rec_boat_counts_1km, colour = region1))+
  geom_point()+
  theme.figure()
```



### Building GAMs to test the effects on the Urchin to snail ratio {.tabset}

#### Model with 1 km recreational boats

```{r Total 1km model with no spatial autocorrelation term}
total_mod_1km <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_1km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```


```{r 1km model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1km)
coef(total_mod_1km)

plot(total_mod_1km)
summary(total_mod_1km)
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1km residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1km <- residuals(total_mod_1km, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)

nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1km, lw)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_total_mod_1km, lw_inverse)
```
</div>
It appears that the residuals are autocorrelated. Will need to address this.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1km concurvity, echo=FALSE}
#concurvity(total_mod_1km, full = TRUE)
concurvity(total_mod_1km, full = FALSE)
```
</div>
Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.



#### Model with 1.5 km recreational boats

```{r Total 1.5km model with no spatial autocorrelation term}
total_mod_1.5km <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_1.5km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```

```{r 1.5 km model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1.5km, rep = 500)

plot(total_mod_1.5km, residuals = TRUE)
summary(total_mod_1.5km)
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1.5km residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1.5km <- residuals(total_mod_1.5km, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)

nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1.5km, lw)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_total_mod_1.5km, lw_inverse)
```
</div>
It appears that the residuals are autocorrelated. Will need to address this.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1.5km concurvity, echo=FALSE}
concurvity(total_mod_1.5km, full = FALSE)
```
</div>
Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.

#### Model with 2 km recreational boats

```{r Total 2km model with no spatial autocorrelation}
total_mod_2km <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_2km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```

```{r 2 km model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_2km, rep = 500)


plot(total_mod_2km)
summary(total_mod_2km)
```

##### Testing the residuals for autocorrelation
<div class="fold o">
```{r Testing 2km residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_2km <- residuals(total_mod_2km, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)

nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10) 
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_2km, lw)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_total_mod_2km, lw_inverse)
```
</div>
It appears that the residuals are autocorrelated. Will need to address this.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 2km concurvity, echo=FALSE}
concurvity(total_mod_2km, full = FALSE)
```
</div>
Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.

### Building GAMs with averaged points to test the effects on the urchin to snail ratio {.tabset}

The idea of averaging the transects at the same sites is to account for the 
spatial autocorrelation.

#### Model with 1km recreational boats and avg transects

```{r Total 1km avg model with no spatial autocorrelation}
total_mod_1km_avg <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_1km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```

```{r 1km avg model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1km_avg)

plot(total_mod_1km_avg)
summary(total_mod_1km_avg)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1km avg residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1km_avg <- residuals(total_mod_1km_avg, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1km_avg, lw_avg)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix_avg <- as.matrix(dist(extracted_coords_for_auto_test_projected_avg))
inverse_dist_matrix_avg <- 1 / (dist_matrix_avg + 0.001)
diag(inverse_dist_matrix_avg) <- 0
lw_inverse_avg <- mat2listw(inverse_dist_matrix_avg, style = "W")
moran.test(residuals_total_mod_1km_avg, lw_inverse_avg)
```
</div>  
No evidence of autocorrelation of the residuals.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1km avg concurvity, echo=FALSE}
concurvity(total_mod_1km_avg, full = FALSE)
```
</div>  

Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.

#### Model with 1.5km recreational boats and avg transects

```{r Total 1.5km avg model with no spatial autocorrelation}
total_mod_1.5km_avg <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_1.5km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```


```{r 1.5 km avg model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1.5km_avg)
#shapiro.test(total_mod_2km$residuals)
##normal

plot(total_mod_1.5km_avg)
summary(total_mod_1.5km_avg)
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1.5km avg residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1.5km_avg <- residuals(total_mod_1.5km_avg, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1.5km_avg, lw_avg)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix_avg <- as.matrix(dist(extracted_coords_for_auto_test_projected_avg))
inverse_dist_matrix_avg <- 1 / (dist_matrix_avg + 0.001)
diag(inverse_dist_matrix_avg) <- 0
lw_inverse_avg <- mat2listw(inverse_dist_matrix_avg, style = "W")
moran.test(residuals_total_mod_1.5km_avg, lw_inverse_avg)
```
</div>
No evidence of autocorrelation of the residuals.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1.5km avg concurvity, echo=FALSE}
concurvity(total_mod_1.5km_avg, full = FALSE)
```
</div>
Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.

#### Model with 2km recreational boats and avg transects

```{r Total 2km avg model with no spatial autocorrelation}
total_mod_2km_avg <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_2km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```

```{r 2 km avg model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_2km_avg)

plot(total_mod_2km_avg)
summary(total_mod_2km_avg)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 2km avg residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_2km_avg <- residuals(total_mod_2km_avg, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_2km_avg, lw_avg)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix_avg <- as.matrix(dist(extracted_coords_for_auto_test_projected_avg))
inverse_dist_matrix_avg <- 1 / (dist_matrix_avg + 0.001)
diag(inverse_dist_matrix_avg) <- 0
lw_inverse_avg <- mat2listw(inverse_dist_matrix_avg, style = "W")
moran.test(residuals_total_mod_2km_avg, lw_inverse_avg)
```
</div>
No evidence of autocorrelation of the residuals.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 2km avg concurvity, echo=FALSE}
concurvity(total_mod_2km_avg, full = FALSE)
```
</div>
Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.


### Building GAMs to test the effects on the Urchin to snail ratio (without mean apparent fishing hours) {.tabset}

Redoing the GAMs without the mean apparent fishing hours because the concurvity 
is very high.

#### Model with 1 km recreational boats (no mean app hours)

```{r Total 1km model no app hours with no spatial autocorrelation term}
total_mod_1km_no_app_hours <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_1km_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```

```{r 1km no app hours model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1km_no_app_hours)

plot(total_mod_1km_no_app_hours)
summary(total_mod_1km_no_app_hours)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1km no app hours residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1km_no_app_hours <- residuals(total_mod_1km_no_app_hours, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)

nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1km_no_app_hours, lw)
```
</div>
It appears that the residuals are autocorrelated. Will need to address this.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1km no app hours concurvity, echo=FALSE}
concurvity(total_mod_1km_no_app_hours, full = FALSE)
```
</div>

Some of the variables appear to still have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. 



#### Model with 1.5 km recreational boats (no mean app hours)

```{r Total 1.5km model no app hours with no spatial autocorrelation term}
total_mod_1.5km_no_app_hours <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_1.5km_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```

```{r 1.5 km no app hours model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1.5km_no_app_hours)

plot(total_mod_1.5km_no_app_hours, residuals = TRUE)
summary(total_mod_1.5km_no_app_hours)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1.5km no app hours residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1.5km_no_app_hours <- residuals(total_mod_1.5km_no_app_hours, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)

nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1.5km_no_app_hours, lw)
```
</div>
It appears that the residuals are autocorrelated. Will need to address this.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1.5km no app hours concurvity, echo=FALSE}
concurvity(total_mod_1.5km_no_app_hours, full = FALSE)
```
</div>
Some of the variables appear to still have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots.

#### Model with 2 km recreational boats (no mean app hours)

```{r Total 2km no app hours model with no spatial autocorrelation}
total_mod_2km_no_app_hours <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_2km_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```

```{r 2 km no app hours model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_2km_no_app_hours)


plot(total_mod_2km_no_app_hours)
summary(total_mod_2km_no_app_hours)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 2km no app hours residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_2km_no_app_hours <- residuals(total_mod_2km_no_app_hours, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)

nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10) 
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_2km_no_app_hours, lw)
```
</div>
It appears that the residuals are autocorrelated. Will need to address this.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 2km no app hours concurvity, echo=FALSE}
concurvity(total_mod_2km_no_app_hours, full = TRUE)
```
</div>
Some of the variables appear to still have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. 



### Building GAMs with averaged points to test the effects on the urchin to snail ratio (without mean apparent fishing hours) {.tabset}

The idea of averaging the transects at the same sites is to account for the 
spatial autocorrelation.

#### Model with 1km recreational boats and avg transects (no app hours)

```{r Total 1km avg no app hours model with no spatial autocorrelation}
total_mod_1km_avg_no_app_hours <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_1km_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```

```{r 1km avg no app hours model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1km_avg_no_app_hours)

plot(total_mod_1km_avg_no_app_hours)
summary(total_mod_1km_avg_no_app_hours)
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1km avg no app hours residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1km_avg_no_app_hours <- residuals(total_mod_1km_avg_no_app_hours, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1km_avg_no_app_hours, lw_avg)
```
</div>
No evidence of autocorrelation of the residuals.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1km avg no app hours concurvity, echo=FALSE}
concurvity(total_mod_1km_avg_no_app_hours, full = FALSE)
```
</div>
Some of the variables appear to still have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. 

#### Model with 1.5km recreational boats and avg transects (no app hours)

```{r Total 1.5km avg no app hours model with no spatial autocorrelation}
total_mod_1.5km_avg_no_app_hours <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_1.5km_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```


```{r 1.5 km avg no app hours model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1.5km_avg_no_app_hours)

plot(total_mod_1.5km_avg_no_app_hours)
summary(total_mod_1.5km_avg_no_app_hours)
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1.5km avg no app hours residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1.5km_avg_no_app_hours <- residuals(total_mod_1.5km_avg_no_app_hours, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1.5km_avg_no_app_hours, lw_avg)
```
</div>
No evidence of autocorrelation of the residuals.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1.5km avg no app hours concurvity, echo=FALSE}
concurvity(total_mod_1.5km_avg_no_app_hours, full = FALSE)
```
</div>
Some of the variables appear to still have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots.

#### Model with 2km recreational boats and avg transects (no mean app hours)

```{r Total 2km avg no app hours model with no spatial autocorrelation}
total_mod_2km_avg_no_app_hours <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_2km_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```

```{r 2 km avg no app hours model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_2km_avg_no_app_hours)

plot(total_mod_2km_avg_no_app_hours)
summary(total_mod_2km_avg_no_app_hours)
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 2km no app hours avg residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_2km_avg_no_app_hours <- residuals(total_mod_2km_avg_no_app_hours, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_2km_avg_no_app_hours, lw_avg)
```
</div>
No evidence of autocorrelation of the residuals.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 2km avg no app hours concurvity, echo=FALSE}
concurvity(total_mod_2km_avg_no_app_hours, full = FALSE)
```
</div>
Some of the variables appear to still have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots.

### Building GAMs on urch sn ratio with TOTAL boats observed instead of rec boats fishing {.tabset}

#### Model with 1 km total boats

```{r Total 1km total boats model with no spatial autocorrelation term}
total_mod_1km_total_boats <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_total_boat_counts_1km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```


```{r 1km total boats model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1km_total_boats)
coef(total_mod_1km_total_boats)
#shapiro.test(total_mod_1km_total_boats$residuals)

plot(total_mod_1km_total_boats)
summary(total_mod_1km_total_boats)
```
When using total boats that term seems to have more of an effect than rec boats 
fishing only.

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1km total boats residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1km_total_boats <- residuals(total_mod_1km_total_boats, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)

nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1km_total_boats, lw)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_total_mod_1km_total_boats, lw_inverse)
```
</div>
It appears that the residuals are autocorrelated. Will need to address this.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1km total boats concurvity, echo=FALSE}
concurvity(total_mod_1km_total_boats, full = FALSE)
```
</div>
Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.


#### Model with 1.5 km total boats

```{r Total 1.5km total boats model with no spatial autocorrelation term}
total_mod_1.5km_total_boats <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_total_boat_counts_1.5km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```


```{r 1.5km total boats model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1.5km_total_boats)
coef(total_mod_1.5km_total_boats)
#shapiro.test(total_mod_1.5km_total_boats$residuals)

plot(total_mod_1.5km_total_boats)
summary(total_mod_1.5km_total_boats)
```
When using total boats that term seems to have more of an effect than rec boats 
fishing only.

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1.5km total boats residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1.5km_total_boats <- residuals(total_mod_1.5km_total_boats, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)

nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1.5km_total_boats, lw)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_total_mod_1.5km_total_boats, lw_inverse)
```
</div>
It appears that the residuals are autocorrelated. Will need to address this.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1.5km total boats concurvity, echo=FALSE}
concurvity(total_mod_1.5km_total_boats, full = FALSE)
```
</div>
Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.

#### Model with 2 km total boats

```{r Total 2km total boats model with no spatial autocorrelation term}
total_mod_2km_total_boats <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_total_boat_counts_2km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```


```{r 2km total boats model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_2km_total_boats)
coef(total_mod_2km_total_boats)
#shapiro.test(total_mod_1km_total_boats$residuals)

plot(total_mod_2km_total_boats)
summary(total_mod_2km_total_boats)
```
When using total boats that term seems to have more of an effect than rec boats 
fishing only.

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 2km total boats residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_2km_total_boats <- residuals(total_mod_2km_total_boats, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)

nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_2km_total_boats, lw)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_total_mod_2km_total_boats, lw_inverse)
```
</div>
It appears that the residuals are autocorrelated. Will need to address this.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 2km total boats concurvity, echo=FALSE}
concurvity(total_mod_2km_total_boats, full = FALSE)
```
</div>
Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.




### Building GAMs on the urch sn ratio with TOTAL boats and averaged transects{.tabset}

#### Distribution of average urch sn ratio

```{r Histogram of average urch sn ratio}
ggplot(env_mod_data_avg, aes(x = urch_sn_ratio_avg))+
  geom_histogram(binwidth = 0.55)+
  theme.figure()

#shapiro.test(env_mod_data_avg$urch_sn_ratio_avg)
#Says that average urch sn ratio is normal
```


#### Model with 1 km total boats avg

```{r Total 1km avg total boats model}
total_mod_1km_total_boats_avg <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_total_boat_counts_1km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = FALSE)
```


```{r 1km total boats avg model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1km_total_boats_avg)
coef(total_mod_1km_total_boats_avg)
#shapiro.test(total_mod_1km_total_boats_avg$residuals)

plot(total_mod_1km_total_boats_avg)
summary(total_mod_1km_total_boats_avg)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1km total boats avg residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1km_total_boats_avg <- residuals(total_mod_1km_total_boats_avg, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1km_total_boats_avg, lw_avg)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix_avg <- as.matrix(dist(extracted_coords_for_auto_test_projected_avg))
inverse_dist_matrix_avg <- 1 / (dist_matrix_avg + 0.001)
diag(inverse_dist_matrix_avg) <- 0
lw_inverse_avg <- mat2listw(inverse_dist_matrix_avg, style = "W")
moran.test(residuals_total_mod_1km_total_boats_avg, lw_inverse_avg)
```
</div>
No spatial autocorrelation.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1km total boats avg concurvity, echo=FALSE}
concurvity(total_mod_1km_total_boats_avg, full = FALSE)
#concurvity(total_mod_1km_total_boats_avg, full = TRUE)
```
</div>
Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.

#### Model with 1.5 km total boats avg

```{r Total 1.5km avg total boats model}
total_mod_1.5km_total_boats_avg <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_total_boat_counts_1.5km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```


```{r 1.5km total boats avg model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1.5km_total_boats_avg)
coef(total_mod_1.5km_total_boats_avg)
#shapiro.test(total_mod_1km_total_boats$residuals)

plot(total_mod_1.5km_total_boats_avg)
summary(total_mod_1.5km_total_boats_avg)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1.5km total boats avg residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1.5km_total_boats_avg <- residuals(total_mod_1.5km_total_boats_avg, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1.5km_total_boats_avg, lw_avg)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix_avg <- as.matrix(dist(extracted_coords_for_auto_test_projected_avg))
inverse_dist_matrix_avg <- 1 / (dist_matrix_avg + 0.001)
diag(inverse_dist_matrix_avg) <- 0
lw_inverse_avg <- mat2listw(inverse_dist_matrix_avg, style = "W")
moran.test(residuals_total_mod_1.5km_total_boats_avg, lw_inverse_avg)
```
</div>
No spatial autocorrelation.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1.5km total boats avg concurvity, echo=FALSE}
concurvity(total_mod_1.5km_total_boats_avg, full = FALSE)
#concurvity(total_mod_1.5km_total_boats_avg, full = TRUE)
```
</div>
Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.


#### Model with 2 km total boats avg

```{r Total 2km avg total boats model}
total_mod_2km_total_boats_avg <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_total_boat_counts_2km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```


```{r 2km total boats avg model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_2km_total_boats_avg)
coef(total_mod_2km_total_boats_avg)
#shapiro.test(total_mod_2km_total_boats_avg$residuals)

plot(total_mod_2km_total_boats_avg)
summary(total_mod_2km_total_boats_avg)
```
Might need to omit mean apparent fishing hours as this improves the residuals
and concurvity.

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 2km total boats avg residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_2km_total_boats_avg <- residuals(total_mod_2km_total_boats_avg, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_2km_total_boats_avg, lw_avg)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix_avg <- as.matrix(dist(extracted_coords_for_auto_test_projected_avg))
inverse_dist_matrix_avg <- 1 / (dist_matrix_avg + 0.001)
diag(inverse_dist_matrix_avg) <- 0
lw_inverse_avg <- mat2listw(inverse_dist_matrix_avg, style = "W")
moran.test(residuals_total_mod_2km_total_boats_avg, lw_inverse_avg)
```
</div>
No spatial autocorrelation.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 2km total boats avg concurvity, echo=FALSE}
concurvity(total_mod_2km_total_boats_avg, full = FALSE)
#concurvity(total_mod_2km_total_boats_avg, full = TRUE)
```
</div>
Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.



### Building GAMs on the urch sn ratio with TOTAL boats, avg transects, and no app hours {.tabset}

#### Model with 1 km total boats avg and no app hours

```{r Total 1km avg no app hours total boats model}
total_mod_1km_total_boats_avg_no_hours <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_total_boat_counts_1km_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = FALSE)
```


```{r 1km total boats avg no app hours model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1km_total_boats_avg_no_hours)
coef(total_mod_1km_total_boats_avg_no_hours)
#shapiro.test(total_mod_1km_total_boats_avg_no_hours$residuals)

plot(total_mod_1km_total_boats_avg_no_hours)
summary(total_mod_1km_total_boats_avg_no_hours)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1km total boats avg no app hours residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1km_total_boats_avg_no_hours <- residuals(total_mod_1km_total_boats_avg_no_hours, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1km_total_boats_avg_no_hours, lw_avg)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix_avg <- as.matrix(dist(extracted_coords_for_auto_test_projected_avg))
inverse_dist_matrix_avg <- 1 / (dist_matrix_avg + 0.001)
diag(inverse_dist_matrix_avg) <- 0
lw_inverse_avg <- mat2listw(inverse_dist_matrix_avg, style = "W")
moran.test(residuals_total_mod_1km_total_boats_avg_no_hours, lw_inverse_avg)
```
</div>
No spatial autocorrelation.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1km total boats avg no app hours concurvity, echo=FALSE}
concurvity(total_mod_1km_total_boats_avg_no_hours, full = FALSE)
#concurvity(total_mod_1km_total_boats_avg_no_hours, full = TRUE)
```
</div>
Still high concurvity but it is improved slightly and the other ones are not as
high as the mean apparent fishing hours were.


#### Model with 1.5 km total boats avg and no app hours

```{r Total 1.5km avg no app hours total boats model}
total_mod_1.5km_total_boats_avg_no_hours <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_total_boat_counts_1.5km_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = FALSE)
```


```{r 1.5km total boats avg no app hours model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1.5km_total_boats_avg_no_hours, rep = 500)
coef(total_mod_1.5km_total_boats_avg_no_hours)
#shapiro.test(total_mod_1.5km_total_boats_avg_no_hours$residuals)

plot(total_mod_1.5km_total_boats_avg_no_hours)
summary(total_mod_1.5km_total_boats_avg_no_hours)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1.5km total boats avg no app hours residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1.5km_total_boats_avg_no_hours <- residuals(total_mod_1.5km_total_boats_avg_no_hours, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1.5km_total_boats_avg_no_hours, lw_avg)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix_avg <- as.matrix(dist(extracted_coords_for_auto_test_projected_avg))
inverse_dist_matrix_avg <- 1 / (dist_matrix_avg + 0.001)
diag(inverse_dist_matrix_avg) <- 0
lw_inverse_avg <- mat2listw(inverse_dist_matrix_avg, style = "W")
moran.test(residuals_total_mod_1.5km_total_boats_avg_no_hours, lw_inverse_avg)
```
</div>
No spatial autocorrelation.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1.5km total boats avg no app hours concurvity, echo=FALSE}
concurvity(total_mod_1.5km_total_boats_avg_no_hours, full = FALSE)
#concurvity(total_mod_1.5km_total_boats_avg_no_hours, full = TRUE)
```
</div>
Still high concurvity but it is improved slightly and the other ones are not as
high as the mean apparent fishing hours were.


#### Model with 2 km total boats avg

```{r Total 2km avg no app hours total boats model}
total_mod_2km_total_boats_avg_no_hours <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_total_boat_counts_2km_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = FALSE)
```


```{r 2km total boats avg no app hours model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_2km_total_boats_avg_no_hours, rep = 500)
coef(total_mod_2km_total_boats_avg_no_hours)
#shapiro.test(total_mod_2km_total_boats_avg_no_hours$residuals)

plot(total_mod_2km_total_boats_avg_no_hours)
summary(total_mod_2km_total_boats_avg_no_hours)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 2km total boats avg no app hours residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_2km_total_boats_avg_no_hours <- residuals(total_mod_2km_total_boats_avg_no_hours, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_2km_total_boats_avg_no_hours, lw_avg)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix_avg <- as.matrix(dist(extracted_coords_for_auto_test_projected_avg))
inverse_dist_matrix_avg <- 1 / (dist_matrix_avg + 0.001)
diag(inverse_dist_matrix_avg) <- 0
lw_inverse_avg <- mat2listw(inverse_dist_matrix_avg, style = "W")
moran.test(residuals_total_mod_2km_total_boats_avg_no_hours, lw_inverse_avg)
```
</div>
No spatial autocorrelation.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 2km total boats avg no app hours concurvity, echo=FALSE}
concurvity(total_mod_2km_total_boats_avg_no_hours, full = FALSE)
#concurvity(total_mod_2km_total_boats_avg_no_hours, full = TRUE)
```
</div>
Still high concurvity but it is improved slightly and the other ones are not as
high as the mean apparent fishing hours were.

### Building GAMs on the urch sn ratio with TOTAL boats and RE of transect/site

This method appears to present some issues. All of the other variables become
insignificant and site code is significant. Additionally, the deviance explained
rises to ~90-95 % which seems like the model is simply fitting for every
individual point.

#### Model with 1 km total boats and RE of transect

```{r Total 1km total boats RE transect model}
total_mod_1km_total_boats_transect <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_total_boat_counts_1km_scale)+
                   s(mean_app_fishing_hours_scale)+
                     s(site_code, bs = "re"),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```


```{r 1km total boats model transect output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1km_total_boats_transect)
coef(total_mod_1km_total_boats_transect)
#shapiro.test(total_mod_1km_total_boats_transect$residuals)

plot(total_mod_1km_total_boats_transect)
summary(total_mod_1km_total_boats_transect)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1km total boats transect residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1km_total_boats_transect <- residuals(total_mod_1km_total_boats_transect, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)

nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1km_total_boats_transect, lw)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_total_mod_1km_total_boats_transect, lw_inverse)
```
</div>
No autocorrelation.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1km total boats transect concurvity, echo=FALSE}
concurvity(total_mod_1km_total_boats_transect, full = FALSE)
```
</div>
Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.



## Modelling the snail counts{.tabset}

### Building GAMs on the snail counts with TOTAL boats (negative binomial)

#### Model with 1 km total boats

```{r Modelling the snail counts 1km total with a total model}
total_mod_snail_counts_1km_nb <- gam(Turban_Snails ~ s(sst_scale)+
                               s(norm_openness_scale)+
                              s(mean_total_boat_counts_1km_scale)+
                                s(mean_app_fishing_hours_scale)+
                               s(Urchins_scale),
                             data = env_mod_data,
                             family = nb(),
                             method = "REML",
                             select = TRUE)
```


```{r Snail 1km total boats model output, echo=FALSE}
set.seed(10)
gam.check(total_mod_snail_counts_1km_nb, rep = 500)
plot(total_mod_snail_counts_1km_nb, rug = TRUE)
     #residuals = TRUE, pch = 1, shade = TRUE,
     #shift = coef(total_mod_snail_counts)[1])
summary(total_mod_snail_counts_1km_nb)
shapiro.test(total_mod_snail_counts_1km_nb$residuals)

#Using nb seems improved from the poisson and quasi-poisson in terms of the residuals but still 
#has some problematic stuff going on
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing snail mod 1km total residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_snail_counts_1km_nb <- residuals(total_mod_snail_counts_1km_nb, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_snail_counts_1km_nb, lw)
```
Some autocorrelation present.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing snail 1km total boats nb concurvity, echo=FALSE}
concurvity(total_mod_snail_counts_1km_nb, full = FALSE)
#concurvity(total_mod_snail_counts_1km_nb, full = TRUE)
```
</div>

Some high concurvity so might need to eliminate certain variables.

#### Model with 1.5 km total boats

```{r Modelling the snail counts 1.5km total with a total model}
total_mod_snail_counts_1.5km_nb <- gam(Turban_Snails ~ s(sst_scale)+
                               s(norm_openness_scale)+
                              s(mean_total_boat_counts_1.5km_scale)+
                                s(mean_app_fishing_hours_scale)+
                               s(Urchins_scale),
                             data = env_mod_data,
                             family = nb(),
                             method = "REML",
                             select = TRUE)

```

```{r Snail 1.5km total boats model output, echo=FALSE}
set.seed(10)
gam.check(total_mod_snail_counts_1.5km_nb, rep = 500)
plot(total_mod_snail_counts_1.5km_nb, rug = TRUE)
     #residuals = TRUE, pch = 1, shade = TRUE,
     #shift = coef(total_mod_snail_counts)[1])
summary(total_mod_snail_counts_1.5km_nb)
shapiro.test(total_mod_snail_counts_1.5km_nb$residuals)

#Using nb seems improved from the poisson and quasi-poisson in terms of the residuals but still 
#has some problematic stuff going on
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing snail mod 1.5km total residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_snail_counts_1.5km_nb <- residuals(total_mod_snail_counts_1.5km_nb, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_snail_counts_1.5km_nb, lw)
```
Some autocorrelation present.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing snail 1.5km total boats nb concurvity, echo=FALSE}
concurvity(total_mod_snail_counts_1.5km_nb, full = FALSE)
#concurvity(total_mod_snail_counts_1.5km_nb, full = TRUE)
```
</div>

#### Model with 2 km total boats

```{r Modelling the snail counts 2km total with a total model}
total_mod_snail_counts_2km_nb <- gam(log(Turban_Snails + 1) ~ s(sst_scale)+
                               s(norm_openness_scale)+
                              s(mean_total_boat_counts_2km_scale)+
                                s(mean_app_fishing_hours_scale)+
                               s(Urchins_scale),
                             data = env_mod_data,
                             family = gaussian(),
                             method = "REML",
                             select = TRUE)
```

```{r Snail 2km total boats model output, echo=FALSE}
set.seed(10)
gam.check(total_mod_snail_counts_2km_nb, rep = 500)
plot(total_mod_snail_counts_2km_nb, rug = TRUE)
     #residuals = TRUE, pch = 1, shade = TRUE,
     #shift = coef(total_mod_snail_counts)[1])
summary(total_mod_snail_counts_2km_nb)
shapiro.test(total_mod_snail_counts_2km_nb$residuals)

#Using nb seems improved from the poisson and quasi-poisson in terms of the residuals but still 
#has some problematic stuff going on
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing snail mod 2km total residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_snail_counts_2km_nb <- residuals(total_mod_snail_counts_2km_nb, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_snail_counts_2km_nb, lw)
```
Some autocorrelation present.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 2km rec boats concurvity, echo=FALSE}
concurvity(total_mod_snail_counts_2km_nb, full = FALSE)
#concurvity(total_mod_snail_counts_2km_nb, full = TRUE)
```
</div>




### Building GAMs on the snail counts with TOTAL boats (log transformed)

#### Model with 1 km total boats

```{r Modelling the snail counts 1km log total with a total model}
total_mod_snail_counts_1km_log <- gam(log(Turban_Snails + 1) ~ s(sst_scale)+
                               s(norm_openness_scale)+
                              s(mean_total_boat_counts_1km_scale)+
                                s(mean_app_fishing_hours_scale)+
                               s(Urchins_scale),
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```


```{r Snail 1km log total boats model output, echo=FALSE}
set.seed(10)
gam.check(total_mod_snail_counts_1km_log, rep = 500)
plot(total_mod_snail_counts_1km_log, rug = TRUE)
     #residuals = TRUE, pch = 1, shade = TRUE,
     #shift = coef(total_mod_snail_counts)[1])
summary(total_mod_snail_counts_1km_log)
shapiro.test(total_mod_snail_counts_1km_log$residuals)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing snail mod 1km log total residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_snail_counts_1km_log <- residuals(total_mod_snail_counts_1km_log, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_snail_counts_1km_log, lw)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_total_mod_snail_counts_1km_log, lw_inverse)
```
Some autocorrelation present. Only significant with the inverse distance method.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing snail 1km total log boats concurvity, echo=FALSE}
concurvity(total_mod_snail_counts_1km_log, full = FALSE)
#concurvity(total_mod_snail_counts_1km_log, full = TRUE)
```
</div>

Some high concurvity so might need to eliminate certain variables.

#### Model with 1.5 km total boats

```{r Modelling the snail counts 1.5km log total with a total model}
total_mod_snail_counts_1.5km_log <- gam(log(Turban_Snails + 1) ~ s(sst_scale)+
                               s(norm_openness_scale)+
                              s(mean_total_boat_counts_1.5km_scale)+
                                s(mean_app_fishing_hours_scale)+
                               s(Urchins_scale),
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```


```{r Snail 1.5km log total boats model output, echo=FALSE}
set.seed(10)
gam.check(total_mod_snail_counts_1.5km_log, rep = 500)
plot(total_mod_snail_counts_1.5km_log, rug = TRUE)
     #residuals = TRUE, pch = 1, shade = TRUE,
     #shift = coef(total_mod_snail_counts)[1])
summary(total_mod_snail_counts_1.5km_log)
shapiro.test(total_mod_snail_counts_1.5km_log$residuals)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing snail mod 1.5km log total residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_snail_counts_1.5km_log <- residuals(total_mod_snail_counts_1.5km_log, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_snail_counts_1.5km_log, lw)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_total_mod_snail_counts_1.5km_log, lw_inverse)
```
Some autocorrelation present. Only significant with the inverse distance method.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing snail 1.5km total boats log concurvity, echo=FALSE}
concurvity(total_mod_snail_counts_1.5km_log, full = FALSE)
#concurvity(total_mod_snail_counts_1.5km_log, full = TRUE)
```
</div>

Some high concurvity so might need to eliminate certain variables.

#### Model with 2 km total boats

```{r Modelling the snail counts 2km log total with a total model}
total_mod_snail_counts_2km_log <- gam(log(Turban_Snails + 1) ~ s(sst_scale)+
                               s(norm_openness_scale)+
                              s(mean_total_boat_counts_2km_scale)+
                                s(mean_app_fishing_hours_scale)+
                               s(Urchins_scale),
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

<div class="fold o"> 
```{r Snail 2km log total boats model output, echo=FALSE}
set.seed(10)
gam.check(total_mod_snail_counts_2km_log, rep = 500)
plot(total_mod_snail_counts_2km_log, rug = TRUE)
     #residuals = TRUE, pch = 1, shade = TRUE,
     #shift = coef(total_mod_snail_counts)[1])
summary(total_mod_snail_counts_2km_log)
shapiro.test(total_mod_snail_counts_2km_log$residuals)
```
</div>

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing snail mod 2km log total residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_snail_counts_2km_log <- residuals(total_mod_snail_counts_2km_log, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_snail_counts_2km_log, lw)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_total_mod_snail_counts_2km_log, lw_inverse)
```
No autocorrelation present.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing snail 2km total boats concurvity, echo=FALSE}
concurvity(total_mod_snail_counts_2km_log, full = FALSE)
#concurvity(total_mod_snail_counts_2km_log, full = TRUE)
```
</div>

Some high concurvity so might need to eliminate certain variables.



### Building GAMs on the snail counts with TOTAL boats and no app hours (log transformed)

#### Model with 1 km total boats

```{r Modelling the snail counts 1km log total no app hours with a total model}
total_mod_snail_counts_1km_log_no_app_hours <- gam(log(Turban_Snails + 1) ~ s(sst_scale)+
                               s(norm_openness_scale)+
                              s(mean_total_boat_counts_1km_scale)+
                               s(Urchins_scale),
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

<div class="fold o"> 
```{r Snail 1km log total no app hours boats model output, echo=FALSE}
set.seed(10)
gam.check(total_mod_snail_counts_1km_log_no_app_hours, rep = 500)
plot(total_mod_snail_counts_1km_log_no_app_hours, rug = TRUE)
     #residuals = TRUE, pch = 1, shade = TRUE,
     #shift = coef(total_mod_snail_counts)[1])
summary(total_mod_snail_counts_1km_log_no_app_hours)
shapiro.test(total_mod_snail_counts_1km_log_no_app_hours$residuals)
```
</div>

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing snail mod 1km log total no app hours residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_snail_counts_1km_log_no_app_hours <- residuals(total_mod_snail_counts_1km_log_no_app_hours, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_snail_counts_1km_log_no_app_hours, lw)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_total_mod_snail_counts_1km_log_no_app_hours, lw_inverse)
```
Some autocorrelation present. Only significant with the inverse distance method.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing snail 1km total log no app hours boats concurvity, echo=FALSE}
concurvity(total_mod_snail_counts_1km_log_no_app_hours, full = FALSE)
#concurvity(total_mod_snail_counts_1km_log_no_app_hours, full = TRUE)
```
</div>

Concurvity slightly improved without the app hours.

#### Model with 1.5 km total boats

```{r Modelling the snail counts 1.5km log total no app hours with a total model}
total_mod_snail_counts_1.5km_log_no_app_hours <- gam(log(Turban_Snails + 1) ~ s(sst_scale)+
                               s(norm_openness_scale)+
                              s(mean_total_boat_counts_1.5km_scale)+
                               s(Urchins_scale),
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

<div class="fold o"> 
```{r Snail 1.5km log total no app hours boats model output, echo=FALSE}
set.seed(10)
gam.check(total_mod_snail_counts_1.5km_log_no_app_hours, rep = 500)
plot(total_mod_snail_counts_1.5km_log_no_app_hours, rug = TRUE)
     #residuals = TRUE, pch = 1, shade = TRUE,
     #shift = coef(total_mod_snail_counts)[1])
summary(total_mod_snail_counts_1.5km_log_no_app_hours)
shapiro.test(total_mod_snail_counts_1.5km_log_no_app_hours$residuals)
```
</div>

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing snail mod 1.5km log total no app hours residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_snail_counts_1.5km_log_no_app_hours <- residuals(total_mod_snail_counts_1.5km_log_no_app_hours, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_snail_counts_1.5km_log_no_app_hours, lw)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_total_mod_snail_counts_1.5km_log_no_app_hours, lw_inverse)
```
Some autocorrelation present. Only significant with the inverse distance method.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing snail 1.5km total log no app hours boats concurvity, echo=FALSE}
concurvity(total_mod_snail_counts_1.5km_log_no_app_hours, full = FALSE)
#concurvity(total_mod_snail_counts_1.5km_log_no_app_hours, full = TRUE)
```
</div>

Concurvity slightly improved without the app hours.

#### Model with 2 km total boats

```{r Modelling the snail counts 2km log total no app hours with a total model}
total_mod_snail_counts_2km_log_no_app_hours <- gam(log(Turban_Snails + 1) ~ s(sst_scale)+
                               s(norm_openness_scale)+
                              s(mean_total_boat_counts_2km_scale)+
                               s(Urchins_scale),
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```


```{r Snail 2km log total no app hours boats model output, echo=FALSE}
set.seed(10)
gam.check(total_mod_snail_counts_2km_log_no_app_hours, rep = 500)
plot(total_mod_snail_counts_2km_log_no_app_hours, rug = TRUE)
     #residuals = TRUE, pch = 1, shade = TRUE,
     #shift = coef(total_mod_snail_counts)[1])
summary(total_mod_snail_counts_2km_log_no_app_hours)
shapiro.test(total_mod_snail_counts_2km_log_no_app_hours$residuals)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing snail mod 2km log total no app hours residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_snail_counts_2km_log_no_app_hours <- residuals(total_mod_snail_counts_2km_log_no_app_hours, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_snail_counts_2km_log_no_app_hours, lw)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_total_mod_snail_counts_2km_log_no_app_hours, lw_inverse)
```
No autocorrelation.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing snail 2km total log no app hours boats concurvity, echo=FALSE}
concurvity(total_mod_snail_counts_2km_log_no_app_hours, full = FALSE)
#concurvity(total_mod_snail_counts_2km_log_no_app_hours, full = TRUE)
```
</div>

Concurvity slightly improved without the app hours.


### Building GAMs on the snail counts with TOTAL boats and averaged points (log transformed)

The idea of averaging the transects at the same sites is to account for the 
spatial autocorrelation.

#### Looking at the distribution of avg snail counts

```{r Histogram of avg snail counts}
ggplot(env_mod_data_avg, aes(x = snails_avg))+
  geom_histogram(binwidth = 75)+
  theme.figure()
```
The average snails show a similar distribution to the full snail counts, but 
they are continuous values so I cannot use the poisson or neg bin. Also cannot
use the gamma because there are some zeros. Will try with Tweedie.

#### Snail model with 1km total boats and avg transects

```{r Modelling the snail counts 1km total avg log with a total model}
total_mod_snail_counts_1km_avg_log <- gam(log(snails_avg + 1) ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_total_boat_counts_1km_scale)+
                   s(mean_app_fishing_hours_scale)+
                     s(urchins_avg_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'ML',
               select = TRUE)
```

```{r snail 1km avg log model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_snail_counts_1km_avg_log, rep = 500)
shapiro.test(total_mod_snail_counts_1km_avg_log$residuals)

plot(total_mod_snail_counts_1km_avg_log)
summary(total_mod_snail_counts_1km_avg_log)
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing snail 1km avg log residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_snail_counts_1km_avg_log <- residuals(total_mod_snail_counts_1km_avg_log, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_snail_counts_1km_avg_log, lw_avg)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix_avg <- as.matrix(dist(extracted_coords_for_auto_test_projected_avg))
inverse_dist_matrix_avg <- 1 / (dist_matrix_avg + 0.001)
diag(inverse_dist_matrix_avg) <- 0
lw_inverse_avg <- mat2listw(inverse_dist_matrix_avg, style = "W")
moran.test(residuals_total_mod_snail_counts_1km_avg_log, lw_inverse_avg)
```
</div>  
No evidence of autocorrelation of the residuals.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing snail 1km avg log concurvity, echo=FALSE}
concurvity(total_mod_snail_counts_1km_avg_log, full = FALSE)
#concurvity(total_mod_snail_counts_1km_avg_log, full = TRUE)
```
</div>  

Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.

#### Snail model with 1.5km total boats and avg transects

```{r Modelling the snail counts 1.5km total avg log with a total model}
total_mod_snail_counts_1.5km_avg_log <- gam(log(snails_avg + 1) ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_total_boat_counts_1.5km_scale)+
                   s(mean_app_fishing_hours_scale)+
                     s(urchins_avg_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'ML',
               select = TRUE)
```

```{r snail 1.5km avg log model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_snail_counts_1.5km_avg_log, rep = 500)
shapiro.test(total_mod_snail_counts_1.5km_avg_log$residuals)

plot(total_mod_snail_counts_1.5km_avg_log)
summary(total_mod_snail_counts_1.5km_avg_log)
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing snail 1.5km avg log residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_snail_counts_1.5km_avg_log <- residuals(total_mod_snail_counts_1.5km_avg_log, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_snail_counts_1.5km_avg_log, lw_avg)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix_avg <- as.matrix(dist(extracted_coords_for_auto_test_projected_avg))
inverse_dist_matrix_avg <- 1 / (dist_matrix_avg + 0.001)
diag(inverse_dist_matrix_avg) <- 0
lw_inverse_avg <- mat2listw(inverse_dist_matrix_avg, style = "W")
moran.test(residuals_total_mod_snail_counts_1.5km_avg_log, lw_inverse_avg)
```
</div>  
No evidence of autocorrelation of the residuals.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing snail 1.5km avg log concurvity, echo=FALSE}
concurvity(total_mod_snail_counts_1.5km_avg_log, full = FALSE)
#concurvity(total_mod_snail_counts_1.5km_avg_log, full = TRUE)
```
</div>  

Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.

#### Snail model with 2km total boats and avg transects

```{r Modelling the snail counts 2km total avg log with a total model}
total_mod_snail_counts_2km_avg_log <- gam(log(snails_avg + 1) ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_total_boat_counts_2km_scale)+
                   s(mean_app_fishing_hours_scale)+
                     s(urchins_avg_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'ML',
               select = TRUE)
```

```{r snail 2km avg log model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_snail_counts_2km_avg_log, rep = 500)
shapiro.test(total_mod_snail_counts_2km_avg_log$residuals)

plot(total_mod_snail_counts_2km_avg_log)
summary(total_mod_snail_counts_2km_avg_log)
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing snail 2km avg log residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_snail_counts_2km_avg_log <- residuals(total_mod_snail_counts_2km_avg_log, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_snail_counts_2km_avg_log, lw_avg)

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix_avg <- as.matrix(dist(extracted_coords_for_auto_test_projected_avg))
inverse_dist_matrix_avg <- 1 / (dist_matrix_avg + 0.001)
diag(inverse_dist_matrix_avg) <- 0
lw_inverse_avg <- mat2listw(inverse_dist_matrix_avg, style = "W")
moran.test(residuals_total_mod_snail_counts_2km_avg_log, lw_inverse_avg)
```
</div>  
No evidence of autocorrelation of the residuals.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing snail 2km avg nb concurvity, echo=FALSE}
concurvity(total_mod_snail_counts_2km_avg_log, full = FALSE)
#concurvity(total_mod_snail_counts_2km_avg_log, full = TRUE)
```
</div>  

Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.