---
title: "RLS Modelling"
author: "Alec"
date: "2024-11-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(mgcv)
library(DHARMa)
library(mgcViz)
library(ggeffects)
library(gratia)
library(geoR)
library(geosphere)
library(gstat)
library(sf)
library(car)
library(MuMIn)
library(lme4)
library(flextable)
```




#Modelling the environmental variables

##Loading in the data
```{r Loading in the data}
env_mod_data <- read_csv(here("./03_clean_data/urch_sn_ratio_mod_data.csv"))
```

```{r Scaling the explanatory variables and converting substrate type to factor}
env_mod_data$sst_scale <- scale(env_mod_data$sst)
env_mod_data$openness_scale <- scale(env_mod_data$openness)
env_mod_data$norm_openness_scale <- scale(env_mod_data$norm_openness)
env_mod_data$mean_app_fishing_hours_scale <- scale(env_mod_data$mean_app_fishing_hours)
env_mod_data$mean_rec_boat_counts_500m_scale <- scale(env_mod_data$mean_rec_boat_counts_500m)
env_mod_data$mean_rec_boat_counts_1km_scale <- scale(env_mod_data$mean_rec_boat_counts_1km)
env_mod_data$mean_rec_boat_counts_1.5km_scale <- scale(env_mod_data$mean_rec_boat_counts_1.5km)
env_mod_data$mean_rec_boat_counts_750m_scale <- scale(env_mod_data$mean_rec_boat_counts_750m)
env_mod_data$mean_rec_boat_counts_2km_scale <- scale(env_mod_data$mean_rec_boat_counts_2km)
env_mod_data$dominant_score <- as.factor(env_mod_data$dominant_score)
env_mod_data$site_code <- as.factor(env_mod_data$site_code)
levels(env_mod_data$dominant_score) <- list(loose = "3", solid = "4")

regions <- c("region1", "region2", "region3", "region4")
env_mod_data[regions] <- lapply(env_mod_data[regions], factor)
```

###Testing the data to see if it is spatially correlated

####Moran's I with transects at same site separate
```{r Using Moran's I test on the data with geosphere}
##Method setting transects at same site to 1 m apart

sphere_coords <- cbind(env_mod_data$longitude, env_mod_data$latitude)
sphere_dist <- distm(sphere_coords, fun = distGeo)

sphere_dist[sphere_dist == 0] <- 1
inv_sphere_dist <- 1 / sphere_dist
diag(inv_sphere_dist) <- 0
  
ape::Moran.I(env_mod_data$urch_sn_ratio, weight = inv_sphere_dist, scaled = TRUE)
```

####Moran's I with transects at same site averaged
```{r Making a df with averaged urch_sn ratios}
urch_sn_ratio_averaged <- env_mod_data%>%
  group_by(site_code, latitude, longitude)%>%
  summarise(avg_urch_sn_ratio = mean(urch_sn_ratio))
```

```{r calculating Moran's I with the averaged points}
sphere_coords_avg <- cbind(urch_sn_ratio_averaged$longitude, urch_sn_ratio_averaged$latitude)

sphere_dist_avg <- distm(sphere_coords_avg, fun = distGeo)

inv_sphere_dist_avg <- 1 / sphere_dist_avg

diag(inv_sphere_dist_avg) <- 0

ape::Moran.I(urch_sn_ratio_averaged$avg_urch_sn_ratio, weight = inv_sphere_dist_avg, scaled = TRUE)
```



####Using a semi-variogram to see what type of spatial autocorrelation factor to use

#####With averaged points
```{r Transforming the coordinates into meters for the variogram}
# Using the averaged data
urch_sn_ratio_averaged_sf <- st_as_sf(urch_sn_ratio_averaged, coords = c("longitude", "latitude"), crs = 4326)
urch_sn_ratio_averaged_projected <- st_transform(urch_sn_ratio_averaged_sf, crs = 3857)
```

```{r Plotting a semi-variogram with the meter distances for the averaged transects}
vario_meters <- variogram(avg_urch_sn_ratio ~ 1, data = urch_sn_ratio_averaged_projected)
vario_meters_df <- as.data.frame(vario_meters)
plot(vario_meters)

#ggplot(vario_meters_df, aes(x = dist, y = gamma)) +
 # geom_line() +
  #geom_point() +
  #scale_x_continuous(breaks = seq(0, 15000, by = 1000)) + # Custom x-axis ticks
  #labs(x = "Distance (m)", y = "Semivariance", title = "Variogram") +
  #theme_minimal()

```

```{r Fitting a variogram model}
vario_meters_mod <- fit.variogram(vario_meters, 
                                  model = vgm(nugget = 0.1, range = 2000, psill = 0.75, 
                                              model = c("Exp", "Sph", "Gau")),
                                  fit.sills = TRUE, fit.ranges = TRUE)
vario_meters_mod
plot(vario_meters, vario_meters_mod)
```

#####With non averaged points

```{r Variogram with the non-averaged points}
urch_sn_coords_var_2 <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
urch_sn_coords_var_2_projected <- st_transform(urch_sn_coords_var_2, crs = 3857)

vario_2 <- variogram(urch_sn_ratio ~ 1, data = urch_sn_coords_var_2_projected)
vario_2_df <- as.data.frame(vario_2)
plot(vario_2)

#ggplot(vario_2_df, aes(x = dist, y = gamma)) +
 # geom_line() +
#  geom_point() +
 # scale_x_continuous(breaks = seq(0, 15000, by = 1000)) + # Custom x-axis ticks
  #labs(x = "Distance (m)", y = "Semivariance", title = "Variogram") +
  #theme_minimal()
```

```{r Fitting a variogram model}
vario_2_mod <- fit.variogram(vario_2,
                             model = vgm(1, model = c("Exp", "Sph", "Gau"), 
                                         nugget = 0.1, range = 1000, psill = 1),
                             fit.sills = TRUE, fit.ranges = TRUE)
vario_2_mod
plot(vario_2, vario_2_mod)
```


#####With shifted points
In order to use the corSpatial functions within a gamm there cannot be sites at the exact same lat and long so 
i have shifted the points at the same site by one meter or so to make the function run.

```{r Making a dataframe with sites at the same location shifted slightly}
env_mod_data_shifted <- env_mod_data%>%
    mutate(latitude = if_else(row_number() == 2, 48.82896, latitude))%>%
  mutate(latitude = if_else(row_number() == 5, 48.85746, latitude))%>%
  mutate(latitude = if_else(row_number() == 7, 48.85827, latitude))%>%
  mutate(latitude = if_else(row_number() == 21, 48.91608, latitude))%>%
   mutate(latitude = if_else(row_number() == 23, 48.83858, latitude))%>%
   mutate(latitude = if_else(row_number() == 25, 48.90711, latitude))%>%
  mutate(latitude = if_else(row_number() == 30, 48.82734, latitude))%>%
  mutate(latitude = if_else(row_number() == 32, 48.95024, latitude))%>%
   mutate(latitude = if_else(row_number() == 34, 48.95509, latitude))
```

```{r Converting the shifted points into meters}
env_mod_sf <- st_as_sf(env_mod_data_shifted, coords = c("longitude", "latitude"), crs = 4326)
env_mod_sf_projected <- st_transform(env_mod_sf, crs = 3857)

projected_coords <- st_coordinates(env_mod_sf_projected)
env_mod_data_shifted$longitude_proj <- projected_coords[ , 1]
env_mod_data_shifted$latitude_proj <- projected_coords[ , 2]
```

```{r Plotting an experimental variogram for the data that is shifted by one meter}
vario_shifted <- variogram(urch_sn_ratio ~ 1, data = env_mod_sf_projected)
plot(vario_shifted)
```

```{r Fitting a variogram model to the experimental variogram}
vario_shifted_mod <- fit.variogram(vario_shifted, model = vgm(nugget = 1, model = c("Exp", "Sph", "Gau"), range = 1000, psill = 1), fit.sills = TRUE, fit.ranges = TRUE)
vario_shifted_mod
plot(vario_shifted, vario_shifted_mod)
```

##Plotting EVs against urch sn ratio

```{r ratio against openness}
ggplot(env_mod_data, aes(x = dominant_score, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against openness}
ggplot(env_mod_data, aes(x = norm_openness, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against sst}
ggplot(env_mod_data, aes(x = sst, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against substrate type}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_2km, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r}
ggplot(env_mod_data, aes(x = mean_app_fishing_hours, y = urch_sn_ratio))+
  geom_point()
```

```{r}
ggplot(env_mod_data, aes(x = region1, y = mean_rec_boat_counts_1.5km))+
  geom_boxplot()
```


start with env models of counts
then 


##Building the GAM

###No spatial autocorrelation term and using norm openness

####temp scaled, norm openness scaled, dominant score, TRUE

```{r Making a gam with no spatial autocorrelation term to start}
env_mod_T <- gam(urch_sn_ratio ~ s(sst_scale) + s(norm_openness_scale) + dominant_score,
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(env_mod_T)

plot(env_mod_T)
```

```{r Checking the model fit with DHARMa}
env_mod_T_resids <- simulateResiduals(env_mod_T)
plot(env_mod_T_resids)
```



#####temp, norm openness scale, TRUE

```{r Making a gam with no spatial autocorrelation term to start}
env_mod_no_dom <- gam(urch_sn_ratio ~ s(sst_scale) + s(norm_openness_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(env_mod_no_dom)

plot(env_mod_no_dom)
```

```{r Checking the model fit with DHARMa}
env_mod_no_dom_resids <- simulateResiduals(env_mod_no_dom)
plot(env_mod_no_dom_resids)
```
Problem when running the model without the substrate type, Why?

####Smaller models within the overall environmental one
Gavin said maybe this is not the best way to do it as you could be biasing your
p values since summary doesn't know you already did selection.

Will need to update to norm openness scaled

#####temp, norm openness, dominant score, FALSE
```{r Making a gam with no spatial autocorrelation term to start}
env_mod_F <- gam(urch_sn_ratio ~ s(sst_scale) + s(norm_openness) + dominant_score,
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = FALSE)

set.seed(13)
gam.check(env_mod_F)

plot(env_mod_F)
```


```{r Checking the model fit with DHARMa}
env_mod_F_resids <- simulateResiduals(env_mod_F)
plot(env_mod_F_resids)
```
All good


#####temp, dominant score, FALSE
```{r Making a gam with no spatial autocorrelation and no fetch}
env_mod_no_fetch <- gam(urch_sn_ratio ~ s(sst_scale) + dominant_score, data = env_mod_data,
                        method = 'REML',
                        family = gaussian(link = "identity"),
                        select = FALSE)

gam.check(env_mod_no_fetch, rep = 500)
plot(env_mod_no_fetch)
```

```{r}
env_mod_no_fetch_resid <- simulateResiduals(env_mod_no_fetch)
plot(env_mod_no_fetch_resid)
```
Possible problem with this model according to DHARMa

#####norm openness scaled, dominant score, FALSE
```{r Making a gam with no spatial autocorrelation and no temperature}
env_mod_no_temp <- gam(urch_sn_ratio ~ s(norm_openness_scale) + dominant_score, data = env_mod_data,
                        method = 'REML',
                        family = gaussian(link = "identity"),
                        select = FALSE)

gam.check(env_mod_no_temp, rep = 500)
plot(env_mod_no_temp)
```

```{r}
env_mod_no_temp_resid <- simulateResiduals(env_mod_no_temp)
plot(env_mod_no_temp_resid)
```



###With spatial autocorrelation term

####temp scaled, norm openness scaled, dom, lat long gp interaction, TRUE

```{r Building the gam with the spatial autocorrelation term}
####This is likely overfitting the model and just saying that a smooth of lat and long explains the variation in urch ratio

env_mod_sp_gp <- gam(urch_sn_ratio ~ s(sst_scale) + s(norm_openness_scale) + dominant_score + s(latitude, longitude, bs = "gp", k = 10),
                  data = env_mod_data, 
                  family = gaussian(link = "identity"),
                  method = 'REML',
                  select = TRUE)   #In this case I have added the lat long as an interaction term directly in the model

set.seed(10)
gam.check(env_mod_sp_gp, rep = 500)
plot(env_mod_sp_gp)
```

```{r DHARMa for model fit}
env_mod_sp_gp_resid <- simulateResiduals(env_mod_sp_gp)
plot(env_mod_sp_gp_resid)
```
All good


####temp scaled, norm openness scaled, dom, lat long gp interaction, TRUE, shifted data point

```{r Building the gam with the spatial autocorrelation term and shifted points}
env_mod_sp_gp_shifted <- gam(urch_sn_ratio ~ s(sst_scale) + s(norm_openness_scale) + dominant_score + s(latitude, longitude, bs = "gp", k = 10),
                  data = env_mod_data_shifted, 
                  family = gaussian(link = "identity"),
                  method = 'REML',
                  select = TRUE)   #In this case I have added the lat long as an interaction term directly in the model

set.seed(10)
gam.check(env_mod_sp_gp, rep = 500)
plot(env_mod_sp_gp)
```

```{r DHARMa for model fit}
env_mod_sp_gp_shifted_resid <- simulateResiduals(env_mod_sp_gp_shifted)
plot(env_mod_sp_gp_shifted_resid)
```
All good

####gamm, temp scaled, norm openness scaled, dom, corExp, TRUE shifted points

```{r Building gamm with spatial autocorrelation term corExp}
env_mod_sp_cor_exp <- gamm(urch_sn_ratio ~ s(sst_scale) + s(norm_openness_scale) + dominant_score, 
                        data = env_mod_data_shifted,
                       family = gaussian(link = "identity"),
                       correlation = corExp(form = ~ longitude_proj + latitude_proj, nugget = FALSE, value = 639), 
                       method = 'REML',
                       select = TRUE)


gam.check(env_mod_sp_cor_exp$gam, rep = 500)
plot(env_mod_sp_cor_exp$gam)
```

```{r DHARMa for model fit}
env_mod_sp_cor_exp_resid <- simulateResiduals(env_mod_sp_cor_exp$lme)
plot(env_mod_sp_cor_exp_resid)
```
Problem with DHARMa for gamms

####gam, temp scaled, norm openness scaled, dom, re of site code, TRUE

```{r Building gam with random effect of site to account for spatial autocorrelation}
env_mod_re_site_code <- gam(urch_sn_ratio ~ s(sst_scale) + s(norm_openness_scale) + dominant_score + s(site_code, bs = "re"), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'REML',
                       select = TRUE)

gam.check(env_mod_re_site_code, rep = 500)
plot(env_mod_re_site_code)
```

```{r DHARMa for model fit}
env_mod_re_site_code_resid <- simulateResiduals(env_mod_re_site_code)
plot(env_mod_re_site_code_resid)
```


####gam, temp scaled, norm openness scaled, dom, random effect of region M1, TRUE

```{r Building a gam with random effect of region M1}
env_mod_re_region_M1 <- gam(urch_sn_ratio ~ s(sst_scale) + s(norm_openness_scale) + dominant_score + s(region1, bs = "re"), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'REML',
                       select = TRUE)

gam.check(env_mod_re_region_M1, rep = 500)
plot(env_mod_re_region_M1)
```

```{r DHARMa for model fit}
env_mod_re_region_M1_resid <- simulateResiduals(env_mod_re_region_M1)
plot(env_mod_re_region_M1_resid)
```

####gam, temp scaled, norm openness scaled, dom, random effect of region M2, TRUE

```{r Building a gam with random effect of region M2}
env_mod_re_region_M2 <- gam(urch_sn_ratio ~ s(sst_scale) + s(norm_openness_scale) + dominant_score + s(region2, bs = "re"), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'REML',
                       select = TRUE)

gam.check(env_mod_re_region_M2, rep = 500)
plot(env_mod_re_region_M2)
```

```{r DHARMa for model fit}
env_mod_re_region_M2_resid <- simulateResiduals(env_mod_re_region_M2)
plot(env_mod_re_region_M2_resid)
```


####gam, temp scaled, norm openness scaled, dom, random effect of region M3, TRUE

```{r Building a gam with random effect of region M3}
env_mod_re_region_M3 <- gam(urch_sn_ratio ~ s(sst_scale) + s(norm_openness_scale) + dominant_score + s(region3, bs = "re"), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'REML',
                       select = TRUE)

gam.check(env_mod_re_region_M3, rep = 500)
plot(env_mod_re_region_M3)
```

```{r DHARMa for model fit}
env_mod_re_region_M3_resid <- simulateResiduals(env_mod_re_region_M3)
plot(env_mod_re_region_M3_resid)
```

####gam, temp scaled, norm openness scaled, dom, random effect of region M4, TRUE

```{r Building a gam with random effect of region M3}
env_mod_re_region_M4 <- gam(urch_sn_ratio ~ s(sst_scale) + s(norm_openness_scale) + dominant_score + s(region4, bs = "re"), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'REML',
                       select = TRUE)

gam.check(env_mod_re_region_M4, rep = 500)
plot(env_mod_re_region_M4)
```
```{r DHARMa for model fit}
env_mod_re_region_M4_resid <- simulateResiduals(env_mod_re_region_M4)
plot(env_mod_re_region_M4_resid)
```


###Testing the models for concurvity of variables

####env_mod_T

```{r testing for multicollinearity}
concurvity(env_mod_T, full = TRUE)
```
```{r }
concurvity(env_mod_T, full = FALSE)
```

```{r}
ggplot(env_mod_data, aes(x = sst_scale, y = norm_openness_scale))+
  geom_point()
```
####env_mod_sp_cor_exp

```{r}
concurvity(env_mod_sp_cor_exp$gam, full = TRUE)
concurvity(env_mod_sp_cor_exp$gam, full = FALSE)
```
####env_mod_re_site_code

```{r}
concurvity(env_mod_re_site_code, full = FALSE)
```

####env_mod_re_region_M1

```{r}
concurvity(env_mod_re_region_M1, full = TRUE)
concurvity(env_mod_re_region_M1, full = FALSE)
```

####env_mod_re_region_M2

```{r}
concurvity(env_mod_re_region_M2, full = TRUE)
concurvity(env_mod_re_region_M2, full = FALSE)
```

####env_mod_re_region_M3

```{r}
concurvity(env_mod_re_region_M3, full = TRUE)
concurvity(env_mod_re_region_M3, full = FALSE)
```

####env_mod_re_region_M4

```{r}
concurvity(env_mod_re_region_M4, full = TRUE)
concurvity(env_mod_re_region_M4, full = FALSE)
```

###AIC with the different models

```{r Comparing the models with an AIC}
AIC(env_mod_T,
    #env_mod_sp_gp, 
    #env_mod_sp_gp_shifted,
    env_mod_sp_cor_exp$lme,
    env_mod_re_site_code,
    env_mod_re_region_M1,
    env_mod_re_region_M2, 
    env_mod_re_region_M3,
    env_mod_re_region_M4
    )%>%
arrange(AIC)%>%
  mutate(delta_AIC = AIC - AIC[1])
```

###Trying with fishing variables now

```{r Plot of urch sn ratio vs rec boats}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_750m, y = urch_sn_ratio))+
  geom_point()
```

```{r plot of urch sn ratio vs rec boat counts}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_500m, y = urch_sn_ratio))+
  geom_point()
```


####gam, rec boats 500m scaled, mean gfw scaled, TRUE
```{r Building a gam with fishing variables}
fish_mod <- gam(urch_sn_ratio ~ s(mean_rec_boat_counts_500m_scale) + s(mean_app_fishing_hours_scale), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'ML',
                       select = TRUE)

gam.check(fish_mod, rep = 500)
plot(fish_mod)
```

```{r DHARMa for model fit}
fish_mod_resids <- simulateResiduals(fish_mod)
plot(fish_mod_resids)
```

####gam, rec boats 750m scaled, mean gfw scaled, TRUE
```{r Building a gam with fishing variables}
fish_mod_750m <- gam(urch_sn_ratio ~ s(mean_rec_boat_counts_750m_scale) + s(mean_app_fishing_hours_scale), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'REML',
                       select = TRUE)

gam.check(fish_mod_750m, rep = 500)
plot(fish_mod_750m)
```

```{r DHARMa for model fit}
fish_mod_750m_resids <- simulateResiduals(fish_mod_750m)
plot(fish_mod_750m_resids)
```

####gam, rec boats 1km scaled, mean gfw scaled, TRUE
```{r Building a gam with fishing variables}
fish_mod_1km <- gam(urch_sn_ratio ~ s(mean_rec_boat_counts_1km_scale) + s(mean_app_fishing_hours_scale), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'ML',
                       select = TRUE)

gam.check(fish_mod_1km, rep = 500)
plot(fish_mod_1km)
```

```{r DHARMa for model fit}
fish_mod_1km_resids <- simulateResiduals(fish_mod_1km)
plot(fish_mod_1km_resids)
```

####gam, rec boats 1.5km scaled, mean gfw scaled, TRUE

```{r Building a gam with fishing variables}
fish_mod_1.5km <- gam(urch_sn_ratio ~ s(mean_rec_boat_counts_1.5km_scale) + s(mean_app_fishing_hours_scale), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'ML',
                       select = TRUE)

gam.check(fish_mod_1.5km, rep = 500)
plot(fish_mod_1.5km)
```

```{r DHARMa for model fit}
fish_mod_1.5km_resids <- simulateResiduals(fish_mod_1.5km)
plot(fish_mod_1.5km_resids)
```

####gam, rec boats 2km scaled, mean gfw scaled, TRUE

```{r Building a gam with fishing variables}
fish_mod_2km <- gam(urch_sn_ratio ~ s(mean_rec_boat_counts_2km_scale) + s(mean_app_fishing_hours_scale), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'ML',
                       select = TRUE)

gam.check(fish_mod_2km, rep = 500)
plot(fish_mod_2km)
summary(fish_mod_2km)
```

```{r DHARMa for model fit}
fish_mod_2km_resids <- simulateResiduals(fish_mod_2km)
plot(fish_mod_2km_resids)
```



###AIC of fishing models

```{r Comparing the models with an AIC}
AIC(fish_mod,
    fish_mod_750m,
    fish_mod_1km,
    fish_mod_1.5km,
    fish_mod_2km
    )%>%
arrange(AIC)%>%
  mutate(delta_AIC = AIC - AIC[1])
```
indicates similar effects with adding distance ---- some problems with shorter distances as there are more small counts

###Total mod with all variables

####gam, temp scaled, norm openness scaled, dom, rec boats 500m scaled, mean gfw scaled, TRUE

```{r Total model with no spatial autocorrelation}
total_mod <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   dominant_score + 
                   s(mean_rec_boat_counts_500m_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod)

plot(total_mod)
```

```{r DHARMa for model fit}
total_mod_resids <- simulateResiduals(total_mod)
plot(total_mod_resids)
```

####gam, temp scaled, norm openness scaled, dom, rec boats 500m scaled, mean gfw scaled, re of region M1, TRUE

```{r Total model with random effect of region 1}
total_mod_re_region1 <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   dominant_score + 
                   s(mean_rec_boat_counts_500m_scale)+
                   s(mean_app_fishing_hours_scale)+
                   s(region1, bs = "re"),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_re_region1)

plot(total_mod_re_region1)
```

```{r DHARMa for model fit}
total_mod_re_region1_resids <- simulateResiduals(total_mod_re_region1)
plot(total_mod_re_region1_resids)
```

####gam, temp scaled, norm openness scaled, dom, rec boats 500m scaled, mean gfw scaled, re of region M2, TRUE

```{r Total model with random effect of region 2}
total_mod_re_region2 <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   dominant_score + 
                   s(mean_rec_boat_counts_500m_scale)+
                   s(mean_app_fishing_hours_scale)+
                   s(region2, bs = "re"),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_re_region2)

plot(total_mod_re_region2)
```

```{r DHARMa for model fit}
total_mod_re_region2_resids <- simulateResiduals(total_mod_re_region2)
plot(total_mod_re_region2_resids)
```



####gam, temp scaled, norm openness scaled, dom, rec boats 2km scaled, mean gfw scaled, TRUE

```{r Total model with no spatial autocorrelation}
total_mod_2km <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   dominant_score + 
                   s(mean_rec_boat_counts_2km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_2km)
#shapiro.test(total_mod_2km$residuals)
##normal

plot(total_mod_2km)
summary(total_mod_2km)
as_flextable(total_mod_2km)
```

```{r DHARMa for model fit}
total_mod_2km_resids <- simulateResiduals(total_mod_2km)
plot(total_mod_2km_resids)
```

```{r Testing concurvity}
concurvity(total_mod_2km, full = FALSE)
```

#####gam, subset of total 2km without stuff that has no impact

```{r}
total_mod_2km_no_extra <- gam(urch_sn_ratio ~ s(sst_scale)+
                                s(norm_openness_scale)+
                   s(mean_rec_boat_counts_2km_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_2km_no_extra)
#shapiro.test(total_mod_2km_no_extra$residuals)
##normal

plot(total_mod_2km_no_extra)
summary(total_mod_2km_no_extra)
as_flextable(total_mod_2km_no_extra)
```

```{r DHARMa for model fit}
total_mod_2km_no_extra_resids <- simulateResiduals(total_mod_2km_no_extra)
plot(total_mod_2km_no_extra_resids)
```

```{r Testing concurvity}
concurvity(total_mod_2km_no_extra, full = FALSE)
```




####gam, temp scaled, norm openness scaled, dom, rec boats 1km scaled, mean gfw scaled, TRUE

```{r Total model with no spatial autocorrelation}
total_mod_1km <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   dominant_score + 
                   s(mean_rec_boat_counts_1km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_1km)
#shapiro.test(total_mod_1km$residuals)
##slightly not normal

plot(total_mod_1km)
summary(total_mod_1km)
as_flextable(total_mod_1km)
```

```{r DHARMa for model fit}
total_mod_1km_resids <- simulateResiduals(total_mod_1km)
plot(total_mod_1km_resids)
```

```{r Testing concurvity}
concurvity(total_mod_1km, full = FALSE)
```

#####gam, subset of total 1km without stuff that has no impact

```{r}
total_mod_1km_no_extra <- gam(urch_sn_ratio ~ s(sst_scale)+
                                s(norm_openness_scale)+
                   s(mean_rec_boat_counts_1km_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_1km_no_extra)
#shapiro.test(total_mod_1km_no_extra$residuals)
##normal

plot(total_mod_1km_no_extra)
summary(total_mod_1km_no_extra)
as_flextable(total_mod_1km_no_extra)
```

```{r DHARMa for model fit}
total_mod_1km_no_extra_resids <- simulateResiduals(total_mod_1km_no_extra)
plot(total_mod_1km_no_extra_resids)
```

```{r Testing concurvity}
concurvity(total_mod_1km_no_extra, full = FALSE)
```


####gam, temp scaled, norm openness scaled, dom, rec boats 1.5km scaled, mean gfw scaled, TRUE

```{r Total model with no spatial autocorrelation}
total_mod_1.5km <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   dominant_score + 
                   s(mean_rec_boat_counts_1.5km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_1.5km)
#shapiro.test(total_mod_1.5km$residuals)
##slightly not normal

plot(total_mod_1.5km)
summary(total_mod_1.5km)
as_flextable(total_mod_1.5km)
```

```{r DHARMa for model fit}
total_mod_1.5km_resids <- simulateResiduals(total_mod_1.5km)
plot(total_mod_1.5km_resids)
```

```{r Testing concurvity}
concurvity(total_mod_1.5km, full = FALSE)
```

#####gam, subset of total 1.5km without stuff that has no impact

```{r}
total_mod_1.5km_no_extra <- gam(urch_sn_ratio ~ s(sst_scale)+
                                s(norm_openness_scale)+
                   s(mean_rec_boat_counts_1.5km_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_1.5km_no_extra)
#shapiro.test(total_mod_1.5km_no_extra$residuals)
##slightly not normal

plot(total_mod_1.5km_no_extra)
summary(total_mod_1.5km_no_extra)
as_flextable(total_mod_1.5km_no_extra)
```

```{r DHARMa for model fit}
total_mod_1.5km_no_extra_resids <- simulateResiduals(total_mod_1.5km_no_extra)
plot(total_mod_1.5km_no_extra_resids)
```

```{r Testing concurvity}
concurvity(total_mod_1.5km_no_extra, full = FALSE)
```

####gam, temp scaled, norm openness scaled, dom, rec boats 2km scaled, mean gfw scaled, re of region M1, TRUE

```{r Total model with random effect of region 1}
total_mod_re_region1_2km <- gam(urch_sn_ratio ~ s(sst_scale)+
                   s(norm_openness_scale) +
                     dominant_score+
                   s(mean_rec_boat_counts_2km_scale)+
                     s(mean_app_fishing_hours_scale)+
                     s(region1, bs = "re"),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = FALSE)

set.seed(13)
gam.check(total_mod_re_region1_2km)

plot(total_mod_re_region1_2km)
summary(total_mod_re_region1_2km)
```


```{r DHARMa for model fit}
total_mod_re_region1_2km_resids <- simulateResiduals(total_mod_re_region1_2km)
plot(total_mod_re_region1_2km_resids)
```
####gam, temp scaled, norm openness scaled, dom, rec boats 500m scaled, mean gfw scaled, re of region M1, TRUE

```{r Total model with random effect of region 1}
total_mod_re_region1_test2 <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   #dominant_score + 
                   s(mean_rec_boat_counts_2km_scale)
                   #s(mean_app_fishing_hours_scale)
                   ,
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_re_region1_test2)

plot(total_mod_re_region1_test2)
summary(total_mod_re_region1_test2)
```
theres a regional effect but havent sampled all regions equally

####gam, temp scaled, norm openness scaled, dom, rec boats 2km scaled, mean gfw scaled, re of region M2, TRUE

```{r Total model with random effect of region 2}
total_mod_re_region2_2km <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   dominant_score + 
                   s(mean_rec_boat_counts_2km_scale)+
                   s(mean_app_fishing_hours_scale)+
                   s(region2, bs = "re"),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_re_region2_2km)

plot(total_mod_re_region2_2km)
```

```{r DHARMa for model fit}
total_mod_re_region2_2km_resids <- simulateResiduals(total_mod_re_region2_2km)
plot(total_mod_re_region2_2km_resids)
```


####gamm, temp scaled, norm openness scaled, dom, mean rec boats 2km scaled, mean apparent fishing hours scaled, corExp, TRUE shifted points

```{r Building total gamm with spatial autocorrelation term corExp}
total_mod_sp_cor_exp <- gamm(urch_sn_ratio ~ s(sst_scale) +
                                   s(norm_openness_scale) +
                                   dominant_score +
                                   s(mean_rec_boat_counts_2km_scale) +
                                   s(mean_app_fishing_hours_scale), 
                        data = env_mod_data_shifted,
                       family = gaussian(link = "identity"),
                       correlation = corExp(form = ~ longitude_proj + latitude_proj, nugget = FALSE, value = 639), 
                       method = 'REML',
                       select = TRUE)


gam.check(total_mod_sp_cor_exp$gam, rep = 500)
plot(total_mod_sp_cor_exp$gam)
summary(total_mod_sp_cor_exp$lme)
```

```{r DHARMa for model fit}
env_mod_sp_cor_exp_total_resid <- simulateResiduals(env_mod_sp_cor_exp_total$lme)
plot(env_mod_sp_cor_exp_total_resid)
```







###Testing for interaction

####gam, temp scaled, norm openness scaled, dom, rec boats 2km scaled, mean gfw scaled, TRUE

```{r Running a model with an interaction}
total_mod_1km_int <- gam(urch_sn_ratio ~ s(sst_scale) + 
                 #  norm_openness_scale:mean_rec_boat_counts_2km_scale  +
                      s(norm_openness_scale) +
                      s(mean_rec_boat_counts_750m_scale)+
                 #  ti(norm_openness_scale, mean_rec_boat_counts_1km_scale, k= 8)+
                   dominant_score + 
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE) 

plot(total_mod_1km_int)
summary(total_mod_1km_int)
```

```{r}
vis.gam(total_mod_1km_int, theta = -45, view = c("norm_openness_scale", "mean_rec_boat_counts_1km_scale"))
?vis.gam
```


####Using ANOVA to see if its worth including the interaction

```{r Comparing models with ANOVA}
anova(total_mod_2km, total_mod_2km_int)
```
Adding the model does not significantly change the model fit. So no point to include.



###AIC with all of the models run

```{r Starting with normal AIC which assumes infinite sample size}
AIC(env_mod_T,
    #env_mod_sp_gp, 
    #env_mod_sp_gp_shifted,
    env_mod_sp_cor_exp$lme,
   # env_mod_re_site_code,
    env_mod_re_region_M1,
    env_mod_re_region_M2, 
    env_mod_re_region_M3,
    env_mod_re_region_M4,
    fish_mod,
    fish_mod_1km,
    fish_mod_1.5km,
    fish_mod_750m,
    fish_mod_2km,
    total_mod,
   total_mod_2km,
    total_mod_re_region1,
    total_mod_re_region2,
   total_mod_re_region1_2km,
   total_mod_re_region2_2km,
   total_mod_sp_cor_exp
    )%>%
arrange(AIC)%>%
  mutate(delta_AIC = AIC - AIC[1])
```

```{r Using corrected AIC because sample size is small}
AICc(env_mod_T,
    #env_mod_sp_gp, 
    #env_mod_sp_gp_shifted,
    env_mod_sp_cor_exp$lme,
    #env_mod_re_site_code,
    env_mod_re_region_M1,
    env_mod_re_region_M2, 
    env_mod_re_region_M3,
    env_mod_re_region_M4,
    fish_mod,
    fish_mod_1km,
    fish_mod_1.5km,
    fish_mod_750m,
    fish_mod_2km,
    total_mod,
    total_mod_2km,
    total_mod_re_region1,
    total_mod_re_region2,
     total_mod_re_region1_2km,
   total_mod_re_region2_2km,
   total_mod_sp_cor_exp
    )%>%
arrange(AICc)%>%
  mutate(delta_AICc = AICc - AICc[1])
```

###Testing if the best models are significantly different

```{r Using ANOVA to test the best models}
anova(total_mod_re_region1, total_mod_re_region1_2km, env_mod_re_region_M1)
```


##Building a glm of the final model to compare

###glm, temp scaled, norm openness scaled, dom, rec boats 2km scaled, mean gfw scaled

```{r GLM of the main model to see if there is much of a difference}
total_mod_2km_glm <- lm(urch_sn_ratio ~ sst_scale + 
                   norm_openness_scale +
                   dominant_score + 
                   mean_rec_boat_counts_2km_scale+
                   mean_app_fishing_hours_scale,
               data = env_mod_data)

set.seed(13)
#shapiro.test(total_mod_2km$residuals)
##normal

plot(total_mod_2km_glm)
summary(total_mod_2km_glm)
```

```{r DHARMa for model fit}
total_mod_2km_glm_resids <- simulateResiduals(total_mod_2km_glm)
plot(total_mod_2km_glm_resids)
```


##Modelling Urchin Counts

###Plotting EVs against urchin counts

##Plotting EVs against urch sn ratio

```{r urchins against openness}
ggplot(env_mod_data, aes(x = dominant_score, y = Urchins, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r urchins against openness}
ggplot(env_mod_data, aes(x = norm_openness, y = Urchins, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r urchins against sst}
ggplot(env_mod_data, aes(x = sst, y = Urchins, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r urchins against rec boat counts type}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_1.5km, y = Urchins, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r}
ggplot(env_mod_data, aes(x = mean_app_fishing_hours, y = Urchins))+
  geom_point()
```

None of the explanatory variables appear to have much of a relationship with urchin counts


###gam, temp scaled, openness scaled, dom substrate

```{r Modelling the urchin counts with a simple environmental model}
env_mod_urchin_counts <- gam(Urchins ~ s(sst_scale)+
                               s(norm_openness_scale),
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
set.seed(10)
gam.check(env_mod_urchin_counts, rep = 500)
plot(env_mod_urchin_counts)
```

```{r DHARMa for model fit}
env_mod_urchin_counts_resids <- simulateResiduals(env_mod_urchin_counts)
plot(env_mod_urchin_counts_resids)
```

###gam, temp scaled, openness scaled, dom substrate, mean rec 1.5km scaled, mean gfw scaled, TRUE

```{r Modelling the urchin counts with a simple environmental model}
total_mod_urchin_counts <- gam(Urchins ~ s(sst_scale)+
                               s(norm_openness_scale)+
                               dominant_score+
                               s(mean_rec_boat_counts_1.5km_scale)+
                               s(mean_app_fishing_hours_scale),
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
set.seed(10)
gam.check(total_mod_urchin_counts, rep = 500)
plot(total_mod_urchin_counts)
```

```{r DHARMa for model fit}
total_mod_urchin_counts_resids <- simulateResiduals(total_mod_urchin_counts)
plot(total_mod_urchin_counts_resids)
```


##Modelling the snail counts

###Plotting the EVs against the snail counts

##Plotting EVs against urch sn ratio

```{r snails against substrate}
ggplot(env_mod_data, aes(x = dominant_score, y = Turban_Snails, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r snails against openness}
ggplot(env_mod_data, aes(x = norm_openness, y = Turban_Snails, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r snails against sst}
ggplot(env_mod_data, aes(x = sst, y = Turban_Snails, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r snails against rec boats}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_2km, y = Turban_Snails, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r snails against gfw}
ggplot(env_mod_data, aes(x = mean_app_fishing_hours, y = Turban_Snails))+
  geom_point()+
  theme.figure()
```

