---
title: "RLS modelling results"
author: "Alec"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
   toc: true
---
   <script src="hideOutput.js"></script> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(mgcv)
library(DHARMa)
library(mgcViz)
library(ggeffects)
library(gratia)
library(geoR)
library(geosphere)
library(gstat)
library(sf)
library(car)
library(MuMIn)
library(lme4)
library(flextable)
library(spdep)
```

# Modelling of red sea urchins, red turban snails, and their ratio in Barkley Sound {.tabset}

## Loading in the data and other setup

For using the data I have downloaded clean data from my preliminary explorations
that contains all of the exlanantory and response variables that I would like to
use. I have also made a separate dataset where I have averaged the values for 
urchin, snail counts etc. for transects occurring at the same sites. I have also 
made a theme for all subsequent figures

```{r Loading in the data, echo=FALSE}
env_mod_data <- read_csv(here("./03_clean_data/urch_sn_ratio_mod_data.csv"))
```

```{r Scaling the explanatory variables and converting substrate type to factor, include=FALSE}
env_mod_data$sst_scale <- scale(env_mod_data$sst)
env_mod_data$openness_scale <- scale(env_mod_data$openness)
env_mod_data$norm_openness_scale <- scale(env_mod_data$norm_openness)
env_mod_data$mean_app_fishing_hours_scale <- scale(env_mod_data$mean_app_fishing_hours)
env_mod_data$mean_rec_boat_counts_500m_scale <- scale(env_mod_data$mean_rec_boat_counts_500m)
env_mod_data$mean_rec_boat_counts_1km_scale <- scale(env_mod_data$mean_rec_boat_counts_1km)
env_mod_data$mean_rec_boat_counts_1.5km_scale <- scale(env_mod_data$mean_rec_boat_counts_1.5km)
env_mod_data$mean_rec_boat_counts_750m_scale <- scale(env_mod_data$mean_rec_boat_counts_750m)
env_mod_data$mean_rec_boat_counts_2km_scale <- scale(env_mod_data$mean_rec_boat_counts_2km)
env_mod_data$dominant_score <- as.factor(env_mod_data$dominant_score)
env_mod_data$site_code <- as.factor(env_mod_data$site_code)
env_mod_data$Urchins_scale <- scale(env_mod_data$Urchins)
env_mod_data$Turban_Snails_scale <- scale(env_mod_data$Turban_Snails)
#Testing out the total boat counts
env_mod_data$total_boat_counts_1.5km_scale <- scale(env_mod_data$mean_total_boat_counts_1.5km)
levels(env_mod_data$dominant_score) <- list(loose = "3", solid = "4")

regions <- c("region1", "region2", "region3", "region4")
env_mod_data[regions] <- lapply(env_mod_data[regions], factor)
```


```{r Creating a dataset with averaged values for urchin/snail counts, and urch sn ratio, include=FALSE}
env_mod_data_avg <- env_mod_data%>%
  group_by(site_code, site_name, latitude, longitude, sst,
           openness, norm_openness, region1, region2, region3, region4, 
           mean_app_fishing_hours, mean_rec_boat_counts_500m, mean_rec_boat_counts_1km, 
           mean_rec_boat_counts_1.5km, mean_rec_boat_counts_750m, mean_rec_boat_counts_2km,
           sst_scale, openness_scale, norm_openness_scale, mean_app_fishing_hours_scale,
           mean_rec_boat_counts_500m_scale, mean_rec_boat_counts_1km_scale, 
           mean_rec_boat_counts_1.5km_scale, mean_rec_boat_counts_750m_scale,
           mean_rec_boat_counts_2km_scale)%>%
  summarise(urchins_avg = mean(Urchins),
            snails_avg = mean(Turban_Snails),
            urch_sn_ratio_avg = mean(urch_sn_ratio),
            depth_avg = mean(depth))
```


```{r Making theme for all subsequent figures, echo=FALSE}
theme.figure <- function (){
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    #axis.line = element_line(colour = "black"),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12, face = "plain"),
    legend.text = element_text(size = 10, face = "plain"),
    legend.title = element_text(size = 14, face = "bold")) 
}

```

## Modelling the log urchin to snail ratio {.tabset}

### Plotting the EVs against urchin to snail ratio

```{r ratio against substrate score, echo=FALSE}
ggplot(env_mod_data, aes(x = dominant_score, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against openness,echo=FALSE}
ggplot(env_mod_data, aes(x = norm_openness, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against sst, echo=FALSE}
ggplot(env_mod_data, aes(x = sst, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against mean rec boat counts 1.5km, echo=FALSE}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_1.5km, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against mean rec boat counts 1km, echo=FALSE}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_1km, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against mean rec boat counts 2km, echo=FALSE}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_2km, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against mean apparent fishing hours, echo=FALSE}
ggplot(env_mod_data, aes(x = mean_app_fishing_hours, y = urch_sn_ratio))+
  geom_point()+
  theme.figure()
```

### Building GAMs to test the effects on the Urchin to snail ratio {.tabset}

#### Model with 1 km recreational boats

```{r Total 1km model with no spatial autocorrelation term}
total_mod_1km <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_1km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```


```{r 1km model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1km)

plot(total_mod_1km)
summary(total_mod_1km)
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1km residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1km <- residuals(total_mod_1km, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)

nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1km, lw)
```
</div>
It appears that the residuals are autocorrelated. Will need to address this.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1km concurvity, echo=FALSE}
concurvity(total_mod_1km, full = FALSE)
```
</div>
Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.



#### Model with 1.5 km recreational boats

```{r Total 1.5km model with no spatial autocorrelation term}
total_mod_1.5km <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_1.5km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```

```{r 1.5 km model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1.5km)

plot(total_mod_1.5km, residuals = TRUE)
summary(total_mod_1.5km)
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1.5km residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1.5km <- residuals(total_mod_1.5km, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)

nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1.5km, lw)
```
</div>
It appears that the residuals are autocorrelated. Will need to address this.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1.5km concurvity, echo=FALSE}
concurvity(total_mod_1.5km, full = FALSE)
```
</div>
Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.

#### Model with 2 km recreational boats

```{r Total 2km model with no spatial autocorrelation}
total_mod_2km <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_2km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```

```{r 2 km model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_2km)


plot(total_mod_2km)
summary(total_mod_2km)
```

##### Testing the residuals for autocorrelation
<div class="fold o">
```{r Testing 2km residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_2km <- residuals(total_mod_2km, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)

nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10) 
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_2km, lw)
```
</div>
It appears that the residuals are autocorrelated. Will need to address this.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 2km concurvity, echo=FALSE}
concurvity(total_mod_2km, full = FALSE)
```
</div>
Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.

### Building GAMs with averaged points to test the effects on the urchin to snail ratio {.tabset}

The idea of averaging the transects at the same sites is to account for the 
spatial autocorrelation.

#### Model with 1km recreational boats and avg transects

```{r Total 1km avg model with no spatial autocorrelation}
total_mod_1km_avg <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_1km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```

```{r 1km avg model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1km_avg)

plot(total_mod_1km_avg)
summary(total_mod_1km_avg)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1km avg residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1km_avg <- residuals(total_mod_1km_avg, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1km_avg, lw_avg)
```
</div>  
No evidence of autocorrelation of the residuals.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1km avg concurvity, echo=FALSE}
concurvity(total_mod_1km_avg, full = FALSE)
```
</div>  

Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.

#### Model with 1.5km recreational boats and avg transects

```{r Total 1.5km avg model with no spatial autocorrelation}
total_mod_1.5km_avg <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_1.5km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```


```{r 1.5 km avg model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1.5km_avg)
#shapiro.test(total_mod_2km$residuals)
##normal

plot(total_mod_1.5km_avg)
summary(total_mod_1.5km_avg)
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1.5km avg residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1.5km_avg <- residuals(total_mod_1.5km_avg, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1.5km_avg, lw_avg)
```
</div>
No evidence of autocorrelation of the residuals.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1.5km avg concurvity, echo=FALSE}
concurvity(total_mod_1.5km_avg, full = FALSE)
```
</div>
Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.

#### Model with 2km recreational boats and avg transects

```{r Total 2km avg model with no spatial autocorrelation}
total_mod_2km_avg <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_2km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```

```{r 2 km avg model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_2km_avg)

plot(total_mod_2km_avg)
summary(total_mod_2km_avg)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 2km avg residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_2km_avg <- residuals(total_mod_2km_avg, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_2km_avg, lw_avg)
```
</div>
No evidence of autocorrelation of the residuals.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 2km avg concurvity, echo=FALSE}
concurvity(total_mod_2km_avg, full = FALSE)
```
</div>
Some of the variables appear to have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. May need to get
rid of the mean apparent fishing hours however. Since the concurvity was very
high.


### Building GAMs to test the effects on the Urchin to snail ratio (without mean apparent fishing hours) {.tabset}

Redoing the GAMs without the mean apparent fishing hours because the concurvity 
is very high.

#### Model with 1 km recreational boats (no mean app hours)

```{r Total 1km model no app hours with no spatial autocorrelation term}
total_mod_1km_no_app_hours <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_1km_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```

```{r 1km no app hours model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1km_no_app_hours)

plot(total_mod_1km_no_app_hours)
summary(total_mod_1km_no_app_hours)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1km no app hours residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1km_no_app_hours <- residuals(total_mod_1km_no_app_hours, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)

nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1km_no_app_hours, lw)
```
</div>
It appears that the residuals are autocorrelated. Will need to address this.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1km no app hours concurvity, echo=FALSE}
concurvity(total_mod_1km_no_app_hours, full = FALSE)
```
</div>

Some of the variables appear to still have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. 



#### Model with 1.5 km recreational boats (no mean app hours)

```{r Total 1.5km model no app hours with no spatial autocorrelation term}
total_mod_1.5km_no_app_hours <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_1.5km_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```

```{r 1.5 km no app hours model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1.5km_no_app_hours)

plot(total_mod_1.5km_no_app_hours, residuals = TRUE)
summary(total_mod_1.5km_no_app_hours)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1.5km no app hours residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1.5km_no_app_hours <- residuals(total_mod_1.5km_no_app_hours, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)

nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1.5km_no_app_hours, lw)
```
</div>
It appears that the residuals are autocorrelated. Will need to address this.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1.5km no app hours concurvity, echo=FALSE}
concurvity(total_mod_1.5km_no_app_hours, full = FALSE)
```
</div>
Some of the variables appear to still have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots.

#### Model with 2 km recreational boats (no mean app hours)

```{r Total 2km no app hours model with no spatial autocorrelation}
total_mod_2km_no_app_hours <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_2km_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```

```{r 2 km no app hours model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_2km_no_app_hours)


plot(total_mod_2km_no_app_hours)
summary(total_mod_2km_no_app_hours)
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 2km no app hours residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_2km_no_app_hours <- residuals(total_mod_2km_no_app_hours, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)

nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10) 
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_2km_no_app_hours, lw)
```
</div>
It appears that the residuals are autocorrelated. Will need to address this.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 2km no app hours concurvity, echo=FALSE}
concurvity(total_mod_2km_no_app_hours, full = FALSE)
```
</div>
Some of the variables appear to still have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. 



### Building GAMs with averaged points to test the effects on the urchin to snail ratio (without mean apparent fishing hours) {.tabset}

The idea of averaging the transects at the same sites is to account for the 
spatial autocorrelation.

#### Model with 1km recreational boats and avg transects (no app hours)

```{r Total 1km avg no app hours model with no spatial autocorrelation}
total_mod_1km_avg_no_app_hours <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_1km_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```

```{r 1km avg no app hours model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1km_avg_no_app_hours)

plot(total_mod_1km_avg_no_app_hours)
summary(total_mod_1km_avg_no_app_hours)
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1km avg no app hours residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1km_avg_no_app_hours <- residuals(total_mod_1km_avg_no_app_hours, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1km_avg_no_app_hours, lw_avg)
```
</div>
No evidence of autocorrelation of the residuals.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1km avg no app hours concurvity, echo=FALSE}
concurvity(total_mod_1km_avg_no_app_hours, full = FALSE)
```
</div>
Some of the variables appear to still have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots. 

#### Model with 1.5km recreational boats and avg transects (no app hours)

```{r Total 1.5km avg no app hours model with no spatial autocorrelation}
total_mod_1.5km_avg_no_app_hours <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_1.5km_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```


```{r 1.5 km avg no app hours model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_1.5km_avg_no_app_hours)

plot(total_mod_1.5km_avg_no_app_hours)
summary(total_mod_1.5km_avg_no_app_hours)
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 1.5km avg no app hours residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_1.5km_avg_no_app_hours <- residuals(total_mod_1.5km_avg_no_app_hours, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1.5km_avg_no_app_hours, lw_avg)
```
</div>
No evidence of autocorrelation of the residuals.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 1.5km avg no app hours concurvity, echo=FALSE}
concurvity(total_mod_1.5km_avg_no_app_hours, full = FALSE)
```
</div>
Some of the variables appear to still have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots.

#### Model with 2km recreational boats and avg transects (no mean app hours)

```{r Total 2km avg no app hours model with no spatial autocorrelation}
total_mod_2km_avg_no_app_hours <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_2km_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)
```

```{r 2 km avg no app hours model output, echo=FALSE}
set.seed(13)
gam.check(total_mod_2km_avg_no_app_hours)

plot(total_mod_2km_avg_no_app_hours)
summary(total_mod_2km_avg_no_app_hours)
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing 2km no app hours avg residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_2km_avg_no_app_hours <- residuals(total_mod_2km_avg_no_app_hours, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_avg <- st_transform(coords_for_auto_test_avg, crs = 3857)
extracted_coords_for_auto_test_projected_avg <- st_coordinates(coords_for_auto_test_projected_avg)

nb_avg <- dnearneigh(extracted_coords_for_auto_test_projected_avg, d1 = 0, d2 = 1000)
#picked 1000 m for distance so that it has some neighbours for enough sites

lw_avg <- nb2listw(nb_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_2km_avg_no_app_hours, lw_avg)
```
</div>
No evidence of autocorrelation of the residuals.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing 2km avg no app hours concurvity, echo=FALSE}
concurvity(total_mod_2km_avg_no_app_hours, full = FALSE)
```
</div>
Some of the variables appear to still have high concurvity scores in the worst case,
but confidence intervals seem stable for the model plots.