---
title: "RLS Modelling"
author: "Alec"
date: "2024-11-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(mgcv)
library(DHARMa)
library(mgcViz)
library(ggeffects)
library(gratia)
library(geoR)
library(geosphere)
library(gstat)
library(sf)
library(car)
library(MuMIn)
library(lme4)
library(flextable)
library(spdep)
```

Jan. 22. I have updated the data with the flights with fog ommitted from the data.
If this results in large problems the original data is available in the backup csv.
Except for the total boats columns but if these are needed then i can just reload the 
DFO data and not exclude the flights with fog.


#Modelling the environmental variables

##Loading in the data
```{r Loading in the data}
env_mod_data <- read_csv(here("./03_clean_data/urch_sn_ratio_mod_data.csv"))
```

```{r Scaling the explanatory variables and converting substrate type to factor}
env_mod_data$sst_scale <- scale(env_mod_data$sst)
env_mod_data$openness_scale <- scale(env_mod_data$openness)
env_mod_data$norm_openness_scale <- scale(env_mod_data$norm_openness)
env_mod_data$mean_app_fishing_hours_scale <- scale(env_mod_data$mean_app_fishing_hours)
env_mod_data$mean_rec_boat_counts_500m_scale <- scale(env_mod_data$mean_rec_boat_counts_500m)
env_mod_data$mean_rec_boat_counts_1km_scale <- scale(env_mod_data$mean_rec_boat_counts_1km)
env_mod_data$mean_rec_boat_counts_1.5km_scale <- scale(env_mod_data$mean_rec_boat_counts_1.5km)
env_mod_data$mean_rec_boat_counts_750m_scale <- scale(env_mod_data$mean_rec_boat_counts_750m)
env_mod_data$mean_rec_boat_counts_2km_scale <- scale(env_mod_data$mean_rec_boat_counts_2km)
env_mod_data$dominant_score <- as.factor(env_mod_data$dominant_score)
env_mod_data$site_code <- as.factor(env_mod_data$site_code)
env_mod_data$Urchins_scale <- scale(env_mod_data$Urchins)
env_mod_data$Turban_Snails_scale <- scale(env_mod_data$Turban_Snails)
#Testing out the total boat counts
env_mod_data$total_boat_counts_1.5km_scale <- scale(env_mod_data$mean_total_boat_counts_1.5km)
levels(env_mod_data$dominant_score) <- list(loose = "3", solid = "4")

regions <- c("region1", "region2", "region3", "region4")
env_mod_data[regions] <- lapply(env_mod_data[regions], factor)
```


```{r Creating a dataset with averaged values for urchin/snail counts, and urch sn ratio}
env_mod_data_avg <- env_mod_data%>%
  group_by(site_code, site_name, latitude, longitude, sst,
           openness, norm_openness, region1, region2, region3, region4, 
           mean_app_fishing_hours, mean_rec_boat_counts_500m, mean_rec_boat_counts_1km, 
           mean_rec_boat_counts_1.5km, mean_rec_boat_counts_750m, mean_rec_boat_counts_2km,
           sst_scale, openness_scale, norm_openness_scale, mean_app_fishing_hours_scale,
           mean_rec_boat_counts_500m_scale, mean_rec_boat_counts_1km_scale, 
           mean_rec_boat_counts_1.5km_scale, mean_rec_boat_counts_750m_scale,
           mean_rec_boat_counts_2km_scale)%>%
  summarise(urchins_avg = mean(Urchins),
            snails_avg = mean(Turban_Snails),
            urch_sn_ratio_avg = mean(urch_sn_ratio),
            depth_avg = mean(depth))
```

```{r Creating a subset dataset for the sites with urchin size info}
env_mod_data_biomass <- urch_sn_ratio_mod_2024%>%
  left_join(max_biomass_df, by = c("site_code", "depth"))%>%
  #left_join(inbetween_biomass_df, c("site_code", "depth"))%>%  #the inbetween categories are wrong because they start between 0 and 2.5
    #filter(!site_code %in% c("BMSC15", "BMSC26"))%>%
  replace(is.na(.), 0)%>%
  select(-c(site_name.y, latitude.x, longitude.x, visibility.y, hour.y))


```

```{r Writing as a csv}
write_csv(env_mod_data_biomass,
          here("./03_clean_data/urchin_biomass_data_updated_jan_25.csv"))
```


###Testing the data to see if it is spatially correlated

####Moran's I with transects at same site separate
```{r Using Moran's I test on the data with geosphere}
##Method setting transects at same site to 1 m apart

sphere_coords <- cbind(env_mod_data$longitude, env_mod_data$latitude)
sphere_dist <- distm(sphere_coords, fun = distGeo)

sphere_dist[sphere_dist == 0] <- 1
inv_sphere_dist <- 1 / sphere_dist
diag(inv_sphere_dist) <- 0
  
ape::Moran.I(env_mod_data$urch_sn_ratio, weight = inv_sphere_dist, scaled = TRUE)

ape::Moran.I(total_mod_1.5km_data_with_resids$`total_mod_1.5km$residuals`, weight = inv_sphere_dist, scaled = TRUE)
```

####Moran's I with transects at same site averaged
```{r Making a df with averaged urch_sn ratios}
urch_sn_ratio_averaged <- env_mod_data%>%
  group_by(site_code, latitude, longitude)%>%
  summarise(avg_urch_sn_ratio = mean(urch_sn_ratio))
```

```{r calculating Moran's I with the averaged points}
sphere_coords_avg <- cbind(urch_sn_ratio_averaged$longitude, urch_sn_ratio_averaged$latitude)

sphere_dist_avg <- distm(sphere_coords_avg, fun = distGeo)

inv_sphere_dist_avg <- 1 / sphere_dist_avg

diag(inv_sphere_dist_avg) <- 0

ape::Moran.I(urch_sn_ratio_averaged$avg_urch_sn_ratio, weight = inv_sphere_dist_avg, scaled = TRUE)
```



####Using a semi-variogram to see what type of spatial autocorrelation factor to use

#####With averaged points
```{r Transforming the coordinates into meters for the variogram}
# Using the averaged data
urch_sn_ratio_averaged_sf <- st_as_sf(urch_sn_ratio_averaged, coords = c("longitude", "latitude"), crs = 4326)
urch_sn_ratio_averaged_projected <- st_transform(urch_sn_ratio_averaged_sf, crs = 3857)
```

```{r Plotting a semi-variogram with the meter distances for the averaged transects}
vario_meters <- variogram(avg_urch_sn_ratio ~ 1, data = urch_sn_ratio_averaged_projected)
vario_meters_df <- as.data.frame(vario_meters)
plot(vario_meters)

#ggplot(vario_meters_df, aes(x = dist, y = gamma)) +
 # geom_line() +
  #geom_point() +
  #scale_x_continuous(breaks = seq(0, 15000, by = 1000)) + # Custom x-axis ticks
  #labs(x = "Distance (m)", y = "Semivariance", title = "Variogram") +
  #theme_minimal()

```

```{r Fitting a variogram model}
vario_meters_mod <- fit.variogram(vario_meters, 
                                  model = vgm(nugget = 0.1, range = 2000, psill = 0.75, 
                                              model = c("Exp", "Sph", "Gau")),
                                  fit.sills = TRUE, fit.ranges = TRUE)
vario_meters_mod
plot(vario_meters, vario_meters_mod)
```

#####With non averaged points

```{r Variogram with the non-averaged points}
urch_sn_coords_var_2 <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
urch_sn_coords_var_2_projected <- st_transform(urch_sn_coords_var_2, crs = 3857)

vario_2 <- variogram(urch_sn_ratio ~ 1, data = urch_sn_coords_var_2_projected)
vario_2_df <- as.data.frame(vario_2)
plot(vario_2)

#ggplot(vario_2_df, aes(x = dist, y = gamma)) +
 # geom_line() +
#  geom_point() +
 # scale_x_continuous(breaks = seq(0, 15000, by = 1000)) + # Custom x-axis ticks
  #labs(x = "Distance (m)", y = "Semivariance", title = "Variogram") +
  #theme_minimal()
```

```{r Fitting a variogram model}
vario_2_mod <- fit.variogram(vario_2,
                             model = vgm(1, model = c("Exp", "Sph", "Gau"), 
                                         nugget = 0.1, range = 1000, psill = 1),
                             fit.sills = TRUE, fit.ranges = TRUE)
vario_2_mod
plot(vario_2, vario_2_mod)
```


#####With shifted points
In order to use the corSpatial functions within a gamm there cannot be sites at the exact same lat and long so 
i have shifted the points at the same site by one meter or so to make the function run.

```{r Making a dataframe with sites at the same location shifted slightly}
env_mod_data_shifted <- env_mod_data%>%
    mutate(latitude = if_else(row_number() == 2, 48.82896, latitude))%>%
  mutate(latitude = if_else(row_number() == 5, 48.85746, latitude))%>%
  mutate(latitude = if_else(row_number() == 7, 48.85827, latitude))%>%
  mutate(latitude = if_else(row_number() == 21, 48.91608, latitude))%>%
   mutate(latitude = if_else(row_number() == 23, 48.83858, latitude))%>%
   mutate(latitude = if_else(row_number() == 25, 48.90711, latitude))%>%
  mutate(latitude = if_else(row_number() == 30, 48.82734, latitude))%>%
  mutate(latitude = if_else(row_number() == 32, 48.95024, latitude))%>%
   mutate(latitude = if_else(row_number() == 34, 48.95509, latitude))
```

```{r Converting the shifted points into meters}
env_mod_sf <- st_as_sf(env_mod_data_shifted, coords = c("longitude", "latitude"), crs = 4326)
env_mod_sf_projected <- st_transform(env_mod_sf, crs = 3857)

projected_coords <- st_coordinates(env_mod_sf_projected)
env_mod_data_shifted$longitude_proj <- projected_coords[ , 1]
env_mod_data_shifted$latitude_proj <- projected_coords[ , 2]
```

```{r Plotting an experimental variogram for the data that is shifted by one meter}
vario_shifted <- variogram(urch_sn_ratio ~ 1, data = env_mod_sf_projected)
plot(vario_shifted)
```

```{r Fitting a variogram model to the experimental variogram}
vario_shifted_mod <- fit.variogram(vario_shifted, model = vgm(nugget = 1, model = c("Exp", "Sph", "Gau"), range = 1000, psill = 1), fit.sills = TRUE, fit.ranges = TRUE)
vario_shifted_mod
plot(vario_shifted, vario_shifted_mod)
```

##Plotting EVs against urch sn ratio

```{r ratio against substrate score}
ggplot(env_mod_data, aes(x = dominant_score, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against openness}
ggplot(env_mod_data, aes(x = norm_openness, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against sst}
ggplot(env_mod_data, aes(x = sst, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against mean rec boat counts 1.5km}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_1.5km, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against mean rec boat counts 1km}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_1km, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r ratio against mean rec boat counts 2km}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_2km, y = urch_sn_ratio, colour = region1))+
  geom_point()+
  theme.figure()
```

```{r}
ggplot(env_mod_data, aes(x = mean_app_fishing_hours, y = urch_sn_ratio))+
  geom_point()
```

```{r}
ggplot(env_mod_data, aes(x = region1, y = mean_rec_boat_counts_1.5km))+
  geom_boxplot()
```


start with env models of counts
then 


##Building the GAM

###No spatial autocorrelation term and using norm openness

####temp scaled, norm openness scaled, dominant score, TRUE

```{r Making a gam with no spatial autocorrelation term to start}
env_mod_T <- gam(urch_sn_ratio ~ s(sst_scale) + s(norm_openness_scale) + dominant_score,
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(env_mod_T)

plot(env_mod_T)
```

```{r Checking the model fit with DHARMa}
env_mod_T_resids <- simulateResiduals(env_mod_T)
plot(env_mod_T_resids)
```



#####temp, norm openness scale, TRUE

```{r Making a gam with no spatial autocorrelation term to start}
env_mod_no_dom <- gam(urch_sn_ratio ~ s(sst_scale) + s(norm_openness_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(env_mod_no_dom)

plot(env_mod_no_dom)
```

```{r Checking the model fit with DHARMa}
env_mod_no_dom_resids <- simulateResiduals(env_mod_no_dom)
plot(env_mod_no_dom_resids)
```
Problem when running the model without the substrate type, Why?

####Smaller models within the overall environmental one
Gavin said maybe this is not the best way to do it as you could be biasing your
p values since summary doesn't know you already did selection.

Will need to update to norm openness scaled

#####temp, norm openness, dominant score, FALSE
```{r Making a gam with no spatial autocorrelation term to start}
env_mod_F <- gam(urch_sn_ratio ~ s(sst_scale) + s(norm_openness) + dominant_score,
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = FALSE)

set.seed(13)
gam.check(env_mod_F)

plot(env_mod_F)
```


```{r Checking the model fit with DHARMa}
env_mod_F_resids <- simulateResiduals(env_mod_F)
plot(env_mod_F_resids)
```
All good


#####temp, dominant score, FALSE
```{r Making a gam with no spatial autocorrelation and no fetch}
env_mod_no_fetch <- gam(urch_sn_ratio ~ s(sst_scale) + dominant_score, data = env_mod_data,
                        method = 'REML',
                        family = gaussian(link = "identity"),
                        select = FALSE)

gam.check(env_mod_no_fetch, rep = 500)
plot(env_mod_no_fetch)
```

```{r}
env_mod_no_fetch_resid <- simulateResiduals(env_mod_no_fetch)
plot(env_mod_no_fetch_resid)
```
Possible problem with this model according to DHARMa

#####norm openness scaled, dominant score, FALSE
```{r Making a gam with no spatial autocorrelation and no temperature}
env_mod_no_temp <- gam(urch_sn_ratio ~ s(norm_openness_scale) + dominant_score, data = env_mod_data,
                        method = 'REML',
                        family = gaussian(link = "identity"),
                        select = FALSE)

gam.check(env_mod_no_temp, rep = 500)
plot(env_mod_no_temp)
```

```{r}
env_mod_no_temp_resid <- simulateResiduals(env_mod_no_temp)
plot(env_mod_no_temp_resid)
```



###With spatial autocorrelation term

####temp scaled, norm openness scaled, dom, lat long gp interaction, TRUE

```{r Building the gam with the spatial autocorrelation term}
####This is likely overfitting the model and just saying that a smooth of lat and long explains the variation in urch ratio

env_mod_sp_gp <- gam(urchin_dens ~ s(sst_scale) + s(norm_openness_scale) + dominant_score + s(latitude, longitude, bs = "gp", k = 10),
                  data = env_mod_data, 
                  family = gaussian(link = "identity"),
                  method = 'REML',
                  select = TRUE)   #In this case I have added the lat long as an interaction term directly in the model

set.seed(10)
gam.check(env_mod_sp_gp, rep = 500)
plot(env_mod_sp_gp)
summary(env_mod_sp_gp)
```

```{r DHARMa for model fit}
env_mod_sp_gp_resid <- simulateResiduals(env_mod_sp_gp)
plot(env_mod_sp_gp_resid)
```
All good


####temp scaled, norm openness scaled, dom, lat long gp interaction, TRUE, shifted data point

```{r Building the gam with the spatial autocorrelation term and shifted points}
env_mod_sp_gp_shifted <- gam(urch_sn_ratio ~ s(sst_scale) + s(norm_openness_scale) + dominant_score + s(latitude, longitude, bs = "gp", k = 10),
                  data = env_mod_data_shifted, 
                  family = gaussian(link = "identity"),
                  method = 'REML',
                  select = TRUE)   #In this case I have added the lat long as an interaction term directly in the model

set.seed(10)
gam.check(env_mod_sp_gp, rep = 500)
plot(env_mod_sp_gp)
```

```{r DHARMa for model fit}
env_mod_sp_gp_shifted_resid <- simulateResiduals(env_mod_sp_gp_shifted)
plot(env_mod_sp_gp_shifted_resid)
```
All good

####gamm, temp scaled, norm openness scaled, dom, corExp, TRUE shifted points

```{r Building gamm with spatial autocorrelation term corExp}
env_mod_sp_cor_exp <- gamm(urch_sn_ratio ~ s(sst_scale) + s(norm_openness_scale) + dominant_score, 
                        data = env_mod_data_shifted,
                       family = gaussian(link = "identity"),
                       correlation = corExp(form = ~ longitude_proj + latitude_proj, nugget = FALSE, value = 639), 
                       method = 'REML',
                       select = TRUE)


gam.check(env_mod_sp_cor_exp$gam, rep = 500)
plot(env_mod_sp_cor_exp$gam)
```

```{r DHARMa for model fit}
env_mod_sp_cor_exp_resid <- simulateResiduals(env_mod_sp_cor_exp$lme)
plot(env_mod_sp_cor_exp_resid)
```
Problem with DHARMa for gamms

####gam, temp scaled, norm openness scaled, dom, re of site code, TRUE

```{r Building gam with random effect of site to account for spatial autocorrelation}
env_mod_re_site_code <- gam(urch_sn_ratio ~ s(sst_scale) + s(norm_openness_scale) + dominant_score + s(site_code, bs = "re"), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'REML',
                       select = TRUE)

gam.check(env_mod_re_site_code, rep = 500)
plot(env_mod_re_site_code)
```

```{r DHARMa for model fit}
env_mod_re_site_code_resid <- simulateResiduals(env_mod_re_site_code)
plot(env_mod_re_site_code_resid)
```


####gam, temp scaled, norm openness scaled, dom, random effect of region M1, TRUE

```{r Building a gam with random effect of region M1}
env_mod_re_region_M1 <- gam(urch_sn_ratio ~ s(sst_scale) + s(norm_openness_scale) + dominant_score + s(region1, bs = "re"), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'REML',
                       select = TRUE)

gam.check(env_mod_re_region_M1, rep = 500)
plot(env_mod_re_region_M1)
```

```{r DHARMa for model fit}
env_mod_re_region_M1_resid <- simulateResiduals(env_mod_re_region_M1)
plot(env_mod_re_region_M1_resid)
```

####gam, temp scaled, norm openness scaled, dom, random effect of region M2, TRUE

```{r Building a gam with random effect of region M2}
env_mod_re_region_M2 <- gam(urch_sn_ratio ~ s(sst_scale) + s(norm_openness_scale) + dominant_score + s(region2, bs = "re"), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'REML',
                       select = TRUE)

gam.check(env_mod_re_region_M2, rep = 500)
plot(env_mod_re_region_M2)
```

```{r DHARMa for model fit}
env_mod_re_region_M2_resid <- simulateResiduals(env_mod_re_region_M2)
plot(env_mod_re_region_M2_resid)
```


####gam, temp scaled, norm openness scaled, dom, random effect of region M3, TRUE

```{r Building a gam with random effect of region M3}
env_mod_re_region_M3 <- gam(urch_sn_ratio ~ s(sst_scale) + s(norm_openness_scale) + dominant_score + s(region3, bs = "re"), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'REML',
                       select = TRUE)

gam.check(env_mod_re_region_M3, rep = 500)
plot(env_mod_re_region_M3)
```

```{r DHARMa for model fit}
env_mod_re_region_M3_resid <- simulateResiduals(env_mod_re_region_M3)
plot(env_mod_re_region_M3_resid)
```

####gam, temp scaled, norm openness scaled, dom, random effect of region M4, TRUE

```{r Building a gam with random effect of region M3}
env_mod_re_region_M4 <- gam(urch_sn_ratio ~ s(sst_scale) + s(norm_openness_scale) + dominant_score + s(region4, bs = "re"), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'REML',
                       select = TRUE)

gam.check(env_mod_re_region_M4, rep = 500)
plot(env_mod_re_region_M4)
```
```{r DHARMa for model fit}
env_mod_re_region_M4_resid <- simulateResiduals(env_mod_re_region_M4)
plot(env_mod_re_region_M4_resid)
```


###Testing the models for concurvity of variables

####env_mod_T

```{r testing for multicollinearity}
concurvity(env_mod_T, full = TRUE)
```
```{r }
concurvity(env_mod_T, full = FALSE)
```

```{r}
ggplot(env_mod_data, aes(x = sst_scale, y = norm_openness_scale))+
  geom_point()
```
####env_mod_sp_cor_exp

```{r}
concurvity(env_mod_sp_cor_exp$gam, full = TRUE)
concurvity(env_mod_sp_cor_exp$gam, full = FALSE)
```
####env_mod_re_site_code

```{r}
concurvity(env_mod_re_site_code, full = FALSE)
```

####env_mod_re_region_M1

```{r}
concurvity(env_mod_re_region_M1, full = TRUE)
concurvity(env_mod_re_region_M1, full = FALSE)
```

####env_mod_re_region_M2

```{r}
concurvity(env_mod_re_region_M2, full = TRUE)
concurvity(env_mod_re_region_M2, full = FALSE)
```

####env_mod_re_region_M3

```{r}
concurvity(env_mod_re_region_M3, full = TRUE)
concurvity(env_mod_re_region_M3, full = FALSE)
```

####env_mod_re_region_M4

```{r}
concurvity(env_mod_re_region_M4, full = TRUE)
concurvity(env_mod_re_region_M4, full = FALSE)
```

###AIC with the different models

```{r Comparing the models with an AIC}
AIC(env_mod_T,
    #env_mod_sp_gp, 
    #env_mod_sp_gp_shifted,
    env_mod_sp_cor_exp$lme,
    env_mod_re_site_code,
    env_mod_re_region_M1,
    env_mod_re_region_M2, 
    env_mod_re_region_M3,
    env_mod_re_region_M4
    )%>%
arrange(AIC)%>%
  mutate(delta_AIC = AIC - AIC[1])
```

###Trying with fishing variables now

```{r Plot of urch sn ratio vs rec boats}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_750m, y = urch_sn_ratio))+
  geom_point()
```

```{r plot of urch sn ratio vs rec boat counts}
ggplot(env_mod_data, aes(x = mean_rec_boat_counts_500m, y = urch_sn_ratio))+
  geom_point()
```


####gam, rec boats 500m scaled, mean gfw scaled, TRUE
```{r Building a gam with fishing variables}
fish_mod <- gam(urch_sn_ratio ~ s(mean_rec_boat_counts_500m_scale) + s(mean_app_fishing_hours_scale), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'ML',
                       select = TRUE)

gam.check(fish_mod, rep = 500)
plot(fish_mod)
```

```{r DHARMa for model fit}
fish_mod_resids <- simulateResiduals(fish_mod)
plot(fish_mod_resids)
```

####gam, rec boats 750m scaled, mean gfw scaled, TRUE
```{r Building a gam with fishing variables}
fish_mod_750m <- gam(urch_sn_ratio ~ s(mean_rec_boat_counts_750m_scale) + s(mean_app_fishing_hours_scale), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'REML',
                       select = TRUE)

gam.check(fish_mod_750m, rep = 500)
plot(fish_mod_750m)
```

```{r DHARMa for model fit}
fish_mod_750m_resids <- simulateResiduals(fish_mod_750m)
plot(fish_mod_750m_resids)
```

####gam, rec boats 1km scaled, mean gfw scaled, TRUE
```{r Building a gam with fishing variables}
fish_mod_1km <- gam(urch_sn_ratio ~ s(mean_rec_boat_counts_1km_scale) + s(mean_app_fishing_hours_scale), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'ML',
                       select = TRUE)

gam.check(fish_mod_1km, rep = 500)
plot(fish_mod_1km)
```

```{r DHARMa for model fit}
fish_mod_1km_resids <- simulateResiduals(fish_mod_1km)
plot(fish_mod_1km_resids)
```

####gam, rec boats 1.5km scaled, mean gfw scaled, TRUE

```{r Building a gam with fishing variables}
fish_mod_1.5km <- gam(urch_sn_ratio ~ s(mean_rec_boat_counts_1.5km_scale) + s(mean_app_fishing_hours_scale), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'ML',
                       select = TRUE)

gam.check(fish_mod_1.5km, rep = 500)
plot(fish_mod_1.5km)
```

```{r DHARMa for model fit}
fish_mod_1.5km_resids <- simulateResiduals(fish_mod_1.5km)
plot(fish_mod_1.5km_resids)
```

####gam, rec boats 2km scaled, mean gfw scaled, TRUE

```{r Building a gam with fishing variables}
fish_mod_2km <- gam(urch_sn_ratio ~ s(mean_rec_boat_counts_2km_scale) + s(mean_app_fishing_hours_scale), 
                        data = env_mod_data,
                       family = gaussian(link = "identity"),
                       method = 'ML',
                       select = TRUE)

gam.check(fish_mod_2km, rep = 500)
plot(fish_mod_2km)
summary(fish_mod_2km)
```

```{r DHARMa for model fit}
fish_mod_2km_resids <- simulateResiduals(fish_mod_2km)
plot(fish_mod_2km_resids)
```



###AIC of fishing models

```{r Comparing the models with an AIC}
AIC(fish_mod,
    fish_mod_750m,
    fish_mod_1km,
    fish_mod_1.5km,
    fish_mod_2km
    )%>%
arrange(AIC)%>%
  mutate(delta_AIC = AIC - AIC[1])
```
indicates similar effects with adding distance ---- some problems with shorter distances as there are more small counts

###Total mod with all variables

####gam, temp scaled, norm openness scaled, dom, rec boats 500m scaled, mean gfw scaled, TRUE

```{r Total model with no spatial autocorrelation}
total_mod <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   dominant_score + 
                   s(mean_rec_boat_counts_500m_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod)

plot(total_mod)
```

```{r DHARMa for model fit}
total_mod_resids <- simulateResiduals(total_mod)
plot(total_mod_resids)
```

####gam, temp scaled, norm openness scaled, dom, rec boats 500m scaled, mean gfw scaled, re of region M1, TRUE

```{r Total model with random effect of region 1}
total_mod_re_region1 <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   dominant_score + 
                   s(mean_rec_boat_counts_500m_scale)+
                   s(mean_app_fishing_hours_scale)+
                   s(region1, bs = "re"),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_re_region1)

plot(total_mod_re_region1)
```

```{r DHARMa for model fit}
total_mod_re_region1_resids <- simulateResiduals(total_mod_re_region1)
plot(total_mod_re_region1_resids)
```

####gam, temp scaled, norm openness scaled, dom, rec boats 500m scaled, mean gfw scaled, re of region M2, TRUE

```{r Total model with random effect of region 2}
total_mod_re_region2 <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   dominant_score + 
                   s(mean_rec_boat_counts_500m_scale)+
                   s(mean_app_fishing_hours_scale)+
                   s(region2, bs = "re"),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_re_region2)

plot(total_mod_re_region2)
```

```{r DHARMa for model fit}
total_mod_re_region2_resids <- simulateResiduals(total_mod_re_region2)
plot(total_mod_re_region2_resids)
```



####gam, temp scaled, norm openness scaled, dom, rec boats 2km scaled, mean gfw scaled, TRUE

```{r Total model with no spatial autocorrelation}
total_mod_2km <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                  # dominant_score + 
                   s(mean_rec_boat_counts_2km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_2km)
#shapiro.test(total_mod_2km$residuals)
##normal

plot(total_mod_2km)
summary(total_mod_2km)
as_flextable(total_mod_2km)
```

```{r DHARMa for model fit}
total_mod_2km_resids <- simulateResiduals(total_mod_2km)
plot(total_mod_2km_resids)
```

```{r Testing residuals for spatial correlation}
#extract residuals from model
residuals_total_mod_2km <- residuals(total_mod_2km, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_2km <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_2km <- st_transform(coords_for_auto_test_2km, crs = 3857)
extracted_coords_for_auto_test_projected_2km <- st_coordinates(coords_for_auto_test_projected_2km)

nb_2km <- dnearneigh(extracted_coords_for_auto_test_projected_2km, d1 = 0, d2 = 10)
lw_2km <- nb2listw(nb_2km, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_2km, lw_2km)


###testing with inverse distance matrix to ensure that the Moran's I value is accurate
dist_matrix_total_mod_2km <- as.matrix(dist(extracted_coords_for_auto_test_projected_2km))
inverse_dist_matrix_total_mod_2km <- 1 / (dist_matrix_total_mod_2km + 0.001)
diag(inverse_dist_matrix_total_mod_2km) <- 0
lw_2km_inverse <- mat2listw(inverse_dist_matrix_total_mod_2km, style = "W")
moran.test(residuals_total_mod_2km, lw_2km_inverse)
```


```{r Testing concurvity}
concurvity(total_mod_2km, full = FALSE)
```

#####gam, subset of total 2km without stuff that has no impact

```{r}
total_mod_2km_no_extra <- gam(urch_sn_ratio ~ s(sst_scale)+
                                s(norm_openness_scale)+
                   s(mean_rec_boat_counts_2km_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_2km_no_extra)
#shapiro.test(total_mod_2km_no_extra$residuals)
##normal

plot(total_mod_2km_no_extra)
summary(total_mod_2km_no_extra)
as_flextable(total_mod_2km_no_extra)
```

```{r DHARMa for model fit}
total_mod_2km_no_extra_resids <- simulateResiduals(total_mod_2km_no_extra)
plot(total_mod_2km_no_extra_resids)
```

```{r Testing concurvity}
concurvity(total_mod_2km_no_extra, full = FALSE)
```




####gam, temp scaled, norm openness scaled, dom, rec boats 1km scaled, mean gfw scaled, TRUE

```{r Total model with no spatial autocorrelation}
total_mod_1km <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                  dominant_score + 
                   s(mean_rec_boat_counts_1km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'ML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_1km)
#shapiro.test(total_mod_1km$residuals)
##slightly not normal

plot(total_mod_1km)
summary(total_mod_1km)
#as_flextable(total_mod_1km)
```

```{r DHARMa for model fit}
total_mod_1km_resids <- simulateResiduals(total_mod_1km)
plot(total_mod_1km_resids)
```

```{r Testing concurvity}
concurvity(total_mod_1km, full = FALSE)
```

```{r Testing residuals for spatial correlation}
#extract residuals from model
residuals_total_mod_1km <- residuals(total_mod_1km, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_1km <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_1km <- st_transform(coords_for_auto_test_1km, crs = 3857)
extracted_coords_for_auto_test_projected_1km <- st_coordinates(coords_for_auto_test_projected_1km)

nb_1km <- dnearneigh(extracted_coords_for_auto_test_projected_1km, d1 = 0, d2 = 10)
lw_1km <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1km, lw_1km)
```


```{r Total model with no spatial autocorrelation}
total_mod_1km_no_dom <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                  #dominant_score + 
                   s(mean_rec_boat_counts_1km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'ML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_1km)
#shapiro.test(total_mod_1km$residuals)
##slightly not normal

plot(total_mod_1km)
summary(total_mod_1km)
#as_flextable(total_mod_1km)
```

```{r}
AIC(total_mod_1km,
    total_mod_1km_no_dom)
```



#####gam, subset of total 1km without stuff that has no impact

```{r}
total_mod_1km_no_extra <- gam(urch_sn_ratio ~ s(sst_scale)+
                                s(norm_openness_scale)+
                   s(mean_rec_boat_counts_1km_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_1km_no_extra)
#shapiro.test(total_mod_1km_no_extra$residuals)
##normal

plot(total_mod_1km_no_extra)
summary(total_mod_1km_no_extra)
as_flextable(total_mod_1km_no_extra)
```

```{r DHARMa for model fit}
total_mod_1km_no_extra_resids <- simulateResiduals(total_mod_1km_no_extra)
plot(total_mod_1km_no_extra_resids)
```

```{r Testing concurvity}
concurvity(total_mod_1km_no_extra, full = FALSE)
```


####gam, temp scaled, norm openness scaled, dom, rec boats 1.5km scaled, mean gfw scaled, TRUE

```{r Total model with no spatial autocorrelation}
total_mod_1.5km <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                  # dominant_score + 
                   s(mean_rec_boat_counts_1.5km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_1.5km)
shapiro.test(total_mod_1.5km$residuals)
##slightly not normal

plot(total_mod_1.5km, residuals = TRUE)
summary(total_mod_1.5km)
#as_flextable(total_mod_1.5km)
```


```{r Total model with no spatial autocorrelation}
total_mod_1.5km_total_boats <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   dominant_score + 
                   s(total_boat_counts_1.5km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_1.5km_total_boats)
shapiro.test(total_mod_1.5km_total_boats$residuals)
##slightly not normal

plot(total_mod_1.5km_total_boats, residuals = TRUE)
summary(total_mod_1.5km_total_boats)
#as_flextable(total_mod_1.5km)
```


```{r DHARMa for model fit}
total_mod_1.5km_resids <- simulateResiduals(total_mod_1.5km)
plot(total_mod_1.5km_resids)
```

```{r}
total_mod_1.5km_data_with_resids <- cbind(env_mod_data, total_mod_1.5km$residuals)
```


```{r Testing concurvity}
concurvity(total_mod_1.5km, full = FALSE)
```

```{r Testing residuals for spatial correlation}
#extract residuals from model
residuals_total_mod_1.5km <- residuals(total_mod_1.5km, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_1.5km <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_1.5km <- st_transform(coords_for_auto_test_1.5km, crs = 3857)
extracted_coords_for_auto_test_projected_1.5km <- st_coordinates(coords_for_auto_test_projected_1.5km)

nb_1.5km <- dnearneigh(extracted_coords_for_auto_test_projected_1.5km, d1 = 0, d2 = 10)
lw_1.5km <- nb2listw(nb_1.5km, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1.5km, lw_1.5km)
```

#####gam, subset of total 1.5km without stuff that has no impact

```{r}
total_mod_1.5km_no_extra <- gam(urch_sn_ratio ~ s(sst_scale)+
                                s(norm_openness_scale)+
                   s(mean_rec_boat_counts_1.5km_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_1.5km_no_extra)
#shapiro.test(total_mod_1.5km_no_extra$residuals)
##slightly not normal

plot(total_mod_1.5km_no_extra)
summary(total_mod_1.5km_no_extra)
as_flextable(total_mod_1.5km_no_extra)
```

```{r DHARMa for model fit}
total_mod_1.5km_no_extra_resids <- simulateResiduals(total_mod_1.5km_no_extra)
plot(total_mod_1.5km_no_extra_resids)
```

```{r Testing concurvity}
concurvity(total_mod_1.5km_no_extra, full = FALSE)
```


###Total mod with no dom


###total mod with all variables and region 1 or corExp

####gam, temp scaled, norm openness scaled, dom, rec boats 2km scaled, mean gfw scaled, re of region M1, TRUE

```{r Total model with random effect of region 1}
total_mod_re_region1_2km <- gam(urch_sn_ratio ~ s(sst_scale)+
                   s(norm_openness_scale) +
                     dominant_score+
                   s(mean_rec_boat_counts_2km_scale)+
                     s(mean_app_fishing_hours_scale)+
                     s(region1, bs = "re"),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_re_region1_2km)

plot(total_mod_re_region1_2km)
summary(total_mod_re_region1_2km)
```

```{r Testing residuals for spatial correlation}
#extract residuals from model
residuals_total_mod_2km_region_1 <- residuals(total_mod_re_region1_2km, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_2km_region_1 <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_2km_region_1 <- st_transform(coords_for_auto_test_2km_region_1, crs = 3857)
extracted_coords_for_auto_test_projected_2km_region_1 <- st_coordinates(coords_for_auto_test_projected_2km_region_1)

nb_2km_region_1 <- dnearneigh(extracted_coords_for_auto_test_projected_2km_region_1, d1 = 0, d2 = 10)
lw_2km_region_1 <- nb2listw(nb_2km_region_1, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_2km_region_1, lw_2km_region_1)
```

```{r DHARMa for model fit}
total_mod_re_region1_2km_resids <- simulateResiduals(total_mod_re_region1_2km)
plot(total_mod_re_region1_2km_resids)
```

####gam, temp scaled, norm openness scaled, dom, rec boats 1.5km scaled, mean gfw scaled, re of region M1, TRUE

```{r Total model with random effect of region 1}
total_mod_re_region1_1.5km <- gam(urch_sn_ratio ~ s(sst_scale)+
                   s(norm_openness_scale) +
                     dominant_score+
                   s(mean_rec_boat_counts_1.5km_scale)+
                     s(mean_app_fishing_hours_scale)+
                     s(region1, bs = "re"),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_re_region1_1.5km)

plot(total_mod_re_region1_1.5km)
summary(total_mod_re_region1_1.5km)
```

```{r Testing residuals for spatial correlation}
#extract residuals from model
residuals_total_mod_1.5km_region_1 <- residuals(total_mod_re_region1_1.5km, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_1.5km_region_1 <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_1.5km_region_1 <- st_transform(coords_for_auto_test_1.5km_region_1, crs = 3857)
extracted_coords_for_auto_test_projected_1.5km_region_1 <- st_coordinates(coords_for_auto_test_projected_1.5km_region_1)

nb_1.5km_region_1 <- dnearneigh(extracted_coords_for_auto_test_projected_1.5km_region_1, d1 = 0, d2 = 10)
lw_1.5km_region_1 <- nb2listw(nb_1.5km_region_1, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1.5km_region_1, lw_1.5km_region_1)
```

####gam, temp scaled, norm openness scaled, dom, rec boats 1km scaled, mean gfw scaled, re of region M1, TRUE

```{r Total model with random effect of region 1}
total_mod_re_region1_1km <- gam(urch_sn_ratio ~ s(sst_scale)+
                   s(norm_openness_scale) +
                     dominant_score+
                   s(mean_rec_boat_counts_1km_scale)+
                     s(mean_app_fishing_hours_scale)+
                     s(region1, bs = "re"),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_re_region1_1km)

plot(total_mod_re_region1_1km)
summary(total_mod_re_region1_1km)
```

```{r Testing residuals for spatial correlation}
#extract residuals from model
residuals_total_mod_1km_region_1 <- residuals(total_mod_re_region1_1km, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_1km_region_1 <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_1km_region_1 <- st_transform(coords_for_auto_test_1km_region_1, crs = 3857)
extracted_coords_for_auto_test_projected_1km_region_1 <- st_coordinates(coords_for_auto_test_projected_1km_region_1)

nb_1km_region_1 <- dnearneigh(extracted_coords_for_auto_test_projected_1km_region_1, d1 = 0, d2 = 10)
lw_1km_region_1 <- nb2listw(nb_1km_region_1, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1km_region_1, lw_1km_region_1)
```



####gamm, temp scaled, norm openness scaled, dom, rec boats 1.5km scaled, mean gfw scaled, corExp, TRUE

```{r}
total_mod_1.5km_corExp <- gamm(urch_sn_ratio ~ s(sst_scale) + 
                                 s(norm_openness_scale) +
                                 dominant_score+
                                 s(mean_rec_boat_counts_1.5km_scale)+
                                 s(mean_app_fishing_hours_scale), 
                        data = env_mod_data_shifted,
                       family = gaussian(link = "identity"),
                       correlation = corExp(form = ~ longitude_proj + latitude_proj, nugget = FALSE), 
                       method = 'REML',
                       select = TRUE)

set.seed(13)
gam.check(total_mod_1.5km_corExp$gam)
#shapiro.test(total_mod_1km$residuals)
##slightly not normal

plot(total_mod_1.5km_corExp$gam)
summary(total_mod_1.5km_corExp$lme)
summary(total_mod_1.5km_corExp$gam)
#as_flextable(total_mod_1km)
```

```{r Testing residuals for spatial correlation}
#extract residuals from model
residuals_total_mod_1.5km_corExp <- residuals(total_mod_1.5km_corExp$gam, "pearson")

#creating the spatial weights matrix
coords_for_auto_test_1.5km_corExp <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_1.5km_corExp <- st_transform(coords_for_auto_test_1.5km_corExp, crs = 3857)
extracted_coords_for_auto_test_projected_1.5km_corExp <- st_coordinates(coords_for_auto_test_projected_1.5km_corExp)

nb_1.5km_corExp <- dnearneigh(extracted_coords_for_auto_test_projected_1.5km_corExp, d1 = 0, d2 = 15)
lw_1.5_corExp <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1.5km_corExp, lw_1.5_corExp)
```

###Plots of EVs between regions

```{r Boxplot of the mean rec fishing between regions}
ggplot(env_mod_data, aes(x = region1, y = mean_rec_boat_counts_1km))+
  geom_boxplot()+
  theme.figure()

ggplot(env_mod_data, aes(x = region1, y = mean_rec_boat_counts_1.5km))+
  geom_boxplot()+
  theme.figure()

ggplot(env_mod_data, aes(x = region1, y = mean_rec_boat_counts_2km))+
  geom_boxplot()+
  theme.figure()
```

```{r Boxplot of mean apparent fishing hours between regions }
ggplot(env_mod_data, aes(x = region1, y = mean_app_fishing_hours))+
  geom_boxplot()+
  theme.figure()
```

```{r Boxplot of normalized openness between regions }
ggplot(env_mod_data, aes(x = region1, y = norm_openness))+
  geom_boxplot()+
  theme.figure()
```


####gam, temp scaled, norm openness scaled, dom, rec boats 500m scaled, mean gfw scaled, re of region M1, TRUE

```{r Total model with random effect of region 1}
total_mod_re_region1_test2 <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   #dominant_score + 
                   s(mean_rec_boat_counts_2km_scale)
                   #s(mean_app_fishing_hours_scale)
                   ,
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_re_region1_test2)

plot(total_mod_re_region1_test2)
summary(total_mod_re_region1_test2)
```
theres a regional effect but havent sampled all regions equally

####gam, temp scaled, norm openness scaled, dom, rec boats 2km scaled, mean gfw scaled, re of region M2, TRUE

```{r Total model with random effect of region 2}
total_mod_re_region2_2km <- gam(urch_sn_ratio ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   dominant_score + 
                   s(mean_rec_boat_counts_2km_scale)+
                   s(mean_app_fishing_hours_scale)+
                   s(region2, bs = "re"),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_re_region2_2km)

plot(total_mod_re_region2_2km)
```

```{r DHARMa for model fit}
total_mod_re_region2_2km_resids <- simulateResiduals(total_mod_re_region2_2km)
plot(total_mod_re_region2_2km_resids)
```


####gamm, temp scaled, norm openness scaled, dom, mean rec boats 2km scaled, mean apparent fishing hours scaled, corExp, TRUE shifted points

```{r Building total gamm with spatial autocorrelation term corExp}
total_mod_sp_cor_exp <- gamm(urch_sn_ratio ~ s(sst_scale) +
                                   s(norm_openness_scale) +
                                   dominant_score +
                                   s(mean_rec_boat_counts_2km_scale) +
                                   s(mean_app_fishing_hours_scale), 
                        data = env_mod_data_shifted,
                       family = gaussian(link = "identity"),
                       correlation = corExp(form = ~ longitude_proj + latitude_proj, nugget = FALSE, value = 639), 
                       method = 'REML',
                       select = TRUE)


gam.check(total_mod_sp_cor_exp$gam, rep = 500)
plot(total_mod_sp_cor_exp$gam)
summary(total_mod_sp_cor_exp$lme)
```

```{r DHARMa for model fit}
env_mod_sp_cor_exp_total_resid <- simulateResiduals(env_mod_sp_cor_exp_total$lme)
plot(env_mod_sp_cor_exp_total_resid)
```







###Testing for interaction

####gam, temp scaled, norm openness scaled, dom, rec boats 2km scaled, mean gfw scaled, TRUE

```{r Running a model with an interaction}
total_mod_1km_int <- gam(urch_sn_ratio ~ s(sst_scale) + 
                 #  norm_openness_scale:mean_rec_boat_counts_2km_scale  +
                      s(norm_openness_scale) +
                      s(mean_rec_boat_counts_750m_scale)+
                 #  ti(norm_openness_scale, mean_rec_boat_counts_1km_scale, k= 8)+
                   dominant_score + 
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE) 

plot(total_mod_1km_int)
summary(total_mod_1km_int)
```

```{r}
vis.gam(total_mod_1km_int, theta = -45, view = c("norm_openness_scale", "mean_rec_boat_counts_1km_scale"))
?vis.gam
```


####Using ANOVA to see if its worth including the interaction

```{r Comparing models with ANOVA}
anova(total_mod_2km, total_mod_2km_int)
```
Adding the model does not significantly change the model fit. So no point to include.



###AIC with all of the models run

```{r Starting with normal AIC which assumes infinite sample size}
AIC(env_mod_T,
    #env_mod_sp_gp, 
    #env_mod_sp_gp_shifted,
    env_mod_sp_cor_exp$lme,
   # env_mod_re_site_code,
    env_mod_re_region_M1,
    env_mod_re_region_M2, 
    env_mod_re_region_M3,
    env_mod_re_region_M4,
    fish_mod,
    fish_mod_1km,
    fish_mod_1.5km,
    fish_mod_750m,
    fish_mod_2km,
    total_mod,
   total_mod_2km,
    total_mod_re_region1,
    total_mod_re_region2,
   total_mod_re_region1_2km,
   total_mod_re_region2_2km,
   total_mod_sp_cor_exp
    )%>%
arrange(AIC)%>%
  mutate(delta_AIC = AIC - AIC[1])
```

```{r Using corrected AIC because sample size is small}
AICc(env_mod_T,
    #env_mod_sp_gp, 
    #env_mod_sp_gp_shifted,
    env_mod_sp_cor_exp$lme,
    #env_mod_re_site_code,
    env_mod_re_region_M1,
    env_mod_re_region_M2, 
    env_mod_re_region_M3,
    env_mod_re_region_M4,
    fish_mod,
    fish_mod_1km,
    fish_mod_1.5km,
    fish_mod_750m,
    fish_mod_2km,
    total_mod,
    total_mod_2km,
    total_mod_re_region1,
    total_mod_re_region2,
     total_mod_re_region1_2km,
   total_mod_re_region2_2km,
   total_mod_sp_cor_exp
    )%>%
arrange(AICc)%>%
  mutate(delta_AICc = AICc - AICc[1])
```

###Testing if the best models are significantly different

```{r Using ANOVA to test the best models}
anova(total_mod_re_region1, total_mod_re_region1_2km, env_mod_re_region_M1)
```


##Building a glm of the final model to compare

###glm, temp scaled, norm openness scaled, dom, rec boats 2km scaled, mean gfw scaled

```{r GLM of the main model to see if there is much of a difference}
total_mod_2km_glm <- lm(urch_sn_ratio ~ sst_scale + 
                   norm_openness_scale +
                   dominant_score + 
                   mean_rec_boat_counts_2km_scale+
                   mean_app_fishing_hours_scale,
               data = env_mod_data)

set.seed(13)
#shapiro.test(total_mod_2km$residuals)
##normal

plot(total_mod_2km_glm)
summary(total_mod_2km_glm)
```

```{r DHARMa for model fit}
total_mod_2km_glm_resids <- simulateResiduals(total_mod_2km_glm)
plot(total_mod_2km_glm_resids)
```


##Modelling Urchin Counts

###Plotting EVs against urchin counts

```{r urchins against openness}
(urchins_dom_substrate_plot <- ggplot(env_mod_data, aes(x = dominant_score, y = Urchins, colour = region1))+
  geom_point()+
    labs(x = "\nDominant substrate type", y = "Red urchin counts\n",
       colour = "Region")+
  theme.figure())
```

```{r urchins against openness}
(urchins_norm_openness_plot <- ggplot(env_mod_data, aes(x = norm_openness, y = Urchins, colour = region1))+
  geom_point()+
      labs(x = "\nNormalized openness", y = "Red urchin counts\n",
       colour = "Region")+
  theme.figure())
```

```{r urchins against sst}
(urchins_sst_plot <- ggplot(env_mod_data, aes(x = sst, y = Urchins, colour = region1))+
  geom_point()+
      labs(x = "\nSea surface temperature (ºC)", y = "Red urchin counts\n",
       colour = "Region")+
  theme.figure())
```

```{r urchins against rec boat counts type}
(urchins_mean_rec_boats_1.5km_plot <- ggplot(env_mod_data, aes(x = mean_rec_boat_counts_1.5km, y = Urchins, colour = region1))+
  geom_point()+
      labs(x = "\nMean # of recreational boats fishing (1.5km radius)", y = "Red urchin counts\n",
       colour = "Region")+
  theme.figure())
```

```{r}
(urchins_mean_app_fishing_hours_plot <- ggplot(env_mod_data, aes(x = mean_app_fishing_hours, y = Urchins,
                         colour = region1))+
  geom_point()+
      labs(x = "\nMean apparent fishing hours", y = "Red urchin counts\n",
       colour = "Region")+
  theme.figure())
```

None of the explanatory variables appear to have much of a relationship with urchin counts

```{r Combining into one figure, fig.height=8, fig.width=12}
(urchin_counts_evs_figure <- ggarrange(urchins_dom_substrate_plot, urchins_norm_openness_plot,
                                      urchins_sst_plot, urchins_mean_app_fishing_hours_plot, 
                                      urchins_mean_rec_boats_1.5km_plot,
                                     labels = c("A", "B", "C", "D", "E"),
                                     common.legend = TRUE,
                                     legend = "bottom",
                                     ncol = 3,
                                     nrow = 2))
#annotate_figure(snail_counts_evs_figure, left = textGrob("Common y-axis", rot = 90, vjust = 1, gp = gpar(cex = 1.3)))
```

```{r Saving the multi-panelled plot}
png(filename = here("./04_plots/report_figures/urchin_counts_evs.png"), width = 12, height = 8, 
    units = "in", pointsize = 15, res = 600)

urchin_counts_evs_figure

dev.off()
```

```{r Looking at distribution of Urchin counts for modelling}
hist(env_mod_data_biomass$total_site_biomass)
shapiro.test(env_mod_data$Urchins)
#Urchins are normally distributed
```


###gam, temp scaled, openness scaled, dom substrate

```{r Modelling the urchin counts with a simple environmental model}
env_mod_urchin_counts <- gam(Urchins ~ s(sst_scale)+
                               s(norm_openness_scale),
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
set.seed(10)
gam.check(env_mod_urchin_counts, rep = 500)
plot(env_mod_urchin_counts)
```

```{r DHARMa for model fit}
env_mod_urchin_counts_resids <- simulateResiduals(env_mod_urchin_counts)
plot(env_mod_urchin_counts_resids)
```

###gam, temp scaled, openness scaled, dom substrate, mean rec 1.5km scaled, mean gfw scaled, TRUE

```{r Modelling the urchin counts with a total model}
total_mod_urchin_counts <- gam(Urchins ~ #s(sst_scale)+
                              # s(norm_openness_scale)+
                                # s(Turban_Snails_scale)+
                              # dominant_score+
                              # s(mean_rec_boat_counts_1.5km_scale)+
                              # s(mean_app_fishing_hours_scale)+
                                 s(total_site_biomass),
                             data = env_mod_data_biomass,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
set.seed(10)
gam.check(total_mod_urchin_counts, rep = 500)
plot(total_mod_urchin_counts)
summary(total_mod_urchin_counts)

#shapiro.test(total_mod_urchin_counts$residuals)
##residuals normal
```

```{r DHARMa for model fit}
total_mod_urchin_counts_resids <- simulateResiduals(total_mod_urchin_counts)
plot(total_mod_urchin_counts_resids)
```


##Modelling the snail counts

###Plotting the EVs against the snail counts

```{r snails against substrate}
(snails_dom_substrate_plot <- ggplot(env_mod_data, aes(x = dominant_score, y = Turban_Snails))+
  geom_boxplot()+
  labs(x = "Dominant substrate type", y = "Turban Snail counts",
       colour = "Region")+
  theme.figure())
```

```{r snails against openness}
(snails_norm_openness_plot <- ggplot(env_mod_data, aes(x = norm_openness, y = Turban_Snails, colour = region1))+
  geom_point()+
   labs(x = "Normalized openness", y = "Turban Snail counts",
       colour = "Region")+
  theme.figure())
```
```{r snails against openness}
ggplot(env_mod_data, aes(x = norm_openness, y = Turban_Snails, colour = region1))+
  geom_point()+
   labs(x = "Normalized openness", y = "Turban snails",
       colour = "Region")+
  theme.figure()
```
```{r snails against openness}
ggplot(env_mod_data, aes(x = norm_openness, y = Urchins, colour = region1))+
  geom_point()+
   labs(x = "Normalized openness", y = "urchins",
       colour = "Region")+
  theme.figure()
```

```{r snails against sst}
(snails_sst_plot <- ggplot(env_mod_data, aes(x = sst, y = Turban_Snails, colour = region1))+
  geom_point()+
   labs(x = "Sea surface temperature (ºC)", y = "Turban Snail counts",
       colour = "Region")+
  theme.figure())
```

```{r snails against rec boats}
(snails_mean_rec_boats_plot <- ggplot(env_mod_data, aes(x = mean_rec_boat_counts_1.5km, y = Turban_Snails, colour = region1))+
  geom_point()+
   labs(x = "Mean # of recreational boats fishing (1.5km radius)", y = "Turban Snail counts",
       colour = "Region")+
  theme.figure())
```

```{r snails against gfw}
(snails_app_fishing_hours_plot <- ggplot(env_mod_data, aes(x = mean_app_fishing_hours, y = Turban_Snails,
                         colour = region1))+
  geom_point()+
   labs(x = "Mean apparent fishing hours", y = "Turban Snail counts",
       colour = "Region")+
  theme.figure())
```

```{r Combining into one figure, fig.height=8, fig.width=12}
(snail_counts_evs_figure <- ggarrange(snails_dom_substrate_plot, snails_norm_openness_plot,
                                     snails_sst_plot, snails_app_fishing_hours_plot,
                                     snails_mean_rec_boats_plot,
                                     labels = c("A", "B", "C", "D", "E"),
                                     common.legend = TRUE,
                                     legend = c("bottom"),
                                     nrow = 2,
                                     ncol = 3))
#annotate_figure(snail_counts_evs_figure, left = textGrob("Common y-axis", rot = 90, vjust = 1, gp = gpar(cex = 1.3)))
```

```{r Saving the multi-panelled plot}
png(filename = here("./04_plots/report_figures/snail_counts_evs.png"), width = 12, height = 8, 
    units = "in", pointsize = 15, res = 600)

snail_counts_evs_figure

dev.off()
```



###Snail count histogram

```{r Histogram of snail counts}
ggplot(env_mod_data, aes(x = Turban_Snails))+
  geom_histogram(binwidth = 75)+
  theme.figure()
```

The snail counts show a Poisson or 

```{r}
mean(env_mod_data$Turban_Snails)
var(env_mod_data$Turban_Snails)

# Table of observed frequencies
obs_freq <- table(env_mod_data$Turban_Snails)
# Expected frequencies based on Poisson distribution
lambda <- mean(env_mod_data$Turban_Snails)
exp_freq <- dpois(as.numeric(names(obs_freq)), lambda) * length(env_mod_data$Turban_Snails)

# Chi-squared test
chisq_test <- chisq.test(obs_freq, p = exp_freq, rescale.p = TRUE)
print(chisq_test)

#The snail counts are likely not poisson distributed
```


###gam, temp scaled, openness scaled, dom substrate

```{r Modelling the snail counts with a simple environmental model}
env_mod_snail_counts <- gam(Turban_Snails ~ s(sst_scale)+
                               s(norm_openness_scale),
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
set.seed(10)
gam.check(env_mod_snail_counts, rep = 500)
plot(env_mod_snail_counts)
```

```{r DHARMa for model fit}
env_mod_snail_counts_resids <- simulateResiduals(env_mod_snail_counts)
plot(env_mod_snail_counts_resids)
```

###gam, temp scaled, openness scaled, dom substrate, mean rec 1.5km scaled, mean gfw scaled, TRUE
```{r}
total_mod_snail_counts <- gam(Turban_Snails ~ s(sst_scale)+
                               s(norm_openness_scale)+
                              dominant_score+
                              s(mean_rec_boat_counts_1.5km_scale)+
                              s(mean_app_fishing_hours_scale)+
                              
                                s(total_site_biomass),
                             data = env_mod_data_biomass,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
set.seed(10)
gam.check(total_mod_snail_counts, rep = 500)
plot(total_mod_snail_counts)
summary(total_mod_snail_counts)
```

```{r Modelling the snail counts with a total model}
total_mod_snail_counts_poisson <- gam(Turban_Snails ~ s(sst_scale, k = 5)+
                               s(norm_openness_scale, k = 5)+
                              s(mean_rec_boat_counts_1.5km_scale, k = 5)+
                              s(mean_app_fishing_hours_scale, k = 5) +
                               s(Urchins_scale, k = 5),
                             data = env_mod_data,
                             family = poisson(link = "log"),
                             method = "ML",
                             select = TRUE)
set.seed(10)
gam.check(total_mod_snail_counts_poisson, rep = 500)
plot(total_mod_snail_counts_poisson, rug = TRUE)
     #residuals = TRUE, pch = 1, shade = TRUE,
     #shift = coef(total_mod_snail_counts)[1])
summary(total_mod_snail_counts_poisson)
#shapiro.test(total_mod_snail_counts$residuals)


#Poisson does not seem to match the distribution well
```

```{r Checking the Poisson model for overdispersion}
total_mod_snail_counts_poisson_resids <- simulateResiduals(total_mod_snail_counts_poisson)
testDispersion(total_mod_snail_counts_poisson_resids)
```
It appears that there is high overdispersion in the data so Poisson may not be best.
When including the urchins with lower k overdispersion goes away.
Making all variables k = 5 maintains high overdispersion.

```{r Modelling the snail counts with a total model}
total_mod_snail_counts_quasipoisson <- gam(Turban_Snails ~ s(sst_scale)+
                               s(norm_openness_scale)+
                              s(mean_rec_boat_counts_1.5km_scale)+
                                 s(mean_app_fishing_hours_scale) +
                               s(Urchins_scale),
                             data = env_mod_data,
                             family = quasipoisson(link = "log"),
                             method = "ML",
                             select = TRUE)
set.seed(10)
gam.check(total_mod_snail_counts_quasipoisson, rep = 500)
plot(total_mod_snail_counts_quasipoisson, rug = TRUE)
     #residuals = TRUE, pch = 1, shade = TRUE,
     #shift = coef(total_mod_snail_counts)[1])
summary(total_mod_snail_counts_quasipoisson)
#shapiro.test(total_mod_snail_counts$residuals)


#Poisson does not seem to match the distribution well
```

```{r Checking the Quasipoisson model for overdispersion}
total_mod_snail_counts_quasipoisson_resids <- simulateResiduals(total_mod_snail_counts_quasipoisson)
testDispersion(total_mod_snail_counts_quasipoisson_resids)
```

```{r Modelling the snail counts with a total model}
total_mod_snail_counts_nb <- gam(Turban_Snails ~ s(sst_scale)+
                               s(norm_openness_scale)+
                              s(mean_rec_boat_counts_1.5km_scale)+
                                s(mean_app_fishing_hours_scale)+
                               s(Urchins_scale),
                             data = env_mod_data,
                             family = nb(),
                             method = "ML",
                             select = TRUE)
set.seed(10)
gam.check(total_mod_snail_counts_nb, rep = 500)
plot(total_mod_snail_counts_nb, rug = TRUE)
     #residuals = TRUE, pch = 1, shade = TRUE,
     #shift = coef(total_mod_snail_counts)[1])
summary(total_mod_snail_counts_nb)
shapiro.test(total_mod_snail_counts_nb$residuals)

#Using nb seems improved from the poisson and quasi-poisson in terms of the residuals but still 
#has some problematic stuff going on
```

```{r Testing  residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_total_mod_snail_counts_nb <- residuals(total_mod_snail_counts_nb, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_snail_counts_nb, lw)
```
There is some significant spatial autocorrelation

```{r Testing 1km no app hours concurvity, echo=FALSE}
concurvity(total_mod_snail_counts_nb, full = TRUE)
```

```{r}
AIC(total_mod_snail_counts_nb, 
    total_mod_snail_counts_poisson,
    total_mod_snail_counts_quasipoisson
    )%>%
arrange(AIC)%>%
  mutate(delta_AIC = AIC - AIC[1])
```


##Re-doing analysis with averaged transects

####gam, temp scaled, norm openness scaled, dom, rec boats 2km scaled, mean gfw scaled, TRUE, AVG

```{r Total model with no spatial autocorrelation}
total_mod_2km_avg <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_2km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_2km_avg)
#shapiro.test(total_mod_2km$residuals)
##normal

plot(total_mod_2km_avg)
summary(total_mod_2km_avg)
```

Some outliers observed, retesting with outliers removed to see if 
they have a large impact on the model

```{r Testing residuals for spatial correlation}
#extract residuals from model
residuals_total_mod_2km_avg <- residuals(total_mod_2km_avg, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_2km_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_2km_avg <- st_transform(coords_for_auto_test_2km_avg, crs = 3857)
extracted_coords_for_auto_test_projected_2km_avg <- st_coordinates(coords_for_auto_test_projected_2km_avg)

nb_2km_avg <- dnearneigh(extracted_coords_for_auto_test_projected_2km_avg, d1 = 0, d2 = 1000)
lw_2km_avg <- nb2listw(nb_2km_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_2km_avg, lw_2km_avg)


###testing with inverse distance matrix to ensure that the Moran's I value is accurate
dist_matrix_total_mod_2km_avg <- as.matrix(dist(extracted_coords_for_auto_test_projected_2km_avg))
inverse_dist_matrix_total_mod_2km_avg <- 1 / (dist_matrix_total_mod_2km_avg + 0.001)
diag(inverse_dist_matrix_total_mod_2km_avg) <- 0
lw_2km_inverse_avg <- mat2listw(inverse_dist_matrix_total_mod_2km_avg, style = "W")
moran.test(residuals_total_mod_2km_avg, lw_2km_inverse_avg)
```



```{r Rerunning the model without the outliers}
total_mod_2km_avg_no_outliers <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                
                   s(mean_rec_boat_counts_2km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data_avg[-c(7, 19), ],
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_2km_avg_no_outliers)
#shapiro.test(total_mod_2km$residuals)
##normal

plot(total_mod_2km_avg_no_outliers)
summary(total_mod_2km_avg_no_outliers)
```

Trend remains largely the same without the outliers and openness is still 
significant so can do analysis with them in so we are not cherry-picking the
sites.

####gam, temp scaled, norm openness scaled, dom, rec boats 1.5km scaled, mean gfw scaled, TRUE, AVG

```{r Total model with no spatial autocorrelation}
total_mod_1.5km_avg <- gam(urch_sn_ratio_avg ~ s(sst_scale) + 
                   s(norm_openness_scale) +
                   s(mean_rec_boat_counts_1.5km_scale)+
                   s(mean_app_fishing_hours_scale),
               data = env_mod_data_avg,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(total_mod_1.5km_avg)
#shapiro.test(total_mod_2km$residuals)
##normal

plot(total_mod_1.5km_avg)
summary(total_mod_1.5km_avg)
```


```{r Testing residuals for spatial correlation}
#extract residuals from model
residuals_total_mod_1.5km_avg <- residuals(total_mod_1.5km_avg, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test_1.5km_avg <- st_as_sf(env_mod_data_avg, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected_1.5km_avg <- st_transform(coords_for_auto_test_1.5km_avg, crs = 3857)
extracted_coords_for_auto_test_projected_1.5km_avg <- st_coordinates(coords_for_auto_test_projected_1.5km_avg)

nb_1.5km_avg <- dnearneigh(extracted_coords_for_auto_test_projected_1.5km_avg, d1 = 0, d2 = 1000)
lw_1.5km_avg <- nb2listw(nb_1.5km_avg, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_total_mod_1.5km_avg, lw_1.5km_avg)
```
