---
title: "RLS Modelling"
author: "Alec"
date: "2024-11-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(mgcv)
library(DHARMa)
library(mgcViz)
library(ggeffects)
library(gratia)
library(geoR)
library(geosphere)
library(gstat)
library(sf)
```




#Modelling the environmental variables

##Loading in the data with old fetch data
```{r Loading in the data}
env_mod_data <- read_csv(here("./03_clean_data/urch_sn_ratio_mod_data.csv"))

#I added columns for site and wider location in excel before downloading the df
```

```{r Scaling the explanatory variables and converting substrate type to factor}
env_mod_data$sst_scale <- scale(env_mod_data$sst)
env_mod_data$avg_fetch_scale <- scale(env_mod_data$avg_fetch)
env_mod_data$dominant_score <- as.factor(env_mod_data$dominant_score)
levels(env_mod_data$dominant_score) <- list(loose = "3", solid = "4")
```


##Plotting EVs against urch sn ratio

```{r ratio against fetch}
ggplot(env_mod_data, aes(x = avg_fetch_scale, y = urch_sn_ratio))+
  geom_point()+
  theme.figure()
```

```{r ratio against sst}
ggplot(env_mod_data, aes(x = sst, y = urch_sn_ratio))+
  geom_point()+
  theme.figure()
```

```{r ratio against substrate type}
ggplot(env_mod_data, aes(x = dominant_score, y = urch_sn_ratio))+
  geom_point()+
  theme.figure()
```

##Building the GAM

###No spatial autocorrelation term
```{r Making a gam with no spatial autocorrelation term to start}
env_mod <- gam(urch_sn_ratio ~ s(sst_scale) + s(avg_fetch_scale) + dominant_score,
               data = env_mod_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(env_mod, rep = 500)

plot(env_mod)
```


```{r Checking the model fit with DHARMa}
env_mod_resids <- simulateResiduals(env_mod)
plot(env_mod_resids)
```
All good




```{r Making a gam with no spatial autocorrelation and no fetch}
env_mod_no_fetch <- gam(urch_sn_ratio ~ s(sst_scale) + dominant_score, data = env_mod_data,
                        method = 'REML',
                        family = gaussian(link = "identity"),
                        select = FALSE)

gam.check(env_mod_no_fetch, rep = 500)
plot(env_mod_no_fetch)
```

```{r}
env_mod_no_fetch_resid <- simulateResiduals(env_mod_no_fetch)
plot(env_mod_no_fetch_resid)
```
Possible problem with this model according to DHARMa

###Testing the data to see if it is spatially correlated

####Moran's I with transects at same site separate
```{r Using Moran's I test on the data with geosphere}
##Method setting transects at same site to 1 m apart

sphere_coords <- cbind(env_mod_data$longitude, env_mod_data$latitude)
sphere_dist <- distm(sphere_coords, fun = distGeo)

sphere_dist[sphere_dist == 0] <- 1
inv_sphere_dist <- 1 / sphere_dist
diag(inv_sphere_dist) <- 0
  
ape::Moran.I(env_mod_data$urch_sn_ratio, weight = inv_sphere_dist, scaled = TRUE)
```

####Moran's I with transects at same site averaged
```{r Making a df with averaged urch_sn ratios}
urch_sn_ratio_averaged <- env_mod_data%>%
  group_by(site_code, latitude, longitude)%>%
  summarise(avg_urch_sn_ratio = mean(urch_sn_ratio))
```

```{r calculating Moran's I with the averaged points}
sphere_coords_avg <- cbind(urch_sn_ratio_averaged$longitude, urch_sn_ratio_averaged$latitude)

sphere_dist_avg <- distm(sphere_coords_avg, fun = distGeo)

inv_sphere_dist_avg <- 1 / sphere_dist_avg

diag(inv_sphere_dist_avg) <- 0

ape::Moran.I(urch_sn_ratio_averaged$avg_urch_sn_ratio, weight = inv_sphere_dist_avg, scaled = TRUE)
```



####Using a semi-variogram to see what type of spatial autocorrelation factor to use

#####With averaged points
```{r Transforming the coordinates into meters for the variogram}
# Using the averaged data
urch_sn_ratio_averaged_sf <- st_as_sf(urch_sn_ratio_averaged, coords = c("longitude", "latitude"), crs = 4326)
urch_sn_ratio_averaged_projected <- st_transform(urch_sn_ratio_averaged_sf, crs = 3857)
```

```{r Plotting a semi-variogram with the meter distances for the averaged transects}
vario_meters <- variogram(avg_urch_sn_ratio ~ 1, data = urch_sn_ratio_averaged_projected)
vario_meters_df <- as.data.frame(vario_meters)
plot(vario_meters)

ggplot(vario_meters_df, aes(x = dist, y = gamma)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = seq(0, 15000, by = 1000)) + # Custom x-axis ticks
  labs(x = "Distance (m)", y = "Semivariance", title = "Variogram") +
  theme_minimal()

```

```{r Fitting a variogram model}
vario_meters_mod <- fit.variogram(vario_meters, 
                                  model = vgm(nugget = 0.1, range = 2000, psill = 0.75, 
                                              model = c("Exp", "Sph", "Gau")),
                                  fit.sills = TRUE, fit.ranges = TRUE)
vario_meters_mod
plot(vario_meters, vario_meters_mod)
```

#####With non averaged points

```{r Variogram with the non-averaged points}
urch_sn_coords_var_2 <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
urch_sn_coords_var_2_projected <- st_transform(urch_sn_coords_var_2, crs = 3857)

vario_2 <- variogram(urch_sn_ratio ~ 1, data = urch_sn_coords_var_2_projected)
vario_2_df <- as.data.frame(vario_2)
plot(vario_2)

ggplot(vario_2_df, aes(x = dist, y = gamma)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = seq(0, 15000, by = 1000)) + # Custom x-axis ticks
  labs(x = "Distance (m)", y = "Semivariance", title = "Variogram") +
  theme_minimal()
```

```{r Fitting a variogram model}
vario_2_mod <- fit.variogram(vario_2, model = vgm(1, model = c("Exp", "Sph", "Gau"), range = 1000, psill = 1), fit.sills = TRUE, fit.ranges = TRUE)
vario_2_mod
summary(vario_2_mod)
plot(vario_2, vario_2_mod)
```


#####With shifted points

```{r Plotting an experimental variogram for the data that is shifted by one meter}
vario_shifted <- variogram(urch_sn_ratio ~ 1, data = env_mod_sf_projected)
plot(vario_shifted)
```

```{r Fitting a variogram model to the experimental variogram}
vario_shifted_mod <- fit.variogram(vario_shifted, model = vgm(nugget = 1, model = c("Exp", "Sph", "Gau"), range = 1000, psill = 1), fit.sills = TRUE, fit.ranges = TRUE)
vario_shifted_mod
plot(vario_shifted, vario_shifted_mod)
```



###With spatial autocorrelation term

```{r Building the gam with the spatial autocorrelation term}
####This is likely overfitting the model and just saying that a smooth of lat and long explains the variation in urch ratio

env_mod_sp_gp <- gam(urch_sn_ratio ~ s(sst_scale) + s(avg_fetch_scale) + dominant_score + s(latitude, longitude, bs = "gp", k = 10),
                  data = env_mod_data, 
                  family = gaussian(link = "identity"),
                  method = 'REML',
                  select = TRUE)   #In this case I have added the lat long as an interaction term directly in the model

set.seed(10)
gam.check(env_mod_sp_gp, rep = 500)
plot(env_mod_sp_gp)
```

```{r DHARMa for model fit}
env_mod_sp_gp_resid <- simulateResiduals(env_mod_sp_gp)
plot(env_mod_sp_gp_resid)
```
All good

```{r Making a dataframe with sites at the same location shifted slightly}
env_mod_data_shifted <- env_mod_data%>%
    mutate(latitude = if_else(row_number() == 2, 48.82896, latitude))%>%
  mutate(latitude = if_else(row_number() == 5, 48.85746, latitude))%>%
  mutate(latitude = if_else(row_number() == 7, 48.85827, latitude))%>%
  mutate(latitude = if_else(row_number() == 21, 48.91608, latitude))%>%
   mutate(latitude = if_else(row_number() == 23, 48.83858, latitude))%>%
   mutate(latitude = if_else(row_number() == 25, 48.90711, latitude))%>%
  mutate(latitude = if_else(row_number() == 30, 48.82734, latitude))%>%
  mutate(latitude = if_else(row_number() == 32, 48.95024, latitude))%>%
   mutate(latitude = if_else(row_number() == 34, 48.95509, latitude))
```

```{r Converting the shifted points into meters}
env_mod_sf <- st_as_sf(env_mod_data_shifted, coords = c("longitude", "latitude"), crs = 4326)
env_mod_sf_projected <- st_transform(env_mod_sf, crs = 3857)

projected_coords <- st_coordinates(env_mod_sf_projected)
env_mod_data_shifted$longitude_proj <- projected_coords[ , 1]
env_mod_data_shifted$latitude_proj <- projected_coords[ , 2]
```


```{r Building the gam with the spatial autocorrelation term and shifted points}
env_mod_sp_gp_shifted <- gam(urch_sn_ratio ~ s(sst_scale) + s(avg_fetch_scale) + dominant_score + s(latitude, longitude, bs = "gp", k = 10),
                  data = env_mod_data_shifted, 
                  family = gaussian(link = "identity"),
                  method = 'REML',
                  select = TRUE)   #In this case I have added the lat long as an interaction term directly in the model

set.seed(10)
gam.check(env_mod_sp_gp, rep = 500)
plot(env_mod_sp_gp)
```

```{r DHARMa for model fit}
env_mod_sp_gp_shifted_resid <- simulateResiduals(env_mod_sp_gp_shifted)
plot(env_mod_sp_gp_shifted_resid)
```
All good

```{r Building gamm with spatial autocorrelation term corSpher}
env_mod_sp_cor_Spher <- gamm(urch_sn_ratio ~ s(sst_scale) + s(avg_fetch_scale) + dominant_score, 
                        data = env_mod_data_shifted,
                       family = gaussian(link = "identity"),
                       correlation = corSpher(form = ~ longitude + latitude), 
                       method = 'REML',
                       select = TRUE)
?corSpher
gam.check(env_mod_sp_cor_Spher$gam, rep = 500)
plot(env_mod_sp_cor_Spher$gam)
summary(env_mod_sp_cor_Spher$lme)
```

```{r DHARMa for model fit}
env_mod_sp_cor_Spher_resid <- simulateResiduals(fittedModel = env_mod_sp_cor_Spher$lme)
plot(env_mod_sp_cor_Spher_resid)
```
Problem with DHARMa for gamms

```{r Building gamm with spatial autocorrelation term corExp}
env_mod_sp_cor_exp <- gamm(urch_sn_ratio ~ s(sst_scale) + s(avg_fetch_scale) + dominant_score, 
                        data = env_mod_data_shifted,
                       family = gaussian(link = "identity"),
                       correlation = corExp(form = ~ longitude_proj + latitude_proj, nugget = FALSE, value = 639), 
                       method = 'REML',
                       select = TRUE)


gam.check(env_mod_sp_cor_exp$gam, rep = 500)
plot(env_mod_sp_cor_exp$gam)
summary(env_mod_sp_cor_exp$lme)
```

```{r DHARMa for model fit}
env_mod_sp_cor_exp_resid <- simulateResiduals(env_mod_sp_cor_exp$lme)
plot(env_mod_sp_cor_exp_resid)
```
Problem with DHARMa for gamms

###AIC with the different models

```{r Comparing the models with an AIC}
AIC(env_mod, env_mod_no_fetch, env_mod_sp_gp, env_mod_sp_gp_shifted,
    env_mod_sp_cor_Spher$lme, 
    env_mod_sp_cor_exp$lme
    )%>%
arrange(AIC)%>%
  mutate(delta_AIC = AIC - AIC[1])
```

```{r Comparing the models with unshifted points in an AIC}
AIC(env_mod, env_mod_no_fetch, env_mod_sp_gp)%>%
arrange(AIC)%>%
  mutate(delta_AIC = AIC - AIC[1])
```

```{r Comparing the models with shifted points in an AIC}
AIC(env_mod, env_mod_no_fetch, env_mod_sp_gp_shifted,
    env_mod_sp_cor_Spher$lme, 
    env_mod_sp_cor_exp$lme
    )%>%
arrange(AIC)%>%
  mutate(delta_AIC = AIC - AIC[1])
```


#Modelling with new fetch data

##Loading in the data
```{r Loading in the data}
env_mod_calc_data <- read_csv(here("./03_clean_data/urch_sn_ratio_mod_calc_data.csv"))
```

```{r Scaling the explanatory variables and converting substrate type to factor}
env_mod_calc_data$sst_scale <- scale(env_mod_calc_data$sst)
env_mod_calc_data$avg_fetch_scale <- scale(env_mod_calc_data$avg_fetch)
env_mod_calc_data$dominant_score <- as.factor(env_mod_calc_data$dominant_score)
levels(env_mod_calc_data$dominant_score) <- list(loose = "3", solid = "4")
```

##Plotting EVs against urch sn ratio

```{r ratio against fetch}
ggplot(env_mod_calc_data, aes(x = avg_fetch_scale, y = urch_sn_ratio))+
  geom_point()+
  theme.figure()
```


##Building the GAM

###No spatial autocorrelation term
```{r Making a gam with no spatial autocorrelation term to start with calc data}
env_mod_calc <- gam(urch_sn_ratio ~ s(sst_scale) + s(avg_fetch_scale) + dominant_score,
               data = env_mod_calc_data,
               family = gaussian(link = "identity"),
               method = 'REML',
               select = TRUE)

set.seed(13)
gam.check(env_mod, rep = 500)

plot(env_mod)
```


```{r Checking the model fit with DHARMa}
env_mod_calc_resids <- simulateResiduals(env_mod_calc)
plot(env_mod_calc_resids)
```
All good




```{r Building the model, this time with fetch fully removed}
env_mod_no_fetch_calc <- gam(urch_sn_ratio ~ s(sst_scale) + dominant_score,
                        data = env_mod_calc_data,
                        method = 'REML',
                        family = gaussian(link = "identity"),
                        select = FALSE)

gam.check(env_mod_no_fetch_calc, rep = 500)
```

```{r}
env_mod_no_fetch_calc_resid <- simulateResiduals(env_mod_no_fetch_calc)
plot(env_mod_no_fetch_calc_resid)
```
Possible error based on DHARMa output

###With autocorrelation effect

```{r Building the gam with the spatial autocorrelation term}
env_mod_sp_gp_calc <- gam(urch_sn_ratio ~ s(sst_scale) + s(avg_fetch_scale) + dominant_score + s(latitude, longitude, bs = "gp", k = 10),
                  data = env_mod_calc_data, 
                  family = gaussian(link = "identity"),
                  method = 'REML',
                  select = TRUE)   #In this case I have added the lat long as an interaction term directly in the model

set.seed(10)
gam.check(env_mod_sp_gp_calc, rep = 500)
plot(env_mod_sp_gp_calc)
```

```{r DHARMa for model fit}
env_mod_sp_gp_calc_resid <- simulateResiduals(env_mod_sp_gp_calc)
plot(env_mod_sp_gp_calc_resid)
```
All good


```{r Making a dataframe with sites at the same location shifted slightly}
env_mod_data_shifted_calc <- env_mod_calc_data%>%
    mutate(latitude = if_else(row_number() == 2, 48.82896, latitude))%>%
  mutate(latitude = if_else(row_number() == 5, 48.85746, latitude))%>%
  mutate(latitude = if_else(row_number() == 7, 48.85827, latitude))%>%
  mutate(latitude = if_else(row_number() == 21, 48.91608, latitude))%>%
   mutate(latitude = if_else(row_number() == 23, 48.83858, latitude))%>%
   mutate(latitude = if_else(row_number() == 25, 48.90711, latitude))%>%
  mutate(latitude = if_else(row_number() == 30, 48.82734, latitude))%>%
  mutate(latitude = if_else(row_number() == 32, 48.95024, latitude))%>%
   mutate(latitude = if_else(row_number() == 34, 48.95508, latitude))
```


```{r Building the gam with the spatial autocorrelation term with calc and shifted}
env_mod_sp_gp_calc_shifted <- gam(urch_sn_ratio ~ s(sst_scale) + s(avg_fetch_scale) + dominant_score + s(latitude, longitude, bs = "gp", k = 10),
                  data = env_mod_data_shifted_calc, 
                  family = gaussian(link = "identity"),
                  method = 'REML',
                  select = TRUE)   #In this case I have added the lat long as an interaction term directly in the model

set.seed(10)
gam.check(env_mod_sp_gp_calc_shifted, rep = 500)
plot(env_mod_sp_gp_calc_shifted)
```
```{r DHARMa for model fit}
env_mod_sp_gp_calc_shifted_resid <- simulateResiduals(env_mod_sp_gp_calc_shifted)
plot(env_mod_sp_gp_calc_shifted_resid)
```
All good


```{r Building gamm with spatial autocorrelation term corSpher with calc}
env_mod_sp_cor_Spher_calc <- gamm(urch_sn_ratio ~ s(sst_scale) + s(avg_fetch_scale) + dominant_score, 
                        data = env_mod_data_shifted_calc,
                       family = gaussian(link = "identity"),
                       correlation = corSpher(form = ~ longitude + latitude), 
                       method = 'REML',
                       select = TRUE)

gam.check(env_mod_sp_cor_Spher_calc$gam, rep = 500)
plot(env_mod_sp_cor_Spher_calc$gam)
```



```{r Building gamm with spatial autocorrelation term corExp with calc}
env_mod_sp_cor_exp_calc <- gamm(urch_sn_ratio ~ s(sst_scale) + s(avg_fetch_scale) + dominant_score, 
                        data = env_mod_data_shifted_calc,
                       family = gaussian(link = "identity"),
                       correlation = corExp(form = ~ longitude + latitude), 
                       method = 'REML',
                       select = TRUE)


gam.check(env_mod_sp_cor_exp_calc$gam, rep = 500)
plot(env_mod_sp_cor_exp_calc$gam)
```

###AIC with the different models for calc

```{r Comparing the calc models with an AIC}
AIC(env_mod_calc, env_mod_no_fetch_calc, env_mod_sp_gp_calc, env_mod_sp_gp_calc_shifted,
    env_mod_sp_cor_Spher_calc$lme, 
    env_mod_sp_cor_exp_calc$lme
    )%>%
arrange(AIC)%>%
  mutate(delta_AIC = AIC - AIC[1])
```


```{r Comparing the models with unshifted points in an AIC}
AIC(env_mod_calc, env_mod_no_fetch_calc, env_mod_sp_gp_calc)%>%
arrange(AIC)%>%
  mutate(delta_AIC = AIC - AIC[1])
```


```{r Comparing the calc models with shifted points in an AIC}
AIC(env_mod_calc, env_mod_no_fetch_calc, env_mod_sp_gp_calc_shifted,
    env_mod_sp_cor_Spher_calc$lme, 
    env_mod_sp_cor_exp_calc$lme
    )%>%
arrange(AIC)%>%
  mutate(delta_AIC = AIC - AIC[1])
```