---
title: "RLS Modelling for Amanda"
author: "Alec"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
   toc: true
---

   <script src="hideOutput.js"></script> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(mgcv)
library(DHARMa)
library(mgcViz)
library(ggeffects)
library(gratia)
library(geoR)
library(geosphere)
library(gstat)
library(sf)
library(car)
library(MuMIn)
library(lme4)
library(spdep)
library(ggpubr)
library(glmmTMB)
library(emmeans)
library(conflicted)
library(performance)
library(ufs)

conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
```


```{r Loading in the data, include=FALSE}
env_mod_data <- read_csv(here("./03_clean_data/urch_sn_ratio_mod_data.csv"))
```

```{r Scaling the explanatory variables and converting substrate type to factor, include=FALSE}
env_mod_data$sst_scale <- scale(env_mod_data$sst)
env_mod_data$openness_scale <- scale(env_mod_data$openness)
env_mod_data$norm_openness_scale <- scale(env_mod_data$norm_openness)
env_mod_data$mean_app_fishing_hours_scale <- scale(env_mod_data$mean_app_fishing_hours)
env_mod_data$mean_rec_boat_counts_500m_scale <- scale(env_mod_data$mean_rec_boat_counts_500m)
env_mod_data$mean_rec_boat_counts_1km_scale <- scale(env_mod_data$mean_rec_boat_counts_1km)
env_mod_data$mean_rec_boat_counts_1.5km_scale <- scale(env_mod_data$mean_rec_boat_counts_1.5km)
env_mod_data$mean_rec_boat_counts_750m_scale <- scale(env_mod_data$mean_rec_boat_counts_750m)
env_mod_data$mean_rec_boat_counts_2km_scale <- scale(env_mod_data$mean_rec_boat_counts_2km)
env_mod_data$dominant_score <- as.factor(env_mod_data$dominant_score)
env_mod_data$site_code <- as.factor(env_mod_data$site_code)
env_mod_data$Urchins_scale <- scale(env_mod_data$Urchins)
env_mod_data$Turban_Snails_scale <- scale(env_mod_data$Turban_Snails)
#Total boat counts
env_mod_data$mean_total_boat_counts_1km_scale <- scale(env_mod_data$mean_total_boat_counts_1km)
env_mod_data$mean_total_boat_counts_1.5km_scale <- scale(env_mod_data$mean_total_boat_counts_1.5km)
env_mod_data$mean_total_boat_counts_2km_scale <- scale(env_mod_data$mean_total_boat_counts_2km)

levels(env_mod_data$dominant_score) <- list(loose = "3", solid = "4")

regions <- c("region1", "region2", "region3", "region4")
env_mod_data[regions] <- lapply(env_mod_data[regions], factor)

## Reordering the factor levels of region1 so that Deer group bamfield is the
#intercept and renaming for plotting
levels(env_mod_data$region1) <- list(DGB = "Deer group/Bamfield",
                                     BG = "Broken Group",
                                     DIEC = "Deep Imperial Eagle Channel",
                                     SR = "Sarita" )


##Adding a column to the env mod data that is the snail to urchin ratio
env_mod_data <- env_mod_data%>%
  mutate(sn_urch_ratio = log10((Turban_Snails + 1) / (Urchins + 1)),
         urchin_dens = (Urchins / 100),
         snail_dens = (Turban_Snails / 100))
```

```{r Creating a factor variable of protection status for the modelling data, include=FALSE}
#In this chunk I am making three objects with the various protection statuses
#of the different sites. protect_status has all sites protected from fishing
#to some degree (RCA, ecological reserve, finfish closure area) and 
#protect_status_park includes the sites within the Broken Group Park that are
#not within the RCA.
protect_status <- c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 
                    0, 0, 0, 1, 1, 0, 0, 0, 0, 0, 1, 1, 1, 1, 0)

protect_status_park <- c(0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0, 0, 1, 1, 1, 1, 0)

#kelp_forest <- c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)


  env_mod_data$protect_status <- as.factor(protect_status)
levels(env_mod_data$protect_status) <- c("No Protection", "Protected")

env_mod_data$protect_status_park <- as.factor(protect_status_park)
levels(env_mod_data$protect_status_park) <- c("No Protection", "Protected")


#Reordering so easier to see which sites are protected in the DF
########This needs to be fixed for the final markdown
env_mod_data <- env_mod_data[, c(seq(1, 11, by = 1), 46, 47, seq(12, 45, by = 1))]
env_mod_data <- env_mod_data[, c(seq(1, 10, by = 1), 46, seq(11, 45), 47 )]

#Will need to confirm which areas were kelp forests based on the photos or
#talking with Jasmin but likely will just be hosie and raymond

#env_mod_data$kelp_forest <- as.factor(kelp_forest)
#levels(env_mod_data$kelp_forest) <- c("No Kelp", "Kelp Forest")
```

```{r Making theme for all subsequent figures, echo=FALSE}
theme.figure <- function (){
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    #axis.line = element_line(colour = "black"),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12, face = "plain"),
    legend.text = element_text(size = 10, face = "plain"),
    legend.title = element_text(size = 14, face = "bold")) 
}

```

# Modelling{.tabset}

#### Testing the association of the categorical variables with chi-square test

```{r Making a contingency table for protection status and region1}
fisher.test(table(unique_site_df$protect_status, unique_site_df$region1))
```
Appears that region and protection status are significantly associated

#### Testing the assumption of equal variance of urchins between regions

```{r}
leveneTest(Urchins ~ region1, data = env_mod_data)
```


## Ecology models{.tabset}

### Red Turban Snail Ecology Models{.tabset}

#### Model 1: Ecology model on the snail counts with temperature, openness, and bottom type

```{r Turban snail ecology/environmental model}
ecology_snail_mod <- gam(log(Turban_Snails + 1) ~ s(sst_scale)+
                               s(norm_openness_scale)+
                           dominant_score,
                                # s(region1, bs = "re"), #Random effect of region does not have an impact on the snails
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

```{r Turban snail ecology/environmental model output, echo=FALSE}
gam.check(ecology_snail_mod)
plot(ecology_snail_mod, all.terms = TRUE)
summary(ecology_snail_mod)
```

Initially did this model with openness as a smooth term but fitting produced a linear effect of openness so re fit model with linear openness.

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail ecology model residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_ecology_snail_mod <- residuals(ecology_snail_mod, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_ecology_snail_mod, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_ecology_snail_mod, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation for both methods when using a two sided test.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing turban snail ecology model for concurvity, echo=FALSE}
concurvity(ecology_snail_mod, full = FALSE)
concurvity(ecology_snail_mod, full = TRUE)
```
</div>

Temperature and openness seem to be somewhat concurve. Temp having no impact on
the model (being removed by select).


#### Model 2: Ecology model on the snail counts with temperature, openness, and bottom type (linear openness)

```{r Turban snail ecology/environmental model linear}
ecology_snail_mod_linear <- gam(log(snail_dens + 0.01) ~ s(sst_scale)+
                               norm_openness_scale+
                           dominant_score,
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

```{r Turban snail ecology/environmental model linear output, echo=FALSE}
gam.check(ecology_snail_mod_linear)
plot(ecology_snail_mod_linear, all.terms = TRUE)
summary(ecology_snail_mod_linear)
```
Results showing the same trends with normalized openness included as a linear term.
Ultimately the results of the environmental model for turban snail abundance/counts
indicate the turban snails are not controlled by temperature or substrate type,
but they are strongly controlled by the openness gradient, with less snails in
more exposed sites. 

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail ecology model linear residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_ecology_snail_mod_linear <- residuals(ecology_snail_mod_linear, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_ecology_snail_mod_linear, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_ecology_snail_mod_linear, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation for both methods when using a two sided test.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing turban snail ecology model linear for concurvity, echo=FALSE}
concurvity(ecology_snail_mod_linear, full = FALSE)
concurvity(ecology_snail_mod_linear, full = TRUE)
```
</div>

Concurvity seems to be improved when having openness as a linear term. Maybe not
and it is just how the function calculates it differently or omits it.


##### Extracting the predicted values for snails based on openness for plotting

```{r extracting predicted snail counts based on openness values}
snail_intercept <- coef(ecology_snail_mod_linear)[1]
snail_op_slope <- coef(ecology_snail_mod_linear)["norm_openness_scale"]

#Calculating the expected values based on the equation
snail_op_pred_data <- env_mod_data
snail_op_pred_data$predicted_snails <- snail_intercept + (snail_op_slope * snail_op_pred_data$norm_openness_scale)
```


```{r Using the methods from Isabel to extract the predicted values}
#Making new dataframe for predicted values

new_data_openness <- expand.grid(
  norm_openness_scale = env_mod_data$norm_openness_scale,
  sst_scale = mean(env_mod_data$sst_scale),
  dominant_score = env_mod_data$dominant_score)

new_data_openness$dominant_score <- as.factor(new_data_openness$dominant_score)

#Predicting the numbers of turban snails and the confidence intervals
predictions_openness <- predict.gam(ecology_snail_mod_linear,
                                    type = "response",
                                    se.fit = TRUE)

new_data_openness_with_preds <- env_mod_data
new_data_openness_with_preds$snail_predictions <- predictions_openness$fit
new_data_openness_with_preds$lower_ci <- (predictions_openness$fit - 1.96 * predictions_openness$se.fit) 
new_data_openness_with_preds$upper_ci <- (predictions_openness$fit + 1.96 * predictions_openness$se.fit)
new_data_openness_with_preds$snail_predictions_normal <- (exp(new_data_openness_with_preds$snail_predictions) - 1)
new_data_openness_with_preds$lower_ci_normal <- (exp(predictions_openness$fit - 1.96 * predictions_openness$se.fit) - 1)
new_data_openness_with_preds$upper_ci_normal <- (exp(predictions_openness$fit + 1.96 * predictions_openness$se.fit) - 1)

ggplot(new_data_openness_with_preds, aes(x = norm_openness, Turban_Snails))+
  geom_line(aes(y = snail_predictions))+
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci))

ggplot(new_data_openness_with_preds, aes(x = norm_openness, y = log(Turban_Snails +1)))+
  geom_point()+
  geom_smooth(method = "lm", data = new_data_openness_with_preds, mapping = aes(x = norm_openness, y = snail_predictions))+
  theme.figure.report()

ggplot(new_data_openness_with_preds, aes(x = norm_openness, y = snail_predictions))+
  geom_point()+
  geom_smooth(method = "lm", data = new_data_openness_with_preds, mapping = aes(x = norm_openness, y = snail_predictions))+
  theme.figure.report()
  
ggplot(new_data_openness_with_preds, aes(x = norm_openness, log(Turban_Snails +1)))+
  geom_point()+
  geom_smooth(method = "lm")+
  theme.figure.report()

ggplot(new_data_openness_with_preds, aes(x = norm_openness, y = Turban_Snails))+
  geom_point()+
  geom_smooth(method = "gam", data = new_data_openness_with_preds, mapping = aes(x = norm_openness, y = snail_predictions_normal))+
  theme.figure.report()
```


```{r Using methods from stickleback DS, include=FALSE, echo=FALSE}
#Creating summary stats for the model
sums_stats_snail_ecol_gam <- ggpredict(ecology_snail_mod_linear,
                                       terms = c("norm_openness_scale"))
  dplyr::rename("norm_openness_scale" = x,
               "log_snails" = predicted)%>%
  as.data.frame()
```


#### Model 3: Ecology model on the snail counts with temperature, openness, and bottom type and random effect (linear openness)

```{r Turban snail ecology/environmental model linear re}
ecology_snail_mod_linear_re <- gam(log(Turban_Snails + 1) ~ s(sst_scale)+
                               norm_openness_scale +
                           dominant_score+
                                s(region1, bs = "re"),
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

```{r Turban snail ecology/environmental model linear re output, echo=FALSE}
gam.check(ecology_snail_mod_linear_re)
plot(ecology_snail_mod_linear_re, all.terms = TRUE)
summary(ecology_snail_mod_linear_re)
```
Results showing the same trends with normalized openness included as a linear term.
Ultimately the results of the environmental model for turban snail abundance/counts
indicate the turban snails are not controlled by temperature or substrate type,
but they are strongly controlled by the openness gradient, with less snails in
more exposed sites. 

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail ecology model linear re residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_ecology_snail_mod_linear_re <- residuals(ecology_snail_mod_linear_re, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_ecology_snail_mod_linear_re, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_ecology_snail_mod_linear_re, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation for both methods when using a two sided test.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing turban snail ecology model linear re for concurvity, echo=FALSE}
concurvity(ecology_snail_mod_linear_re, full = FALSE)
concurvity(ecology_snail_mod_linear_re, full = TRUE)
```
</div>

Random effect of region is highly concurve.

#### Model 4: Ecology model on the snail counts with temperature, openness, and bottom type and fixed effect of region (linear openness)

```{r Turban snail ecology/environmental model linear FE}
ecology_snail_mod_linear_fe <- gam(log(Turban_Snails + 1) ~ s(sst_scale)+
                               norm_openness_scale +
                           dominant_score+
                                region1,
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

```{r Turban snail ecology/environmental model linear fe output, echo=FALSE}
gam.check(ecology_snail_mod_linear_fe)
plot(ecology_snail_mod_linear_fe, all.terms = TRUE)
summary(ecology_snail_mod_linear_fe)
```
Results showing the same trends with normalized openness included as a linear term.
Region as a fixed effect does not seem to be significant.

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail ecology model linear fe residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_ecology_snail_mod_linear_fe <- residuals(ecology_snail_mod_linear_fe, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_ecology_snail_mod_linear_fe, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_ecology_snail_mod_linear_fe, lw_inverse,
           alternative = "two.sided")
```
</div>

Slight autocorrelation for inverse distance method

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing turban snail ecology model linear fe for concurvity, echo=FALSE}
concurvity(ecology_snail_mod_linear_fe, full = FALSE)
concurvity(ecology_snail_mod_linear_fe, full = TRUE)
```
</div>

Temp and para are highly concurve.


#### Comparing Turban snail ecology models

```{r AIC of Turban snail ecology models}
AICc(ecology_snail_mod_linear,
    ecology_snail_mod_linear_re,
    ecology_snail_mod_linear_fe)%>%
  arrange(AICc)%>%
  mutate(delta_AICc = AICc - AICc[1])
```

AIC indicates that it does not improve the model by adding a random effect of 
region. The random effect is basically being removed from the model by the 
select = TRUE argument. This means that for the snail ecology model I will 
pick the more parsimonius model without the random effect.

Also, having region as a fixed effect makes the AIC much worse.
Trying with both methods (REML and ML) the model with the fixed effect is much
worse.

### Red Sea urchin ecology models{.tabset}

#### Model 5: Ecology model on the urchin counts with temperature, openness, and bottom type

```{r Urchin ecology/environmental model}
ecology_urchin_mod <- gam(urchin_dens ~ s(sst_scale)+
                               s(norm_openness_scale)+
                           dominant_score,
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

```{r Urchin ecology/environmental model output, echo=FALSE}
set.seed(299)
gam.check(ecology_urchin_mod, rep = 500)
plot(ecology_urchin_mod, all.terms = TRUE)
summary(ecology_urchin_mod)
shapiro.test(ecology_urchin_mod$residuals)
```
Tried with other distributions as well but gaussian appears to be the best.
Confirm this with Amanda.

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Urchin ecology model residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_ecology_urchin_mod <- residuals(ecology_urchin_mod, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_ecology_urchin_mod, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_ecology_urchin_mod, lw_inverse,
           alternative = "two.sided")
```
</div>

Some positive autocorrelation but only with the inverse distance method. The 
neighbour method has no autocorrelation.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing urchin ecology model for concurvity, echo=FALSE}
concurvity(ecology_urchin_mod, full = FALSE)
concurvity(ecology_urchin_mod, full = TRUE)
```
</div>

Concurvity is high for temperature and openness. Both are being removed by the 
double penalty.


#### Model 6: Ecology model on the urchin counts with temperature, openness, and bottom type (random effect of region)

```{r Urchin ecology/environmental model region re}
ecology_urchin_mod_re <- gam(urchin_dens ~ s(sst_scale)+
                               s(norm_openness_scale)+
                           dominant_score+
                                s(region1, bs = "re"),
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

```{r Urchin ecology/environmental model region re output, echo=FALSE}
set.seed(299)
gam.check(ecology_urchin_mod_re, rep = 500)
plot(ecology_urchin_mod_re, all.terms = TRUE)
summary(ecology_urchin_mod_re)
shapiro.test(ecology_urchin_mod_re$residuals)
variance_comp(ecology_urchin_mod_re)
```
Gaussian looks ok. When adding the random effect of region the openness becomes
borderline and negative. 

```{r Extracting predictions for the dominant score}
pred_urchin_dom_score <- tidy(ecology_urchin_mod_re, parametric = TRUE)
pred_urchin_dom_score
pred_urchin_dom_score$term <- as.factor(pred_urchin_dom_score$term)
levels(pred_urchin_dom_score$term) <- c("loose", "solid")

#pulling out intercept to adjust for this when plotting
dom_substrate_urchin_intercept <- pred_urchin_dom_score%>%
  filter(term == "loose")%>%
  pull(estimate)

#adjusting the non-intercept term to include the intercept
pred_urchin_dom_score <- pred_urchin_dom_score%>%
  mutate(adj_estimate = ifelse(term != "loose", estimate + dom_substrate_urchin_intercept, estimate))
```

```{r Trying to predict the urchin counts from openness}
urchin_ecology_data_for_openness <- data.frame(
  dominant_score = factor("loose", levels = levels(env_mod_data$dominant_score)),
  sst_scale = mean(env_mod_data$sst_scale),
  norm_openness_scale = seq(min(env_mod_data$norm_openness_scale), max(env_mod_data$norm_openness_scale), length.out = 100),
  region1 = factor("DGB", levels = levels(env_mod_data$region1))
)

#getting mean and sd for norm_openness
norm_openness_mean <- mean(env_mod_data$norm_openness)
norm_openness_sd <- sd(env_mod_data$norm_openness)

#predicting the data
predicted_openness_effect_urchin <- predict(ecology_urchin_mod_re, newdata = urchin_ecology_data_for_openness, se.fit = TRUE)
urchin_ecology_data_for_openness$fit <- predicted_openness_effect_urchin$fit
urchin_ecology_data_for_openness$upper <- predicted_openness_effect_urchin$fit + 1.96 * predicted_openness_effect_urchin$se.fit
urchin_ecology_data_for_openness$lower <- predicted_openness_effect_urchin$fit - 1.96 * predicted_openness_effect_urchin$se.fit
urchin_ecology_data_for_openness$norm_openness_original <- urchin_ecology_data_for_openness$norm_openness_scale * norm_openness_sd + norm_openness_mean
```

```{r predicting urchin density based on dominant seafloor}
urchin_ecology_data_for_seafloor <- data.frame(
  dominant_score = levels(env_mod_data$dominant_score),
  sst_scale = mean(env_mod_data$sst_scale),
  norm_openness_scale = mean(env_mod_data$norm_openness_scale),
  region1 = factor("DGB", levels = levels(env_mod_data$region1))
)


#predicting the data
predicted_sf_effect_urchin <- predict(ecology_urchin_mod_re, newdata = urchin_ecology_data_for_seafloor, se.fit = TRUE)
urchin_ecology_data_for_seafloor$fit <- predicted_sf_effect_urchin$fit
urchin_ecology_data_for_seafloor$upper <- predicted_sf_effect_urchin$fit + 1.96 * predicted_sf_effect_urchin$se.fit
urchin_ecology_data_for_seafloor$lower <- predicted_sf_effect_urchin$fit - 1.96 * predicted_sf_effect_urchin$se.fit
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Urchin ecology model region re residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_ecology_urchin_mod_re <- residuals(ecology_urchin_mod_re, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_ecology_urchin_mod_re, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_ecology_urchin_mod_re, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing urchin ecology model re for concurvity, echo=FALSE}
concurvity(ecology_urchin_mod_re, full = FALSE)
concurvity(ecology_urchin_mod_re, full = TRUE)
```
</div>

Temp and region are basically fully concurve.


#### Model 7: Ecology model on the urchin counts with temperature, openness, and bottom type (fixed effect of region)

```{r Urchin ecology/environmental model region fe}
ecology_urchin_mod_fe <- gam(urchin_dens ~ s(sst_scale)+
                               s(norm_openness_scale)+
                           dominant_score+
                                region1,
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

```{r Urchin ecology/environmental model region fe output, echo=FALSE}
set.seed(299)
gam.check(ecology_urchin_mod_fe, rep = 500)
plot(ecology_urchin_mod_fe, all.terms = TRUE)
summary(ecology_urchin_mod_fe)
shapiro.test(ecology_urchin_mod_fe$residuals)
```
Gaussian looks ok. When adding the fixed effect of region the openness becomes
borderline and negative. Also dominant substrate type becomes significant with 
more urchins on solid substrate.

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Urchin ecology model region fe residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_ecology_urchin_mod_fe <- residuals(ecology_urchin_mod_fe, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_ecology_urchin_mod_fe, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_ecology_urchin_mod_fe, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing urchin ecology model fe for concurvity, echo=FALSE}
concurvity(ecology_urchin_mod_fe, full = FALSE)
concurvity(ecology_urchin_mod_fe, full = TRUE)
```
</div>

Temp is very concurve.


##### Testing the model for VIF stuff
```{r}
vif(ecology_urchin_mod_fe)

performance::check_collinearity(ecology_urchin_mod_fe,
                   component = c("all"))

fisher.test(table(env_mod_data$dominant_score, env_mod_data$region1))
#Not significant
```

#### Comparing the urchin ecology models

```{r AIC of Urchin ecology models}
AICc(ecology_urchin_mod,
     ecology_urchin_mod_re,
     ecology_urchin_mod_fe)%>%
  arrange(AICc)%>%
  mutate(delta_AICc = AICc - AICc[1])
```
When testing with REML, the model with the random effect is much better than 
the model with no random effect. When comparing with ML, the model with the 
fixed effect is much better than the model without the fixed effect. In both 
the model with the fixed effect is better than the model with the random effect.

When repeating with urchin density and REML the fixed and random are basically the same.

## Fishing models{.tabset}

### Confirming the validity of protection status with DFO flyover data

For the protection status categories, I have grouped them as sites that 
experience at least a ban on fishing for part of the year as protected and all
others as non-protected. This means that the sites that are in the Broken group
but are not within the RCA are counted as not protected.

In order to confirm/support our assumptions of which sites are protected,
I am using the DFO flyover data to compare the numbers of boats around the 
different sites.

Based on the boxplots below it seems evident that the protected sites selected
are experiencing less fishing pressure and less overall human/boat pressure.

#### Plot indicating which regions protected and non-protected sites are located in

It seems that our current region scheme has no protected sites located within the deer group/bamfield,
will need to look into if this is a problem.

```{r Plot indicating the groupings of protected/non-protected and region, echo=FALSE}
ggplot(env_mod_data, aes(y = region1, x = protect_status))+
  geom_jitter(width = 0.2, height = 0.2)+
  labs(x = "Protection Status", y = "Region")+
  theme.figure()
```



#### Boxplots of rec boat counts between protection levels

```{r Boxplots of the rec boats at the three distances between protection levels, echo=FALSE}
ggplot(env_mod_data, aes(x = protect_status, y = mean_rec_boat_counts_1km))+
  geom_boxplot()+
  geom_jitter()+
  theme.figure()

ggplot(env_mod_data, aes(x = protect_status, y = mean_rec_boat_counts_1.5km))+
  geom_boxplot()+
  geom_jitter()+
  theme.figure()

ggplot(env_mod_data, aes(x = protect_status, y = mean_rec_boat_counts_2km))+
  geom_boxplot()+
  geom_jitter()+
  theme.figure()
```

#### Boxplots of total boat counts between protection levels

```{r Boxplots of the total boats at the three distances between protection levels, echo=FALSE}
ggplot(env_mod_data, aes(x = protect_status, y = mean_total_boat_counts_1km))+
  geom_boxplot()+
  geom_jitter()+
  theme.figure()

ggplot(env_mod_data, aes(x = protect_status, y = mean_total_boat_counts_1.5km))+
  geom_boxplot()+
  geom_jitter()+
  theme.figure()

ggplot(env_mod_data, aes(x = protect_status, y = mean_total_boat_counts_2km))+
  geom_boxplot()+
  geom_jitter()+
  theme.figure()
```

#### Comparing the rec boat counts between sites in the broken group that have been labelled as protected or non-protected

```{r Comparing the rec boat counts between sites in the broken group that have been labelled as protected or non-protected, echo=FALSE}
ggplot(env_mod_data[c(8, 9, 10, 11, 12, 13),], aes(x = protect_status, y = mean_rec_boat_counts_1km))+
  geom_boxplot()+
  geom_point()+
  theme.figure()

ggplot(env_mod_data[c(8, 9, 10, 11, 12, 13),], aes(x = protect_status, y = mean_rec_boat_counts_1.5km))+
  geom_boxplot()+
  geom_point()+
  theme.figure()

ggplot(env_mod_data[c(8, 9, 10, 11, 12, 13),], aes(x = protect_status, y = mean_rec_boat_counts_2km))+
  geom_boxplot()+
  geom_point()+
  theme.figure()
```

#### Comparing the total boat counts between sites in the broken group that have been labelled as protected or non-protected

```{r Comparing the total boat counts between sites in the broken group that have been labelled as protected or non-protected, echo=FALSE}
ggplot(env_mod_data[c(8, 9, 10, 11, 12, 13),], aes(x = protect_status, y = mean_total_boat_counts_1km))+
  geom_boxplot()+
  geom_point()+
  theme.figure()

ggplot(env_mod_data[c(8, 9, 10, 11, 12, 13),], aes(x = protect_status, y = mean_total_boat_counts_1.5km))+
  geom_boxplot()+
  geom_point()+
  theme.figure()

ggplot(env_mod_data[c(8, 9, 10, 11, 12, 13),], aes(x = protect_status, y = mean_total_boat_counts_2km))+
  geom_boxplot()+
  geom_point()+
  theme.figure()
```



#### Testing the differences in rec boat counts between protected and non-protected with a glm

```{r Making a subset df with only one observation for each site}
unique_site_df <- env_mod_data%>%
  ungroup()%>%
  select(site_code, protect_status,mean_rec_boat_counts_1km, mean_rec_boat_counts_1.5km, mean_rec_boat_counts_2km, region1)%>%
  distinct()
```


```{r Testing the variance between protection status}
leveneTest(mean_rec_boat_counts_1km ~ protect_status, data = unique_site_df)
leveneTest(mean_rec_boat_counts_1.5km ~ protect_status, data = unique_site_df)
leveneTest(mean_rec_boat_counts_2km ~ protect_status, data = unique_site_df)
```
Equal variance for 1km, but not for 1.5 or 2

```{r Normality of rec boat counts within protected sites}
#Making data frame with only sites that are protected
rec_boat_protected <- subset(unique_site_df, protect_status == "Protected")

shapiro.test(rec_boat_protected$mean_rec_boat_counts_1km)
hist(rec_boat_protected$mean_rec_boat_counts_1km)

shapiro.test(rec_boat_protected$mean_rec_boat_counts_1.5km)
hist(rec_boat_protected$mean_rec_boat_counts_1.5km)

shapiro.test(rec_boat_protected$mean_rec_boat_counts_2km)
hist(rec_boat_protected$mean_rec_boat_counts_2km)
```
Only the 1.5 km boat counts are normal

```{r Normality of rec boat counts within non-protected sites}
#Making data frame with only sites that are non-protected
rec_boat_non_protected <- subset(unique_site_df, protect_status == "No Protection")

shapiro.test(rec_boat_non_protected$mean_rec_boat_counts_1km)
hist(rec_boat_non_protected$mean_rec_boat_counts_1km)

shapiro.test(rec_boat_non_protected$mean_rec_boat_counts_1.5km)
hist(rec_boat_non_protected$mean_rec_boat_counts_1.5km)

shapiro.test(rec_boat_non_protected$mean_rec_boat_counts_2km)
hist(rec_boat_non_protected$mean_rec_boat_counts_2km)
#None of the distributions are normal
```
2km is normal but the others are not

##### Doing statistical test of differences in rec boat counts between protection status

Since the data do not all fit the assumptions of normality or equal variances we
will need to use wilcoxon rank sum tests to compare.

```{r Wilcoxon rank sum tests of mean rec boat counts between protected and non-protected sites}
wilcox_1km <- wilcox.test(mean_rec_boat_counts_1km ~ protect_status, data = unique_site_df, alternative = "t")
wilcox_1.5km <- wilcox.test(mean_rec_boat_counts_1.5km ~ protect_status, data = unique_site_df, alternative = "t")
wilcox_2km <- wilcox.test(mean_rec_boat_counts_2km ~ protect_status, data = unique_site_df, alternative = "t")

print(wilcox_1km)
print(wilcox_1.5km)
print(wilcox_2km)
```
All of the wilcoxon rank sum tests indicate that there is a highly significant
difference in the number of rec fishing boats between protected and non-protected
sites. Based on the boxplots we can say that there are significantly more rec
boats at non-protected sites than protected sites.

```{r Mean rec boat counts 1km glm}
rec_boat_1km_mod <- glmmTMB(mean_rec_boat_counts_1km ~ protect_status,
                            data = env_mod_data,
                             family = gaussian(link = "identity"))
```





```{r Snail count fishing model with protection linear output, echo=FALSE}
#Checking resids
rec_boat_1km_mod_resids <- simulateResiduals(rec_boat_1km_mod)
plot(rec_boat_1km_mod_resids)

hist(residuals(rec_boat_1km_mod))
shapiro.test(residuals(rec_boat_1km_mod))

#Looking at model results
summary(rec_boat_1km_mod)
```

### Red Turban snail fishing models{.tabset}

#### Plotting counts of snails between protected and non-protected

```{r Plotting counts of snails between protected and non-protected, echo=FALSE}
ggplot(env_mod_data, aes(x = protect_status, y = Turban_Snails))+
  geom_boxplot()+
  geom_jitter(height = 0)+
  labs(x = "Protection Status", y = "Turban Snail counts")+
  theme.figure()
```

Seeing a weak trend of more snails in the protected sites when compared to the non-
protected sites.



#### Model 8: Fishing model on the snail counts with protection and openness (glm)

For this model we are seeing that protection status has a non-significant
positive effect on the log snail counts

```{r Checking variance between protected and non-protected for turban snails, echo=FALSE}
leveneTest(sqrt(Turban_Snails) ~ protect_status, data = env_mod_data)
```
Variance is equal for raw turban snail counts but significantly different for 
log transformed turban snail counts

```{r Snail count fishing  model with protection and openness linear}
snail_fishing_mod_linear <- glmmTMB(log(Turban_Snails +1) ~ norm_openness_scale +
                                      protect_status,
                                    data = env_mod_data,
                                    family = gaussian(link = "identity"))
```

```{r Snail count fishing model with protection and openness linear output, echo=FALSE}
#Checking resids
snail_fishing_mod_linear_resids <- simulateResiduals(snail_fishing_mod_linear)
plot(snail_fishing_mod_linear_resids)

hist(residuals(snail_fishing_mod_linear))
shapiro.test(residuals(snail_fishing_mod_linear))

#residuals_test <- residuals(snail_fishing_mod_linear, type = "pearson")
#fitted_values_test <- fitted(snail_fishing_mod_linear)

#ggplot(data = data.frame(fitted_values_test, residuals_test), aes(x = fitted_values_test, y = residuals_test)) +
#  geom_point() +
#  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
#  labs(x = "Fitted Values", y = "Pearson Residuals", title = "Residuals vs Fitted") +
#  theme_minimal()

#Looking at model results
summary(snail_fishing_mod_linear)
```


```{r Plotting the linear snail fishing model with protection and openness, echo=FALSE}
pred_snail_fishing_mod_linear_openness <- ggpredict(snail_fishing_mod_linear, terms = c("norm_openness_scale", "protect_status"))


plot(pred_snail_fishing_mod_linear_openness)


pred_snail_fishing_mod_linear_protect_status <- ggpredict(snail_fishing_mod_linear, terms = "protect_status")
#plot(pred_snail_fishing_mod_linear_protect_status)

ggplot(pred_snail_fishing_mod_linear_protect_status, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Turban Snail Counts", x = "Protection Status")+
  theme.figure()

#Using emmeans to represent the protection status
#plot(emmeans(snail_fishing_mod_linear, ~ protect_status, comparisons = TRUE))
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail fishing model linear residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_snail_fishing_mod_linear <- residuals(snail_fishing_mod_linear, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_snail_fishing_mod_linear, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_snail_fishing_mod_linear, lw_inverse,
           alternative = "two.sided")
```
</div>
No autocorrelation with neighbours but it is significant with inverse distance.


#### Model 9: Fishing model on the snail counts with protection and openness and random effect (glmm)

For this model we are seeing that protection status has a non-significant
positive effect on the log snail counts

```{r Checking variance between protected and non-protected for turban snails re, include=FALSE,echo=FALSE}
leveneTest(log(Turban_Snails + 1) ~ protect_status, data = env_mod_data)
```
Variance is equal for raw turban snail counts but significantly different for 
log transformed turban snail counts

```{r Snail count fishing  model with protection and openness linear}
snail_fishing_mod_linear_re <- glmmTMB(log(Turban_Snails +1) ~ norm_openness_scale +
                                      protect_status + (1|region1),
                                    data = env_mod_data,
                                    family = gaussian(link = "identity"))
```

```{r Snail count fishing model with protection and openness linear re output, echo=FALSE}
#Checking resids
snail_fishing_mod_linear_re_resids <- simulateResiduals(snail_fishing_mod_linear_re)
plot(snail_fishing_mod_linear_re_resids)

hist(residuals(snail_fishing_mod_linear_re))
shapiro.test(residuals(snail_fishing_mod_linear_re))

#residuals_test <- residuals(snail_fishing_mod_linear, type = "pearson")
#fitted_values_test <- fitted(snail_fishing_mod_linear)

#ggplot(data = data.frame(fitted_values_test, residuals_test), aes(x = fitted_values_test, y = residuals_test)) +
#  geom_point() +
#  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
#  labs(x = "Fitted Values", y = "Pearson Residuals", title = "Residuals vs Fitted") +
#  theme_minimal()

#Looking at model results
summary(snail_fishing_mod_linear_re)
```


```{r Plotting the linear snail fishing model with protection and openness re, echo=FALSE}
pred_snail_fishing_mod_linear_re_openness <- ggpredict(snail_fishing_mod_linear_re, terms = c("norm_openness_scale", "protect_status"))


plot(pred_snail_fishing_mod_linear_re_openness)


pred_snail_fishing_mod_linear_re_protect_status <- ggpredict(snail_fishing_mod_linear_re, terms = "protect_status")
#plot(pred_snail_fishing_mod_linear_protect_status)

ggplot(pred_snail_fishing_mod_linear_re_protect_status, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Turban Snail Counts", x = "Protection Status")+
  theme.figure()

#Using emmeans to represent the protection status
#plot(emmeans(snail_fishing_mod_linear, ~ protect_status, comparisons = TRUE))
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail fishing model linear re residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_snail_fishing_mod_linear_re <- residuals(snail_fishing_mod_linear_re, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_snail_fishing_mod_linear_re, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_snail_fishing_mod_linear_re, lw_inverse,
           alternative = "two.sided")
```
</div>
No autocorrelation with neighbours but it is significant with inverse distance.

#### Comparing Turban snail fishing models with openness

```{r AIC of the Turban snail fishing models with openness}
AICc(snail_fishing_mod_linear, 
    snail_fishing_mod_linear_re)%>%
  arrange(AICc)%>%
  mutate(delta_AICc = AICc - AICc[1])

AICc(snail_fishing_mod_linear_no_op,
     ecology_snail_mod_linear)
```

```{r ANOVA of the turban snail fishing models with openness}
anova(snail_fishing_mod_linear, snail_fishing_mod_linear_re)
```

It appears that the random effect of region is not explaining any of the variance
in the data and both AICc and ANOVA prefer the model without the random effect.

#### Model 10: Fishing model on the snail counts with just protection (glm)

```{r Testing the snail counts for equal variance among protection levels}
leveneTest(snail_dens ~ protect_status, data = env_mod_data)
#normal is good
leveneTest(log(snail_dens + 0.01) ~ protect_status, data = env_mod_data)
#Log transformed is not equal variance so will try with square root transform

leveneTest(sqrt(snail_dens) ~ protect_status, data = env_mod_data)
#There is equal variance with the square rooted values
```


```{r Snail count fishing model with protection linear}
snail_fishing_mod_linear_no_op <- glmmTMB(sqrt(snail_dens) ~
                                      protect_status,
                                    data = env_mod_data,
                                    family = gaussian(link = "identity"))
```

```{r Snail count fishing model with protection linear output, echo=FALSE}
#Checking resids
snail_fishing_mod_linear_no_op_resids <- simulateResiduals(snail_fishing_mod_linear_no_op)
plot(snail_fishing_mod_linear_no_op_resids)

hist(residuals(snail_fishing_mod_linear_no_op))
shapiro.test(residuals(snail_fishing_mod_linear_no_op))

#residuals_test <- residuals(snail_fishing_mod_linear, type = "pearson")
#fitted_values_test <- fitted(snail_fishing_mod_linear)

#ggplot(data = data.frame(fitted_values_test, residuals_test), aes(x = fitted_values_test, y = residuals_test)) +
#  geom_point() +
#  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
#  labs(x = "Fitted Values", y = "Pearson Residuals", title = "Residuals vs Fitted") +
#  theme_minimal()

#Looking at model results
summary(snail_fishing_mod_linear_no_op)
performance::r2(snail_fishing_mod_linear_no_op)
```


```{r Plotting the linear snail fishing model with protection, echo=FALSE}
pred_snail_fishing_mod_linear_no_op_protect_status <- ggpredict(snail_fishing_mod_linear_no_op, terms = "protect_status")
#plot(pred_snail_fishing_mod_linear_protect_status)

ggplot(pred_snail_fishing_mod_linear_no_op_protect_status, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Turban Snail Density (n/m2)", x = "Protection Status")+
  theme.figure()

#Using emmeans to represent the protection status
#plot(emmeans(snail_fishing_mod_linear, ~ protect_status, comparisons = TRUE))
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail fishing model linear no openness residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_snail_fishing_mod_linear_no_op <- residuals(snail_fishing_mod_linear_no_op, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_snail_fishing_mod_linear_no_op, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_snail_fishing_mod_linear_no_op, lw_inverse,
           alternative = "two.sided")
```
</div>
No autocorrelation.

#### Model 11: Fishing model on the snail counts with protection and a random effect (glmm)

```{r Snail count fishing model with protection linear re}
snail_fishing_mod_linear_no_op_re <- glmmTMB(sqrt(snail_dens) ~
                                      protect_status + (1|region1),
                                    data = env_mod_data,
                                    family = gaussian(link = "identity"))
```

```{r Snail count fishing model with protection linear re output, echo=FALSE}
#Checking resids
snail_fishing_mod_linear_no_op_re_resids <- simulateResiduals(snail_fishing_mod_linear_no_op_re)
plot(snail_fishing_mod_linear_no_op_re_resids)

hist(residuals(snail_fishing_mod_linear_no_op_re))
shapiro.test(residuals(snail_fishing_mod_linear_no_op_re))

#residuals_test <- residuals(snail_fishing_mod_linear, type = "pearson")
#fitted_values_test <- fitted(snail_fishing_mod_linear)

#ggplot(data = data.frame(fitted_values_test, residuals_test), aes(x = fitted_values_test, y = residuals_test)) +
#  geom_point() +
#  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
#  labs(x = "Fitted Values", y = "Pearson Residuals", title = "Residuals vs Fitted") +
#  theme_minimal()

#Looking at model results
summary(snail_fishing_mod_linear_no_op_re)
```


```{r Plotting the linear snail fishing model with protection re, echo=FALSE}
pred_snail_fishing_mod_linear_no_op_re_protect_status <- ggpredict(snail_fishing_mod_linear_no_op_re, terms = "protect_status")
#plot(pred_snail_fishing_mod_linear_protect_status)

ggplot(pred_snail_fishing_mod_linear_no_op_re_protect_status, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Turban Snail Counts", x = "Protection Status")+
  theme.figure()

#Using emmeans to represent the protection status
#plot(emmeans(snail_fishing_mod_linear, ~ protect_status, comparisons = TRUE))
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail fishing model linear no openness re residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_snail_fishing_mod_linear_no_op_re <- residuals(snail_fishing_mod_linear_no_op_re, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_snail_fishing_mod_linear_no_op_re, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_snail_fishing_mod_linear_no_op_re, lw_inverse,
           alternative = "two.sided")
```
</div>
No autocorrelation.


#### Model 12: Fishing model on the snail counts with protection and a fixed effect of region (glm)

```{r Snail count fishing model with protection linear fe}
snail_fishing_mod_linear_no_op_fe <- glmmTMB(sqrt(Turban_Snails) ~
                                      protect_status + region1,
                                    data = env_mod_data,
                                    family = gaussian(link = "identity"))
```

```{r Snail count fishing model with protection linear fe output, echo=FALSE}
#Checking resids
snail_fishing_mod_linear_no_op_fe_resids <- simulateResiduals(snail_fishing_mod_linear_no_op_fe)
plot(snail_fishing_mod_linear_no_op_fe_resids)

hist(residuals(snail_fishing_mod_linear_no_op_fe))
shapiro.test(residuals(snail_fishing_mod_linear_no_op_fe))

#residuals_test <- residuals(snail_fishing_mod_linear, type = "pearson")
#fitted_values_test <- fitted(snail_fishing_mod_linear)

#ggplot(data = data.frame(fitted_values_test, residuals_test), aes(x = fitted_values_test, y = residuals_test)) +
#  geom_point() +
#  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
#  labs(x = "Fitted Values", y = "Pearson Residuals", title = "Residuals vs Fitted") +
#  theme_minimal()

#Looking at model results
summary(snail_fishing_mod_linear_no_op_fe)
```


```{r Plotting the linear snail fishing model with protection fe, echo=FALSE}
pred_snail_fishing_mod_linear_no_op_fe_protect_status <- ggpredict(snail_fishing_mod_linear_no_op_fe, terms = "protect_status")
#plot(pred_snail_fishing_mod_linear_protect_status)

ggplot(pred_snail_fishing_mod_linear_no_op_fe_protect_status, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Turban Snail Counts", x = "Protection Status")+
  theme.figure()

#Using emmeans to represent the protection status
#plot(emmeans(snail_fishing_mod_linear, ~ protect_status, comparisons = TRUE))
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail fishing model linear no openness fe residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_snail_fishing_mod_linear_no_op_fe <- residuals(snail_fishing_mod_linear_no_op_fe, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_snail_fishing_mod_linear_no_op_fe, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_snail_fishing_mod_linear_no_op_fe, lw_inverse,
           alternative = "two.sided")
```
</div>
No autocorrelation.

#### Comparing Turban Snail fishing models with no openness

```{r AIC of the Turban snail fishing models}
AICc(snail_fishing_mod_linear_no_op, 
    snail_fishing_mod_linear_no_op_re,
    snail_fishing_mod_linear_no_op_fe)%>%
  arrange(AICc)%>%
  mutate(delta_AICc = AICc - AICc[1])
```

AIC indicates that the simpler model is better. Adding the additional term adds
~2 AIC which I believe means that it is not improving the model at all, and it 
is just getting a 2 point penalty applied.

### Red Sea Urchin fishing models{.tabset}

#### Plot of urchin counts between protection status

```{r Plot of number of urchins between protected and non-protected sites}
ggplot(env_mod_data, aes(x = protect_status, y = Urchins))+
  geom_boxplot()+
  geom_jitter()+
  labs(x = "Protection Status", y = "Red Sea urchin counts")+
  theme.figure()
```


#### Model 13: Fishing model on the urchin counts (glm)

```{r Checking variance between protected and non-protected for urchins, echo=FALSE}
leveneTest(Urchins ~ protect_status, data = env_mod_data)
```
Equal variance between protected and non-protected sites so assumption of 
homogeneity of variance is met.


```{r Urchin fishing model glm}
urchin_fishing_mod_glm <- glmmTMB(Urchins ~ protect_status,
                               family = gaussian(link = "identity"),
                                   data = env_mod_data)
```

```{r Urchin fishing mod glm output, echo=FALSE}
#Checking resids
urchin_fishing_mod_glm_resids <- simulateResiduals(urchin_fishing_mod_glm)
plot(urchin_fishing_mod_glm_resids)

hist(residuals(urchin_fishing_mod_glm))
shapiro.test(residuals(urchin_fishing_mod_glm))

#Looking at model results
summary(urchin_fishing_mod_glm)
```
Residuals look ok for just using a gaussian distribution. Seeing a non-significant
negative effect of protection status on red sea urchins. That is, there are less
red sea urchins in protected sites.


```{r Plotting the effect of protection status on urchin counts, echo=FALSE}
pred_urchin_fishing_mod_glm <- ggpredict(urchin_fishing_mod_glm, terms = "protect_status")

ggplot(pred_urchin_fishing_mod_glm, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(x = "Protection Status", y = "Predicted Urchin Counts")+
  theme.figure()
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Urchin fishing model glm residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_urchin_fishing_mod_glm <- residuals(urchin_fishing_mod_glm, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 325)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_urchin_fishing_mod_glm, lw,
           alternative = "two.sided")


#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_urchin_fishing_mod_glm, lw_inverse,
           alternative = "two.sided")
```
</div>
Seeing some autocorrelation in the inverse method but not the neighbour method.


#### Model 14: Fishing model on the urchin counts with random effect (glmm)


```{r Urchin fishing model glmm}
urchin_fishing_mod_glmm <- glmmTMB(urchin_dens ~ protect_status + (1|region1),
                               family = gaussian(link = "identity"),
                                   data = env_mod_data)
```

```{r Urchin fishing mod glmm output, echo=FALSE}
#Checking resids
urchin_fishing_mod_glmm_resids <- simulateResiduals(urchin_fishing_mod_glmm)
plot(urchin_fishing_mod_glmm_resids)

hist(residuals(urchin_fishing_mod_glmm))
shapiro.test(residuals(urchin_fishing_mod_glmm))

#Looking at model results
summary(urchin_fishing_mod_glmm)
model_summary <- summary(urchin_fishing_mod_glmm)

z_values <- model_summary$coefficients$cond[, "z value"]

#Looking at some of the variance metrics for the model
performance::r2(urchin_fishing_mod_glmm)
performance::icc(urchin_fishing_mod_glmm)
```
Residuals look ok for just using a gaussian distribution. Seeing a non-significant
negative effect of protection status on red sea urchins. That is, there are less
red sea urchins in protected sites.


```{r Plotting the effect of protection status on urchin counts glmm, echo=FALSE}
pred_urchin_fishing_mod_glmm <- ggpredict(urchin_fishing_mod_glmm, terms = c("protect_status", "region1"))

ggplot(pred_urchin_fishing_mod_glmm, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(x = "Protection Status", y = "Predicted Urchin Counts")+
  theme.figure()
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Urchin fishing model glmm residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_urchin_fishing_mod_glmm <- residuals(urchin_fishing_mod_glmm, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_urchin_fishing_mod_glmm, lw,
           alternative = "two.sided")


#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_urchin_fishing_mod_glmm, lw_inverse,
           alternative = "two.sided")
```
</div>
No autocorrelation with random effect included.

#### Model 15: Fishing model on the urchin counts with fixed effect (glm)


```{r Urchin fishing model fe glm}
urchin_fishing_mod_glm_fe <- glmmTMB(Urchins ~ protect_status + region1,
                               family = gaussian(link = "identity"),
                                   data = env_mod_data)
```

```{r Urchin fishing mod fe glm output, echo=FALSE}
#Checking resids
urchin_fishing_mod_glm_fe_resids <- simulateResiduals(urchin_fishing_mod_glm_fe)
plot(urchin_fishing_mod_glm_fe_resids)

hist(residuals(urchin_fishing_mod_glm_fe))
shapiro.test(residuals(urchin_fishing_mod_glm_fe))

#Looking at model results
summary(urchin_fishing_mod_glm_fe)
```
Residuals look ok for just using a gaussian distribution. Seeing a non-significant
negative effect of protection status on red sea urchins. That is, there are less
red sea urchins in protected sites. Main effects that are evident is regional 
effect with broken group and sarita having lower urchins than bamfield.


```{r Plotting the effect of protection status on urchin counts glm fe, echo=FALSE}
pred_urchin_fishing_mod_glm_fe <- ggpredict(urchin_fishing_mod_glm_fe, terms = c("protect_status", "region1"))


ggplot(pred_urchin_fishing_mod_glm_fe, aes(x = group, y = predicted, colour = x))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(x = "Protection Status", y = "Predicted Urchin Counts")+
  theme.figure()
```


```{r Making the data in a way that it would plot}
#renaming group to region1
pred_urchin_fishing_mod_glm_fe$region1 <- pred_urchin_fishing_mod_glm_fe$group
pred_urchin_fishing_mod_glm_fe$protect_status <- pred_urchin_fishing_mod_glm_fe$x

#combining with env_mod_data for plotting the faceted boxplots
plotting_pred_urchin_fishing_mod_glm_fe <- env_mod_data%>%
  left_join(pred_urchin_fishing_mod_glm_fe, by = c("protect_status", "region1"))

#Adding a row with the info for the bamfield protected estimates so that it plots it
plotting_pred_urchin_fishing_mod_glm_fe[36,] <- plotting_pred_urchin_fishing_mod_glm_fe[35,]

#Changing the row data to be what the bamfield protected estimates are
plotting_pred_urchin_fishing_mod_glm_fe <- plotting_pred_urchin_fishing_mod_glm_fe%>%
 mutate(predicted = if_else(row_number() == 36, 383.59993, predicted))%>%
  mutate(std.error = if_else(row_number() == 36, 89.97562, std.error))%>%
  mutate(conf.low = if_else(row_number() == 36, 207.25095, conf.low))%>%
  mutate(conf.high = if_else(row_number() == 36, 559.9489, conf.high))%>%
  mutate(x = if_else(row_number() == 36, "Protected", x))%>%
  #Making the urchins counts NA so that it doesnt plot a boxplot for bamfield protected
  mutate(Urchins = if_else(row_number() == 36, NA, Urchins))
```


##### Making subsets of the predicted values
```{r Making subsets for the predicted urchin count values for protected/non-protected}
urchin_counts_protected_preds <- subset(pred_urchin_fishing_mod_glm_fe,
                                        x == "Protected")

urchin_counts_non_protected_preds <- subset(pred_urchin_fishing_mod_glm_fe,
                                            x == "No Protection")


```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Urchin fishing model fe glm residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_urchin_fishing_mod_glm_fe <- residuals(urchin_fishing_mod_glm_fe, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_urchin_fishing_mod_glm_fe, lw,
           alternative = "two.sided")


#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_urchin_fishing_mod_glm_fe, lw_inverse,
           alternative = "two.sided")
```
</div>
No autocorrelation with fixed effect included.



##### Testing the model for VIF stuff

```{r}
vif(urchin_fishing_mod_glm_fe)

performance::check_collinearity(urchin_fishing_mod_glm_fe,
                   component = c("all"))
```

#### Comparing urchin fishing models

```{r AIC of the urchin fishing models}
AICc(urchin_fishing_mod_glm,
    urchin_fishing_mod_glmm, 
    urchin_fishing_mod_glm_fe)%>%
  arrange(AICc)%>%
  mutate(delta_AICc = AICc - AICc[1])
```

Model with the random effect included has slightly worse AIC, but it is within
2 points even with an additional degree of freedom. The random effect model also
eliminates the autocorrelation. 
Model with the region included is the best based on AIC.


#### Testing the model with a subset without the deer group sites

```{r Subsetting the data to exclude deer group}
env_mod_data_wo_deer <- env_mod_data%>%
  filter(region1 != "Deer group/Bamfield")
```


```{r Checking variance between protected and non-protected for urchins, echo=FALSE}
leveneTest(sqrt(Urchins) ~ protect_status, data = env_mod_data_wo_deer)
```
Equal variance between protected and non-protected sites so assumption of 
homogeneity of variance is met.


```{r Urchin fishing model glmm no deer}
urchin_fishing_mod_glmm_no_deer <- glmmTMB(Urchins ~ protect_status + (1|region1),
                               family = gaussian(link = "identity"),
                                   data = env_mod_data_wo_deer)
```

```{r Urchin fishing mod glmm output no deer, echo=FALSE}
#Checking resids
urchin_fishing_mod_glmm_no_deer_resids <- simulateResiduals(urchin_fishing_mod_glmm_no_deer)
plot(urchin_fishing_mod_glmm_no_deer_resids)

hist(residuals(urchin_fishing_mod_glmm_no_deer))
shapiro.test(residuals(urchin_fishing_mod_glmm_no_deer))

#Looking at model results
summary(urchin_fishing_mod_glmm_no_deer)
```

## Urchin Snail ratio models{.tabset}

### Model 16: Urchin Snail Ratio fishing model (glm)

```{r Checking variance between protected and non-protected for urch sn ratio re, echo=FALSE}
leveneTest(urch_sn_ratio ~ protect_status, data = env_mod_data)
```
Variance is equal between protection status.


```{r Urchin ratio protection GLM model}
urch_sn_ratio_protection_model_glm <- glmmTMB(urch_sn_ratio ~ protect_status, 
                          data = env_mod_data,
                          family = gaussian(link = "identity"))
```

```{r Urchin sn ratio protection mod glm output, echo=FALSE}
#Checking resids
urch_sn_ratio_protection_model_glm_resids <- simulateResiduals(urch_sn_ratio_protection_model_glm)
plot(urch_sn_ratio_protection_model_glm_resids)


hist(residuals(urch_sn_ratio_protection_model_glm))
shapiro.test(residuals(urch_sn_ratio_protection_model_glm))

#Looking at model results
summary(urch_sn_ratio_protection_model_glm)
```

```{r Plotting the effect of protection status on urchin ratio in glm, echo=FALSE}
pred_urch_sn_ratio_protection_model_glm <- ggpredict(urch_sn_ratio_protection_model_glm, terms = "protect_status")

ggplot(pred_urch_sn_ratio_protection_model_glm, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Log Urchin to Snail ratio", x = "Protection Status")+
  theme.figure()
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing urch sn ratio model linear residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_urch_sn_ratio_protection_model_glm <- residuals(urch_sn_ratio_protection_model_glm, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_urch_sn_ratio_protection_model_glm, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_urch_sn_ratio_protection_model_glm, lw_inverse,
           alternative = "two.sided")
```
</div>

Only some correlation with the inverse distance method.

### Model 17: Urchin Snail Ratio fishing model with random effect (glmm)



```{r Urchin ratio protection re GLMM model}
urch_sn_ratio_protection_model_glmm <- glmmTMB(urch_sn_ratio ~ protect_status +
                          (1|region1), 
                          data = env_mod_data,
                          family = gaussian(link = "identity"))
```

```{r Urchin sn ratio protection mod re glmm output, echo=FALSE}
#Checking resids
urch_sn_ratio_protection_model_glmm_resids <- simulateResiduals(urch_sn_ratio_protection_model_glmm)
plot(urch_sn_ratio_protection_model_glmm_resids)

#Plot residuals of random effects 
urch_sn_ratio_protection_model_glmm_ranef_values <- ranef(urch_sn_ratio_protection_model_glmm)$cond$region1
urch_sn_ratio_protection_model_glmm_ranef_values_numeric <- urch_sn_ratio_protection_model_glmm_ranef_values$`(Intercept)`

hist(residuals(urch_sn_ratio_protection_model_glmm))
shapiro.test(residuals(urch_sn_ratio_protection_model_glmm))

#Looking at model results
summary(urch_sn_ratio_protection_model_glmm)
performance::r2(urch_sn_ratio_protection_model_glmm)
```

This model is fitting a separate intercept for each level of the random effect.

```{r Plotting the effect of protection status on urchin ratio in re glmm, echo=FALSE}
pred_urch_sn_ratio_protection_model_glmm <- ggpredict(urch_sn_ratio_protection_model_glmm, terms = "protect_status")

ggplot(pred_urch_sn_ratio_protection_model_glmm, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Log Urchin to Snail ratio", x = "Protection Status")+
  theme.figure()
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing urch sn ratio model linear re residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_urch_sn_ratio_protection_model_glmm <- residuals(urch_sn_ratio_protection_model_glmm, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_urch_sn_ratio_protection_model_glmm, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_urch_sn_ratio_protection_model_glmm, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation.

### Model 18: Urchin Snail Ratio fishing model with fixed effect (glm)


```{r Urchin ratio protection fe GLM model}
urch_sn_ratio_protection_model_glm_fe <- glmmTMB(urch_sn_ratio ~ protect_status +
                          region1, 
                          data = env_mod_data,
                          family = gaussian(link = "identity"))
```

```{r Urchin sn ratio protection mod fe glm output, echo=FALSE}
#Checking resids
urch_sn_ratio_protection_model_glm_fe_resids <- simulateResiduals(urch_sn_ratio_protection_model_glm_fe)
plot(urch_sn_ratio_protection_model_glm_fe_resids)

hist(residuals(urch_sn_ratio_protection_model_glm_fe))
shapiro.test(residuals(urch_sn_ratio_protection_model_glm_fe))

#Looking at model results
summary(urch_sn_ratio_protection_model_glm_fe)
```


```{r Plotting the effect of protection status on urchin ratio in fe glm, echo=FALSE}
pred_urch_sn_ratio_protection_model_glm_fe <- ggpredict(urch_sn_ratio_protection_model_glm_fe, terms = c("protect_status", "region1"))

ggplot(pred_urch_sn_ratio_protection_model_glm_fe, aes(x = group, y = predicted, colour = x))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Log Urchin to Snail ratio", x = "Protection Status")+
  theme.figure()
```

```{r Making subsets for the predicted values for protected/non-protected}
urch_sn_ratio_protected_preds <- subset(pred_urch_sn_ratio_protection_model_glm_fe,
                                        x == "Protected")

urch_sn_ratio_non_protected_preds <- subset(pred_urch_sn_ratio_protection_model_glm_fe,
                                            x == "No Protection")
```

```{r Making the data in a way that it would plot ursh sn ratio}
#renaming group to region1
pred_urch_sn_ratio_protection_model_glm_fe$region1 <- pred_urch_sn_ratio_protection_model_glm_fe$group
pred_urch_sn_ratio_protection_model_glm_fe$protect_status <- pred_urch_sn_ratio_protection_model_glm_fe$x

#combining with env_mod_data for plotting the faceted boxplots
plotting_pred_urch_sn_ratio_protection_model_glm_fe <- env_mod_data%>%
  left_join(pred_urch_sn_ratio_protection_model_glm_fe, by = c("protect_status", "region1"))

#Adding a row with the info for the bamfield protected estimates so that it plots it
plotting_pred_urch_sn_ratio_protection_model_glm_fe[36,] <- plotting_pred_urch_sn_ratio_protection_model_glm_fe[35,]

#Changing the row data to be what the bamfield protected estimates are
plotting_pred_urch_sn_ratio_protection_model_glm_fe <- plotting_pred_urch_sn_ratio_protection_model_glm_fe%>%
 mutate(predicted = if_else(row_number() == 36, 0.4249653, predicted))%>%
  mutate(std.error = if_else(row_number() == 36, 0.4572245, std.error))%>%
  mutate(conf.low = if_else(row_number() == 36, -0.4711782, conf.low))%>%
  mutate(conf.high = if_else(row_number() == 36, 1.3211087, conf.high))%>%
  mutate(x = if_else(row_number() == 36, "Protected", x))%>%
  #Making the urchins counts NA so that it doesnt plot a boxplot for bamfield protected
  mutate(urch_sn_ratio = if_else(row_number() == 36, NA, urch_sn_ratio))
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing urch sn ratio model linear fe residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_urch_sn_ratio_protection_model_glm_fe <- residuals(urch_sn_ratio_protection_model_glm_fe, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_urch_sn_ratio_protection_model_glm_fe, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_urch_sn_ratio_protection_model_glm_fe, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation.


##### Testing for the VIF of the model
```{r}
vif(urch_sn_ratio_protection_model_glm_fe)

performance::check_collinearity(urch_sn_ratio_protection_model_glm_fe,
                   component = c("all"))
```


### Comparing Urchin Snail ratio models without openness

```{r AIC of the urchin snail ratio models without openness}
AICc(urch_sn_ratio_protection_model_glm,
    urch_sn_ratio_protection_model_glmm,
    urch_sn_ratio_protection_model_glm_fe)%>%
  arrange(AICc)%>%
  mutate(delta_AICc = AICc - AICc[1])
```
AIC suggests that the model with the random effect included may be slightly 
better than the model without region as it has a lower AIC even with 
an additional degree of freedom.
AIC suggests that the model with the fixed effect of region is much better
than the other two models.


### Testing the model without the deer group sites

```{r Urchin ratio protection re GLMM model no deer}
urch_sn_ratio_protection_model_glmm_no_deer <- glmmTMB(urch_sn_ratio ~ protect_status,
                         # (1|region1), 
                          data = env_mod_data_wo_deer,
                          family = gaussian(link = "identity"))
```

```{r Urchin sn ratio protection mod re glmm output no deer, echo=FALSE}
#Checking resids
urch_sn_ratio_protection_model_glmm_no_deer_resids <- simulateResiduals(urch_sn_ratio_protection_model_glmm_no_deer)
plot(urch_sn_ratio_protection_model_glmm_no_deer_resids)

#Plot residuals of random effects 
urch_sn_ratio_protection_model_glmm_no_deer_ranef_values <- ranef(urch_sn_ratio_protection_model_glmm_no_deer)$cond$region1
urch_sn_ratio_protection_model_glmm_no_deer_ranef_values_numeric <- urch_sn_ratio_protection_model_glmm_no_deer_ranef_values$`(Intercept)`

hist(residuals(urch_sn_ratio_protection_model_glmm_no_deer))
shapiro.test(residuals(urch_sn_ratio_protection_model_glmm_no_deer))

#Looking at model results
summary(urch_sn_ratio_protection_model_glmm_no_deer)
```

### Model 14: Urchin snail ratio model with openness (glm)

```{r Urchin ratio protection and openness GLM model}
urch_sn_ratio_protection_open_model_glm <- glmmTMB(urch_sn_ratio ~ 
                                                     protect_status + 
                                                     norm_openness_scale, 
                          data = env_mod_data,
                          family = gaussian(link = "identity"))
```

```{r Urchin sn ratio protection and openness mod glm output, echo=FALSE}
#Checking resids
urch_sn_ratio_protection_open_model_glm_resids <- simulateResiduals(urch_sn_ratio_protection_open_model_glm)
plot(urch_sn_ratio_protection_open_model_glm_resids)


hist(residuals(urch_sn_ratio_protection_open_model_glm))
shapiro.test(residuals(urch_sn_ratio_protection_open_model_glm))

#Looking at model results
summary(urch_sn_ratio_protection_open_model_glm)
```

```{r Plotting the effect of protection status and openness on urchin ratio in glm, echo=FALSE}
pred_urch_sn_ratio_protection_open_model_glm <- ggpredict(urch_sn_ratio_protection_open_model_glm, terms = c("norm_openness_scale", "protect_status"))

plot(pred_urch_sn_ratio_protection_open_model_glm)

ggplot(pred_urch_sn_ratio_protection_open_model_glm, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Log Urchin to Snail ratio", x = "Protection Status")+
  theme.figure()


pred_urch_sn_ratio_protection_open_model_glm_protect_status <- ggpredict(urch_sn_ratio_protection_open_model_glm, terms = "protect_status")

ggplot(pred_urch_sn_ratio_protection_open_model_glm_protect_status, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Log Urchin to Snail ratio", x = "Protection Status")+
  theme.figure()
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing urch sn ratio model linear with openness residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_urch_sn_ratio_protection_open_model_glm <- residuals(urch_sn_ratio_protection_open_model_glm, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_urch_sn_ratio_protection_open_model_glm, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_urch_sn_ratio_protection_open_model_glm, lw_inverse,
           alternative = "two.sided")
```
</div>

Evidence of autocorrelation with both methods.

### Model 15: Urchin snail ratio model with openness and random effect (glmm)

```{r Urchin ratio protection and openness re GLMM model}
urch_sn_ratio_protection_open_model_glmm <- glmmTMB(urch_sn_ratio ~ 
                                                     protect_status + 
                                                     norm_openness_scale+
                                                      (1|region1), 
                          data = env_mod_data,
                          family = tweedie(link = "log"))
```

```{r Urchin sn ratio protection and openness re mod glmm output, echo=FALSE}
#Checking resids
urch_sn_ratio_protection_open_model_glmm_resids <- simulateResiduals(urch_sn_ratio_protection_open_model_glmm)
plot(urch_sn_ratio_protection_open_model_glmm_resids)


hist(residuals(urch_sn_ratio_protection_open_model_glmm))
shapiro.test(residuals(urch_sn_ratio_protection_open_model_glmm))

#Looking at model results
summary(urch_sn_ratio_protection_open_model_glmm)
```

```{r Plotting the effect of protection status and openness on urchin ratio in glm, echo=FALSE}
pred_urch_sn_ratio_protection_open_model_glmm <- ggpredict(urch_sn_ratio_protection_open_model_glmm, terms = c("norm_openness_scale", "protect_status"))

plot(pred_urch_sn_ratio_protection_open_model_glmm)


pred_urch_sn_ratio_protection_open_model_glmm_protect_status <- ggpredict(urch_sn_ratio_protection_open_model_glmm, terms = "protect_status")

ggplot(pred_urch_sn_ratio_protection_open_model_glmm_protect_status, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Log Urchin to Snail ratio", x = "Protection Status")+
  theme.figure()
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing urch sn ratio model linear with openness re residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_urch_sn_ratio_protection_open_model_glmm <- residuals(urch_sn_ratio_protection_open_model_glmm, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_urch_sn_ratio_protection_open_model_glmm, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_urch_sn_ratio_protection_open_model_glmm, lw_inverse,
           alternative = "two.sided")
```
</div>

Evidence of autocorrelation with both methods.



## Snail to Urchin models{.tabset}

### Model 20: Snail urchin ratio model (glm)

```{r Checking variance between protected and non-protected for sn urch ratio, echo=FALSE}
leveneTest(sn_urch_ratio ~ protect_status, data = env_mod_data)
```
Variance is equal between protection status.

```{r Snail ratio protection GLM model}
sn_urch_ratio_protection_model_glm <- glmmTMB(sn_urch_ratio ~ protect_status, 
                          data = env_mod_data,
                          family = gaussian(link = "identity"))
```

```{r Sn urch ratio protection mod glm output, echo=FALSE}
#Checking resids
sn_urch_ratio_protection_model_glm_resids <- simulateResiduals(sn_urch_ratio_protection_model_glm)
plot(sn_urch_ratio_protection_model_glm_resids)


hist(residuals(sn_urch_ratio_protection_model_glm))
shapiro.test(residuals(sn_urch_ratio_protection_model_glm))

#Looking at model results
summary(sn_urch_ratio_protection_model_glm)
```

```{r Plotting the effect of protection status on snail ratio in glm, echo=FALSE}
pred_sn_urch_ratio_protection_model_glm <- ggpredict(sn_urch_ratio_protection_model_glm, terms = "protect_status")

ggplot(pred_sn_urch_ratio_protection_model_glm, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Log Snail to Urchin ratio", x = "Protection Status")+
  theme.figure()
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing sn urch ratio model linear residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_sn_urch_ratio_protection_model_glm <- residuals(sn_urch_ratio_protection_model_glm, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_sn_urch_ratio_protection_model_glm, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_sn_urch_ratio_protection_model_glm, lw_inverse,
           alternative = "two.sided")
```
</div>

Only some correlation with the inverse distance method.



### Model 21: Snail urchin ratio fishing model (glmm)
```{r Snail ratio protection re GLMM model}
sn_urch_ratio_protection_model_glmm <- glmmTMB(sn_urch_ratio ~ protect_status +
                          (1|region1), 
                          data = env_mod_data,
                          family = gaussian(link = "identity"))
```

```{r Snail urchin ratio protection mod re glmm output, echo=FALSE}
#Checking resids
sn_urch_ratio_protection_model_glmm_resids <- simulateResiduals(sn_urch_ratio_protection_model_glmm)
plot(sn_urch_ratio_protection_model_glmm_resids)

#Plot residuals of random effects 
sn_urch_ratio_protection_model_glmm_ranef_values <- ranef(sn_urch_ratio_protection_model_glmm)$cond$region1
sn_urch_ratio_protection_model_glmm_ranef_values_numeric <- sn_urch_ratio_protection_model_glmm_ranef_values$`(Intercept)`

hist(residuals(sn_urch_ratio_protection_model_glmm))
shapiro.test(residuals(sn_urch_ratio_protection_model_glmm))

#Looking at model results
summary(sn_urch_ratio_protection_model_glmm)
performance::r2(sn_urch_ratio_protection_model_glmm)

#Checking for significance of random effect
performance::icc(sn_urch_ratio_protection_model_glmm)
```

This model is fitting a separate intercept for each level of the random effect.

```{r Plotting the effect of protection status on snail ratio in re glmm, echo=FALSE}
pred_sn_urch_ratio_protection_model_glmm <- ggpredict(sn_urch_ratio_protection_model_glmm, terms = "protect_status")

ggplot(pred_sn_urch_ratio_protection_model_glmm, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Log Snail to Urchin ratio", x = "Protection Status")+
  theme.figure()
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing sn urch ratio model linear re residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_sn_urch_ratio_protection_model_glmm <- residuals(sn_urch_ratio_protection_model_glmm, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_sn_urch_ratio_protection_model_glmm, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_sn_urch_ratio_protection_model_glmm, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation.


### Model 22: Snail urchin ratio fishing model with fixed effect (glm)

```{r Snail ratio protection fe GLM model}
sn_urch_ratio_protection_model_glm_fe <- glmmTMB(sn_urch_ratio ~ protect_status +
                          region1, 
                          data = env_mod_data,
                          family = gaussian(link = "identity"))
```

```{r Snail urch ratio protection mod fe glm output, echo=FALSE}
#Checking resids
sn_urch_ratio_protection_model_glm_fe_resids <- simulateResiduals(sn_urch_ratio_protection_model_glm_fe)
plot(sn_urch_ratio_protection_model_glm_fe_resids)

hist(residuals(sn_urch_ratio_protection_model_glm_fe))
shapiro.test(residuals(sn_urch_ratio_protection_model_glm_fe))

#Looking at model results
summary(sn_urch_ratio_protection_model_glm_fe)
```


```{r Plotting the effect of protection status on snail ratio in fe glm, echo=FALSE}
pred_sn_urch_ratio_protection_model_glm_fe <- ggpredict(sn_urch_ratio_protection_model_glm_fe, terms = c("protect_status", "region1"))

ggplot(pred_sn_urch_ratio_protection_model_glm_fe, aes(x = group, y = predicted, colour = x))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Log Snail to Urchin ratio", x = "Protection Status")+
  theme.figure()
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing sn urch ratio model linear fe residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_sn_urch_ratio_protection_model_glm_fe <- residuals(sn_urch_ratio_protection_model_glm_fe, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_sn_urch_ratio_protection_model_glm_fe, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_sn_urch_ratio_protection_model_glm_fe, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation.


##### Testing for the VIF of the model
```{r Testing for multicollinearity of the predictors snail to urchin}
#vif(sn_urch_ratio_protection_model_glm_fe)

performance::check_collinearity(sn_urch_ratio_protection_model_glm_fe,
                   component = c("all"))
```

Correlation does not seem to be problematic.


### Comparing the snail to urchin ratio models

```{r AIC of the snail urchin ratio models without openness}
AICc(sn_urch_ratio_protection_model_glm,
     sn_urch_ratio_protection_model_glmm)%>%
    # sn_urch_ratio_protection_model_glm_fe)%>%
  arrange(AICc)%>%
  mutate(delta_AICc = AICc - AICc[1])
```


## Testing effect of region on the different variables

### Testing region on snail abundance

```{r Running an ANOVA and testing for assumptions}
snail_counts_region_an <- lm(log(Turban_Snails + 1) ~ region1, data = env_mod_data)
shapiro.test(residuals(snail_counts_region_an))
leveneTest(log(Turban_Snails + 1) ~ region1, data = env_mod_data)
#Variances are not equal so will need to use Welch's ANOVA

snail_counts_region_welch <- oneway(log(Turban_Snails + 1) ~ region1, data = env_mod_data, var.equal = FALSE, posthoc = 'games-howell')
?oneway.test

```


```{r Testing normality of snail counts between regions}
#Making data frames with only sites from each region
turban_snails_SR <- subset(env_mod_data, region1 == "SR")
turban_snails_DGB <- subset(env_mod_data, region1 == "DGB")
turban_snails_DIEC <- subset(env_mod_data, region1 == "DIEC")
turban_snails_BG <- subset(env_mod_data, region1 == "BG")

#testing normality for each subset
shapiro.test(log(turban_snails_SR$Turban_Snails + 1))
shapiro.test(log(turban_snails_DIEC$Turban_Snails + 1))
shapiro.test(log(turban_snails_DGB$Turban_Snails + 1))
shapiro.test(log(turban_snails_BG$Turban_Snails + 1))
```
All are normal with a log transformation

```{r Testing for equal variance of snail counts between regions}
leveneTest(log(Turban_Snails + 1) ~ region1, data = env_mod_data)
```
When log transforming the variance become unequal


### Testing region on urchin abundance

```{r Running an ANOVA on urchins and testing for assumptions}
urchin_counts_region_an <- lm(Urchins ~ region1, data = env_mod_data)
shapiro.test(residuals(urchin_counts_region_an))
leveneTest(Urchins ~ region1, data = env_mod_data)

#Looking at output
anova(urchin_counts_region_an)
summary(urchin_counts_region_an)

#TUkey Pairwise test
TukeyHSD(aov(urchin_counts_region_an))
```
Assumptions of ANova are met and indicate significant differences

### Testing region on snail urchin ratio

```{r Running an ANOVA on snail urchin ratio and testing for assumptions}
snail_urchin_ratio_region_an <- lm(sn_urch_ratio ~ region1, data = env_mod_data)
shapiro.test(residuals(snail_urchin_ratio_region_an))
leveneTest(sn_urch_ratio ~ region1, data = env_mod_data)

#Looking at output
anova(snail_urchin_ratio_region_an)
summary(snail_urchin_ratio_region_an)

#Tukey pairwise test
TukeyHSD(aov(snail_urchin_ratio_region_an))

```

Assumptions of ANOVA met and indicate significant differences