---
title: "RLS Modelling for Amanda"
author: "Alec"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
   toc: true
---

   <script src="hideOutput.js"></script> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(mgcv)
library(DHARMa)
library(mgcViz)
library(ggeffects)
library(gratia)
library(geoR)
library(geosphere)
library(gstat)
library(sf)
library(car)
library(MuMIn)
library(lme4)
library(spdep)
library(ggpubr)
library(glmmTMB)
library(emmeans)
```


```{r Loading in the data, include=FALSE}
env_mod_data <- read_csv(here("./03_clean_data/urch_sn_ratio_mod_data.csv"))
```

```{r Scaling the explanatory variables and converting substrate type to factor, include=FALSE}
env_mod_data$sst_scale <- scale(env_mod_data$sst)
env_mod_data$openness_scale <- scale(env_mod_data$openness)
env_mod_data$norm_openness_scale <- scale(env_mod_data$norm_openness)
env_mod_data$mean_app_fishing_hours_scale <- scale(env_mod_data$mean_app_fishing_hours)
env_mod_data$mean_rec_boat_counts_500m_scale <- scale(env_mod_data$mean_rec_boat_counts_500m)
env_mod_data$mean_rec_boat_counts_1km_scale <- scale(env_mod_data$mean_rec_boat_counts_1km)
env_mod_data$mean_rec_boat_counts_1.5km_scale <- scale(env_mod_data$mean_rec_boat_counts_1.5km)
env_mod_data$mean_rec_boat_counts_750m_scale <- scale(env_mod_data$mean_rec_boat_counts_750m)
env_mod_data$mean_rec_boat_counts_2km_scale <- scale(env_mod_data$mean_rec_boat_counts_2km)
env_mod_data$dominant_score <- as.factor(env_mod_data$dominant_score)
env_mod_data$site_code <- as.factor(env_mod_data$site_code)
env_mod_data$Urchins_scale <- scale(env_mod_data$Urchins)
env_mod_data$Turban_Snails_scale <- scale(env_mod_data$Turban_Snails)
#Total boat counts
env_mod_data$mean_total_boat_counts_1km_scale <- scale(env_mod_data$mean_total_boat_counts_1km)
env_mod_data$mean_total_boat_counts_1.5km_scale <- scale(env_mod_data$mean_total_boat_counts_1.5km)
env_mod_data$mean_total_boat_counts_2km_scale <- scale(env_mod_data$mean_total_boat_counts_2km)

levels(env_mod_data$dominant_score) <- list(loose = "3", solid = "4")

regions <- c("region1", "region2", "region3", "region4")
env_mod_data[regions] <- lapply(env_mod_data[regions], factor)
```

```{r Creating a factor variable of protection status for the modelling data, include=FALSE}
#In this chunk I am making three objects with the various protection statuses
#of the different sites. protect_status has all sites protected from fishing
#to some degree (RCA, ecological reserve, finfish closure area) and 
#protect_status_park includes the sites within the Broken Group Park that are
#not within the RCA.
protect_status <- c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 
                    0, 0, 0, 1, 1, 0, 0, 0, 0, 0, 1, 1, 1, 1, 0)

protect_status_park <- c(0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0, 0, 1, 1, 1, 1, 0)

#kelp_forest <- c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)


  env_mod_data$protect_status <- as.factor(protect_status)
levels(env_mod_data$protect_status) <- c("No Protection", "Protected")

env_mod_data$protect_status_park <- as.factor(protect_status_park)
levels(env_mod_data$protect_status_park) <- c("No Protection", "Protected")


#Reordering so easier to see which sites are protected in the DF
env_mod_data <- env_mod_data[, c(seq(1, 11, by = 1), 44, 45, seq(12, 43, by = 1))]


#Will need to confirm which areas were kelp forests based on the photos or
#talking with Jasmin but likely will just be hosie and raymond

#env_mod_data$kelp_forest <- as.factor(kelp_forest)
#levels(env_mod_data$kelp_forest) <- c("No Kelp", "Kelp Forest")
```

```{r Making theme for all subsequent figures, echo=FALSE}
theme.figure <- function (){
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    #axis.line = element_line(colour = "black"),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12, face = "plain"),
    legend.text = element_text(size = 10, face = "plain"),
    legend.title = element_text(size = 14, face = "bold")) 
}

```

# Modelling{.tabset}

## Ecology models{.tabset}

### Red Turban Snail Ecology Models{.tabset}

#### Model 1: Ecology model on the snail counts with temperature, openness, and bottom type

```{r Turban snail ecology/environmental model}
ecology_snail_mod <- gam(log(Turban_Snails + 1) ~ s(sst_scale)+
                               s(norm_openness_scale)+
                           dominant_score,
                                # s(region1, bs = "re"), #Random effect of region does not have an impact on the snails
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

```{r Turban snail ecology/environmental model output, echo=FALSE}
gam.check(ecology_snail_mod)
plot(ecology_snail_mod, all.terms = TRUE)
summary(ecology_snail_mod)
```

Initially did this model with openness as a smooth term but fitting produced a linear effect of openness so re fit model with linear openness.

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail ecology model residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_ecology_snail_mod <- residuals(ecology_snail_mod, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_ecology_snail_mod, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_ecology_snail_mod, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation for both methods when using a two sided test.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing turban snail ecology model for concurvity, echo=FALSE}
concurvity(ecology_snail_mod, full = FALSE)
concurvity(ecology_snail_mod, full = TRUE)
```
</div>

Temperature and openness seem to be somewhat concurve. Temp having no impact on
the model (being removed by select).


#### Model 2: Ecology model on the snail counts with temperature, openness, and bottom type (linear openness)

```{r Turban snail ecology/environmental model linear}
ecology_snail_mod_linear <- gam(log(Turban_Snails + 1) ~ s(sst_scale)+
                               norm_openness_scale +
                           dominant_score,
                                #s(region1, bs = "re"), #Random effect of region does not have an impact on the snails
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

```{r Turban snail ecology/environmental model linear output, echo=FALSE}
gam.check(ecology_snail_mod_linear)
plot(ecology_snail_mod_linear, all.terms = TRUE)
summary(ecology_snail_mod_linear)
```
Results showing the same trends with normalized openness included as a linear term.
Ultimately the results of the environmental model for turban snail abundance/counts
indicate the turban snails are not controlled by temperature or substrate type,
but they are strongly controlled by the openness gradient, with less snails in
more exposed sites. 

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail ecology model linear residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_ecology_snail_mod_linear <- residuals(ecology_snail_mod_linear, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_ecology_snail_mod_linear, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_ecology_snail_mod_linear, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation for both methods when using a two sided test.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing turban snail ecology model linear for concurvity, echo=FALSE}
concurvity(ecology_snail_mod_linear, full = FALSE)
concurvity(ecology_snail_mod_linear, full = TRUE)
```
</div>

Concurvity seems to be improved when having openness as a linear term. Maybe not
and it is just how the function calculates it differently or omits it.


### Red Sea urchin ecology models{.tabset}

#### Model 3: Ecology model on the urchin counts with temperature, openness, and bottom type

```{r Urchin ecology/environmental model}
ecology_urchin_mod <- gam(Urchins ~ s(sst_scale)+
                               s(norm_openness_scale)+
                           dominant_score,
                                #s(region1, bs = "re"),
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

```{r Urchin ecology/environmental model output, echo=FALSE}
set.seed(299)
gam.check(ecology_urchin_mod, rep = 500)
plot(ecology_urchin_mod, all.terms = TRUE)
summary(ecology_urchin_mod)
shapiro.test(ecology_urchin_mod$residuals)
```
Tried with other distributions as well but gaussian appears to be the best.
Confirm this with Amanda.

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Urchin ecology model residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_ecology_urchin_mod <- residuals(ecology_urchin_mod, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_ecology_urchin_mod, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_ecology_urchin_mod, lw_inverse,
           alternative = "two.sided")
```
</div>

Some positive autocorrelation but only with the inverse distance method. The 
neighbour method has no autocorrelation.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing urchin ecology model for concurvity, echo=FALSE}
concurvity(ecology_urchin_mod, full = FALSE)
concurvity(ecology_urchin_mod, full = TRUE)
```
</div>

Concurvity is high for temperature and openness. Both are being removed by the 
double penalty.


#### Model 4: Ecology model on the urchin counts with temperature, openness, and bottom type (random effect of region)

```{r Urchin ecology/environmental model region re}
ecology_urchin_mod_re <- gam(Urchins ~ s(sst_scale)+
                               s(norm_openness_scale)+
                           dominant_score+
                                s(region1, bs = "re"),
                             data = env_mod_data,
                             family = gaussian(link = "identity"),
                             method = "REML",
                             select = TRUE)
```

```{r Urchin ecology/environmental model region re output, echo=FALSE}
set.seed(299)
gam.check(ecology_urchin_mod_re, rep = 500)
plot(ecology_urchin_mod_re, all.terms = TRUE)
summary(ecology_urchin_mod_re)
shapiro.test(ecology_urchin_mod_re$residuals)
```
Gaussian looks ok. When adding the random effect of region the openness becomes
borderline and negative. 

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Urchin ecology model region re residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_ecology_urchin_mod_re <- residuals(ecology_urchin_mod_re, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_ecology_urchin_mod_re, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_ecology_urchin_mod_re, lw_inverse,
           alternative = "two.sided")
```
</div>

No autocorrelation.

##### Testing the model for concurvity
<div class="fold o"> 
```{r Testing urchin ecology model re for concurvity, echo=FALSE}
concurvity(ecology_urchin_mod_re, full = FALSE)
concurvity(ecology_urchin_mod_re, full = TRUE)
```
</div>

Temp and region are basically fully concurve.



## Fishing models{.tabset}

### Confirming the validity of protection status with DFO flyover data

For the protection status categories, I have grouped them as sites that 
experience at least a ban on fishing for part of the year as protected and all
others as non-protected. This means that the sites that are in the Broken group
but are not within the RCA are counted as not protected.

In order to confirm/support our assumptions of which sites are protected,
I am using the DFO flyover data to compare the numbers of boats around the 
different sites.

Based on the boxplots below it seems evident that the protected sites selected
are experiencing less fishing pressure and less overall human/boat pressure.

#### Plot indicating which regions protected and non-protected sites are located in

It seems that our current region scheme has no protected sites located within the deer group/bamfield,
will need to look into if this is a problem.

```{r Plot indicating the groupings of protected/non-protected and region, echo=FALSE}
ggplot(env_mod_data, aes(y = region1, x = protect_status))+
  geom_jitter(width = 0.2, height = 0.2)+
  labs(x = "Protection Status", y = "Region")+
  theme.figure()
```



#### Boxplots of rec boat counts between protection levels

```{r Boxplots of the rec boats at the three distances between protection levels, echo=FALSE}
ggplot(env_mod_data, aes(x = protect_status, y = mean_rec_boat_counts_1km))+
  geom_boxplot()+
  geom_jitter()+
  theme.figure()

ggplot(env_mod_data, aes(x = protect_status, y = mean_rec_boat_counts_1.5km))+
  geom_boxplot()+
  geom_jitter()+
  theme.figure()

ggplot(env_mod_data, aes(x = protect_status, y = mean_rec_boat_counts_2km))+
  geom_boxplot()+
  geom_jitter()+
  theme.figure()
```

#### Boxplots of total boat counts between protection levels

```{r Boxplots of the total boats at the three distances between protection levels, echo=FALSE}
ggplot(env_mod_data, aes(x = protect_status, y = mean_total_boat_counts_1km))+
  geom_boxplot()+
  geom_jitter()+
  theme.figure()

ggplot(env_mod_data, aes(x = protect_status, y = mean_total_boat_counts_1.5km))+
  geom_boxplot()+
  geom_jitter()+
  theme.figure()

ggplot(env_mod_data, aes(x = protect_status, y = mean_total_boat_counts_2km))+
  geom_boxplot()+
  geom_jitter()+
  theme.figure()
```

#### Comparing the rec boat counts between sites in the broken group that have been labelled as protected or non-protected

```{r Comparing the rec boat counts between sites in the broken group that have been labelled as protected or non-protected, echo=FALSE}
ggplot(env_mod_data[c(8, 9, 10, 11, 12, 13),], aes(x = protect_status, y = mean_rec_boat_counts_1km))+
  geom_boxplot()+
  geom_point()+
  theme.figure()

ggplot(env_mod_data[c(8, 9, 10, 11, 12, 13),], aes(x = protect_status, y = mean_rec_boat_counts_1.5km))+
  geom_boxplot()+
  geom_point()+
  theme.figure()

ggplot(env_mod_data[c(8, 9, 10, 11, 12, 13),], aes(x = protect_status, y = mean_rec_boat_counts_2km))+
  geom_boxplot()+
  geom_point()+
  theme.figure()
```

#### Comparing the total boat counts between sites in the broken group that have been labelled as protected or non-protected

```{r Comparing the total boat counts between sites in the broken group that have been labelled as protected or non-protected, echo=FALSE}
ggplot(env_mod_data[c(8, 9, 10, 11, 12, 13),], aes(x = protect_status, y = mean_total_boat_counts_1km))+
  geom_boxplot()+
  geom_point()+
  theme.figure()

ggplot(env_mod_data[c(8, 9, 10, 11, 12, 13),], aes(x = protect_status, y = mean_total_boat_counts_1.5km))+
  geom_boxplot()+
  geom_point()+
  theme.figure()

ggplot(env_mod_data[c(8, 9, 10, 11, 12, 13),], aes(x = protect_status, y = mean_total_boat_counts_2km))+
  geom_boxplot()+
  geom_point()+
  theme.figure()
```


### Red Turban snail fishing models{.tabset}

#### Plotting counts of snails between protected and non-protected

```{r Plotting counts of snails between protected and non-protected, echo=FALSE}
ggplot(env_mod_data, aes(x = protect_status, y = Turban_Snails))+
  geom_boxplot()+
  geom_jitter(height = 0)+
  labs(x = "Protection Status", y = "Turban Snail counts")+
  theme.figure()
```

Seeing a weak trend of more snails in the protected sites when compared to the non-
protected sites.



#### Model 5: Fishing model on the snail counts with protection and openness (glm)

For this model we are seeing that protection status has a non-significant
positive effect on the log snail counts

```{r Checking variance between protected and non-protected for turban snails, echo=FALSE}
leveneTest(Turban_Snails ~ protect_status, data = env_mod_data)
```


```{r Snail count fishing  model with protection and openness linear}
snail_fishing_mod_linear <- glmmTMB(log(Turban_Snails +1) ~ norm_openness_scale +
                                      protect_status,
                                    data = env_mod_data,
                                    family = gaussian(link = "identity"))
```

```{r Snail count fishing model with protection and openness linear output, echo=FALSE}
#Checking resids
snail_fishing_mod_linear_resids <- simulateResiduals(snail_fishing_mod_linear)
plot(snail_fishing_mod_linear_resids)

hist(residuals(snail_fishing_mod_linear))
shapiro.test(residuals(snail_fishing_mod_linear))

#residuals_test <- residuals(snail_fishing_mod_linear, type = "pearson")
#fitted_values_test <- fitted(snail_fishing_mod_linear)

#ggplot(data = data.frame(fitted_values_test, residuals_test), aes(x = fitted_values_test, y = residuals_test)) +
#  geom_point() +
#  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
#  labs(x = "Fitted Values", y = "Pearson Residuals", title = "Residuals vs Fitted") +
#  theme_minimal()

#Looking at model results
summary(snail_fishing_mod_linear)
```


```{r Plotting the linear snail fishing model with protection and openness, echo=FALSE}
pred_snail_fishing_mod_linear_openness <- ggpredict(snail_fishing_mod_linear, terms = c("norm_openness_scale", "protect_status"))


plot(pred_snail_fishing_mod_linear_openness)


pred_snail_fishing_mod_linear_protect_status <- ggpredict(snail_fishing_mod_linear, terms = "protect_status")
#plot(pred_snail_fishing_mod_linear_protect_status)

ggplot(pred_snail_fishing_mod_linear_protect_status, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Turban Snail Counts", x = "Protection Status")+
  theme.figure()

#Using emmeans to represent the protection status
#plot(emmeans(snail_fishing_mod_linear, ~ protect_status, comparisons = TRUE))
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail fishing model linear residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_snail_fishing_mod_linear <- residuals(snail_fishing_mod_linear, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_snail_fishing_mod_linear, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_snail_fishing_mod_linear, lw_inverse,
           alternative = "two.sided")
```
</div>
No autocorrelation with neighbours but it is significant with inverse distance.


#### Model 6: Fishing model on the snail counts with just protection (glm)

```{r Snail count fishing model with protection linear}
snail_fishing_mod_linear_no_op <- glmmTMB(log(Turban_Snails +1) ~
                                      protect_status,
                                    data = env_mod_data,
                                    family = gaussian(link = "identity"))
```

```{r Snail count fishing model with protection linear output, echo=FALSE}
#Checking resids
snail_fishing_mod_linear_no_op_resids <- simulateResiduals(snail_fishing_mod_linear_no_op)
plot(snail_fishing_mod_linear_no_op_resids)

hist(residuals(snail_fishing_mod_linear_no_op))
shapiro.test(residuals(snail_fishing_mod_linear_no_op))

#residuals_test <- residuals(snail_fishing_mod_linear, type = "pearson")
#fitted_values_test <- fitted(snail_fishing_mod_linear)

#ggplot(data = data.frame(fitted_values_test, residuals_test), aes(x = fitted_values_test, y = residuals_test)) +
#  geom_point() +
#  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
#  labs(x = "Fitted Values", y = "Pearson Residuals", title = "Residuals vs Fitted") +
#  theme_minimal()

#Looking at model results
summary(snail_fishing_mod_linear_no_op)
```


```{r Plotting the linear snail fishing model with protection, echo=FALSE}
pred_snail_fishing_mod_linear_no_op_protect_status <- ggpredict(snail_fishing_mod_linear_no_op, terms = "protect_status")
#plot(pred_snail_fishing_mod_linear_protect_status)

ggplot(pred_snail_fishing_mod_linear_no_op_protect_status, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Turban Snail Counts", x = "Protection Status")+
  theme.figure()

#Using emmeans to represent the protection status
#plot(emmeans(snail_fishing_mod_linear, ~ protect_status, comparisons = TRUE))
```


##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Turban snail fishing model linear no openness residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_snail_fishing_mod_linear_no_op <- residuals(snail_fishing_mod_linear_no_op, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_snail_fishing_mod_linear_no_op, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_snail_fishing_mod_linear_no_op, lw_inverse,
           alternative = "two.sided")
```
</div>
No autocorrelation.

### Red Sea Urchin fishing models{.tabset}

#### Plot of urchin counts between protection status

```{r Plot of number of urchins between protected and non-protected sites}
ggplot(env_mod_data, aes(x = protect_status, y = Urchins))+
  geom_boxplot()+
  geom_jitter()+
  labs(x = "Protection Status", y = "Red Sea urchin counts")+
  theme.figure()
```


#### Model 7: Fishing model on the urchin counts (glm)

```{r Checking variance between protected and non-protected for urchins, echo=FALSE}
leveneTest(Urchins ~ protect_status, data = env_mod_data)
```
Equal variance between protected and non-protected sites so assumption of 
homogeneity of variance is met.


```{r Urchin fishing model glm}
urchin_fishing_mod_glm <- glmmTMB(Urchins ~ protect_status,
                               family = gaussian(link = "identity"),
                                   data = env_mod_data)
```

```{r Urchin fishing mod glm output, echo=FALSE}
#Checking resids
urchin_fishing_mod_glm_resids <- simulateResiduals(urchin_fishing_mod_glm)
plot(urchin_fishing_mod_glm_resids)

hist(residuals(urchin_fishing_mod_glm))
shapiro.test(residuals(urchin_fishing_mod_glm))

#Looking at model results
summary(urchin_fishing_mod_glm)
```
Residuals look ok for just using a gaussian distribution. Seeing a non-significant
negative effect of protection status on red sea urchins. That is, there are less
red sea urchins in protected sites.


```{r Plotting the effect of protection status on urchin counts, echo=FALSE}
pred_urchin_fishing_mod_glm <- ggpredict(urchin_fishing_mod_glm, terms = "protect_status")

ggplot(pred_urchin_fishing_mod_glm, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(x = "Protection Status", y = "Predicted Urchin Counts")+
  theme.figure()
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing Urchin fishing model glm residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_urchin_fishing_mod_glm <- residuals(urchin_fishing_mod_glm, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_urchin_fishing_mod_glm, lw,
           alternative = "two.sided")


#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_urchin_fishing_mod_glm, lw_inverse,
           alternative = "two.sided")
```
</div>
Seeing some autocorrelation in the inverse method but not the neighbour method.


## Urchin Snail ratio models{.tabset}

### Model 8: Urchin Snail Ratio fishing model (glm)

```{r Checking variance between protected and non-protected for urch sn ratio, echo=FALSE}
leveneTest(urch_sn_ratio ~ protect_status, data = env_mod_data)
```
Variance is equal between protection status.


```{r Urchin ratio protection GLMM model}
urch_sn_ratio_protection_model_glmm <- glmmTMB(urch_sn_ratio ~ protect_status +
                          (1|region1), 
                          data = env_mod_data,
                          family = gaussian(link = "identity"))
```

```{r Urchin sn ratio protection mod glmm output, echo=FALSE}
#Checking resids
urch_sn_ratio_protection_model_glmm_resids <- simulateResiduals(urch_sn_ratio_protection_model_glmm)
plot(urch_sn_ratio_protection_model_glmm_resids)

#Plot residuals of random effects 
urch_sn_ratio_protection_model_glmm_ranef_values <- ranef(urch_sn_ratio_protection_model_glmm)$cond$region1
urch_sn_ratio_protection_model_glmm_ranef_values_numeric <- urch_sn_ratio_protection_model_glmm_ranef_values$`(Intercept)`

qqnorm(urch_sn_ratio_protection_model_glmm_ranef_values_numeric) #This is fine as far as I know 

hist(residuals(urch_sn_ratio_protection_model_glmm))
shapiro.test(residuals(urch_sn_ratio_protection_model_glmm))

#Looking at model results
summary(urch_sn_ratio_protection_model_glmm)
```

```{r Plotting the effect of protection status on urchin ratio in glm, echo=FALSE}
pred_urch_sn_ratio_protection_model_glmm <- ggpredict(urch_sn_ratio_protection_model_glmm, terms = "protect_status")

ggplot(pred_urch_sn_ratio_protection_model_glmm, aes(x = x, y = predicted))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3)+
  labs(y = "Predicted Log Urchin to Snail ratio", x = "Protection Status")+
  theme.figure()
```

##### Testing the residuals for autocorrelation
<div class="fold o"> 
```{r Testing urch sn ratio model linear residuals for spatial correlation, echo=FALSE, warning=FALSE}
#extract residuals from model
residuals_urch_sn_ratio_protection_model_glmm <- residuals(urch_sn_ratio_protection_model_glmm, type = "pearson")

#creating the spatial weights matrix
coords_for_auto_test <- st_as_sf(env_mod_data, coords = c("longitude", "latitude"), crs = 4326)
coords_for_auto_test_projected <- st_transform(coords_for_auto_test, crs = 3857)
extracted_coords_for_auto_test_projected <- st_coordinates(coords_for_auto_test_projected)


nb <- dnearneigh(extracted_coords_for_auto_test_projected, d1 = 0, d2 = 10)
#picked 10 m for distance so that it includes the transects at same sites

lw <- nb2listw(nb, style = "W", zero.policy = TRUE)


##Testing the residuals with Moran's I test and weight matrix
moran.test(residuals_urch_sn_ratio_protection_model_glmm, lw,
           alternative = "two.sided")

#Testing the residuals with inverse distance matrix as well to confirm
dist_matrix <- as.matrix(dist(extracted_coords_for_auto_test_projected))
inverse_dist_matrix <- 1 / (dist_matrix + 0.001)
diag(inverse_dist_matrix) <- 0
lw_inverse <- mat2listw(inverse_dist_matrix, style = "W")
moran.test(residuals_urch_sn_ratio_protection_model_glmm, lw_inverse,
           alternative = "two.sided")
```